---
title: "dataRetrieval_output_elements"
author: "Emi Fergus"
date: "6/10/2025"
format: html
editor: visual
---

## Issue
There are discrepancies with elements in the crosswalk table and elements returned with USGS's `dataRetrieval`. We will need to update the crosswalk table as a short-term solution because the incorrect schema names will not be updated for awhile most likely.

See .csv file 'result.dr.beta.legacy.profile.csv' to see elements in USGS's crosswalk table compared to elements returned with dataRetrieval.

## USGS dataRetrieval Elements

Comparing fields (elements) returned with `dataRetrieval::readWQPdata` `dataProfile` (see `readWQPdata` help for options for both Legacy and beta WQX3). 'Result' = Legacy; 'ResultWQX3' = WQX3 Beta

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(remotes)
library(devtools)
load_all()

library(tidyverse)
library(readr)
library(knitr)

library(dataRetrieval) # v2.7.18.9002
# Old version of dataRetrieval (temporary)
#remotes::install_github("DOI-USGS/dataRetrieval", ref = "develop", dependencies = TRUE, force = TRUE)
# to update EPATADA package to use this version of the package

```

# Compare elements from legacy and beta web services

## Example query

Example query using dataRetrieval::readWQPdata. USGS site (USGS-440504121435800). Comparing query output from the different versions of dataRetieval (legacy2.0 'resultPhysChem' vs beta3.0'fullPhysChem'). Note: Update (6/10/25 - beta services are returning WQX data (not just USGS) OLD 5/13/2025 the beta version of WQP was only returning USGS data, not WQX data. The Portal team is aware and working on it.

```{r}
#| echo: false
#| message: false
WQPquery2 <- list(siteid = "USGS-440504121435800"#, #Middle Green Lake (1980-1981)   McKenzie River at Springfield ("USGS-440508122581200")
                  #startDate = "1975-01-01",
                  #endDate = "1978-12-31",
                  #characteristicName = "Temperature, water"

)


################
# COMPARE OUTPUT FROM resultPhysChem and fullPhysChem
# LEGACY PORTAL data call (note: 'resultPhysChem' ONLY works with LEGACY)
result.legacy <- dataRetrieval::readWQPdata(WQPquery2,
                                           service = "Result", # beta "ResultWQX3"
                                           dataProfile = "resultPhysChem", #"fullPhysChem",
                                           ignore_attributes = T)

# BETA PORTAL data call (note different service and dataProfile) (n = 183 fields vs n = 141 with legacy fullPhysChem)
result.beta <- dataRetrieval::readWQPdata(WQPquery2,
                                          service = "ResultWQX3", # beta "ResultWQX3"
                                          dataProfile = "fullPhysChem", #"resultPhysChem", #"fullPhysChem",
                                          ignore_attributes = T)
# Make column headings a vector
legacy <- as.data.frame(colnames(result.legacy)) |> 
  rename("legacy" = 1) |> 
  mutate(WqxV2.FieldName = legacy)

beta <- as.data.frame(colnames(result.beta)) |> 
  rename("beta" = 1) |> 
  mutate(FieldName3.0 = beta)

#knitr::kable(legacy)
#knitr::kable(beta)
```

## Join Legacy and Beta field elements using crosswalk table

Read in WQX crosswalk table to join the legacy and beta field elements together to compare. Had to modify the crosswalk table to replace special characters in the legacy column (WqxV2.FieldName) so that they will match fields returned in dataRetrieval. For example, 'ActivityStartTime/Time' in crosswalk table vs 'ActivityStartTime.Time' from dataRetrieval.

```{r}
#| echo: false
#| message: false
## Updated crosswalk table (temporary)
## 4/3/25
wqxcrswlk <- readr::read_csv("C:/Users/efergus/OneDrive - Environmental Protection Agency (EPA)/a_WDIB/TADA/WQP_transition/Crosswalk_tables/Temp_Crswlk/wqxcrswlk_temp.csv")
# Need to create additional key fields so that they don't get 'overwritten'
# Need to rename field names to have "." instead of "/"
cols_wqx2 <- wqxcrswlk$WqxV2.FieldName

wqxcrswlk_mod <- wqxcrswlk |> 
  mutate(WqxV2.FieldName = str_replace_all(WqxV2.FieldName, c('/' = '\\.')))|> 
  mutate(crswlkFieldName3 = FieldName3.0,
         crswlkFieldName2 = WqxV2.FieldName)


# Full Join the legacy table to the crosswalk table - will return all x rows followed by unmatched y rows (this is true for legacy legacy profile e.g., 'LastUpdated' is not in the crosswalk table)
legacy_mod <- wqxcrswlk_mod |> 
  full_join(legacy, join_by(WqxV2.FieldName)) 

# Full join beta table to modified crosswalk table
all_mod <- legacy_mod |> 
  full_join(beta, join_by(FieldName3.0)) |> 
  rename(DR2legacy = legacy, DR3beta = beta) |> 
  select(crswlkFieldName3, crswlkFieldName2, DR2legacy, DR3beta)

# Filter fields returned with readWqpdata profiles (from 385 to 199)
# https://stackoverflow.com/questions/33520854/filtering-data-frame-based-on-na-on-multiple-columns/69395357
# all_red <- all_mod |> 
#   filter_at(vars(legacy,beta), any_vars(!is.na(.))) 

# WRITE MODIFIED TABLE
write_csv(all_mod, "C:/Users/efergus/OneDrive - Environmental Protection Agency (EPA)/a_WDIB/TADA/WQP_transition/dataRetrieval_output/result.dr.beta.legacy.profile.csv")

```

# Comparing elements returned from dataRetrieval versions

## Fields in DR beta but not in DR legacy or in crosswalk table

These are instances that may need to be manually corrected in the crosswalk table. They are fields returned in the DR beta results version but are not associated with a crosswalk field.

```{r}
#| echo: false
beta_only <- all_mod |> 
  filter(is.na(DR2legacy)) |> 
  filter(is.na(crswlkFieldName3)) |> 
  filter(is.na(crswlkFieldName2))

knitr::kable(beta_only)
```

'Result_ResultDetectionCondition' is an old version of the name that did not get updated in the WQP to 'Result_DetectionCondition'. The WQP team won't be able to address this for some time. Lee, Kevin, Hillary, and Emi met and decided in the mean time we will modify the crosswalk table to roll back to this old version of the name to be able to map schema elements between versions. It's easier to make changes to the crosswalk table than to make changes to the WQP schema services.

'GroupSummaryWeight_Measure' and 'GroupSummaryWeight_MeasureUnit' should be linked to 'ResultBiological_GroupSummaryWeightMeasure' and 'ResultBiological_GroupSummaryWeightMeasureUnit'(in WQX3.0 crosswalk).

'DataQuality_ResultComment' might be linked to 'Result_Comment' (in WQX3.0 crosswalk)?

'Activity_StartTimeZone_offset' and 'Activity_StartDateTime' are fields USGS creates in WQP and can be ignored.

## Fields not returned in the DR beta full results profile call but returned in DR legacy AND not in crosswalk

Fields that are not returned in the dataRetrieval beta (readWQPdata - 'fullPhysChem' profile). Many of these issues are because the crosswalk table has 'see Comment Field' for 2.0 elements that did not have a one-to-one match. They could be manually added to a modified crosswalk table. 

```{r}
#| echo: false

legacy_only <- all_mod |> 
  filter(is.na(DR3beta) & is.na(crswlkFieldName3) | crswlkFieldName2 == "see Comment field" & !is.na(DR3beta))

knitr::kable(legacy_only)


```

Sample Collection Method, Data Quality variables, Result Analytical Method and Lab Sample prep are variables that could be manualy assigned a crosswalk with 2.0schema names if appropriate. Some fields being returned might be due to a recently noticed glitch that the DR 2.0 call may be returning some 3.0 fields...but this needs to be explored more. Ex) BinaryObject elements I think are DR 3.0 not DR 2.0 but being returned with the (readWQPdata - 'resultPhysChem' profile).

'LastUpdated' is a field that is returned with the WQP summary service - this may also be a glitch to be returned with a results profile call? 'LastUpdated' could be matched with 'LastChangeDate' in the WQP3.0 schema and DR3.0.

## Fields that may need to be updated in the crosswalk table

Manually created table with fields that may need to be updated

```{r}
#| echo: false

fix <- data.frame(
  crswlkFieldName3 = c("Result_DetectionCondition",NA, NA, "ResultBiological_GroupSummaryWeightMeasure","ResultBiological_GroupSummaryWeightMeasureUnit","SampleCollectionMethod_Description","LastChangeDate", NA),
  crswlkFieldName2 = c("ResultDetectionConditionText",NA, NA,"GroupSummaryCountWeight/MeasureValue", "GroupSummaryCountWeight/MeasureUnitCode","see_Comment", NA, NA),
  DR3 = c("Result_ResultDetectionCondition", "Activity_StartDateTime","Activity_StartTimeZone_offset","GroupSummaryWeight_Measure","GroupSummaryWeight_MeasureUnit","SampleCollectionMethod_Description", "LastChangeDate", NA),
  DR2 = c("ResultDetectionConditionText","ActivityStartDateTime","ActivityStartTime.TimeZoneCode_offset",NA, NA,"SampleCollectionMethod.MethodDescriptionText", NA, "LastUpdated")
    
)

knitr::kable(fix)
```

## Crosswalk 2.0 Field Names with many-to-one

```{r}
#| echo: false
crs2_tochange <- all_mod |> 
  filter(crswlkFieldName2 == "see Comment field")

knitr::kable(crs2_tochange)
```

# Compare elements in Beta across different data profiles (fullPhysChem, narrow, basic)


```{r}
#| echo: false
#| message: false
WQPquery2 <- list(siteid = "USGS-440504121435800"#, #Middle Green Lake (1980-1981)   McKenzie River at Springfield ("USGS-440508122581200")
                  #startDate = "1975-01-01",
                  #endDate = "1978-12-31",
                  #characteristicName = "Temperature, water"

)


################
# COMPARE OUTPUT FROM resultPhysChem and fullPhysChem
# LEGACY PORTAL data call (note: 'resultPhysChem' ONLY works with LEGACY)
result.full <- dataRetrieval::readWQPdata(WQPquery2,
                                           service = "ResultWQX3", # beta "ResultWQX3"
                                           dataProfile = "fullPhysChem", #"fullPhysChem",
                                           ignore_attributes = T)

# BETA PORTAL data call (note different service and dataProfile) (n = 183 fields vs n = 141 with legacy fullPhysChem)
result.basic <- dataRetrieval::readWQPdata(WQPquery2,
                                          service = "ResultWQX3", # beta "ResultWQX3"
                                          dataProfile = "basicPhysChem", #"resultPhysChem", #"fullPhysChem",
                                          ignore_attributes = T)
# Make column headings a vector
full <- as.data.frame(colnames(result.full)) |> 
  rename("full" = 1) |> 
  mutate(FieldName3.0 = full)

basic<- as.data.frame(colnames(result.basic)) |> 
  rename("basic" = 1) |> 
  mutate(FieldName3.0 = basic)

# Full Join the legacy table to the crosswalk table - will return all x rows followed by unmatched y rows (this is true for legacy legacy profile e.g., 'LastUpdated' is not in the crosswalk table)
full_mod <- wqxcrswlk_mod |> 
  full_join(full, join_by(FieldName3.0)) 

# Full join beta table to modified crosswalk table
beta_all_mod <- full_mod |> 
  full_join(basic, join_by(FieldName3.0)) |> 
  rename(BetaFull = full, BetaBasic = basic) |> 
  select(crswlkFieldName3, crswlkFieldName2, BetaFull, BetaBasic)

# Filter fields returned with readWqpdata profiles (from 385 to 199)
# https://stackoverflow.com/questions/33520854/filtering-data-frame-based-on-na-on-multiple-columns/69395357
# all_red <- all_mod |> 
#   filter_at(vars(legacy,beta), any_vars(!is.na(.))) 

# WRITE MODIFIED TABLE
write_csv(beta_all_mod, "C:/Users/efergus/OneDrive - Environmental Protection Agency (EPA)/a_WDIB/TADA/WQP_transition/dataRetrieval_output/result.dr.beta.full.basic.profile.csv")


```



## Other WQP queries using beta services
```{r, echo=FALSE}
# USGS-OR n = 144
WQPquery3 <- list(organization = "CNENVSER",
                  startDate = "2018-01-01", 
                  endDate = "2018-01-31")
result.beta3 <- dataRetrieval::readWQPdata(WQPquery3,
                                          service = "ResultWQX3", # beta "ResultWQX3"
                                          dataProfile = "fullPhysChem",
                                          ignore_attributes = T)
```

```{r, echo=FALSE}
# Kootenai - TP n = 304 obs
WQPquery4 <- list(organization = "KOOTENAI_TRIBE_OF_IDAHO_WQX",
                  startDate = "2006-04-01", 
                  endDate = "2006-05-31",
                  characteristicName = "Total Phosphorus, mixed forms")

result.beta4 <- dataRetrieval::readWQPdata(WQPquery4,
                                          service = "ResultWQX3", # beta "ResultWQX3"
                                          dataProfile = "fullPhysChem",
                                          ignore_attributes = T)
```
