---
title: "Contributing"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Contributing}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Contribute to TADA!

We encourage you to read this project's
[CONTRIBUTING](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)
policy (you are here), its
[LICENSE](https://github.com/USEPA/TADA/blob/develop/LICENSE.md), and
its [README](https://usepa.github.io/EPATADA/).

We're so glad you're thinking about contributing to an EPA open source
project! If you're unsure about anything, just ask --- or submit your
issue or pull request anyway. The worst that can happen is we'll
politely ask you to change something. We appreciate all friendly
contributions.

No matter who you are, if you spot an error, omission, or bug, you're
welcome to open an issue in this repo!

## TADA Working Group Mission

To share and develop R code for evaluating and visualizing Water Quality
Portal (WQP) data more efficiently though collaboration and open-source
programming. This includes working together to find commonalities in
assessment processes across the nation, creating flexible tools that can
be easily customized to work within existing workflows, supporting each
other in learning R, and ensuring products will be accessible to
organizations most in need.

## Package Development

This article will walk through how to contribute to the TADA package via
a pull request workflow. This is also not a complete guide to R package
development (a comprehensive guide is [R
Packages](https://r-pkgs.org/)), instead this is meant as more of a
checklist for the general steps. Several references are included at the
bottom for more information on [R-package](https://r-pkgs.org/)
development and [git](https://happygitwithr.com/) workflows.

## What is GitHub?

GitHub is a third-party website that offers version-controlled
repositories that developers and scientists can use to collaborate on
projects (e.g., software, text, manuscripts, etc.) in real-time. GitHub
also provides social networking features that allow developers to follow
open-source projects, share code and learn how code changes are made
throughout the development process. GitHub is so named because it
utilizes the open-source version control system (VCS) known as Git.
There are multiple ways to interact with GitHub using Git.

-   Option 1: Interact with GitHub using a GUI instead of the command
    line or a web browser

    -   [Getting Started with GitHub
        Desktop](https://docs.github.com/en/desktop/overview/getting-started-with-github-desktop)

-   Option 2: Interact with GitHub using the command line or a web
    browser

    -   [Setting Up Git](https://docs.github.com/articles/set-up-git/)

    -   [Git Basics](https://git-scm.com/book/ch1-3.html)

    -   [Comprehensive Guide: Happy Git and GitHub for the
        useR](https://happygitwithr.com/ "Great and comprehensive guide for Git with an R flair")

## Required Installations

There are several programs that are needed before any work can begin.
With admin access to your computer, you can install all of these,
otherwise create a ticket with your IT group with the following
requests. The links provided assume a Windows computer. Adjustments
might be needed for Mac or Linux OS:

-   Recommend all three:

    -   [R](https://cran.r-project.org/bin/windows/base/)

    -   [RStudio](https://posit.co/download/rstudio-desktop/)

    -   [Rtools](https://cran.r-project.org/bin/windows/Rtools/)

-   Download both:

    -   [GitHub Desktop](https://desktop.github.com/)

    -   [Git](https://git-scm.com/downloads)

External Collaborators (using GitHub Desktop):

Once you have downloaded R, RStudio, Rtools, GitHub Desktop, and Git you
will need to create a GitHub account. Open GitHub Desktop and login in
with your account. In the web browser go the main TADA page
(<https://github.com/USEPA/TADA>) and click on the "Code" button.

Select the "Open With GitHub Desktop" option. When a pop-up window
appears in GitHub Desktop, click "Clone". As you do not have write
access to the TADA repository, you will be prompted to create a fork. In
the "How are you planning to use this fork?" window you should select
"To contribute to the parent project".

To make moving between branches while working in RStudio easy, create an
R Project. Open R Studio and the file menu select "New Project". In the
New Project Wizard select "Existing Directory" and select the file path
for the local repository. Name the R Project "TADA.RProj" as this file
name is already included in the .gitignore. Now when you open TADA.RProj
the file name should match the name of the branch you have selected in
GitHub Desktop. Any changes you make in R Studio and save will be saved
to the selected branch.

Once those have been installed, the following R packages will be needed
for R-package development work:

```{r, eval = F}
install.packages(c("devtools", "rmarkdown"))
```

## Issues

-   If you see an error or have feedback, the best way to let us know is
    to file an [issue](https://github.com/USEPA/TADA/issues).
-   Issues are labeled to help indicate what they are about. For
    example, we are using "Good First Issue" to indicate issues that
    might be good first pickings for your first contribution to this
    open-source project.
-   Pull requests can be directly
    [linked](https://docs.github.com/en/issues/tracking-your-work-with-issues/linking-a-pull-request-to-an-issue)
    to a specific issue. If linked, the Repository Administrators can
    more easily review the pull request and issue at the same time once
    a contributor submits the pull request. The issue can then be closed
    once the pull request is merged.

## Branches and Pull Requests

All new development currently happens in the **develop** branch.

To contribute a specific change or new code, outside contributors can
[create a new
fork](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/proposing-changes-to-your-work-with-pull-requests/creating-and-deleting-branches-within-your-repository),
using the **develop** branch as the source. We recommend that
contributors work from their personal fork, and on one specific "task"
at a time. When complete, you may submit a pull request to request that
your changes be merged from your fork into the TADA develop branch.
Contributors should submit separate pull requests for each "task".

"Tasks" should be small in scope. For example, they may pertain to a bug
fix or update relevant to a single function. A single "task" may also
encompass the same changes made across many functions if needed. Another
example of a single "task" could be to make changes to all documentation
to improve clarity, for example. Furthermore, a task may include
developing a new function, or a series of related functions. In some
cases, tasks can also be synonymous with issues, and the pull requests
can be directly linked to a specific issue (in that case, the Repository
Administrators will review the pull request and issue at the same time
and the issue can be closed once the pull request is merged).

Complete the pull request by detailing all fixes and contributions, and
tagging TADA repo admins who should review the work. For this package,
please tag cristinamullin (Cristina Mullin). Repository Administrators
will review code contributions from both internal EPA and external
collaborators and integrate code commits into source code. This is done
to ensure code stability and consistency and prevent degradation of code
performance. After review, the admin will either accept the submission,
recommend specific improvements to the submission, or in some cases
reject the submission.

To avoid issues, developers contributing code should contact the
repository admins (Cristina Mullin) early in the development process and
maintain contact throughout to help ensure the submission is compatible
with the code base and is a robust addition.

A few tips:

-   GitHub makes it easy to review Pull Requests for changes on each
    file, and to discuss the changes if needed using comments

-   We have set up [automated
    checks](https://github.com/USEPA/TADA/tree/develop/.github/workflows)
    (using Github Actions). Any new code included in Pull Requests must
    pass these checks before it can be merged into develop. These checks
    will run automatically when Pull Requests are submitted, and are
    ancillary to our content review, and help make sure the code is
    robust.

-   We recommend to run devtools::check() often during development.

-   Be descriptive with your commits and commit descriptions, but don't
    over-think it! All changes are tracked. Include comments and details
    within the package code itself to describe what the code does.

-   Always pull in upstream changes from the develop branch before
    adding a new commit on your branch.

-   Directly link Pull Request to Issues they address, if relevant.

## Additional References

-   [R Packages](https://r-pkgs.org/)

-   [testthat](https://testthat.r-lib.org/)

-   [R markdown](https://r4ds.had.co.nz/r-markdown.html)

## Open-Source Code Policy

Effective August 8, 2016, the [OMB Mandate: M-16-21; Federal Source Code
Policy: Achieving Efficiency, Transparency, and Innovation through
Reusable and Open Source
Software](https://www.whitehouse.gov/wp-content/uploads/legacy_drupal_files/omb/memoranda/2016/m_16_21.pdf)
applies to new custom-developed code created or procured by EPA
consistent with the scope and applicability requirements of Office of
Management and Budget's (OMB's) Federal Source Code Policy. In general,
it states that all new custom-developed code by Federal Agencies should
be made available and reusable as open-source code.

The EPA specific implementation of OMB Mandate M-16-21 is addressed in
the [System Life Cycle Management
Procedure](https://www.epa.gov/irmpoli8/policy-procedures-and-guidance-system-life-cycle-management-slcm).
EPA has chosen to use GitHub as its version control system as well as
its inventory of open-source code projects. EPA uses GitHub to inventory
its custom-developed, open-source code and generate the necessary
metadata file that is then posted to code.gov for broad reuse in
compliance with OMB Mandate M-16-21.

If you have any questions or want to read more, check out the [EPA Open
Source Project Repo](https://github.com/USEPA/open-source-projects).

## License

All contributions to this project will be released under the CCO-1.0
license file dedication. By submitting a pull request or issue, you are
agreeing to comply with this waiver of copyright interest.

## Disclaimer

This United States Environmental Protection Agency (EPA) GitHub project
code is provided on an "as is" basis and the user assumes responsibility
for its use. EPA has relinquished control of the information and no
longer has responsibility to protect the integrity, confidentiality, or
availability of the information. Any reference to specific commercial
products, processes, or services by service mark, trademark,
manufacturer, or otherwise, does not constitute or imply their
endorsement, recommendation or favoring by EPA. The EPA seal and logo
shall not be used in any manner to imply endorsement of any commercial
product or activity by EPA or the United States Government.

## Contact

If you have any questions, please reach out to the TADA Team
([mywaterway\@epa.gov](mailto:mywaterway@epa.gov){.email}).
