---
title: "Example Module 2 Workflow"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example Module 2 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
description: An example Module 2 workflow for creating an assessment unit/monitoring location crosswalk. This workflow considers (1) an optional user-supplied crosswalk, (2) any monitoring location data stored in ATTAINS, and (3) the results of TADA_CreateATTAINSAUML prioritized in the order listed to create a comprehensive assessment unit/monitoring location crosswalk.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get Montana TADA Data

Get data bacteria and pH data from Missoula County, Montana.

```{r tada.data}

# get MT data 
tada.MT <- TADA_DataRetrieval(
   startDate = "2020-01-01", endDate = "2022-12-31",
   statecode = "MT",
   characteristicName = c("Escherichia", "Escherichia coli", "pH"),
   county = "Missoula County",
   ask = FALSE)

# clean up data set (minimal)
  tada.MT.clean <- tada.MT %>%
    TADA_RunKeyFlagFunctions() %>%
    TADA_SimpleCensoredMethods()
  
# remove intermediate objects
  rm(tada.MT)

```


## Set up Demo User-Supplied Crosswalk

Get existing data from ATTAINS and subset a few rows to create a demo user-supplied crosswalk.

```{r demo.user, echo = FALSE}

# get crosswalk from ATTAINS
attains.existing.MT <- TADA_GetATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# clean existing crosswalk from ATTAINS to make sure monitoring location IDs are 
# WQP compatible
clean.existing.attains.MT <- TADA_UpdateATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# set up demo user supplied cw (select a few Monitoring Locations from the tada df)
user.supplied.cw <- clean.existing.attains.MT %>%
  dplyr::select(ATTAINS.AssessmentUnitIdentifier, 
                ATTAINS.MonitoringLocationIdentifier) %>%
  dplyr::filter(ATTAINS.MonitoringLocationIdentifier %in% c(
    "MDEQ_WQ_WQX-C04CKFKR05", "MDEQ_WQ_WQX-C04KNDYC01", "MDEQ_WQ_WQX-C04KNDYC02", 
    "MDEQ_WQ_WQX-C04KNDYC04", "MDEQ_WQ_WQX-C04KNDYC54")) %>%
  dplyr::rename(AssessmentUnitIdentifier = ATTAINS.AssessmentUnitIdentifier,
                MonitoringLocationIdentifier = ATTAINS.MonitoringLocationIdentifier)

rm(attains.existing.MT, clean.existing.attains.MT)
```

## Run TADA_CreateAUMLRef

Run TADA_CreateAUMLRef which can incorporate any combination of (1) (optional) user-supplied crosswalk, (2) crosswalk downloaded from ATTAINS, and (3) results from TADA_CreateATTAINSAUMLCrosswalk. The assignments are prioritized in the order listed. Any MonitoringLocations which remain unassigned after checking the user-supplied crosswalk will be subset and run with TADA_CreateATTAINSAUMLCrosswalk to attempt additional AssessmentUnit assignments.

To make function run more efficiently, catchment data are not currently being returned for the user-supplied crosswalk and the crosswalk downloaded from ATTAINS. We have plans to create additional functions/internal reference files to faciliate this in the future. Therefore, any catchment data available for mapping is currently from TADA_CreateATTAINSAUMLCrosswalk assessed MonitoringLocations only.

```{r tada.get.attains}

# use TADA get attains to make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLRef(tada.MT.clean,
                                 au_ref = user.supplied.cw,
                                 org_id = "MTDEQ", 
                                 add_catch = FALSE,
                                 batch_upload = TRUE)

```
## TADA_ViewATTAINS

The TADA_View ATTAINS map function has been edited to shows the results from all three different sources using different color circle markers (not final colors, we are very open to suggestions). The source has also been added to the popup window.
```{r view.map}

TADA_ViewATTAINS(MT.AUMLRef)

```

## TADA_UpdateATTAINSAUMLCrosswalk

After making any necessary changes to the ATTAINS_batchupload data frame generated by TADA_CreateAUMLRef, run TADA_UpdateATTAINSAUMLCrosswalk to add new links to WQP site pages.

```{r add.links}

batch.upload.MT <- MT.AUMLRef$ATTAINS_batchupload %>%
  TADA_UpdateATTAINSAUMLCrosswalk(# selected attains_replace = TRUE because all matches                                     currently in ATTAINS are included in this new                                          crosswalk
                                  attains_replace =TRUE,
                                  batch_upload = TRUE,
                                  wqp_data_links = "add",
                                  #ml ids have already  been corrected if needed
                                  update_mlid = FALSE,
                                  org_id = "MTDEQ")

```


