---
title: "Example Workflow"
author: "TADA Team"
date: "2025-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get Montana TADA Data

Get data bacteria and pH data from Montana.

```{r tada.data}

# get Montna bacteria data 
tada.MT <- TADA_DataRetrieval(
   startDate = "2020-01-01", endDate = "2022-12-31",
   statecode = "MT",
    characteristicName = c("Escherichia", "Escherichia coli", "pH"),
   ask = FALSE)

# review comparable data identifiers
sort(unique(tada.MT$TADA.ComparableDataIdentifier))

# clean up data set (minimal)
  tada.MT.clean <- tada.MT %>%
    TADA_RunKeyFlagFunctions() %>%
    TADA_SimpleCensoredMethods() %>%
  # could separate this out and add to harmonization table?  
  TADA_HarmonizeSynonyms()
  
# review comparable data identifiers
  sort(unique(tada.MT.clean$TADA.ComparableDataIdentifier))

```


## Retrieve existing monitoring location and assessment unit crosswalk from ATTAINS

Get existing data from ATTAINS

```{r existing.attains.crosswalk, echo = FALSE}

attains.existing.MT <- TADA_GetATTAINSAUSiteCrosswalk(org_id = "MTDEQ")
```

## Clean/update existing data from ATTAINS
```{r clean.existing.crosswalk}

clean.existing.attains.MT <- TADA_UpdateMonitoringLocationsInATTAINS(org_id = "MTDEQ")

```


## Run TADA_GetATTAINS

```{r tada.get.attains}

# use TADA get attains to make AU assignments for unassigned MLs
get.attains.MT <- TADA_GetATTAINS(tada.MT.clean, return_nearest = TRUE,
                                  au_ref = clean.existing.attains.MT)

```


o	User will review their user supplied list (either from TADA_GetATTAINSAUSite if they have submitted it to ATTAINS or as their own internal data frame)
•	Run TADA_ViewATTAINS
o	Decide if they want to keep any matches suggested by TADA_GetATTAINS.
o	TT has a map that seems to show these match types, but thinking through how we want to do this in the TADA package side? See TADAShinyJoinToAU tab 2 for more info.
•	Run TADA_UpdateMonitoringLocationsInATTAINS
o	Once a final list is reviewed and a user is comfortable with this, they can run this and have an easy way to submit to ATTAINS if desired.
•	Run TADA_CreateUseAURef (current PR)
o	From the user’s final sites to AU list, they must now decide how to assign Uses to these AUs they have defined.
o	Need to consider the best way to help assign uses to any NEW AUs (or if manual assignment is best). TADA_CreateWaterUseRef attempts to assist with this, but thinking through other ways to visualize this output on a map (Ex. AUs on a map, and we can hover over them to see if Uses have been assigned)

