---
title: "TADA Module 3: Alternative Options to Generate the Criteria and Methodology template for User Input"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA Module 3: Alternative Options to Generate the Criteria and Methodology template for User Input}
  %\VignetteEngine{knitr::rmarkdown}
description: A vignette (under development) showcasing alternative ways to generate the criteria and methodology file.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to download the EPATADA R package from GitHub, access
and parameterize several important functions, and create basic
visualizations with a sample data set.

**Note: EPATADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.**

**If you are interested in contributing to EPATADA development, more
information is available at:**

[**Contributing**](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)

**We welcome collaboration with external partners.**

## Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load the EPATADA R Package using the remotes package.
Dependency packages will also be downloaded automatically from CRAN. You
may be prompted in the console to update dependencies that have more
recent versions available. If you see this prompt, it is recommended to
update all of them (enter 1 into the console).

```{r install_TADA, eval = T, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)

# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r dev_install, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "Create-Module-3-AURef",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the EPATADA R Package
into your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Help pages

All TADA R package functions have their own individual help pages,
listed on the [Function
reference](https://usepa.github.io/EPATADA/reference/index.html) page on
the GitHub site. Users can also access the help page for a given
function in R or RStudio using the following format (example below):
`?[name of TADA function]`

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

## Module 3 Functions in TADA

Disclaimer: The EPATADA Module 3 functions were designed to: (1) assist
users with associating Water Quality Portal monitoring locations with
assessment units and designated uses from ATTAINS and (2) compare Water
Quality Portal results with numeric water quality criteria. EPATADA
functions do not constitute current EPA policy or regulatory
requirements. Organizations may choose to use EPATADA as a a tool in
their decision making processes. Use of EPATADA is not required.

## Get WQP Monitoring Data in Montana Using TADA_DataRetrieval()

Get bacteria and pH data from Missoula County, Montana.

```{r tada-data}
# get MT data
tada.MT <- TADA_DataRetrieval(
  startDate = "2020-01-01",
  endDate = "2022-12-31",
  statecode = "MT",
  characteristicName = c(
    "Escherichia",
    "Escherichia coli",
    "pH"
  ),
  county = "Missoula County",
  ask = FALSE
)

# clean up data set (minimal)
tada.MT.clean <- tada.MT %>%
  TADA_RunKeyFlagFunctions() %>%
  TADA_SimpleCensoredMethods() %>%
  TADA_HarmonizeSynonyms()

# remove intermediate objects
rm(tada.MT)

# or uncomment the code below and load internal copy of TADA df from EPATADA
# tada.MT.clean <- Data_MT_MissoulaCounty
```

# Defining Criteria - Magnitude Methodology

Users can proceed with a few different options for generating their
criteria table

1)  Generate a blank criteria and methods table and fill it out from
    scratch.

2)  Provide a user-supplied criteria and methods table partially/fully
    filled out.

    A.) The default option in this scenario will display all unique
    **TADA.ComparableDataIdentifiers** (or WQP CharacteristicName) in
    your TADA/WQP data frame to ensure you review any missing WQP
    Characteristic, speciation and fraction combinations.

    B.) Alternatively, users can choose to display all unique TADA
    Characteristic name rather than TADA.ComparableDataIdentifier. In
    this scenario, each ATTAINS.ParameterName in the analysis summary
    output will be grouped to any of these TADA/WQP CharacteristicName,
    unless a fraction or speciation is defined.

3)  Users can also choose to provide an autofill option, which will help
    to fill out any missing rows with ATTAINS.ParameterName and
    ATTAINS.UseName that are pulled in from ATTAINS as the default.

    A.) If a user has supplied a list of new or updated use names to AU
    that may not be retrievable from the prior ATTAINS assessment cycle,
    they should provide a useAURef crosswalk table in this function.
    This should only be provided when auto_assign = TRUE.

4)  (Recommended) Go through the step-by-step review process with the 3
    TADA crosswalk reference file generation for TADA_CreateParamRef,
    TADA_CreateUseParamRef, and TADA_CreateMLSummaryRef. This vignette
    does not go through this recommended workflow. Please see
    ExampleMod3Workflow.Rmd for this guided workflow.

Each option will allow the option to append additional rows to summarize
EPA304a recommended standards, if one has been defined. Please contact
the TADA team if you believe there are additional entries or
modification to these defined standards.

## Option A: Fully blank template

A blank template is generated. This can be generated and filled out in
the excel file.

```{r blank template output}
MT.Criteria_blank <- TADA_DefineCriteriaMethodology(
  tada.MT.clean, # remove this as an arg input
  org_id = "MTDEQ", # can remove this too
  # auto_assign = FALSE,
  excel = FALSE
  # uncomment to run the excel file
  # excel = TRUE, overwrite = TRUE
)
```

## Option B: Auto Fill option (Intermediate Tabs are Hidden)

You can also generate auto_assigned value(s) of ATTAINS.ParameterName,
ATTAINS.UseName to TADA.CharacteristicName using default options. Users
should be aware that this will only return rows for any matching values
from a WQP characteristic to ATTAINS parameter alias table. It is likely
that these value(s) will require a thorough review process during each
step of the process with the recommended workflow of
TADA_CreateParamRef, TADA_CreateUseParamRef and TADA_CreateMLSummaryRef.

Users can view the output of these 3 functions in the excel spreadsheet
if desired. They are hidden as default (NOTE TO CONSIDER: Should these
be kept hidden? The goal of this output has been to focus on an
auto_assign option as a quick get around to needing to fill out the
table from scratch, but if users may find it worthwhile, we can always
show these hidden tabs to allow for an easier review process and
updating of tables directly in the excel file.)

```{r auto-fill-output}
MT.Criteria_autofill <- TADA_DefineCriteriaMethodology(
  tada.MT.clean,
  org_id = "MTDEQ",
  auto_assign = TRUE,
  # displayUniqueId = FALSE,
  # excel = FALSE
  # uncomment to run the excel file
  excel = TRUE, overwrite = TRUE
)
```

Users who would like to ensure all Characteristic, Speciation and
Fractions are being considered, can use displayUniqueId = TRUE to show
all unique TADA.ComparableDataIdentifier(s) shown as explicit crosswalk.
Note: This may generate many additional rows if your WQP data results
are not harmonized or if there are many different combinations of
Characteristic, Speciation and Fractions to consider.

```{r auto-fill-display-TADA.ComparableDataIdentifier}
MT.Criteria_autofill_w_uniqueID <- TADA_DefineCriteriaMethodology(
  tada.MT.clean,
  org_id = "MTDEQ",
  auto_assign = TRUE,
  displayUniqueId = TRUE,
  excel = FALSE
  # uncomment to run the excel file
  # excel = TRUE, overwrite = TRUE
)
```

We can also choose to append epa304a recommended standards into the
criteria table for any WQP characteristics in your data frame that are
found.

```{r auto-fill-epa304a-and-display-TADA.ComparableDataIdentifier}
MT.Criteria_autofill_w_uniqueID <- TADA_DefineCriteriaMethodology(
  tada.MT.clean,
  org_id = "MTDEQ", 
  auto_assign = TRUE, 
  displayUniqueId = TRUE, 
  excel = FALSE
  # uncomment to run the excel file
  # excel = TRUE, overwrite = TRUE
)
```

## Option C: User Supplied Table

A user has a completed or partially filled criteria file, let's use
MTDEQ as our example org. MTDEQ should thoroughly review this table and
determine if there are values that needs to be fixed or if there are
missing WQP Char to consider that isn't defined in their criteria and
methods table that they have supplied. Users will be warned how many WQP
Char values are not defined from their user supplied table.

In this first example, a user supplies their own criteria table. The user 
supplied table is prioritized. Any
missing WQP/TADA.CharacteristicName will be matched from ATTAINS based
on the auto_assign = TRUE option.

Note: If a user has an updated list of use names that have been applied
to and have chosen to apply the auto_assign = TRUE input, they should
also provide a useAURef input. Otherwises the uses will be pulled in
from the prior ATTAINS assessment cycle.

```{r user-supplied-epa304a-and-autofill-output}
MT.Criteria_user_supplied_autofill <- TADA_DefineCriteriaMethodology(
  .data = tada.MT.clean,
  criteriaMethods = criteria_table, # user supplied table - all rows are kept from this table
  org_id = "MTDEQ",
  auto_assign = TRUE, 
  useAURef = MT.UseAURef_with_WaterUseRef, 
  displayUniqueId = FALSE,
  epa304a = TRUE,
  excel = FALSE
  # uncomment to run the excel file
  # excel = TRUE, overwrite = TRUE
)
```

Users will need to determine their level of desired grouping of
TADA.CharacteristicName by aggregation. If a user has not gone through
the review process with the 3 crosswalk reference files, they need to
specify what combinations of fraction and speciations falls under an
ATTAINS.ParameterName and ATTAINS.UseName combination. By specifying
displayUniqueId = TRUE, this will display all combinations in the
criteria table output.

```{r display-TADA.ComparableDataIdentifier-user-supplied-epa304a-and-autofill-output}
# Will display all unique rows of TADA.Characteristic Name to ATTAINS ParameterName and ATTAINS UseName
MT.Criteria_user_supplied_autofill2 <- TADA_DefineCriteriaMethodology(
  .data = tada.MT.clean,
  criteriaMethods = criteria_table, # user supplied table - all rows are kept from this table
  org_id = "MTDEQ",
  auto_assign = TRUE, 
  displayUniqueId = TRUE, # will display all unique TADA.ComparableDataIdentifier in this table.
  epa304a = TRUE,
  excel = FALSE
  # uncomment to run the excel file
  # excel = TRUE, overwrite = TRUE
)
```

## Choose a Final Criteria Template, Save and Re-use

```{r save-and-re-use-your-most-up-to-date-criteria-table}
# Save the criteria table of your liking to be used for your next analysis needs.
# TADA_CreateCSV(MT.Criteria_user_supplied_autofill2)

# We can now reuse this criteria table
MT.Criteria_reuse <- TADA_DefineCriteriaMethodology(
  tada.MT.clean,
  criteriaMethods = MT.Criteria_user_supplied_autofill2, # user supplied table - all rows are kept from this table
  org_id = "MTDEQ",
  auto_assign = FALSE,
  displayUniqueId = FALSE,
  excel = FALSE
  # uncomment to run the excel file
  # excel = TRUE, overwrite = TRUE
)
```

Users are recommended to go through each of the 3 reference files one at
a time though in their review process. In this case, a user should
provide a MLSummaryRef file function input and turn the auto_assign
option to FALSE. Please see ExampleMod3Workflow.Rmd vignette for the
step by step process.
