---
title: "Example Module 2 Workflow"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example Module 2 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
description: An example Module 2 workflow focused on A) creating an ATTAINS assessment unit/WQP monitoring location crosswalk, B) reviewing assessment units and monitoring locations on a map, C) submitting an up to date assessment unit/WQP monitoring location crosswalk to ATTAINS, and D) assigning uses to assessment units.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview and Setup

This example Module 2 workflow focuses on the following tasks:

A)  **Creating a ML/AU Crosswalk**: Use **`TADA_CreateAUMLCrosswalk`**
    to establish a crosswalk between ATTAINS assessment units and WQP
    monitoring locations.

B)  **Map Review**: Utilize **`TADA_ViewATTAINS`** to review assessment
    units and monitoring locations on a map, ensuring accuracy and
    completeness.

C)  **Submitting the Crosswalk**: Employ
    **`TADA_UpdateATTAINSAUMLCrosswalk`** to submit an up-to-date
    assessment unit/WQP monitoring location crosswalk to ATTAINS.

D)  **Assigning Uses**: Use **`TADA_CreateUseAURef`** to assign specific
    uses to the assessment units.

### Install and Load the EPATADA R Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
```

Load the remotes library:

```{r load_remotes}
library(remotes)
```

Next, install and load EPATADA using the remotes package. USGS's
dataRetrieval and other TADA R Package dependencies will also be
downloaded automatically from CRAN with the TADA install. If desired,
the development version of dataRetrieval can be downloaded directly from
GitHub (un-comment).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)

# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r dev_install, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "Create-Module-3-AURef",
  dependencies = TRUE
)
```

Finally, use the library() function to load the EPATADA R Package and
other required packages/

```{r library, results = 'hide'}
library(EPATADA)
```

### Get WQP Monitoring Data in MT

Get bacteria and pH data from Missoula County, Montana.

```{r tada.data, eval = F}
# get MT data
tada.MT <- TADA_DataRetrieval(
  startDate = "2020-01-01",
  endDate = "2022-12-31",
  statecode = "MT",
  characteristicName = c(
    "Escherichia",
    "Escherichia coli",
    "pH"
  ),
  county = "Missoula County",
  ask = FALSE
)

# clean up data set (not comprehensive)
tada.MT.clean <- tada.MT %>%
  TADA_RunKeyFlagFunctions() %>%
  TADA_SimpleCensoredMethods() %>%
  TADA_HarmonizeSynonyms() 

# Remove intermediate objects
rm(tada.MT)
```

This example is also included in the EPATADA R package, and can be
loaded into your R environment like this:

```{r load WQP data}
tada.MT.clean <- Data_MT_MissoulaCounty
```

## Step A: Discover ML/AU Matches

The **`TADA_CreateAUMLCrosswalk`** function efficiently creates a
crosswalk between ATTAINS assessment units and WQP monitoring locations.
It uses three prioritized data sources:

1.  **User-Supplied Crosswalk:** An optional crosswalk provided by the
    user (e.g., see example **`user.supplied.cw`** in next code chunk).

2.  **ATTAINS Crosswalk:** Utilizes **`TADA_GetATTAINSAUMLCrosswalk`**
    to incorporate crosswalk information stored by participating
    organizations in ATTAINS.

3.  **ATTAINS Catchments/Geospatial Join:** Employs
    **`TADA_CreateATTAINSAUMLCrosswalk`** to connect monitoring
    locations to assessment units using ATTAINS catchments through a
    geospatial join.

**Process**

-   The function prioritizes these sources in the order listed. It
    automatically attempts to assign unassigned monitoring locations
    using the next available data source. For example, if any monitoring
    locations remain unassigned after checking the user-supplied
    crosswalk, the function will check the ATTAINS crosswalk, then the
    geospatial join, as needed.

-   This entire process is automated within
    **`TADA_CreateAUMLCrosswalk`**, so users do not need to run each
    step individually, though they have the option to do so if desired.

**Current Limitations**

-   To make this function run more efficiently, catchments from the
    ATTAINS geospatial service are not returned for matches found in the
    user-supplied crosswalk or ATTAINS crosswalk (#1 or #2 above).
    ATTAINS catchments are only returned from
    `TADA_CreateATTAINSAUMLCrosswalk` assessed WQP monitoring locations.

**Future Improvements**

-   Plans are underway to develop a separate process to return all
    ATTAINS catchments associated with all WQP monitoring locations in
    the future, to support mapping and crosswalk review.

### Create An Example User-Supplied Crosswalk

Get existing data from ATTAINS using `TADA_GetATTAINSAUMLCrosswalk` and
subset a few rows to create an example user-supplied ATTAINS Assessment
Unit and WQP Monitoring Location crosswalk for demonstration purposes.
If you have your own crosswalk, this step can be skipped.

```{r demo.user, results = 'hide'}
# review valid ATTAINS org IDs
ATTAINS_orgs <- rExpertQuery::EQ_DomainValues("org_id")

# get crosswalk from ATTAINS
attains.existing.MT <- TADA_GetATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# clean existing crosswalk from ATTAINS to make sure WQP monitoring location IDs pulled from ATTAINS are WQP compatible (adds org ID if missing)
clean.existing.attains.MT <- TADA_UpdateATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# create example user supplied crosswalk (select a few Monitoring Locations from the tada df to use in the example for demonstration purposes)
user.supplied.cw <- clean.existing.attains.MT %>%
  dplyr::select(
    ATTAINS.AssessmentUnitIdentifier,
    ATTAINS.MonitoringLocationIdentifier,
    ATTAINS.WaterType
  ) %>%
  dplyr::filter(ATTAINS.MonitoringLocationIdentifier %in% c(
    "MDEQ_WQ_WQX-C04CKFKR05", "MDEQ_WQ_WQX-C04KNDYC01", "MDEQ_WQ_WQX-C04KNDYC02",
    "MDEQ_WQ_WQX-C04KNDYC04", "MDEQ_WQ_WQX-C04KNDYC54"
  )) %>%
  dplyr::rename(
    AssessmentUnitIdentifier = ATTAINS.AssessmentUnitIdentifier,
    MonitoringLocationIdentifier = ATTAINS.MonitoringLocationIdentifier,
    WaterType = ATTAINS.WaterType
  ) %>%
  
  # Add an example new assessment unit for demonstration purposes
  dplyr::bind_rows(c(
    AssessmentUnitIdentifier = "NEW:EX_MDEQ_WQ_WQX",
    MonitoringLocationIdentifier = "NARS_WQX-NWC_MT-10184",
    WaterType = "LAKE, FRESHWATER"
  ))


rm(attains.existing.MT, clean.existing.attains.MT, ATTAINS_orgs)
```

### Run `TADA_CreateAUMLCrosswalk`

```{r tada.get.attains}
# make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLCrosswalk(
  tada.MT.clean,
  au_ref = user.supplied.cw,
  org_id = "MTDEQ",
  add_catch = FALSE,
  nhd_catch = FALSE,
  batch_upload = TRUE)
```

## Step B: Review MLs/AUs on a Map

The **`TADA_ViewATTAINS()`** function creates a map that visually
represents results from three different crosswalk sources using distinct
circle markers:

-   **User-Supplied Crosswalk**: Monitoring locations derived from a
    user-supplied reference are indicated by a user icon.

-   **ATTAINS Crosswalk**: Locations that match with the crosswalk
    downloaded from ATTAINS are marked with a circle that contains a
    check mark.

-   **ATTAINS Catchments/Geospatial Join**: Matches found by joining
    monitoring locations with ATTAINS catchments are shown as solid fill
    circle markers.

Additionally, when a user clicks on any circle marker, a pop-up window
appears displaying the assessment unit crosswalk source for that
particular monitoring location.

```{r view.map}
TADA_ViewATTAINS(MT.AUMLRef, ref_icons = TRUE)
```

## Step C: Upload ML/AU crosswalk to ATTAINS

The **`TADA_UpdateATTAINSAUMLCrosswalk`** function generates an output
that can be batch uploaded to ATTAINS, enabling the creation or updating
of Monitoring Location Identifiers within ATTAINS Assessment Unit
profiles.

**Workflow Example:**

1.  **Access the Element**:\
    Begin by accessing the **`ATTAINS_batchupload`** element from the
    list generated by the **`TADA_CreateAUMLCrosswalk`** function (e.g.
    MT.AUMLRef).

2.  **Review and Modify**:\
    Review the **`ATTAINS_batchupload`** element and make any necessary
    modifications.

3.  **Execute the Function**:\
    Execute **`TADA_UpdateATTAINSAUMLCrosswalk`**. This process adds new
    links to WQP site pages, preparing the data for upload to ATTAINS.

4.  **Choose Update Option**:\
    When running **`TADA_UpdateATTAINSAUMLCrosswalk()`**, with the
    `attains_replace` function input users have the option to:

    -   **Overwrite**: Replace all existing records.

    -   **Append**: Add new Monitoring Location Identifiers to the
        current records.

For more detailed instructions, enter `?TADA_UpdateATTAINSAUMLCrosswalk`
into the console.

```{r add.links}
batch.upload.MT <- MT.AUMLRef$ATTAINS_batchupload %>%
  TADA_UpdateATTAINSAUMLCrosswalk( # selected attains_replace = TRUE because all matches currently in ATTAINS are included in this new crosswalk
    attains_replace = TRUE,
    batch_upload = TRUE,
    wqp_data_links = "add",
    # ml ids have already  been corrected if needed
    update_mlid = FALSE,
    org_id = "MTDEQ"
  )
```

## Step D: Assign Uses to AUs

After mapping the WQP monitoring locations (MLs) to ATTAINS assessment
units (AUs), we will use the `TADA_CreateUseAURef` function to assign
specific uses to these assessment units. Additionally, we will
demonstrate how to automatically assign uses to a new assessment unit,
`NEW:EX_MDEQ_WQ_WQX`, which is not yet present in ATTAINS. This will be
done by associating it with the water type "LAKE, FRESHWATER,"
showcasing the process of assigning uses based on water type.

```{r get.MLAURef}
Final.MT.AUMLRef <- MT.AUMLRef$ATTAINS_crosswalk %>%
  dplyr::mutate(
    ATTAINS.WaterType = 
      dplyr::case_when(
        ATTAINS.AssessmentUnitIdentifier == 
          "NEW:EX_MDEQ_WQ_WQX" ~ "LAKE, FRESHWATER",
        TRUE ~ ATTAINS.WaterType))
```

Now, we will assign uses to each unique assessment unit (AU) in our
finalized monitoring location to assessment unit crosswalk (e.g.,
`Final.MT.AUMLRef`). The process of assigning uses to AUs follows these
prioritized steps:

1.  User-Supplied Crosswalk (Optional): Utilize a crosswalk provided by
    the user that maps specific uses to assessment units.

2.  Existing Uses from ATTAINS: Import existing uses for assessment
    units as defined by ATTAINS organizations.

```{r assign.use.to.AU}
MT.UseAURef <- TADA_CreateUseAURef(
  AUMLRef = Final.MT.AUMLRef, 
  org_id = "MTDEQ")
```

### **Assigning Uses to New AUs**

For any new assessment unit (AU), users must determine how to assign the
appropriate uses. The **`TADA_CreateWaterUseRef`** function is designed
to assist with this process. It imports combinations of water types and
use names associated with the specified organization from ATTAINS. By
doing so, it automatically assigns uses to new assessment units based on
their water type, thereby streamlining the integration of new AUs into
existing frameworks.

For more information on how to customize this function to suit your
needs, enter **`?TADA_CreateWaterUseRef`** into the R console.

```{r assign.use.to.new.AU.with.waterUseRef}
MT.UseAURef_with_WaterUseRef <- 
  TADA_CreateUseAURef(
    waterUseRef = TADA_CreateWaterUseRef(org_id = "MTDEQ"),
    AUMLRef = Final.MT.AUMLRef,
    org_id = "MTDEQ")
```

Users also have the option to manually assign use names to new AUs if
they prefer not to use the TADA_CreateWaterUseRef function. This
approach allows for greater flexibility and customization when
integrating new AUs into existing frameworks.

```{r assign.use.to.new.AU.manual}
MT.UseAURef_manual <- MT.UseAURef %>%
  dplyr::left_join(
    data.frame(
      ATTAINS.UseName = c("Aquatic Life", "Drinking Water"),
      ATTAINS.AssessmentUnitIdentifier = c("NEW:EX_MDEQ_WQ_WQX", "NEW:EX_MDEQ_WQ_WQX")
    ),
    by = ("ATTAINS.AssessmentUnitIdentifier")
  ) %>%
  dplyr::mutate(ATTAINS.UseName = dplyr::coalesce(ATTAINS.UseName.x, ATTAINS.UseName.y)) %>%
  dplyr::select(-ATTAINS.UseName.x, -ATTAINS.UseName.y)
```

## Summary

In conclusion, the "Example Module 2 Workflow" significantly enhances
the ability to maintain up-to-date information on WQP Monitoring
Locations and the uses being assessed at various ATTAINS Assessment
Units. By systematically guiding users through the creation of
crosswalks, map reviews, data submission, and use assignments, this
workflow supports efficient data management and improves
reproducibility.
