---
title: "Example Module 3 Workflow"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example Mod 3 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
description: Recommended TADA module 3 Criteria and Methodology Table generation (under development) showcasing water quality data analysis functions and example workflows.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup-knitr, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = FALSE
)
```

# Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language. The
EPATADA R Package is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to user
feedback. We appreciate your feedback, patience, and interest in these
helpful tools. If you are interested in contributing to TADA
development, more information is available
[here](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html). We
welcome collaboration with external partners!

# Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing the EPATADA R package because it is only
available on GitHub.

```{r setup-remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

```{css, include = F}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

Next, install and load the EPATADA R Package using the remotes package.
Dependency packages will also be downloaded automatically from CRAN. You
may be prompted in the console to update dependencies that have more
recent versions available. If you see this prompt, it is recommended to
update all of them (enter 1 into the console).

```{r install-develop-branch, eval = T, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

```{r install-current-branch, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the EPATADA R Package
into your R session.

```{r load-EPATADA, results = 'hide'}
library(EPATADA)
```

All EPATADA R package functions have their own individual help pages
(see [Function
Reference](https://usepa.github.io/EPATADA/reference/index.html)). Users
can also access function help pages from RStudio by entering
`?[name of TADA function]` into the console (example below).

```{r help-pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

# Introduction to TADA Module 3

This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to create WQP and ATTAINS crosswalks that are needed
prior to defining and capturing organization specific water quality
analysis criteria and methodologies. This is a continuation of an
example workflow with ExampleMod2Workflow.Rmd.

Specifically, this vignette provides an overview of two functions that
can assist users with:

-   Creating a crosswalk (reference table) between ATTAINS parameter
    names and WQP/TADA characteristic names.

-   Creating a crosswalk (reference table) of all unique combinations of
    ATTAINS uses and parameters applicable to ATTAINS Assessment Units
    of interest.

-   Define and summarize your spatial level of analysis (ML versus AU)
    and capture any unique site-specfic WQS criteria needs for your
    organization.

These Module 3 functions are being designed with the flexibility for
users from states, tribes, or territories to input organization-specific
information (and link directly to information in ATTAINS). While TADA
functions help generate these crosswalks, users must review and modify
the tables generated in each step of the process to ensure their
accuracy. Additionally, the TADA team has also incorporated national
recommended Clean Water Act (CWA) 304(a) numeric criteria for optional
use in Module 3 functions by leveraging data from EPA's [Criteria Search
Tool
(CST)](https://www.epa.gov/wqs-tech/state-specific-water-quality-standards-effective-under-clean-water-act-cwa).
In the next vignette, TADAModule3_PartB , which is coming soon, this
will allow users to analyze their data against the 304(a) criteria; and
to easily compare results when when using 304(a) criteria vs. their own
organization's criteria.

# Example data from WQP

Let's start with an example data frame that has gone through data
cleaning, wrangling, harmonization, handling of censored data, removal
of suspect results, and other important TADA Module 1 functions. This
process should be completed by users before utilizing Module 3
functions.

```{r load-data}
# import example data set, example AUML crosswalk and example uses to AU crosswalk.
# load("C:/Users/kwong01/EPATADA/inst/extdata/tada.MT.clean.rda")
# load("C:/Users/kwong01/EPATADA/inst/extdata/Final.MT.AUMLRef.rda")
# load("C:/Users/kwong01/EPATADA/inst/extdata/MT.UseAURef.rda")
```

## Get WQP Monitoring Data in Montana Using TADA_DataRetrieval()

Get bacteria and pH data from Missoula County, Montana.

```{r tada.data}

# get MT data 
tada.MT <- TADA_DataRetrieval(
   startDate = "2020-01-01", 
   endDate = "2022-12-31",
   statecode = "MT",
   characteristicName = c("Escherichia", 
                          "Escherichia coli", 
                          "pH"),
   county = "Missoula County",
   ask = FALSE)

# clean up data set (minimal)
  tada.MT.clean <- tada.MT %>%
    TADA_RunKeyFlagFunctions() %>%
    TADA_SimpleCensoredMethods()
  
# remove intermediate objects
  rm(tada.MT)

```

This example will focus only on Montana. The remainder of this vignette
will walk through how to fill out two crosswalk tables
(TADA_CreateParamRef and TADA_CreateUseParamRef) that are needed before
we can start assigning applicable criteria and methodologies information
for this specific ATTAINS organization (MTDEQ).

# ATTAINS Domains and Allowable Values

A crosswalk between TADA characteristics and ATTAINS parameters is
needed before we can integrate information from these two data sets.
Before creating the TADA/WQP CharacteristicName and ATTAINS parameter
crosswalk table (TADA_CreateParamRef), let's review all the unique
TADA.ComparableDataIdentifier's (characteristic, fraction and speciation
combinations) in the example data. How many results are available for
each within the example data set (Data_NCTC)?

```{r comparabledataID-counts}
# create table with counts of TADA.ComparableDataIdentifiers
TADA_FieldValuesTable(tada.MT.clean, field = "TADA.ComparableDataIdentifier")
```

In this vignette, we will crosswalk each of these
TADA.ComparableDataIdentifier's to ATTAINS parameter names. ATTAINS has
multiple domains, including the parameter names domain, that have
allowable values. Using the R package rExpertQuery, which provides
functions for downloading tidy data from the ATTAINS public web services
(<https://github.com/USEPA/rExpertQuery>), we can review all ATTAINS
domains (see example below).

```{r ATTAINS-domains}
# return ATTAINS domain names
rExpertQuery::EQ_DomainValues()
```

ParameterName is one of the ATTAINS domains that has a specific list of
allowable values. Keep in mind that the ATTAINS parameters domain list
only includes parameters that have been listed as causes (of water
quality impairment) in prior ATTAINS CWA assessment cycles. Let's review
the complete list:

```{r ATTAINS-params}
# return ATTAINS parameter domain values
TADA_TableExport(rExpertQuery::EQ_DomainValues("param_name"))
```

In the next section, we will review which parameters have been listed in
ATTAINS in the past for a specific organization. In order to select a
specific organization in the TADA_CreateParamRef() function, we must
first review the ATTAINS organization domain list to get the ATTAINS
organization id's to use for the function input.

```{r ATTAINS-orgs}
# return ATTAINS organization domain values
TADA_TableExport(rExpertQuery::EQ_DomainValues("org_id"))
```

# TADA_CreateParamRef() Basics

ATTAINS Parameter Names are more general than the TADA Comparable Data
Identifiers. Therefore, we recommend users manually provide a crosswalk.
TADA_CreateParamRef() can help generate this crosswalk, but if desired,
users can provide a pre-filled crosswalk with the required column names
("TADA.ComparableDataIdentifier" and "ATTAINS.ParameterName").
TADA_CreateParamRef() includes an argument input 'auto_assign' to assist
users in doing an exact match between TADA Characteristic Names to
ATTAINS Parameter Name. There are roughly 598/1180 unique ATTAINS
Parameter Names that have an exact match. \*\*It is important to note
that even with these exact matches, there are additional complexities
related to the fraction, speciation and units that may be unique to your
organization. For example, some organizations may consider any of
Nitrate/Nitrite, Ammonia as "Nitrogen" as the ATTAINS.ParameterName and
may not consider the speciation.

Let's go through a few examples.

TADA_CreateParamRef() creates a template which includes all
TADA.ComparableDataIdentifiers from the TADA data frame and provides a
blank column, "ATTAINS.ParameterName" for users to input the
corresponding ATTAINS parameter name.

Setting "excel = TRUE" in TADA_CreateParamRef tells the function to
create an Excel spreadsheet of the reference table for easy user review
and editing. The excel output includes logic to help a user select
ATTAINS parameter names that have been used in the past by the selected
organization. If users do not want to work in Excel, they can use "excel
= FALSE" to return a data frame. The data frame can be edited directly
in R if desired, although there are no TADA-specific functions designed
to facilitate this.

When using TADA_CreateParamRef(), users should specify the
organization(s) of interested in the "org_names" argument. This ensures
that the correct number of rows will be created in the reference table.
When "excel = TRUE", specifying the organization(s) also creates a
separate tab in the Excel file which contains the parameter names used
in prior assessment cycles by the specified ATTAINS organization(s).
This will allow users to decide whether to continue to use the same
ATTAINS parameter names from prior assessments that their
organizations(s) have used in the past, or to use other valid parameter
names from the entire ATTAINS domain list. If there is not a suitable
ATTAINS parameter name to match a TADA.ComparableDataIdentifier, users
should contact the ATTAINS team (attains\@epa.gov) for assistance.

For our first example, let's start of with a single organization. In
this example TADA data frame, this spatial area consists of WQP data
from several different states. We will specify Maryland "MDE_EASP" as
our ATTAINS organization identifier of interest. By default, auto_assign
= "None" which will not populate any entries for the
ATTAINS.ParameterName column, and users will need to manually fill this
crosswalk out.

```{r auto_assign-none}
# create TADA parameter reference table for specified organization
MT.ParamRef_None <- TADA_CreateParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  auto_assign = "None",
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

# export TADA parameter reference table as csv
TADA_TableExport(MT.ParamRef_None)
```

Now, let's go through an example in which users auto-populate the
crosswalk with exact matches. We will use auto_assign = 'All' to
indicate we will use exact matches regardless of if your organization
has used this "ATTAINS.ParameterName" domain value in the past. i.e.
Your organization is fine with using an ATTAINS Parameter Name that has
been submitted to ATTAINS outside of your organization.

```{r auto_assign-All}
# create TADA parameter reference table for specified organization
MT.ParamRef_All <- TADA_CreateParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  auto_assign = "All",
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

# export TADA parameter reference table as csv
TADA_TableExport(MT.ParamRef_All)
```

We can also use auto_assign = 'Org' to indicate we will use exact
matches ONLY if your organization has used this "ATTAINS.ParameterName"
domain value in the past.

```{r auto_assign-Org}
# create TADA parameter reference table for specified organization
MT.ParamRef_Org <- TADA_CreateParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  auto_assign = "Org",
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

# export TADA parameter reference table as csv
TADA_TableExport(MT.ParamRef_Org)
```

The code chunk below demonstrates how the parameter reference table can
be modified in R. Once modifications are complete, we can re-run
TADA_CreateParamRef using the modified parameter reference table as the
input for "paramRef". This will update the ATTAINS.FlagParameterName
column in the parameter reference table (see example below). This column
provides information about whether or not the organization specified in
the "ATTAINS.OrganizationIdentifier" column has listed this parameter as
a cause in prior assessment cycles. It also flags rows where the
TADA.ComparableDataIdentifier as not been assigned an ATTAINS parameter
name.

```{r Manually-Create-ParamRef}
# only run the code chunks below if you make edits to the excel file. Please note you must open and save the excel file at least once to reflect all Excel formula based values.

# downloads_path <- file.path(Sys.getenv("USERPROFILE"), "Downloads", "myfileRef.xlsx")
# ParamRef <- openxlsx::read.xlsx(downloads_path, sheet = "CreateParamRef")

# MT.ParamRef_Excel <- TADA_CreateParamRef(
#   Data_NCTC,
#   org_names = c("MTDEQ"),
#   paramRef = ParamRef,
#   excel = TRUE, overwrite = TRUE
#   )

ParamRef <- MT.ParamRef_None %>%
  dplyr::mutate(ATTAINS.ParameterName = dplyr::case_when(
    grepl("PH_NONE_NONE_NONE", TADA.ComparableDataIdentifier) ~ "PH",
    grepl("ESCHERICHIA COLI", TADA.ComparableDataIdentifier) ~ "ESCHERICHIA COLI (E. COLI)")) %>% 
  dplyr::bind_rows(data.frame(TADA.ComparableDataIdentifier = "PH_NONE_NONE_NONE", ATTAINS.ParameterName = "PH, HIGH", ATTAINS.OrganizationIdentifier = "MTDEQ")) 

MT.ParamRef_Manual <- TADA_CreateParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  paramRef = ParamRef,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(MT.ParamRef_Manual)
```

We can now save and reuse MT.ParamRef_Manual as the user-supplied
paramRef argument input in TADA_CreateParamRef(). When this reference
file is used for the next assessment or analysis cycle, this process of
reusing your pre-filled paramRef table will assist in reviewing new WQX
characteristic names, if any are added, and determine the appropriate
crosswalk for these new entries if needed.

```{r User-Supplied-ParamRef}
MT.ParamRef_User_Supplied <- TADA_CreateParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  paramRef = MT.ParamRef_Manual,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(MT.ParamRef_User_Supplied)
```

QUESTION: Initially, we were thinking we would include all combinations of 
crosswalks (from user supplied and from an auto_assign method), but now we
are leaning towards full prioritization of the user-supplied crosswalk. 

Your user supplied 'paramRef' is prioritized over the 'auto_assign'
argument inputs. The auto_assign option will allow you to fill in any
remaining blanks that you have not filled in. 

```{r finalize-paramRef}
MT.ParamRef_Final <- TADA_CreateParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  paramRef = MT.ParamRef_Manual,
  auto_assign = "All",
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(MT.ParamRef_Final)
```

Once the parameter reference table has been reviewed and modified, it
can be saved and reused for analysis of future TADA data frames.

remove intermediate variable, keep only the crosswalk table that we will
continue to use in the workflow.

```{r remove-intermediate-variables}
rm(ParamRef, MT.ParamRef_All, MT.ParamRef_Manual, MT.ParamRef_None, MT.ParamRef_Org, MT.ParamRef_User_Supplied)
```

# TADA_CreateUseParamRef() Basics

Our first example will pull in all prior ATTAINS uses (ATTAINS.UseNames)
and parameter names (ATTAINS.ParameterName) specific to the specified
ATTAINS organization (defined by the org_names function argument for
TADA_CreateUseParamRef), which in this case is "MTDEQ" (Montana). Users
will review the output and choose which ATTAINS.UseName(s) are
applicable to their analysis. The default functionality is to "Include"
all ATTAINS use_names (see the "IncludeOrExclude" column) that has an
ATTAINS.UseName associated with it. Later in the data analysis process,
users will be asked to define criteria and methodologies that is
applicable to each ATTAINS use_name and parameter name labeled as
"Include". If an ATTAINS.UseName is not applicable, users should choose
"Exclude" for that ATTAINS parameter name and use name.

For any ATTAINS parameter name(s) that have not been used by an
organization in prior assessment cycles, there will be no prior use
names associated with them. In this case, users will be responsible for
manually assigning the appropriate ATTAINS.UseName applicable to their
org under the column 'ATTAINS.UseName', and for inputting additional
rows as needed if the parameter applies to multiples uses. Users can
also choose to 'auto_assign' all unique ATTAINS.UseName by
ATTAINS.OrganizationName to any ATTAINS.ParameterName missing a use name
associated with it.

In the example below, we can see "PH" and "ESCHERICHIA COLI (E. COLI)"
as the ATTAINS.ParameterNames were assessed in prior assessment cycles
for MTDEQ (see ATTAINS.FlagUseName column in NCTC_UseParamRef). "PH,
HIGH" was not listed in ATTAINS for MTDEQ in the prior assessment cycle,
but was included as a parameter name MTDEQ would like to assess for in
this current assessment/analysis cycle. Thus, the ATTAINS.UseName is
left blank for "PH, HIGH".

```{r useparamref-Manual}
MT.UseParamRef_Manual <- TADA_CreateUseParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  paramRef = MT.ParamRef_Final,
  auto_assign = FALSE,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
```

Now, if a user has completed assignments of uses to AUs in the
recommended ExampleMod2Workflow.Rmd, they need to provide this as a
useAURef argument input. This will help to ensure that the uses are
applicable to your most recent assessment/analysis needs. These uses are
extracted based on the uses to water type assignments in your useAURef
and joined to an internal parameter and water type extraction from
rExpertQuery to assign any new uses to parameters.

```{r useParamRef-with-useAURef}
MT.UseParamRef_with_useAURef <- TADA_CreateUseParamRef(
    tada.MT.clean,
    org_id = c("MTDEQ"),
    paramRef = MT.ParamRef_Final, 
    useAURef = MT.UseAURef_with_WaterUseRef, # uses will come from the user supplied reference table.
    auto_assign = FALSE,
    excel = FALSE)

```

If desired, a user can manually assign "PH, HIGH" to any applicable uses
(disclaimer: this is for demonstration purposes only and does not
reflect MTDEQ's criteria and assessment process).

```{r useparamref-Manual-Update}
add_data <- data.frame(
  "ATTAINS.OrganizationIdentifier" = "MTDEQ",
  "ATTAINS.ParameterName" = rep("PH, HIGH", 2),
  "ATTAINS.UseName" = c(
    "Aquatic Life",
    "Agriculture"
  )
)

# The output of this will not reflect changes to the ATTAINS.FlagUseName column. To do so, we need to re run TADA_CreateUseParamRef() with UseParamRef = add_data as an argument input.
UseParamRef <- MT.UseParamRef_Manual %>%
  dplyr::left_join(add_data, by = c("ATTAINS.OrganizationIdentifier", "ATTAINS.ParameterName"), keep = FALSE) %>%
  dplyr::mutate(ATTAINS.UseName = dplyr::coalesce(ATTAINS.UseName.x, ATTAINS.UseName.y)) %>%
  dplyr::select(-c(ATTAINS.UseName.x, ATTAINS.UseName.y)) %>%
  dplyr::mutate(IncludeOrExclude = "Include")

# PH will now reflect the changes
MT.UseParamRef_Manual_Update <- TADA_CreateUseParamRef(
  tada.MT.clean,
  paramRef = MT.ParamRef_Final,
  useParamRef = UseParamRef, # Edits were made to UseParamRef, updates flag column
  org_id = c("MTDEQ"),
  auto_assign = FALSE,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

TADA_TableExport(MT.UseParamRef_Manual_Update)
```

Alternatively, We can choose to assign all unique use names found in
your current useParamRef table, by your organization to those
ATTAINS.ParameterName without any associated ATTAINS.UseName. Users will
need to review this assignment carefully and choose include or exclude
appropriately or include additional rows as needed if there are more
ATTAINS.UseName applicable for that ATTAINS.ParameterName that was not
captured by the auto_assign method.

```{r useparamref-AutoAssign}
MT.UseParamRef_AutoAssign <- TADA_CreateUseParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  paramRef = MT.ParamRef_Final,
  useAURef = MT.UseAURef_with_WaterUseRef,
  auto_assign = TRUE,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(MT.UseParamRef_AutoAssign)
```

Users can also choose to supply their own UseParamRef as an argument
input. This may be useful if your organization has made significant
changes to use names in the current cycle that cannot be retrieved from
the prior ATTAINS assessment cycle or if there are just additional use
names that were unable to be retrieved from the prior assessment cycle.
(Note: TADA functions leverages the ATTAINS assessment profiles which
will not reflect any submitted use name changes to ATTAINS until the
next assessment cycle is approved. In addition, it may not be possible
to retrieve every use and parameter name combination from ATTAINS as it
is dependent on what each organization has submitted to ATTAINS as only
those uses that are impaired may have been submitted.)

In this example, let's use a filled out criteria table for MTDEQ and
extract the use and parameters from their crosswalk. We can see that
MTDEQ has additonal rows for E. Coli that was unable to be retrieved
from ATTAINS. We also see PH was not found in the user supplied
useParamRef table. We will assume MTDEQ has forgotten to include PH in
their table and showcase the importance of considering all readily
available data from your WQP data retrieval that you may need to
consider. We will change this table to "Include" for all PH values.

```{r useparamref-User-Supplied}
# write code to save MTDEQ criteria table into TADA as .Rda and then load it into the R environment.
# load(criteria_table)

MT.UseParamRef_User_Supplied <- TADA_CreateUseParamRef(
  tada.MT.clean,
  org_id = c("MTDEQ"),
  paramRef = MT.ParamRef_Final,
  useParamRef = criteria_table,
  auto_assign = TRUE,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

MT.UseParamRef_User_Supplied <- 
  dplyr::mutate(MT.UseParamRef_User_Supplied, IncludeOrExclude = "Include")

TADA_TableExport(MT.UseParamRef_AutoAssign)
```

remove intermediate variable, keep only the crosswalk table that we will
continue to use in the workflow.

```{r remove-intermediate-variables-useParam}
rm(MT.UseParamRef_AutoAssign, MT.UseParamRef_Manual, MT.UseParamRef_Manual_Update)
```

# TADA_CreateMLSummaryRef()

Now, we will proceed with generating a final summary of analyzing your
WQP data and specifying the spatial level of summary. In this example,
we will showcase an example of an AU and ML level of summary for MTDEQ
analysis and compare the two outputs.

## Define Spatial Summary by Monitoring Location (ML)

Let's start off with a ML summary of MTDEQ data frame.

```{r ML_Summary}
MT.MLSummaryRef_ML <- TADA_CreateMLSummaryRef(
  .data = tada.MT.clean, 
  useParamRef = MT.UseParamRef_AutoAssign,
  org_id = "MTDEQ"
  )
```

## Define Spatial Summary by Monitoring Location (AU)

Now let's compare this to the AU level of summary. To provide the ML to
AU level of summary, this requires an AUML crosswalk, we recommend you
to review the ExampleMod2Workflow vignette for more in depth information
on this crosswalk.

```{r AU_Summary, eval = FALSE}
MT.MLSummaryRef_AU <- TADA_CreateMLSummaryRef(
  .data = tada.MT.clean, 
  useParamRef = MT.UseParamRef_AutoAssign, 
  useAURef = MT.UseAURef_with_WaterUseRef,
  AUMLRef = Final.MT.AUMLRef,
  org_id = "MTDEQ"
  )
```

Compare the nrow for each of the ML versus AU level of summary. The AU
level of summary is further filtered down by the AU to ML crosswalk, the
use to AU crosswalk, and the use to parameter crosswalk based on what
TADA.ComparableDataIdentifier has been collected at those sites and
taking into consideration of what parameter(s) and use(s) have been
assessed for each assessment unit. (Note: Would users find it useful to still include a
parameter and use for a site even if the site does not contain that
TADA.ComparableDataIdentifier - the ATTAINS.ParameterName? This could be
included in the final summary table and labeled as NA or insufficient
data.)

```{r compare-AU-versus-ML-Summary,, eval = FALSE}
nrow(MT.MLSummaryRef_AU)
nrow(MT.MLSummaryRef_ML)
```

# TADA DefineCriteriaMethodology()

Now, lets get to generating the criteria and methodology file for MTDEQ
to fill out. We will showcase how this table will be generated for the
first time using the recommended step-by-step workflow and showcase how
a user can go about updating their criteria and methodology file as
needed.

First, let's show how our step-by-step process all come together.

```{r step-by-step-output}
MT.CriteriaMethods_Blank <- TADA_DefineCriteriaMethodology(
  .data = tada.MT.clean,
  org_id = "MTDEQ",
  MLSummaryRef = MT.MLSummaryRef_ML,
  excel = FALSE
)
```

Now, if MTDEQ, already has a filled out/partially filled out criteria
methods table, let's provide this as an argument input.

Note: Whenever a criteriaMethods argument input value is provided, users
will be warned if there are additional WQP characteristics (or
TADA.ComparableDataIdentifier - see argument input uniqueDataId =
TRUE/FALSE in the R documentation for more information) that are not
captured in their user supplied table.

```{r user-supplied-output}
MT.CriteriaMethods_User_Supplied <- TADA_DefineCriteriaMethodology(
  .data = tada.MT.clean,
  org_id = "MTDEQ",
  MLSummaryRef = MT.MLSummaryRef_ML,
  criteriaMethods = criteria_table,
  excel = FALSE
)
```

Note: When a user provides a MLSummaryRef, it is assumed their criteria
tables parameter crosswalk is based on the paramRef and useParamRef
crosswalk done. However, whenever a criteriaMethods table is provided,
all unique WQP Characteristic (or TADA.ComprableDataIdentifier) will be
shown in the table.

```{r step-by-step-output-with-displayUniqueId}
MT.CriteriaMethods_User_Supplied2 <- TADA_DefineCriteriaMethodology(
  .data = tada.MT.clean,
  org_id = "MTDEQ",
  MLSummaryRef = MT.MLSummaryRef_ML,
  criteriaMethods = criteria_table,
  displayUniqueId = TRUE,
  excel = FALSE
)
```

You can append epa304a recommended standards (if there is one found for
a TADA.CharacteristicName)

```{r append-epa304a-recommended-standards}
MT.CriteriaMethods_User_Supplied3 <- TADA_DefineCriteriaMethodology(
  .data = tada.MT.clean,
  org_id = "MTDEQ",
  MLSummaryRef = MT.MLSummaryRef_ML,
  criteriaMethods = criteria_table,
  displayUniqueId = TRUE,
  epa304a = TRUE,
  excel = FALSE
)
```
