---
title: "Example Module 2 Workflow"
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
editor: visual
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
description: The workflow for creating a crosswalk between ATTAINS assessment units and WQP monitoring locations prioritizes using an optional user-supplied crosswalk first. If that is unavailable, it retrieves a crosswalk submitted to ATTAINS by state or tribal organizations, and if neither is available, it generates one using geographical data to link assessment units with monitoring locations.
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example Module 2 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Install and Load EPATADA Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r installremotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load EPATADA using the remotes package. USGS's
dataRetrieval and other TADA R Package dependencies will also be
downloaded automatically from CRAN with the TADA install. If desired,
the development version of dataRetrieval can be downloaded directly from
GitHub (un-comment).

```{r installTADA, eval = T, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)

# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r devinstall, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "combined-review",
  dependencies = TRUE
)
```

Finally, use the library() function to load the TADA R Package into your
R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Get WQP Monitoring Data in Montana Using TADA_DataRetrieval()

Get bacteria and pH data from Missoula County, Montana.

```{r tadadata}
# get MT data 
tada.MT <- TADA_DataRetrieval(
   startDate = "2020-01-01", 
   endDate = "2022-12-31",
   statecode = "MT",
   characteristicName = c("Escherichia", 
                          "Escherichia coli", 
                          "pH"),
   county = "Missoula County",
   ask = FALSE)

# clean up data set (minimal)
  tada.MT.clean <- tada.MT %>%
    TADA_RunKeyFlagFunctions() %>%
    TADA_SimpleCensoredMethods()
  
# remove intermediate objects
  rm(tada.MT)
```

## Create An Example User-Supplied ATTAINS Assessment Unit and WQP Monitoring Location Crosswalk

Get existing data from ATTAINS and subset a few rows to create an
example user-supplied crosswalk for demonstration purposes.

```{r MTexample, echo = FALSE}
# get crosswalk from ATTAINS
attains.existing.MT <- TADA_GetATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# clean existing crosswalk from ATTAINS to make sure WQP monitoring location IDs pulled from ATTAINS are WQP compatible (adds org ID if missing)
clean.existing.attains.MT <- TADA_UpdateATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# create example user supplied crosswalk (select a few Monitoring Locations from the tada df to use in the example for demonstration purposes)
user.supplied.cw <- clean.existing.attains.MT %>%
  dplyr::select(ATTAINS.AssessmentUnitIdentifier, 
                ATTAINS.MonitoringLocationIdentifier) %>%
  dplyr::filter(ATTAINS.MonitoringLocationIdentifier %in% c(
    "MDEQ_WQ_WQX-C04CKFKR05", "MDEQ_WQ_WQX-C04KNDYC01", "MDEQ_WQ_WQX-C04KNDYC02", 
    "MDEQ_WQ_WQX-C04KNDYC04", "MDEQ_WQ_WQX-C04KNDYC54")) %>%
  dplyr::rename(AssessmentUnitIdentifier = ATTAINS.AssessmentUnitIdentifier,
                MonitoringLocationIdentifier = ATTAINS.MonitoringLocationIdentifier)

rm(attains.existing.MT, clean.existing.attains.MT)
```

## Find Additional ATTAINS Assessment Unit and WQP Monitoring Location Matches Using TADA_CreateAUMLCrosswalk()

The workflow for creating an ATTAINS assessment unit (AU) and Water
Quality Portal (WQP) monitoring location crosswalk involves three steps,
prioritized in the following order:

1.  **Optional User-Supplied Crosswalk**: The first option allows users
    to provide their own crosswalk. This is an optional step where users
    can supply a pre-existing crosswalk that they have created or
    obtained.

2.  **Crosswalk Retrieved Using TADA_CreateATTAINSAUMLCrosswalk
    Function**: If a user-supplied crosswalk is not available, the next
    step is to use the **`TADA_CreateATTAINSAUMLCrosswalk`** function.
    This function retrieves a crosswalk, if available, that has been
    voluntarily submitted to ATTAINS by a state or tribal organization.
    This means the crosswalk has been shared by these organizations to
    facilitate the linking of assessment units to monitoring locations.

3.  **Crosswalk Generated by TADA_CreateATTAINSAUML**: If neither of the
    previous options is available, the final step is to generate a
    crosswalk using the **`TADA_CreateATTAINSAUML`** function. This
    function creates a crosswalk by using the ATTAINS AU catchments to
    join assessment units (AUs) with WQP monitoring locations.
    Essentially, it uses geographical data to establish connections
    between the assessment units and monitoring locations.

The workflow prioritizes these options in the order listed, starting
with the user-supplied crosswalk as the preferred method, followed by
the retrieval of a voluntarily submitted crosswalk, and finally
generating one through geographical associations if needed.

To make this function run more efficiently, and because we already have
assessment unit assignments for some monitoring locations from the
user-supplied crosswalk and the crosswalk downloaded from ATTAINS, the
ATTAINS catchment data from the geospatial service is not currently
being returned for matches in the crosswalks supplied by users or
imported from ATTAINS. We have plans to create additional
functions/internal reference files to facilitate this returning the
associated catchments in the future. Therefore, any catchment data
available for mapping is currently from TADA_CreateATTAINSAUMLCrosswalk
assessed WQP monitoring locations only.

```{r createAUMLcrosswalk}
# use TADA get attains to make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLCrosswalk(tada.MT.clean,
                                 au_ref = user.supplied.cw,
                                 org_id = "MTDEQ", 
                                 add_catch = FALSE,
                                 batch_upload = TRUE)
```

## Generate a Map with TADA_ViewATTAINS

The TADA_ViewATTAINS() map function shows the results from all three
different crosswalk sources (user supplied crosswalk, ATTAINS crosswalk,
and ATTAINS catchments/geospatial join) using different color circle
markers (not final colors, we are very open to suggestions). The source
has also been added to the popup window.

```{r viewmap}
TADA_ViewATTAINS(MT.AUMLRef)
```

## TADA_UpdateATTAINSAUMLCrosswalk

After making any necessary changes to the ATTAINS_batchupload data frame
generated by TADA_CreateAUMLCrosswalk, run
TADA_UpdateATTAINSAUMLCrosswalk to add new links to WQP site pages.

```{r addlinks}
batch.upload.MT <- MT.AUMLRef$ATTAINS_batchupload %>%
  TADA_UpdateATTAINSAUMLCrosswalk(attains_replace = TRUE,
                                  batch_upload = TRUE,
                                  wqp_data_links = "add",
                                  update_mlid = FALSE,
                                  org_id = "MTDEQ")
```
