---
title: "Example Module 2 Workflow"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example Module 2 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
description: An example Module 2 workflow for creating an ATTAINS assessment unit/WQP monitoring location crosswalk. This workflow considers (1) an optional user-supplied crosswalk, (2) any WQP monitoring location data stored in ATTAINS, and (3) the results of TADA_CreateATTAINSAUML prioritized in the order listed to create a comprehensive assessment unit/monitoring location crosswalk.
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = TRUE
)
```

## Install and Load EPATADA Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load EPATADA using the remotes package. USGS's
dataRetrieval and other TADA R Package dependencies will also be
downloaded automatically from CRAN with the TADA install. If desired,
the development version of dataRetrieval can be downloaded directly from
GitHub (un-comment).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)

# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r dev_install, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "combined-review",
  dependencies = TRUE
)
```

Finally, use the library() function to load the TADA R Package into your
R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Get WQP Monitoring Data in Montana Using TADA_DataRetrieval()

Get bacteria and pH data from Missoula County, Montana.

```{r tada.data}

# get MT data 
tada.MT <- TADA_DataRetrieval(
   startDate = "2020-01-01", 
   endDate = "2022-12-31",
   statecode = "MT",
   characteristicName = c("Escherichia", 
                          "Escherichia coli", 
                          "pH"),
   county = "Missoula County",
   ask = FALSE)

# clean up data set (minimal)
  tada.MT.clean <- tada.MT %>%
    TADA_RunKeyFlagFunctions() %>%
    TADA_SimpleCensoredMethods()
  
# remove intermediate objects
  rm(tada.MT)

```

## Create An Example User-Supplied ATTAINS Assessment Unit and WQP Monitoring Location Crosswalk

Get existing data from ATTAINS and subset a few rows to create an
example user-supplied crosswalk for demonstration purposes.

```{r demo.user, echo = FALSE}

# get crosswalk from ATTAINS
attains.existing.MT <- TADA_GetATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# clean existing crosswalk from ATTAINS to make sure WQP monitoring location IDs pulled from ATTAINS are WQP compatible (adds org ID if missing)
clean.existing.attains.MT <- TADA_UpdateATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# create example user supplied crosswalk (select a few Monitoring Locations from the tada df to use in the example for demonstration purposes)
user.supplied.cw <- clean.existing.attains.MT %>%
  dplyr::select(ATTAINS.AssessmentUnitIdentifier, 
                ATTAINS.MonitoringLocationIdentifier) %>%
  dplyr::filter(ATTAINS.MonitoringLocationIdentifier %in% c(
    "MDEQ_WQ_WQX-C04CKFKR05", "MDEQ_WQ_WQX-C04KNDYC01", "MDEQ_WQ_WQX-C04KNDYC02", 
    "MDEQ_WQ_WQX-C04KNDYC04", "MDEQ_WQ_WQX-C04KNDYC54")) %>%
  dplyr::rename(AssessmentUnitIdentifier = ATTAINS.AssessmentUnitIdentifier,
                MonitoringLocationIdentifier = ATTAINS.MonitoringLocationIdentifier)

rm(attains.existing.MT, clean.existing.attains.MT)
```

## Find Additional ATTAINS Assessment Unit and WQP Monitoring Location Matches Using TADA_CreateAUMLCrosswalk()

Run TADA_CreateAUMLCrosswalk which can incorporate any combination of:

1.  (optional) user-supplied crosswalk,

2.  crosswalk downloaded from ATTAINS, and

3.  results from TADA_CreateATTAINSAUMLCrosswalk which leverages the
    ATTAINS catchments to join monitoring locations to assessment units
    via a geospatial join .

The assignments are prioritized in the order listed. Any WQP monitoring
locations which remain unassigned after checking the user-supplied
crosswalk will be subset and run with TADA_CreateATTAINSAUMLCrosswalk
which will attempt to make additional ATTAINS assessment unit
assignments.

To make this function run more efficiently, and because we already have
assessment unit assignments for some monitoring locations from the
user-supplied crosswalk and the crosswalk downloaded from ATTAINS, the
ATTAINS catchment data from the geospatial service is not currently
being returned for matches in the crosswalks supplied by users or
imported from ATTAINS. We have plans to create additional
functions/internal reference files to facilitate this returning the
associated catchments in the future. Therefore, any catchment data
available for mapping is currently from TADA_CreateATTAINSAUMLCrosswalk
assessed WQP monitoring locations only.

```{r tada.get.attains}

# use TADA get attains to make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLCrosswalk(tada.MT.clean,
                                 au_ref = user.supplied.cw,
                                 org_id = "MTDEQ", 
                                 add_catch = FALSE,
                                 batch_upload = TRUE)

```

## Generate a Map with TADA_ViewATTAINS

The TADA_ViewATTAINS() map function shows the results from all three
different crosswalk sources (user supplied crosswalk, ATTAINS crosswalk,
and ATTAINS catchments/geospatial join) using different color circle
markers (not final colors, we are very open to suggestions). The source
has also been added to the popup window.

```{r view.map}

TADA_ViewATTAINS(MT.AUMLRef)

```

## TADA_UpdateATTAINSAUMLCrosswalk

After making any necessary changes to the ATTAINS_batchupload data frame
generated by TADA_CreateAUMLCrosswalk, run
TADA_UpdateATTAINSAUMLCrosswalk to add new links to WQP site pages.

```{r add.links}

batch.upload.MT <- MT.AUMLRef$ATTAINS_batchupload %>%
  TADA_UpdateATTAINSAUMLCrosswalk(# selected attains_replace = TRUE because all matches                                     currently in ATTAINS are included in this new                                          crosswalk
                                  attains_replace =TRUE,
                                  batch_upload = TRUE,
                                  wqp_data_links = "add",
                                  #ml ids have already  been corrected if needed
                                  update_mlid = FALSE,
                                  org_id = "MTDEQ")

```
