---
title: "Example Module 2 Workflow"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example Module 2 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
description: An example Module 2 workflow focused on A) creating an ATTAINS assessment unit/WQP monitoring location crosswalk, B) reviewing assessment units and monitoring locations on a map, C) submitting an up to date assessment unit/WQP monitoring location crosswalk to ATTAINS, and D) assigning uses to assessment units.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = FALSE
)
```

## Overview and Setup

This is an example Module 2 workflow focused on: A) creating an ATTAINS
assessment unit/WQP monitoring location crosswalk using
`TADA_CreateAUMLCrosswalk`, B) reviewing assessment units and monitoring
locations on a map with `TADA_ViewATTAINS`, C) submitting an up to date
assessment unit/WQP monitoring location crosswalk to ATTAINS using
`TADA_UpdateATTAINSAUMLCrosswalk`, and D) assigning uses to assessment
units with `TADA_CreateUseAURef`.

### Install and Load EPATADA Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load EPATADA using the remotes package. USGS's
dataRetrieval and other TADA R Package dependencies will also be
downloaded automatically from CRAN with the TADA install. If desired,
the development version of dataRetrieval can be downloaded directly from
GitHub (un-comment).

```{r install_TADA, eval = T, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)

# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r dev_install, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "Create-Module-3-AURef",
  dependencies = TRUE
)
```

Finally, use the library() function to load the EPATADA R Package and
other required packages/

```{r library, results = 'hide'}
library(EPATADA)
library(dplyr)
```

### Get WQP Monitoring Data in Montana Using TADA_DataRetrieval()

Get bacteria and pH data from Missoula County, Montana.

```{r tada.data}
# get MT data
tada.MT <- TADA_DataRetrieval(
  startDate = "2020-01-01",
  endDate = "2022-12-31",
  statecode = "MT",
  characteristicName = c(
    "Escherichia",
    "Escherichia coli",
    "pH"
  ),
  county = "Missoula County",
  ask = FALSE
)

# clean up data set (not comprehensive)
tada.MT.clean <- tada.MT %>%
  TADA_RunKeyFlagFunctions() %>%
  TADA_SimpleCensoredMethods() %>%
  TADA_HarmonizeSynonyms()

# remove intermediate objects
rm(tada.MT)

# or uncomment the code below to load the example data set
# tada.MT.clean <- Data_MT_MissoulaCounty
```

### Create An Example User-Supplied Crosswalk

Get existing data from ATTAINS using `TADA_GetATTAINSAUMLCrosswalk` and
subset a few rows to create an example user-supplied ATTAINS Assessment
Unit and WQP Monitoring Location crosswalk for demonstration purposes.
If you have your own crosswalk, this step can be skipped.

```{r demo.user, results = 'hide'}
# review valid ATTAINS org IDs
ATTAINS_orgs <- rExpertQuery::EQ_DomainValues("org_id")

# get crosswalk from ATTAINS
attains.existing.MT <- TADA_GetATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# clean existing crosswalk from ATTAINS to make sure WQP monitoring location IDs pulled from ATTAINS are WQP compatible (adds org ID if missing)
clean.existing.attains.MT <- TADA_UpdateATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# create example user supplied crosswalk (select a few Monitoring Locations from the tada df to use in the example for demonstration purposes)
user.supplied.cw <- clean.existing.attains.MT %>%
  dplyr::select(
    ATTAINS.AssessmentUnitIdentifier,
    ATTAINS.MonitoringLocationIdentifier,
    ATTAINS.WaterType
  ) %>%
  dplyr::filter(ATTAINS.MonitoringLocationIdentifier %in% c(
    "MDEQ_WQ_WQX-C04CKFKR05", "MDEQ_WQ_WQX-C04KNDYC01", "MDEQ_WQ_WQX-C04KNDYC02",
    "MDEQ_WQ_WQX-C04KNDYC04", "MDEQ_WQ_WQX-C04KNDYC54"
  )) %>%
  dplyr::rename(
    AssessmentUnitIdentifier = ATTAINS.AssessmentUnitIdentifier,
    MonitoringLocationIdentifier = ATTAINS.MonitoringLocationIdentifier,
    WaterType = ATTAINS.WaterType
  ) %>%
  # Add an example new assessment unit for demonstration purposes.
  dplyr::bind_rows(c(
    AssessmentUnitIdentifier = "NEW:EX_MDEQ_WQ_WQX",
    MonitoringLocationIdentifier = "NARS_WQX-NWC_MT-10184",
    WaterType = "LAKE, FRESHWATER"
  ))


rm(attains.existing.MT, clean.existing.attains.MT, ATTAINS_orgs)
```

## Step A: Discover matches with TADA_CreateAUMLCrosswalk()

`TADA_CreateAUMLCrosswalk` is a new function that can be leveraged to
create an ATTAINS assessment unit/WQP monitoring location crosswalk. It
considers:

1.  an optional user-supplied crosswalk,

2.  the results of `TADA_GetATTAINSAUMLCrosswalk` which brings in any
    WQP monitoring location information stored in ATTAINS by
    participating organizations, and

3.  the results of `TADA_CreateATTAINSAUMLCrosswalk` which leverages the
    ATTAINS catchments to join monitoring locations to assessment units
    via a geospatial join.

The three potential crosswalk data sources are prioritized in the order
listed to create a comprehensive assessment unit/monitoring location
crosswalk. This means that any WQP monitoring locations which remain
unassigned after checking the user-supplied crosswalk will be subset and
the function will then attempt to make additional ATTAINS assessment
unit assignments using `TADA_GetATTAINSAUMLCrosswalk.` The same process
then repeats and the function will then attempt to make additional
ATTAINS assessment unit assignments using
`TADA_CreateATTAINSAUMLCrosswalk`. All of this happens behind the scenes
within the `TADA_CreateAUMLCrosswalk` functions so that users do not
need to run them all separately (though that is still an option if users
prefer).

To make this function run more efficiently, and because we already have
assessment unit assignments for some monitoring locations from the
user-supplied crosswalk and the crosswalk downloaded from ATTAINS
(`TADA_GetATTAINSAUMLCrosswalk`), the ATTAINS catchment data from the
geospatial service is not currently being returned for matches in the
crosswalks supplied by users or imported from ATTAINS. We have plans to
create additional functions/internal reference files to facilitate this
returning the associated catchments in the future. Therefore, any
catchment data available for mapping is currently from
`TADA_CreateATTAINSAUMLCrosswalk` assessed WQP monitoring locations
only.

```{r tada.get.attains}
# use TADA get attains to make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLCrosswalk(tada.MT.clean,
  au_ref = user.supplied.cw,
  org_id = "MTDEQ",
  add_catch = FALSE,
  nhd_catch = TRUE,
  batch_upload = TRUE)
```

## Step B: Generate a Map with TADA_ViewATTAINS

The `TADA_ViewATTAINS()` map function shows the results from all three
different crosswalk sources (user supplied crosswalk, ATTAINS crosswalk,
and ATTAINS catchments/geospatial join) by using different circle
markers. The user icon indicates the monitoring location was from a user
supplied ref. The circle with a check mark signifies a match from the
crosswalk downloaded from ATTAINS (`TADA_GetATTAINSAUMLCrosswalk`).
Matches found by comparing monitoring locations with ATTAINS catchment
and assessment geometry with `TADA_CreateATTAINSAUMLCrosswalk` are
represented by solid fill circle markers. The source is also displayed
in the pop up window that appears when a user clicks on each circle
marker.

```{r view.map}
TADA_ViewATTAINS(MT.AUMLRef, ref_icons = TRUE)
```

## Step C: Update crosswalk in ATTAINS using TADA_UpdateATTAINSAUMLCrosswalk

The output of TADA_UpdateATTAINSAUMLCrosswalk can be uploaded to ATTAINS
to create or update Monitoring Location Identifiers in ATTAINS
Assessment Unit profiles. Users can specify whether all records should
be overwritten (replaced) or if new Monitoring Location Identifiers
should be appended (added) to existing records. Enter
?TADA_UpdateATTAINSAUMLCrosswalk() into the console for more details.

In this example workflow, after making any necessary changes to the
ATTAINS_batchupload data frame generated by TADA_CreateAUMLCrosswalk, we
run TADA_UpdateATTAINSAUMLCrosswalk to add new links to WQP site pages.
This is now ready to upload to ATTAINS.

```{r add.links}
batch.upload.MT <- MT.AUMLRef$ATTAINS_batchupload %>%
  TADA_UpdateATTAINSAUMLCrosswalk( # selected attains_replace = TRUE because all matches currently in ATTAINS are included in this new crosswalk
    attains_replace = TRUE,
    batch_upload = TRUE,
    wqp_data_links = "add",
    # ml ids have already  been corrected if needed
    update_mlid = FALSE,
    org_id = "MTDEQ"
  )
```

## TADA_CreateUseAURef

After assigning ML sites to AU, we will use TADA_CreateUseAURef to
assign uses to these AUs. First, convert the ML sites to AU reference
table to the appropriate format for TADA workflow. We will also assign
the NEW:EX_MDEQ_WQ_WQX assessment unit example to the water type "LAKE,
FRESHWATER" for demonstration on how to auto_assign uses to a new AU
that is not yet in ATTAINS nor has been assigned uses to it yet. To
create the final AU/ML reference will use MT.AUMLRef\$ATTAINS_crosswalk
as it contains ATTAINS.WaterType which is required farther along in the
assessment process.

```{r get.MLAURef}
Final.MT.AUMLRef <- MT.AUMLRef$ATTAINS_crosswalk %>%
  dplyr::mutate(
    ATTAINS.WaterType = dplyr::case_when(
      ATTAINS.AssessmentUnitIdentifier == "NEW:EX_MDEQ_WQ_WQX" ~ "LAKE, FRESHWATER",
      TRUE ~ ATTAINS.WaterType
    )
  )
```

## Step D: Assign uses to AUs with TADA_CreateUseAURef

After assigning WQP MLs to ATTAINS AUs, we will use
`TADA_CreateUseAURef` to assign uses to these AUs. First, we will
convert the ML to AU crosswalk to the appropriate format for this TADA
workflow. We will also assign the *NEW:EX_MDEQ_WQ_WQX* assessment unit
example to the water type *LAKE, FRESHWATER* to demonstrate how to
`auto_assign` uses to a new AU that is not yet in ATTAINS based on water
type.

Now, we can use Final.MT.AUMLRef as our ML to AU crosswalk and proceed
with assigning uses for each unique AU. The assignment of uses to AUs
are prioritized as follows:

1)  (optional) User supplied uses to AU

2)  Imports existing uses by AU from ExpertQuery

```{r assign.use.to.AU}
MT.UseAURef <- TADA_CreateUseAURef(AUMLRef = Final.MT.AUMLRef, org_id = "MTDEQ")
```

# Assign uses to new AUs

For any new AU, users will need to determine how to assign uses to them.

1)  TADA has developed a function called TADA_CreateWaterUseRef which
    imports existing uses by water type from ExpertQuery, and will
    assign any new AU these uses. Users may choose to edit the
    TADA_CreateWaterUseRef according to their preference.

```{r assign.use.to.new.AU.with.waterUseRef}
MT.UseAURef_with_WaterUseRef <- TADA_CreateUseAURef(
  waterUseRef = TADA_CreateWaterUseRef(org_id = "MTDEQ"),
  AUMLRef = Final.MT.AUMLRef,
  org_id = "MTDEQ"
)
```

2)  User may also proceed to do a manual assignment of use names to
    these new AUs if desired rather than using `TADA_CreateWaterUseRef`.

```{r assign.use.to.new.AU.manual}
MT.UseAURef_manual <- MT.UseAURef %>%
  dplyr::left_join(
    data.frame(
      ATTAINS.UseName = c("Aquatic Life", "Drinking Water"),
      ATTAINS.AssessmentUnitIdentifier = c("NEW:EX_MDEQ_WQ_WQX", "NEW:EX_MDEQ_WQ_WQX")
    ),
    by = ("ATTAINS.AssessmentUnitIdentifier")
  ) %>%
  dplyr::mutate(ATTAINS.UseName = dplyr::coalesce(ATTAINS.UseName.x, ATTAINS.UseName.y)) %>%
  dplyr::select(-ATTAINS.UseName.x, -ATTAINS.UseName.y)
```
