---
title: "Example Module 2 Workflow"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Example Module 2 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
description: An example Module 2 workflow focused on A) creating an ATTAINS assessment unit/WQP monitoring location crosswalk, B) reviewing assessment units and monitoring locations on a map, C) submitting an up to date assessment unit/WQP monitoring location crosswalk to ATTAINS, and D) assigning uses to assessment units.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview and Setup

This example Module 2 workflow focuses on the following tasks:

A)  **Creating a ML/AU Crosswalk**: Use **`TADA_CreateAUMLCrosswalk`**
    to establish a crosswalk between ATTAINS assessment units and WQP
    monitoring locations.

B)  **Map Review**: Utilize **`TADA_ViewATTAINS`** to review assessment
    units and monitoring locations on a map, ensuring accuracy and
    completeness.

C)  **Submitting the Crosswalk**: Employ
    **`TADA_UpdateATTAINSAUMLCrosswalk`** to submit an up-to-date
    assessment unit/WQP monitoring location crosswalk to ATTAINS.

D)  **Assigning Uses**: Use **`TADA_CreateUseAURef`** to assign specific
    uses to the assessment units.

### Install and Load EPATADA Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
```

Load the remotes library:

```{r load_remotes}
library(remotes)
```

Next, install and load EPATADA using the remotes package. USGS's
dataRetrieval and other TADA R Package dependencies will also be
downloaded automatically from CRAN with the TADA install. If desired,
the development version of dataRetrieval can be downloaded directly from
GitHub (un-comment).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)

# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r dev_install, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "Create-Module-3-AURef",
  dependencies = TRUE
)
```

Finally, use the library() function to load the EPATADA R Package and
other required packages/

```{r library, results = 'hide'}
library(EPATADA)
```

### Get WQP Monitoring Data in Montana Using TADA_DataRetrieval()

Get bacteria and pH data from Missoula County, Montana.

```{r tada.data, eval = F}
# get MT data
tada.MT <- TADA_DataRetrieval(
  startDate = "2020-01-01",
  endDate = "2022-12-31",
  statecode = "MT",
  characteristicName = c(
    "Escherichia",
    "Escherichia coli",
    "pH"
  ),
  county = "Missoula County",
  ask = FALSE
)

# clean up data set (not comprehensive)
tada.MT.clean <- tada.MT %>%
  TADA_RunKeyFlagFunctions() %>%
  TADA_SimpleCensoredMethods() %>%
  TADA_HarmonizeSynonyms() 

# Remove intermediate objects
rm(tada.MT)
```

This example is also included in the EPATADA R package, and can be
loaded into your R environment like this:

```{r load WQP data}
tada.MT.clean <- Data_MT_MissoulaCounty
```

## Step A: Discover Matches with `TADA_CreateAUMLCrosswalk()`

The **`TADA_CreateAUMLCrosswalk`** function efficiently creates a
crosswalk between ATTAINS assessment units and WQP monitoring locations.
It uses three prioritized data sources:

1.  **User-Supplied Crosswalk:** An optional crosswalk provided by the
    user (e.g., see example **`user.supplied.cw`** in next code chunk).

2.  **ATTAINS Crosswalk:** Utilizes **`TADA_GetATTAINSAUMLCrosswalk`**
    to incorporate crosswalk information stored by participating
    organizations in ATTAINS.

3.  **ATTAINS Catchements/Geospatial Join:** Employs
    **`TADA_CreateATTAINSAUMLCrosswalk`** to connect monitoring
    locations to assessment units using ATTAINS catchments through a
    geospatial join.

**Process**

-   The function prioritizes these sources in the order listed. It
    automatically attempts to assign unassigned monitoring locations
    using the next available data source. For example, if any monitoring
    locations remain unassigned after checking the user-supplied
    crosswalk, the function will check the ATTAINS crosswalk, then the
    geospatial join, as needed.

-   This entire process is automated within
    **`TADA_CreateAUMLCrosswalk`**, so users do not need to run each
    step individually, though they have the option to do so if desired.

**Current Limitations**

-   To make this function run more efficiently, catchments from the
    ATTAINS geospatial service are not returned for matches found in the
    user-supplied crosswalk or ATTAINS crosswalk (#1 or #2 above).
    ATTAINS catchments are only returned from
    `TADA_CreateATTAINSAUMLCrosswalk` assessed WQP monitoring locations.

**Future Improvements**

-   Plans are underway to develop a separate process to return all
    ATTAINS catchments associated with all WQP monitoring locations in
    the future, to support mapping and crosswalk review.

### Create An Example User-Supplied Crosswalk

Get existing data from ATTAINS using `TADA_GetATTAINSAUMLCrosswalk` and
subset a few rows to create an example user-supplied ATTAINS Assessment
Unit and WQP Monitoring Location crosswalk for demonstration purposes.
If you have your own crosswalk, this step can be skipped.

```{r demo.user, results = 'hide'}
# review valid ATTAINS org IDs
ATTAINS_orgs <- rExpertQuery::EQ_DomainValues("org_id")

# get crosswalk from ATTAINS
attains.existing.MT <- TADA_GetATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# clean existing crosswalk from ATTAINS to make sure WQP monitoring location IDs pulled from ATTAINS are WQP compatible (adds org ID if missing)
clean.existing.attains.MT <- TADA_UpdateATTAINSAUMLCrosswalk(org_id = "MTDEQ")

# create example user supplied crosswalk (select a few Monitoring Locations from the tada df to use in the example for demonstration purposes)
user.supplied.cw <- clean.existing.attains.MT %>%
  dplyr::select(
    ATTAINS.AssessmentUnitIdentifier,
    ATTAINS.MonitoringLocationIdentifier,
    ATTAINS.WaterType
  ) %>%
  dplyr::filter(ATTAINS.MonitoringLocationIdentifier %in% c(
    "MDEQ_WQ_WQX-C04CKFKR05", "MDEQ_WQ_WQX-C04KNDYC01", "MDEQ_WQ_WQX-C04KNDYC02",
    "MDEQ_WQ_WQX-C04KNDYC04", "MDEQ_WQ_WQX-C04KNDYC54"
  )) %>%
  dplyr::rename(
    AssessmentUnitIdentifier = ATTAINS.AssessmentUnitIdentifier,
    MonitoringLocationIdentifier = ATTAINS.MonitoringLocationIdentifier,
    WaterType = ATTAINS.WaterType
  ) %>%
  # Add an example new assessment unit for demonstration purposes.
  dplyr::bind_rows(c(
    AssessmentUnitIdentifier = "NEW:EX_MDEQ_WQ_WQX",
    MonitoringLocationIdentifier = "NARS_WQX-NWC_MT-10184",
    WaterType = "LAKE, FRESHWATER"
  ))


rm(attains.existing.MT, clean.existing.attains.MT, ATTAINS_orgs)
```

### Run `TADA_CreateAUMLCrosswalk`

```{r tada.get.attains}
# use TADA get attains to make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLCrosswalk(
  tada.MT.clean,
  au_ref = user.supplied.cw,
  org_id = "MTDEQ",
  add_catch = FALSE,
  nhd_catch = FALSE,
  batch_upload = TRUE)
```

## Step B: Generate a Map with `TADA_ViewATTAINS`

The **`TADA_ViewATTAINS()`** function creates a map that visually
represents results from three different crosswalk sources using distinct
circle markers:

-   **User-Supplied Crosswalk**: Monitoring locations derived from a
    user-supplied reference are indicated by a user icon.

-   **ATTAINS Crosswalk**: Locations that match with the crosswalk
    downloaded from ATTAINS are marked with a circle that contains a
    check mark.

-   **ATTAINS Catchments/Geospatial Join**: Matches found by joining
    monitoring locations with ATTAINS catchments are shown as solid fill
    circle markers.

Additionally, when a user clicks on any circle marker, a pop-up window
appears displaying the assessment unit crosswalk source for that
particular monitoring location.

```{r view.map}
TADA_ViewATTAINS(MT.AUMLRef, ref_icons = TRUE)
```

## Step C: Update crosswalk in ATTAINS using TADA_UpdateATTAINSAUMLCrosswalk

The output of TADA_UpdateATTAINSAUMLCrosswalk can be uploaded to ATTAINS
to create or update Monitoring Location Identifiers in ATTAINS
Assessment Unit profiles. Users can specify whether all records should
be overwritten (replaced) or if new Monitoring Location Identifiers
should be appended (added) to existing records. Enter
?TADA_UpdateATTAINSAUMLCrosswalk() into the console for more details.

In this example workflow, after making any necessary changes to the
ATTAINS_batchupload data frame generated by TADA_CreateAUMLCrosswalk, we
run TADA_UpdateATTAINSAUMLCrosswalk to add new links to WQP site pages.
This is now ready to upload to ATTAINS.

```{r add.links}
batch.upload.MT <- MT.AUMLRef$ATTAINS_batchupload %>%
  TADA_UpdateATTAINSAUMLCrosswalk( # selected attains_replace = TRUE because all matches currently in ATTAINS are included in this new crosswalk
    attains_replace = TRUE,
    batch_upload = TRUE,
    wqp_data_links = "add",
    # ml ids have already  been corrected if needed
    update_mlid = FALSE,
    org_id = "MTDEQ"
  )
```

## TADA_CreateUseAURef

After assigning ML sites to AU, we will use TADA_CreateUseAURef to
assign uses to these AUs. First, convert the ML sites to AU reference
table to the appropriate format for TADA workflow. We will also assign
the NEW:EX_MDEQ_WQ_WQX assessment unit example to the water type "LAKE,
FRESHWATER" for demonstration on how to auto_assign uses to a new AU
that is not yet in ATTAINS nor has been assigned uses to it yet. To
create the final AU/ML reference will use MT.AUMLRef\$ATTAINS_crosswalk
as it contains ATTAINS.WaterType which is required farther along in the
assessment process.

```{r get.MLAURef}
Final.MT.AUMLRef <- MT.AUMLRef$ATTAINS_crosswalk %>%
  dplyr::mutate(
    ATTAINS.WaterType = dplyr::case_when(
      ATTAINS.AssessmentUnitIdentifier == "NEW:EX_MDEQ_WQ_WQX" ~ "LAKE, FRESHWATER",
      TRUE ~ ATTAINS.WaterType
    )
  )
```

## Step D: Assign uses to AUs with TADA_CreateUseAURef

After assigning WQP MLs to ATTAINS AUs, we will use
`TADA_CreateUseAURef` to assign uses to these AUs. First, we will
convert the ML to AU crosswalk to the appropriate format for this TADA
workflow. We will also assign the *NEW:EX_MDEQ_WQ_WQX* assessment unit
example to the water type *LAKE, FRESHWATER* to demonstrate how to
`auto_assign` uses to a new AU that is not yet in ATTAINS based on water
type.

Now, we can use Final.MT.AUMLRef as our ML to AU crosswalk and proceed
with assigning uses for each unique AU. The assignment of uses to AUs
are prioritized as follows:

1)  (optional) User supplied uses to AU

2)  Imports existing uses by AU from ExpertQuery

```{r assign.use.to.AU}
MT.UseAURef <- TADA_CreateUseAURef(AUMLRef = Final.MT.AUMLRef, org_id = "MTDEQ")
```

# Assign uses to new AUs

For any new AU, users will need to determine how to assign uses to them.

1)  TADA has developed a function called TADA_CreateWaterUseRef which
    imports existing uses by water type from ExpertQuery, and will
    assign any new AU these uses. Users may choose to edit the
    TADA_CreateWaterUseRef according to their preference.

```{r assign.use.to.new.AU.with.waterUseRef}
MT.UseAURef_with_WaterUseRef <- TADA_CreateUseAURef(
  waterUseRef = TADA_CreateWaterUseRef(org_id = "MTDEQ"),
  AUMLRef = Final.MT.AUMLRef,
  org_id = "MTDEQ"
)
```

2)  User may also proceed to do a manual assignment of use names to
    these new AUs if desired rather than using `TADA_CreateWaterUseRef`.

```{r assign.use.to.new.AU.manual}
MT.UseAURef_manual <- MT.UseAURef %>%
  dplyr::left_join(
    data.frame(
      ATTAINS.UseName = c("Aquatic Life", "Drinking Water"),
      ATTAINS.AssessmentUnitIdentifier = c("NEW:EX_MDEQ_WQ_WQX", "NEW:EX_MDEQ_WQ_WQX")
    ),
    by = ("ATTAINS.AssessmentUnitIdentifier")
  ) %>%
  dplyr::mutate(ATTAINS.UseName = dplyr::coalesce(ATTAINS.UseName.x, ATTAINS.UseName.y)) %>%
  dplyr::select(-ATTAINS.UseName.x, -ATTAINS.UseName.y)
```
