---
title: "Geospatial Data Integration"
format: html
editor: visual
authors: 
  - name: "Cristina Mullin"
    orcid: 0000-0002-0615-6087
    affiliation: EPA
  - name: "Marc Weber"
    affiliation: EPA
  - name: "Dave Blodgett"
    affiliation: USGS
  - name: "Michael Dumelle"
    affiliation: EPA
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Geospatial Data Integration}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r knitsetup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = F
)
```

```{css, include = F}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

# Overview

Let's walk through how to create an efficient and reproducible workflow
that integrates several R Packages developed by the U.S. Environmental
Protection Agency (EPA) and the U.S. Geological Survey (USGS) to support
water quality programs (such as the Clean Water Act) and geospatial
(watershed or waterbody) level analyses.

This workflow demonstrates potential uses (beyond their original
collection purpose) for publicly available water quality data from WQP.
To start, participants will learn how to use EPA's Tools for Automated
Data Analysis (TADA) R Package to retrieve, wrangle, harmonize, quality
check, visualize and analyze WQP data from multiple organizations.

Next, we will showcase how to bring in other web services and libraries
for easy integration of additional hydrologic and geospatial data. We
then plan to touch briefly on packages that can assist with building
statistical models. Finally, we will demonstrate an example for
analyzing water quality by Assessment Units (AUs), which are state or
tribal nation defined watershed or waterbody areas used for CWA
assessments and reporting water quality conditions to EPA and the
public.

**Intended Audience**

Water Quality eXchange (WQX) and Water Quality Portal (WQP) community,
Clean Water Act (CWA) community (EPA, States and Tribal Nations), water
quality and geospatial data analysts/researchers, EPA/USGS and other
federal agencies.

**Leveraged R Packages**

-   EPA: EPATADA, StreamCatTools, spsurvey, spmodel, SSN2

-   USGS: dataRetrieval, nhdplusTools, hydroloom

-   Fundamental geospatial packages: sf, prism, terra, leaflet and tmap

# Install and load packages

We will be leveraging the EPATADA R Package for WQP data retrieval,
cleaning, visualization and other steps needed to prepare for analysis.
Let's dive into Green Bay, WI!

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load EPATADA using the remotes package. USGS's
dataRetrieval, EPA's StreamCatTools, and other TADA R Package
dependencies will also be downloaded automatically with the TADA
install.

If desired, the development versions of dataRetrieval and StreamCatTools
can be downloaded directly from GitHub (un-comment).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)

# remotes::install_github("USGS-R/dataRetrieval",
#                         dependencies=TRUE)
#
# remotes::install_github("USEPA/StreamCatTools",
#                         dependencies = TRUE)
```

```{r dev_install, eval = F, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the TADA R Package into
your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

# Start timer!

It's go time! Let's time our process.

```{r startime}
# Record start time
start.time <- Sys.time()
```

# WQP data discovery and cleaning

## EPATADA R Package

This is an abbreviated introduction to key TADA Module 1 WQP Data
Discovery and Cleaning functions. Additional functions and a more
detailed example workflow is available
[here](https://usepa.github.io/EPATADA/articles/TADAModule1.html).

### Module 1: Retrieve, wrangle, filter, flag, harmonize

#### Retrieve and autoclean

In this example, we will first use EPA's How's My Waterway (HMW)
application to find an applicable Hydrologic Unit Code (HUC) for our
area of interest - the [Fox River, Green Bay,
WI](https://mywaterway.epa.gov/community/040302040405/monitoring). Next,
let's query the WQP using the identified HUC, state abbreviation, and a
date range. In this example, we'll start by pulling all data available
in the WQP for this HUC 12 in Wisconsin for the last 5 years.

WATERSHED: City of Green Bay-Fox River (040302040405)

```{r getdata}
# Uncomment to query the WQP
GreenBay_FoxRiver <- TADA_DataRetrieval(
  statecode = "WI",
  startDate = "2015-01-01",
  endDate = "2024-12-30",
  huc = c("040302040405"),
  applyautoclean = TRUE,
  ask = FALSE
)

# GreenBay_FoxRiver <- NMCWorkshopData::GreenBay_FoxRiver
```

In the default arguments for TADA_DataRetrieval, applyautoclean = TRUE.
This runs TADA_AutoClean on the newly retrieved data frame.
TADA_AutoClean is a powerful function which performs a variety of tasks
including: (1) creating new "TADA" prefixed columns and and capitalizing
their contents to reduce case sensitivity issues, (2) converts special
characters in value columns, (3) converts latitude and longitude values
to numeric, (4) replaces "meters" with "m", (5) replaces deprecated
characteristic names with current WQX names, (6) harmonizes result and
detection limit units to WQX, TADA or user supplied target units, (7)
converts depths to meters, and (8) creates the column
TADA.ComparableDataIdentifier by concatenating characteristic name,
result sample fraction, method speciation, and result measure unit.

Now, let's use EPATADA functions to review, visualize, and whittle the
returned WQP data down to include only results that are applicable to
our water quality analysis and area of interest.

#### **Duplicates from a single organization**

We can identify data records uploaded by the same organization with the
same date, time, monitoring location, activity type, characteristic
name, fraction, taxonomic name, depth columns, and result value and
flags them as potential duplicates. The data user must determine if the
data records are unique or represent overlap that could cause issues in
analysis. For this example, we will retain only results flagged as
"Unique".

```{r single-org-dups}
# find duplicate results submitted by single org
GreenBay_FoxRiver <- TADA_FindPotentialDuplicatesSingleOrg(GreenBay_FoxRiver)

# retain unique flagged results
GreenBay_FoxRiver <- dplyr::filter(GreenBay_FoxRiver, TADA.SingleOrgDup.Flag == "Unique")
```

#### **Nondetects**

TADA provides some simple methods for dealing with censored results,
such as multiplying the detection limit by a user supplied value or
leaving the result as is.

```{r censored}
# substitude nondetects with 0.5 detection limit, leave overdetects as is
GreenBay_FoxRiver <- TADA_SimpleCensoredMethods(GreenBay_FoxRiver, nd_method = "multiplier", nd_multiplier = 0.5, od_method = "as-is", od_multiplier = "null")
```

#### **Duplicates from multiple organizations**

Two organizations sometimes submit the same exact data to WQP. Filtering
out these duplicates can prevent issues in analysis.

```{r multi-org-dups, eval = F}
# find potential dups multiple orgs
GreenBay_FoxRiver <- TADA_FindPotentialDuplicatesMultipleOrgs(GreenBay_FoxRiver)

# filter out
GreenBay_FoxRiver <- dplyr::filter(GreenBay_FoxRiver, TADA.ResultSelectedMultipleOrgs == "Y")
```

#### **Quality and other issues**

Filter out any remaining irrelevant data, NAs and empty columns.

```{r autofilter}
unique(GreenBay_FoxRiver$TADA.ResultMeasureValueDataTypes.Flag)

sum(is.na(GreenBay_FoxRiver$TADA.ResultMeasureValue))

GreenBay_FoxRiver <- TADA_ConvertSpecialChars(GreenBay_FoxRiver, col = "TADA.ResultMeasureValue", clean = TRUE)
```

Check to make sure there are no more NAs in TADA.ResultMeasureValue.

```{r recheck}
unique(GreenBay_FoxRiver$TADA.ResultMeasureValueDataTypes.Flag)

sum(is.na(GreenBay_FoxRiver$TADA.ResultMeasureValue))

GreenBay_FoxRiver <- GreenBay_FoxRiver |>
  dplyr::filter(!is.na(TADA.ResultMeasureValue))

sum(is.na(GreenBay_FoxRiver$TADA.ResultMeasureValue))
```

Flag or remove QAQC samples and suspect results.

```{r flag}
GreenBay_FoxRiver <- TADA_RunKeyFlagFunctions(GreenBay_FoxRiver, clean = TRUE)
```

Flag results above and below threshold, but do not remove them

```{r thresholds}
GreenBay_FoxRiver <- TADA_FlagAboveThreshold(GreenBay_FoxRiver, clean = FALSE, flaggedonly = FALSE)

GreenBay_FoxRiver <- TADA_FlagBelowThreshold(GreenBay_FoxRiver, clean = FALSE, flaggedonly = FALSE)
```

#### **Harmonize synonyms across characteristic, fraction, and speciation**

```{r harmonize}
GreenBay_FoxRiver <- TADA_HarmonizeSynonyms(GreenBay_FoxRiver)
```

#### **Calculate Total N and Total P from various species and fractions**

```{r TN-TP}
GreenBay_FoxRiver <- TADA_CalculateTotalNP(GreenBay_FoxRiver, daily_agg = "max")
```

#### Review and filter

Review unique characteristic, fraction, and species combinations

```{r table}
GreenBay_FoxRiver_Counts <- TADA_FieldValuesTable(GreenBay_FoxRiver, field = "TADA.ComparableDataIdentifier")

DT::datatable(GreenBay_FoxRiver_Counts, fillContainer = TRUE)
```

Filter to focus on frequently monitored characteristics in example data

```{r subset}
GreenBay_FoxRiver_Subset <- GreenBay_FoxRiver |>
  dplyr::filter(TADA.ComparableDataIdentifier %in%
    c(
      "SPECIFIC CONDUCTANCE_NA_NA_US/CM",
      "PH_NA_NA_NA",
      "TOTAL NITROGEN, MIXED FORMS_UNFILTERED_AS N_MG/L",
      "TOTAL PHOSPHORUS, MIXED FORMS_UNFILTERED_AS P_UG/L",
      "DISSOLVED OXYGEN (DO)_NA_NA_MG/L"
    ))
```

Review organizations for subset

```{r org-review}
# Create pie of results by organization
TADA_FieldValuesPie(GreenBay_FoxRiver_Subset, field = "OrganizationFormalName")
```

### Exploratory visualizations

Generate stats table

```{r stats}
GreenBay_FoxRiver_Subset_Stats <- TADA_Stats(GreenBay_FoxRiver_Subset)

DT::datatable(GreenBay_FoxRiver_Subset_Stats, fillContainer = TRUE)
```

Generate scatterplot

```{r scatterplot}
TADA_TwoCharacteristicScatterplot(GreenBay_FoxRiver_Subset, id_cols = "TADA.ComparableDataIdentifier", groups = c("TOTAL PHOSPHORUS, MIXED FORMS_UNFILTERED_AS P_UG/L", "TOTAL NITROGEN, MIXED FORMS_UNFILTERED_AS N_MG/L"))
```

Generate map

```{r TADA Overview Map}
TADA_OverviewMap(GreenBay_FoxRiver_Subset)
```

#### **Coordinate issues**

```{r Coordinate issues}
# Change coordinate sign if appropriate
GreenBay_FoxRiver <- TADA_FlagCoordinates(GreenBay_FoxRiver_Subset, clean_outsideUSA = "change sign", clean_imprecise = FALSE)

# This df has NA lons from USGS that must be addressed before TADA_MakeSpatial can be run...
sum(is.na(GreenBay_FoxRiver_Subset$LongitudeMeasure))

# Remove rows with NA lons from df
GreenBay_FoxRiver_Subset <- GreenBay_FoxRiver_Subset[!is.na(GreenBay_FoxRiver_Subset$LongitudeMeasure), ]

# Recheck
sum(is.na(GreenBay_FoxRiver_Subset$LongitudeMeasure))
```

# Geospatial data integration

Next, we will show how to integrate additional geospatial data with the
"clean" dataset we just created from the WQP.

## Make spatial

First, leverage TADA_MakeSpatial to transform a WQP dataframe into a
geospatial sf object.

```{r}
GreenBay_FoxRiver_sf <- TADA_MakeSpatial(GreenBay_FoxRiver_Subset)
```

Then create a unique identifier based on shared lat long values and
filter to just the 25 unique locations.

```{r}
GreenBay_FoxRiver_sf$latlon <- paste0(GreenBay_FoxRiver_sf$TADA.LongitudeMeasure, GreenBay_FoxRiver_sf$TADA.LatitudeMeasure)

GreenBay_FoxRiver_sf <- GreenBay_FoxRiver_sf |>
  dplyr::group_by(latlon) |>
  dplyr::mutate(loc_id = dplyr::cur_group_id())

GreenBay_FoxRiver_sf_locs <- GreenBay_FoxRiver_sf |>
  dplyr::filter(!duplicated(loc_id))
```

## Access NHDPlus COMIDs for sites

We use `StreamCatTools` function `sc_get_comid` (which uses an
`nhdplusTools` web service client) to get the comid for each location.

```{r, eval = F}
GreenBay_FoxRiver_sf_locs$COMID <- as.integer(strsplit(StreamCatTools::sc_get_comid(GreenBay_FoxRiver_sf_locs), split = ",")[[1]])

nhdplus_data <- nhdplusTools::subset_nhdplus(GreenBay_FoxRiver_sf_locs$COMID, nhdplus_data = "download")

outlet <- dplyr::filter(nhdplus_data$NHDFlowline_Network, hydroseq == min(hydroseq))

nhdplusTools::plot_nhdplus(bbox = sf::st_bbox(outlet))
plot(sf::st_transform(sf::st_geometry(GreenBay_FoxRiver_sf_locs), 3857), add = TRUE)
```

## dataRetrieval/NLDI, nhdplusTools, hydroloom

Do a network navigation and get NHDPlus for our data. Note that the
network navigation only includes flowline geometry. `nhdplusTools`
subsets all of the NHDPlus.

```{r, eval = F}
all_network <- dataRetrieval::findNLDI(comid = outlet$comid, nav = "UT", distance_km = 500)

# we could select only comids on network
nhdplus_data <- nhdplusTools::subset_nhdplus(comids = as.integer(all_network$UT_flowlines$nhdplus_comid), nhdplus_data = "download", flowline_only = FALSE)

# or we could just get everything in the bbox to be sure we get non-network stuff too!
nhdplus_data <- nhdplusTools::subset_nhdplus(
  bbox = sf::st_bbox(all_network$UT_flowlines),
  nhdplus_data = "download",
  flowline_only = FALSE
)

# see ?nhdplusTools::subset_nhdplus for lots more options!

sapply(nhdplus_data, nrow)

sapply(nhdplus_data, names)
```

### Addressing sites to the network

There are two forms of hydrographic addresses: catchment indexing and
linear referencing. The former is established with a point in polygon
analysis. The latter is more nuanced. The following block shows how to
establish both with the data we just retrieved.

Note that hydroloom is compatible with nhdplus and other attribute
systems. See [hydroloom documentation for
more!](https://doi-usgs.github.io/hydroloom/articles/hydroloom.html)

```{r, eval = F}
GreenBay_FoxRiver_sf_locs <- sf::st_join(
  GreenBay_FoxRiver_sf_locs,
  hydroloom::st_compatibalize(
    dplyr::select(nhdplus_data$CatchmentSP, featureid),
    GreenBay_FoxRiver_sf_locs
  )
)

# NOTE that featureid and comid are the same!!
all(GreenBay_FoxRiver_sf_locs$COMID == GreenBay_FoxRiver_sf_locs$featureid)

(linear_references <- hydroloom::index_points_to_lines(
  nhdplus_data$NHDFlowline_Network,
  GreenBay_FoxRiver_sf_locs
))

GreenBay_FoxRiver_sf_locs <- dplyr::bind_cols(GreenBay_FoxRiver_sf_locs, linear_references)
```

We can take this one step further by indexing points to waterbodies! The
return here tells us what waterbody our locations are near or within.
For on-network waterbodies, it will also include the outlet flowline for
each waterbody.

```{r, eval = F}
all_wb <- dplyr::bind_rows(
  dplyr::select(nhdplus_data$NHDWaterbody, wbid = comid),
  dplyr::select(nhdplus_data$NHDArea, wbid = comid)
)

(waterbody_indexes <- hydroloom::index_points_to_waterbodies(
  sf::st_transform(all_wb, 5070),
  GreenBay_FoxRiver_sf_locs,
  flines = nhdplus_data$NHDFlowline_Network,
  search_radius = units::as_units(1000, "m")
))
```

```{r, eval = F}
par(mar = c(0, 0, 0, 0))
nhdplusTools::plot_nhdplus(
  bbox = sf::st_bbox(GreenBay_FoxRiver_sf),
  cache_data = tempfile(fileext = ".rds")
)
plot(
  sf::st_transform(
    all_wb[all_wb$wbid %in% waterbody_indexes$near_wbid, ],
    3857
  ),
  add = TRUE,
  col = "darkblue", border = NA
)
plot(sf::st_transform(sf::st_geometry(GreenBay_FoxRiver_sf_locs), 3857), add = TRUE, col = "white")
```

There's much much more where that came from. See the pkgdown sites for
[nhdplusTools](https://doi-usgs.github.io/nhdplusTools/) and
[hydroloom](https://doi-usgs.github.io/hydroloom/index.html) for more!

## [StreamCatTools](https://usepa.github.io/StreamCatTools/index.html)

### Accessing watershed information for sites

We can access watershed information for each unique site location,
getting both the landscape data for local catchment or the full upstream
watershed for each particular site using
[StreamCatTools](https://usepa.github.io/StreamCatTools/index.html)

**Discover what StreamCat metrics we might want to use**

```{r, eval = F}
metrics <- StreamCatTools::sc_get_params(param = "metric_names")
print(paste0("A selection of available StreamCat metrics include: ", paste(metrics[1:10], collapse = ", ")))
```

**Discover land cover of watersheds for sites**

We'll pull in all the NLCD categories at the local catchment level for
each location

```{r, eval = F}
GB_FR_NLCD <- StreamCatTools::sc_nlcd(year = "2019", aoi = "cat", comid = GreenBay_FoxRiver_sf_locs$COMID)


GB_FR_Urb <- GB_FR_NLCD |>
  dplyr::mutate(Pct_Urbanized = pcturbop2019cat + pcturbmd2019cat + pcturblo2019cat + pcturbhi2019cat) |>
  dplyr::select(comid, Pct_Urbanized)
GB_FR_Urb
```

**Visualize urbanization for local catchment for each location**

```{r, eval = F}
ggplot2::ggplot(GB_FR_Urb, ggplot2::aes(x = Pct_Urbanized)) +
  ggplot2::geom_density()
```

### Pull in data for modeling

Now we'll just demonstrate pulling in watershed data that we might use
in a modeling exercise as spatial covariates

```{r, eval = F}
ws_data <- StreamCatTools::sc_get_data(metric = "fert,nsurp,nani,manure,IWI", aoi = "cat,ws", comid = GreenBay_FoxRiver_sf_locs$COMID)
```

# Building statistical models

The following information is not integrated into the coding exercise,
but provides information on two additional EPA R packages that can be
leveraged to build statistical models.

**Spatial Dependence**

-   For spatial data, nearby observations tend to be more similar than
    distant observations

-   This phenomena is called *spatial dependence* and can be built into
    statistical models

-   The benefits of incorporating spatial dependence are *significant*
    and include:

    -   More realistic characterization of ecological drivers
    -   More precise predictions at unobserved locations

## `spmodel`

-   The `spmodel` R package makes spatial models accessible via
    straightforward extensions to common modeling functions like `lm()`
    and `glm()`
-   Spatial dependence is based on Euclidean (straight-line) distance
-   Learn more at <https://usepa.github.io/spmodel/>

## `SSN2`

-   Like `spmodel`, `SSN2` extends common modeling functions like `lm()`
    and `glm()`
-   Spatial dependence is based on stream network distance
    (flow-connected, flow-unconnected)
-   `SSN2` is an updated version of `SSN` (`SSN` has been archived)
-   Learn more at <https://usepa.github.io/SSN2/>

# CWA use case

## EPATADA R Package

### Module 2: Geospatial Functions

Additional functions and a more detailed example workflow is available
here: <https://usepa.github.io/EPATADA/articles/TADAModule2.html>

**CWA Use Case**

We do not have time to cover the full CWA assessment process today.
Let's focus on geospatial aspects! For more information about CWA
assessments, see: [Integrated Reporting Memoranda under CWA Sections
303(d), 305(b) and
314](https://www.epa.gov/tmdl/Integrated%20Reporting%20Guidance%20under%20CWA%20Sections%20303%28d%29%2C%20305%28b%29%20and%20314).

**What are Assessment Units?**

Assessment units are the geospatial areas used for spatial analyses of
water quality data. CWA assessment determinations are made by assessment
unit, meaning the entire assessment unit is assessed as either meeting
or not meeting water quality standards (i.e., thresholds or criteria)
for assigned uses.

**How are assessment units delineated?**

Assessment units are typically delineated by using watershed-oriented
collections of stream reaches, often broken down by physical features
like waterfalls, bridge crossings, or changes in land use, to analyze
water quality impairments within a specific area, ensuring data
homogeneity and spatial clarity within the assessment unit. Existing
state and tribal Assessment Units are available from ATTAINS geospatial
services.

One of the first steps in the CWA assessment process is to define
Assessment Units and associate data with them. A major source for water
quality data is the WQP.

-   Assessment Units: state or tribal waterbody geospatial features
    -   These may be lines, areas or points
-   Water Quality Portal Monitoring Locations
    -   These are points

#### **Matching WQP monitoring locations with ATTAINS assessment units**

-   TADA_CreateATTAINSAUMLCrosswalk() automates matching of WQP
    monitoring locations with ATTAINS assessment units that fall within
    (intersect) the same NHDPlus catchment
    ([details](https://usepa.github.io/EPATADA/articles/TADAModule2.html))
-   The function uses high resolution NHDPlus catchments by default
    because 80% of state submitted assessment units in ATTAINS were
    developed based on high res NHD; users can select med-res if
    applicable to their use case

```{r}
WQP_with_ATTAINSonly <- TADA_CreateATTAINSAUMLCrosswalk(GreenBay_FoxRiver_Subset, return_sf = TRUE)

TADA_ViewATTAINS(WQP_with_ATTAINSonly)
```

**Challenges with Automated Approach**

-   Certain NHDPlus high res catchments overlap multiple ATTAINS
    assessment units (state submitted hydrography) which means the sites
    are assigned to both AUs in the current functions. Another challenge
    is that the WQP sites are not always accurate (imprecise
    coordinates). WQP location metadata may also be helpful for
    matching/QAQC'ing waterbody names with ATTAINS waterbody names
    instead of relying solely on the lat/long and geospatial/mapping
    information. Users must manually review associations for accuracy.

#### Analyzing areas outside of ATTAINS assessment unit areas

*See: nhdplusTools, TADA::fetchNHD(), TADA_CreateATTAINSAUMLCrosswalk(),
TADA_ViewATTAINS()*

Some waterbodies have data available in the WQP or from other sources,
but there are no existing Assessment Units in ATTAINS for them. If you
are interested in analyzing WQP data available in areas that fall
outside of the ATTAINS assessment unit areas (using all available data
from the WQP), TADA has included a way to explore this using
TADA_CreateATTAINSAUMLCrosswalk() fill_USGS_catch function input
(leveraging USGS's nhdplusTools and TADA geospatial functions). This is
included for exploratory purposes only. In theory, states and tribal
nations could use the high res catchments as new assessment unit
polygons to assess additional areas where there is WQP data but no
Assessment Unit yet in ATTAINS, but that process is outside of TADA.

For WQP monitoring sites that DO NOT overlap an existing ATTAINS feature
(neither ATTAINS NHDPlus high res catchment snap shot or state submitted
points/lines/polys), there is nothing to use from ATTAINS because these
are areas where there is WQP data but no ATTAINS Assessment Unit yet.

For these, we implemented a solution using NHDPlusTools to pull in
either NHDPlus high res or med res catchments (user can choose, but high
res is the default) and match those with the WQP sites & create new IDs
(essentially creating new AUs that are the catchments that intersect
these WQP sites).

```{r}
WQP_withATTAINSandNHDPluscatchments <- TADA_CreateATTAINSAUMLCrosswalk(GreenBay_FoxRiver_Subset, fill_USGS_catch = TRUE, return_sf = TRUE)

TADA_ViewATTAINS(WQP_withATTAINSandNHDPluscatchments)
```

# Stop timer!

```{r}
end.time <- Sys.time()

end.time - start.time
```

# Contribute

Note: TADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.

If you are interested in contributing to TADA development, more
information is available at:

<https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html>

We welcome collaboration with external partners.

Contribute to EPATADA in a way that helps elevate work you have already
done, broadens the user base of the package, or improves the resource
for all!

**Thank you to our workshop contributors!**

-   EPA: Cristina Mullin (mullin.cristina\@epa.gov), Marc Weber, Hillary
    Marler, Kenny Wong, Michael Dumelle, Shelly Thawley

-   USGS: Dave Blodgett
