---
title: "Nutrient Data: Water Quality Portal Data Discovery and Cleaning"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Nutrient Data: Water Quality Portal Data Discovery and Cleaning}
  %\VignetteEngine{knitr::rmarkdown}
description: An overview of TADA functionalities to clean, harmonize and calculate Total Nitrogen and Total Phosphorus data from the WQP. This vignette will go through a standardized and consistent workflow for a variety of users interested in using WQP data for nutrient analysis. 
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{css, include = F}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

## Overview

This vignette walks through how to use the TADA R Package to discover
and clean (i.e., wrangle, Quality Assure and Quality Control (QAQC), and
harmonize) [Water Quality Portal
(WQP)](https://www.waterqualitydata.us/) data from multiple
organizations.

## Install and Load the EPATADA R Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes")
# Load the remotes library
library(remotes)
```

Next, install and load TADA using the remotes package. TADA R Package
dependencies will also be downloaded automatically from CRAN with the
TADA install. You may be prompted in the console to update dependency
packages that have more recent versions available. If you see this
prompt, it is recommended to update all of them (enter 1 into the
console).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the TADA R Package into
your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Background

This vignette summarizes cross-project efforts within EPA to discover,
harmonize, and analyze nutrient data from the WQP, with a focus on
reproducibly calculating Total Nitrogen (TN) and Total Phosphorus (TP)
from available fraction and sub-species. It outlines shared challenges,
the TADA workflow, and related tools and updates that may affect your
data retrieval and processing.

Multiple projects across EPA have experienced challenges with
harmonization and QAQC of nutrient data from WQP. WQX has a [best
practices document for nutrient data
submission](https://gcc02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fwww.epa.gov%2Fwaterdata%2Fwater-quality-data-upload-wqx&data=05%7C02%7CWong.Kenny%40epa.gov%7C59f010b61c194df4a2de08decadedc10%7C88b378b367484867acf976aacbeca6a7%7C0%7C0%7C639171255679011335%7CUnknown%7CTWFpbGZsb3d8eyJFbXB0eU1hcGkiOnRydWUsIlYiOiIwLjAuMDAwMCIsIlAiOiJXaW4zMiIsIkFOIjoiTWFpbCIsIldUIjoyfQ%3D%3D%7C0%7C%7C%7C&sdata=TfvQIaSAUJeMT7Fdxm%2BoBzpbY4SWpXBn%2FhlJptLlfwU%3D&reserved=0),
but not all data submitters follow it. TADA uses the the logic with ECHO
WQI logic for WQP nutrient data. TADA has further expanded this effort
for additional missing characteristic and speciation combinations that
are available in WQP.

## Case Studies

These projects include:

1)  National Nutrient Inventory (NNI)

2)  Hypoxia Task Force (HTF) Monitoring Group

3)  ECHOS and ICIS integration

4)  Freshwater Explorer

## Relevant TADA functions

-   [TADA_DataRetrieval()](https://usepa.github.io/EPATADA/reference/TADA_DataRetrieval.html)

-   [TADA_HarmonizeSynonyms()](https://usepa.github.io/EPATADA/reference/TADA_HarmonizeSynonyms.html)

-   [TADA_CalculateTotalNP()](https://usepa.github.io/EPATADA/reference/TADA_CalculateTotalNP.html)

## TADA_DataRetrieval

WQP data is retrieved and processed for compatibility with TADA. This
function, **TADA_DataRetrieval**, builds on USGS's dataRetrieval R
package functions. It joins three WQP profiles: Site, Sample Results
(physical/chemical metadata), and Project. In addition, it changes all
data in the Characteristic, Speciation, Fraction, and Unit fields to
uppercase and addresses result values that include special characters.

This function accepts the same inputs as the dataRetrieval `readWQPdata`
function. `readWQPdata` does not restrict the characteristics pulled
from [Water Quality Portal (WQP)](https://www.waterqualitydata.us/).

Data retrieval filters include:

-   startDate

-   endDate

-   characteristicName

-   sampleMedia

-   siteType

-   statecode (see list of possible state and territory abbreviations
    [here](https://www2.census.gov/geo/docs/reference/state.txt))

-   countycode

-   siteid

-   organization

-   project

-   huc

-   characteristicType

-   providers

In addition to these filters, TADA_DataRetrieval accepts additional
geospatial-related filters that are not included in the dataRetrieval
`readWQPdata` function:

-   aoi_sf

-   tribal_area_type

-   tribe_name_parcel (Note: The TADA_TribalOptions function can be used
    to narrow down options for use with this tribe_name_parcel filter
    option. See ?TADA_TribalOptions for more info).

After data is downloaded using the filters above, the default
TADA_DataRetrieval function also automatically runs the
**TADA_AutoClean** function. If desired, users can set applyautoclean =
FALSE in their TADA_DataRetrieval calls.

Tips:

1.  All the query filters for the WQP work as an AND but within the
    fields there are ORs. For example:

    -   Characteristics: If you choose NITROGEN & PHOSPHORUS - it's an
        OR. This means you will retrieve both NITROGEN OR PHOSPHORUS
        data if available.

    -   States: Similarly, if you choose VA and IL, it's an OR. This
        means you will retrieve both VA OR IL data if available.

    -   Combinations of fields are ANDs, such as State/VA AND
        Characteristic/DO". This means you will receive all DO data
        available in VA.

    -   "Characteristic" and "Characteristic Type" also work as an AND.
        This means that the Characteristic must fall within the
        CharacteristicGroup if both filters are being used, if not you
        will get an error. If you would like all available nutrient
        labeled data, it is recommended to query by CharacteristicGroup
        = "Nutrient" and leave "Characteristic" blank.

2.  The "siteid" is a general term WQP uses to describe both Site IDs
    from USGS databases and Monitoring Location Identifiers (from WQX).
    Each monitoring location in the Water Quality Portal (WQP) has a
    unique Monitoring Location Identifier, regardless of the database
    from which it derives. The Monitoring Location Identifier from the
    WQP is the concatenated Organization Identifier plus the Site ID
    number. Site IDs that only include a number are only unique
    identifiers for monitoring locations within USGS NWIS or EPA's WQX
    databases separately.

3.  The aoi_sf and tribal arguments are meant to be used on their own.
    For example, if both an aoi_sf argument and tribal information are
    provided an error is returned because it's unclear what the priority
    location should be for the query. Similarly, aoi_sf and
    tribal_area_type are not meant to be used with location-related
    filters (e.g., statecode, siteid). In these instances a warning is
    returned but the query proceeds by using only the aoi_sf or
    tribal_area_type information.

Additional resources:

-   Review function documentation by entering the following code into
    the console: ?TADA_DataRetrieval

-   [Introduction to the dataRetrieval
    package](https://doi-usgs.github.io/dataRetrieval/)

-   [Water Quality Portal Web Services
    Guide](https://www.waterqualitydata.us/webservices_documentation/)

-   [dataRetrieval
    Tutorial](https://waterdata.usgs.gov/blog/dataretrieval/)

Use the code below to download data from the WQP using
TADA_DataRetrieval. Edit the code chunk below to define your own WQP
query inputs.

Downloads using TADA_DataRetrieval will have the same columns each time,
but be aware that data are uploaded to the Water Quality Portal by
individual organizations, which may or may not follow the same
conventions. Data and metadata quality are not guaranteed! Carefully
explore data to make sure it meets your quality assurance requirements.

Note: TADA_DataRetrieval (by leveraging dataRetrieval), automatically
converts the date times to UTC. It also automatically converts field
formats to dates, datetimes, and numerics based on a standard algorithm.

Enter ?TADA_DataRetrieval into the console to review example queries and
additional information.

To retrieve data from the Water Quality Portal (WQP) using a bounding
box (bbox), we first define a bbox that encloses four HTF-priority sites
in Mississippi. We’ll use this spatial extent to identify existing WQP
data within the area. If broader coverage is needed, we can expand the
bbox to include a wider extent and capture additional results beyond the
initial boundary.

```{r}
MS_WQP_PrioritySites <- TADA_DataRetrieval(
  startDate = "2000-01-01",
  endDate = "2020-12-31",
  # bBox = c(-90.914167, 32.018, -90.696944, 32.444167), # original bbox
  bBox = c(-91.179367, 31.793500, -90.431744, 32.668667), # bbox expanded 25km
  #characteristicName = c("Phosphorus","Nitrogen","Nitrate"),
  characteristicType = "Nutrient", # also try it without this filter (all chars) - check to see if there's any missing nutrient data
  applyautoclean = TRUE,
  ask = FALSE
  )
```

Data submitter may define fraction, speciation and units for nutrient
data differently. TADA aims to capture all these unique identifiers with
the TADA column: TADA.ComparableDataIdentifier which concatenates the
characteristic, speciation, fraction and units into one string. How many
unique TADA.ComparableDataIdentifier do we see?

```{r}
TADA_FieldValuesTable(MS_WQP_PrioritySites, field =
"TADA.ComparableDataIdentifier")

```

This vignette will showcase how to efficiently clean and harmonize
synonyms and calculate total nutrient summation from this data. First,
let's work on cleaning up our WQP data frame. We will start off with
filtering by surface waters only

```{r}
MS_WQP_PrioritySites2 <- TADA_MediaFilter(MS_WQP_PrioritySites,
                                clean = TRUE,
                                surface_water = FALSE,
                                ground_water = TRUE,
                                sediment = TRUE,
                                other = TRUE
)
```

view TADA overview map, see result counts, number of char per site and
site details

```{r}
TADA_OverviewMap(MS_WQP_PrioritySites)
```

clean for invalid coordinates

```{r}
MS_WQP_PrioritySites3 <- TADA_FlagCoordinates(MS_WQP_PrioritySites2, clean_outsideUSA = "no", clean_imprecise = TRUE)
```

run key flagging functions and remove suspect results (result units,
method-speciation flags, fraction flag, MeasureQualifierCode and
Activity Type)

```{r}
MS_WQP_PrioritySites4 <- TADA_RunKeyFlagFunctions(MS_WQP_PrioritySites3, clean = TRUE)
```

identify potential dup sites

```{r}
MS_WQP_PrioritySites5 <- MS_WQP_PrioritySites4 |>
  TADA_FindPotentialDuplicatesSingleOrg() |>
  dplyr::filter(TADA.SingleOrgDup.Flag == "Unique") |>
  TADA_FindPotentialDuplicatesMultipleOrgs(
    dist_buffer = 100,
    org_hierarchy = "none"
  ) |>
  dplyr::filter(TADA.ResultSelectedMultipleOrgs == "Y")
```

proceed with identifying censored data, if any. Define handling of
censored data.

```{r}
MS_WQP_PrioritySites6 <- TADA_SimpleCensoredMethods(MS_WQP_PrioritySites5,
                                                nd_method = "multiplier",
                                                nd_multiplier = 0.5,
                                                od_method = "as-is",
                                                od_multiplier = "null"
)
```

review for censored data

```{r}
TADA_FieldValuesPie(MS_WQP_PrioritySites6, field = "TADA.CensoredData.Flag")
```

harmonize synonyms

```{r}
MS_WQP_PrioritySites7 <- TADA_HarmonizeSynonyms(MS_WQP_PrioritySites6)
```

uses echos logic to append rows for TN & TP summation

```{r}
TotalNP <- TADA_CalculateTotalNP(MS_WQP_PrioritySites7)
```

view final cleaned table

```{r}
TADA_FieldValuesTable(TotalNP, field = "TADA.ComparableDataIdentifier")
```

### Case Study 1: Hypoxia Task Force

For this case study the HTF monitoring group is interested in
determining where to prioritize funds for a list of priority monitoring
sites using NPS grant funds.

Looking at historical 20 years of nutrient data (from 2000 to 2020) for
a certain region of Mississippi, let's determine if there are any
existing overlapping WQP sites with these priority sites of interest. We
are also interested in determining if there are prior water bodies that
have been assessed in prior ATTAINS assessment cycles near these
priority sites.

First, let's look at our cleaned nutrient data on the map. We will run
TADA_CreateAUMLCrosswalk to determine what sites can be matched to
ATTAINS Assessment Units through a geospatial join or if any exists
directly from ATTAINS in this region.

Crosswalk ATTAINS AU and ML sites

```{r}
AUML <- TADA_CreateAUMLCrosswalk(TotalNP, org_id = "21MSWQ")
# View ATTAINS AU and ML sites map
TADA_ViewATTAINS(AUML)
```

Now let's add our priority sites to the map. These lat/longs are
provided from the HTF list of priority sites appendix.

Add dummy variables with priority sites in the WQP data frame to map it
in TADA (edit in excel)

```{r}
extra_sites <- tibble::tibble(
  OrganizationIdentifier = "USGS",
  TADA.MonitoringLocationIdentifier = c("USGS-07288955", "USGS-322023090544500", "USGS-07290000", "21MSWQ_WQX-07290650"),
  TADA.MonitoringLocationName = c("Priority Site 1", "Priority Site 2", "Priority Site 3", "Priority Site 4"),
  TADA.MonitoringLocationTypeName = "River/Stream",
  LatitudeMeasure = c(32.444167, 32.339722, 32.347778, 32.018),
  TADA.LatitudeMeasure = c(32.444167, 32.339722, 32.347778, 32.018),
  LongitudeMeasure = c(-90.914167, -90.9125, -90.696944, -90.877194),
  TADA.LongitudeMeasure = c(-90.914167, -90.9125, -90.696944, -90.877194),
  .source = "overlay"  # optional flag to tell them apart later
) |>
  TADA_CorrectColType()
```

Make sure types match and bind to data

```{r}
MS_WQP_modified <- dplyr::bind_rows(
  TotalNP |>
    dplyr::mutate(.source = ".wqp"),
  extra_sites
) |> TADA_CorrectColType()
```

Identify these priority sites on the map (include one existing AU row as
TADA_CreateAUMLCrosswalk() currently crashes if no AUML from a user
supplied ref is found in the data frame.

```{r}
flag_priority_AURef <- tibble::tibble(
  OrganizationIdentifier = "USGS",
  TADA.MonitoringLocationIdentifier = c("USGS-07290000", "USGS-07288955", "USGS-322023090544500", "USGS-07290000", "21MSWQ_WQX-07290650"),
  TADA.MonitoringLocationName = c("NA", "Priority Site 1", "Priority Site 2", "Priority Site 3", "Priority Site 4"),
  ATTAINS.OrganizationIdentifier = c("21MSWQ", "21MSWQ", "21MSWQ", "21MSWQ", "21MSWQ"),
  ATTAINS.AssessmentUnitIdentifier = c("MS107811", "Flag1", "Flag2", "Flag3", "Flag4"),
  ATTAINS.WaterType = c("STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER")
) |>
  TADA_CorrectColType()
```

recreate the AUML crosswalk with the dummy priority sites

```{r}
AUML2 <- TADA_CreateAUMLCrosswalk(MS_WQP_modified, org_id = "21MSWQ", au_ref = flag_priority_AURef)
# view the map, examine where sites overlap (exact), where there exists nearby sites, and if ATTAINS assessments are found prior.
TADA_ViewATTAINS(AUML2)
```
