---
title: "Nutrient Data: Water Quality Portal Data Discovery and Cleaning"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Nutrient Data: Water Quality Portal Data Discovery and Cleaning}
  %\VignetteEngine{knitr::rmarkdown}
description: An overview of TADA functionalities to clean, harmonize and calculate Total Nitrogen and Total Phosphorus data from the WQP. This vignette will go through a standardized and consistent workflow for a variety of users interested in using WQP data for nutrient analysis. 
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{css, include = F}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

## Overview

This vignette walks through how to use the TADA R Package to discover
and clean (i.e., wrangle, Quality Assure and Quality Control (QAQC), and
harmonize) [Water Quality Portal
(WQP)](https://www.waterqualitydata.us/) data from multiple
organizations.

## Install and Load the EPATADA R Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes")
# Load the remotes library
library(remotes)
```

Next, install and load TADA using the remotes package. TADA R Package
dependencies will also be downloaded automatically from CRAN with the
TADA install. You may be prompted in the console to update dependency
packages that have more recent versions available. If you see this
prompt, it is recommended to update all of them (enter 1 into the
console).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the TADA R Package into
your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Background

This vignette summarizes cross-project efforts within EPA to discover,
harmonize, and analyze nutrient data from the WQP, with a focus on
reproducibly calculating Total Nitrogen (TN) and Total Phosphorus (TP)
from available fraction and sub-species. It outlines shared challenges,
the TADA workflow, and related tools and updates that may affect your
data retrieval and processing.

Multiple projects across EPA have experienced challenges with
harmonization and QAQC of nutrient data from WQP. WQX has a [best
practices document for nutrient data
submission](https://gcc02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fwww.epa.gov%2Fwaterdata%2Fwater-quality-data-upload-wqx&data=05%7C02%7CWong.Kenny%40epa.gov%7C59f010b61c194df4a2de08decadedc10%7C88b378b367484867acf976aacbeca6a7%7C0%7C0%7C639171255679011335%7CUnknown%7CTWFpbGZsb3d8eyJFbXB0eU1hcGkiOnRydWUsIlYiOiIwLjAuMDAwMCIsIlAiOiJXaW4zMiIsIkFOIjoiTWFpbCIsIldUIjoyfQ%3D%3D%7C0%7C%7C%7C&sdata=TfvQIaSAUJeMT7Fdxm%2BoBzpbY4SWpXBn%2FhlJptLlfwU%3D&reserved=0),
but not all data submitters follow it. TADA uses [ECHO Water Quality
Indicator](https://echo.epa.gov/maps/water-quality-indicators) logic for
WQP nutrient data. The TADA team has expanded this effort by adding
additional characteristic and speciation combinations found in the WQP
to the TADA internal reference file for nutrient data handling.

## Case Studies

1)  National Nutrient Inventory (NNI)

2)  Hypoxia Task Force (HTF) Monitoring Group

3)  ECHOS and ICIS integration

4)  Freshwater Explorer

## Relevant TADA functions

-   [TADA_DataRetrieval()](https://usepa.github.io/EPATADA/reference/TADA_DataRetrieval.html)

-   [TADA_HarmonizeSynonyms()](https://usepa.github.io/EPATADA/reference/TADA_HarmonizeSynonyms.html)

-   [TADA_CalculateTotalNP()](https://usepa.github.io/EPATADA/reference/TADA_CalculateTotalNP.html)

## TADA_DataRetrieval

**TADA_DataRetrieval** builds on USGS's dataRetrieval R package
functions to retrieve WQP portal data and process it for compatibility
with TADA. It joins three WQP profiles: Site, Sample Results
(physical/chemical metadata), and Project. In addition, it changes all
data in the Characteristic, Speciation, Fraction, and Unit fields to
uppercase and addresses result values that include special characters.
**TADA_DataRetrieval** accepts the same inputs as the dataRetrieval
`readWQPdata` function.

dataRetrieval and **TADA_DataRetrieval** filters include:

-   startDate

-   endDate

-   characteristicName

-   sampleMedia

-   siteType

-   statecode (see list of possible state and territory abbreviations
    [here](https://www2.census.gov/geo/docs/reference/state.txt))

-   countycode

-   siteid

-   organization

-   project

-   huc

-   characteristicType

-   providers

**TADA_DataRetrieval** accepts additional geospatial-related filters
including:

-   aoi_sf

-   tribal_area_type

-   tribe_name_parcel (Note: The **TADA_TribalOptions** function can be
    used to narrow down options for use with this tribe_name_parcel
    filter option. See **?TADA_TribalOptions** for more info).

After data are downloaded, **TADA_DataRetrieval** function also
automatically runs the **TADA_AutoClean** function. If users do not want
to apply the default TADA data cleaning procedures in
**TADA_AutoClean**, they can set applyautoclean = FALSE in their
**TADA_DataRetrieval** calls.

Tips:

1.  All the **TADA_DataRetrieva**l query params for WQP data work as
    "AND" between fields and "OR" within fields. For example:

    -   characteristicName: If you choose NITROGEN and PHOSPHORUS, both
        NITROGEN and PHOSPHORUS are retrieved.

    -   statecode: If you choose VA and IL, results from either VA or IL
        are retrieved.

    -   characteristicName and statecode: If you choose DISSOLVED OXYGEN
        (DO) and VA, dissolved oxygen results from Virginia are
        retrieved.

    -   characteristicName and characteristicType work together as an
        AND. This means that the selected characteristicName must fall
        within the selected characteristicGroup if both filters are
        being used, Conflicting characteristicNamesand
        characteristicType will cause an error. If you would like all
        available nutrient data, it is recommended to query by
        characteristicGroup = "Nutrient" and leave characteristicName
        blank.

2.  "siteid" is a general term WQP uses to describe both Site IDs (from
    USGS databases) and Monitoring Location Identifiers (from WQX). Each
    monitoring location in the Water Quality Portal (WQP) has a unique
    Monitoring Location Identifier, regardless of its origin. The
    Monitoring Location Identifier from the WQP is generated by
    concatenating the Organization Identifier plus the Site ID number
    from the source database.

3.  The aoi_sf and tribal arguments are not meant to be used
    simultaneously and will result in an error if a user attempt to
    apply them together. For example, if both an aoi_sf argument and
    tribal information are provided an error is returned because the
    function cannot prioritize both location arguments. Similarly,
    aoi_sf and tribal_area_type are not meant to be used with other
    location-related params (e.g., statecode, siteid). In these
    instances a warning is returned but the query proceeds by using only
    the aoi_sf or tribal_area_type information.

Additional resources:

-   Review function documentation by entering the following code into
    the console:

    -   ?TADA_DataRetrieval

    -   ?TADA_AutoClean

-   [Introduction to the dataRetrieval
    package](https://doi-usgs.github.io/dataRetrieval/)

-   [Water Quality Portal Web Services
    Guide](https://www.waterqualitydata.us/webservices_documentation/)

-   [dataRetrieval
    Tutorial](https://waterdata.usgs.gov/blog/dataretrieval/)

Use the code below to download data from the WQP using
TADA_DataRetrieval. Edit the params in the code chunk to define your own
WQP query.

Downloads using **TADA_DataRetrieval** will have the same columns each
time. Be aware that data are uploaded to the Water Quality Portal by
individual organizations, which may or may not follow the same
conventions. Data and metadata quality are not guaranteed! Carefully
explore data to make sure it meets your quality assurance requirements.

Note: **TADA_DataRetrieval** (by leveraging dataRetrieval),
automatically converts the date times to UTC. It also automatically
converts field formats to dates, datetimes, and numerics based on a
standard algorithm.

To retrieve data from the Water Quality Portal (WQP) using a bounding
box (bbox), we first define a bbox that encloses four HTF-priority sites
in Mississippi. We’ll use this spatial extent to identify existing WQP
data within the area. If broader coverage is needed, we can expand the
bbox to include a wider extent and capture additional results beyond the
initial boundary.

```{r}
MS_WQP_PrioritySites <- TADA_DataRetrieval(
  startDate = "2000-01-01",
  endDate = "2020-12-31",
  # bBox = c(-90.914167, 32.018, -90.696944, 32.444167), # original bbox
  bBox = c(-91.179367, 31.793500, -90.431744, 32.668667), # bbox expanded 25km
  #characteristicName = c("Phosphorus","Nitrogen","Nitrate"),
  characteristicType = "Nutrient", # also try it without this filter (all chars) - check to see if there's any missing nutrient data
  applyautoclean = TRUE,
  ask = FALSE
  )
```

Data submitters may define fraction, speciation and units for nutrient
data differently. TADA aims to capture these difference by creating
column TADA.ComparableDataIdentifier which is constructed by
concatenating the characteristic, speciation, fraction and unit into one
character string string. How many unique TADA.ComparableDataIdentifier
do we see in the MS priority sites data set?

```{r}
TADA_FieldValuesTable(MS_WQP_PrioritySites, field =
"TADA.ComparableDataIdentifier")

```

There are
`r length(unique(MS_WQP_PrioritySites$TADA.ComparableDataIdentifier))`
TADA.ComparableDataIdentifiers.

Next, we need to clean and harmonize synonyms and calculate total
nutrient summation. First, let's work on cleaning up our WQP data frame.
We will start off with filtering to retain only surface water data using
**TADA_MediaFilter.** How many non-surface water results were removed
from the data set?

```{r}
MS_WQP_PrioritySites2 <- TADA_MediaFilter(MS_WQP_PrioritySites,
                                clean = TRUE,
                                surface_water = FALSE,
                                ground_water = TRUE,
                                sediment = TRUE,
                                other = TRUE
)
```

There were
`r formatC(nrow(MS_WQP_PrioritySites) - nrow(MS_WQP_PrioritySites2), big.mark = ",")`
non-surface water results removed from the data set.

Nowe, we can take a look at the sites remaining in the data set by using
**TADA_OverviewMap** which displays the sites on a map, using symbology
and pop-ups to describe result counts, number of characteristics per
site and other site details.

Which site has the greatest number of characteristics? Which site was
visited the most? Which site has the greatest number of results?

```{r}
TADA_OverviewMap(MS_WQP_PrioritySites)
```

```{r overview.answers, include=FALSE}

most.chars <- MS_WQP_PrioritySites2 |>
  dplyr::group_by(TADA.MonitoringLocationIdentifier) |>
  dplyr::summarise(nChars = dplyr::n_distinct(TADA.CharacteristicName)) |>
  dplyr::slice_max(nChars)

most.visits <- MS_WQP_PrioritySites2 |>
  dplyr::group_by(TADA.MonitoringLocationIdentifier) |>
  dplyr::summarise(nVisit = dplyr::n_distinct(ActivityStartDate)) |>
  dplyr::slice_max(nVisit)

most.results <- MS_WQP_PrioritySites2 |>
  dplyr::group_by(TADA.MonitoringLocationIdentifier) |>
  dplyr::summarise(nResult = dplyr::n_distinct(ResultIdentifier)) |>
  dplyr::slice_max(nResult)

```

`r most.chars$TADA.MonitoringLocationIdentifier[1]` has the greatest
number of characteristics
(``` r``most.chars$TADA.MonitoringLocationIdentifier[1] ``` ).

`r most.visits$TADA.MonitoringLocationIdentifier[1]` has the greatest
number of results (`rmost.visits$TADA.MonitoringLocationIdentifier[1]`
).

`r most.results$TADA.MonitoringLocationIdentifier[1]` has the greatest
number of results (`rmost.results$TADA.MonitoringLocationIdentifier[1]`
).

After reviewing the overview map, we can move on to some additional data
cleaning steps. Let's start this process by removing results tied to
monitoring locations with imprecise coordinates (less than 3 decimal
digits). How many monitoring locations were identified as having
imprecise coordinates? How many results were removed from the data set
as a result?

```{r}
MS_WQP_PrioritySites3 <- TADA_FlagCoordinates(MS_WQP_PrioritySites2, clean_outsideUSA = "no", clean_imprecise = TRUE)
```

```{r coords.answers, echo=FALSE}

imprecise.coords <- TADA_FlagCoordinates(MS_WQP_PrioritySites2, clean_outsideUSA = "no", clean_imprecise = TRUE, flaggedonly = TRUE)

if (nrow(imprecise.coords) == 0) {
  cat("No monitoring locations with imprecise coordinates were discovered in the data set.")
} else {
  n_sites <- dplyr::n_distinct(imprecise.coords$TADA.MonitoringLocationIdentifier)
  n_rows <- nrow(imprecise.coords)

  cat(
    "There were ", n_rows, " flagged rows representing ",
    n_sites, " unique monitoring locations with imprecise coordinates."
  )
}

```

Now let's identify and handle any censored results. We can start by
using **TADA_IDCensored** to identify censored data and
**TADA_FieldValuesPie** to visualize how many non-censored, over-detect,
and non-detect results are in the data set. How many results are in each
censored data flag category.

```{r}

MS_WQP_PrioritySites4 <- TADA_IDCensoredData(MS_WQP_PrioritySites3)

test <- TADA_FieldValuesPie(MS_WQP_PrioritySites4, field = "TADA.CensoredData.Flag")
```

```{r censored.flag.answers}

cens.answers <- TADA_FieldValuesTable(MS_WQP_PrioritySites4, field = "TADA.CensoredData.Flag")

get_count <- function(df, value_name) {
  x <- df |>
    dplyr::filter(Value == value_name) |>
    dplyr::pull(Count)
  
  if (length(x) == 0) 0 else x
}

cens.values <- c(
  "Non-Detect",
  "Over Detect",
  "Uncensored",
  "Detection condition is missing and required for censored data ID."
)

cens.counts <- setNames(
  lapply(cens.values, \(v) {
    x <- cens.answers |>
      dplyr::filter(Value == v) |>
      dplyr::pull(Count)
    if (length(x) == 0) 0 else x
  }),
  c("n.nd", "n.od", "n.uncens", "n.missing")
)
```

There `r formatC(cens.counts$n.nd)` non-detect results,
`r formatC(cens.counts$n.od)` over detect results,

`r formatC(cens.counts$n.uncens)` results and
`r formatC(cens.counts$n.missing)` results missing detection condition
information.

We've identified the censored records, and now we need to handle them
with **TADA_SimpleCensoredMethods** which allows us to assign simple
methods for handling censored data. We will use the default settings
which means non-detects values will be estimated as 1/2 the lower
detection limit and over detect values will be left as-is.

```{r}
MS_WQP_PrioritySites5 <- TADA_SimpleCensoredMethods(MS_WQP_PrioritySites4)
```

Now that the censored records have been assigned numeric values, we can
use **TADA_HarmonizeSynonyms** to harmonize the data to consistent
target characteristic names, speciation., and fractions.

```{r}
MS_WQP_PrioritySites6 <- TADA_HarmonizeSynonyms(MS_WQP_PrioritySites5)
```

We can compare the comparable data identifiers in the data before and
after harmonization.

```{r comp.ids}

before.comp.ids <- sort(unique(MS_WQP_PrioritySites5$TADA.ComparableDataIdentifier))

after.comp.ids <- sort(unique(MS_WQP_PrioritySites6$TADA.ComparableDataIdentifier))


before.comp <- TADA_FieldValuesTable(MS_WQP_PrioritySites5, field = "TADA.ComparableDataIdentifier")

after.comp <- TADA_FieldValuesTable(MS_WQP_PrioritySites6, field = "TADA.ComparableDataIdentifier")

# will create plot to compare
```

Now we can run key flagging functions and remove results flagged as
suspect based on their result units, method-speciation, fraction,
qualifier codes or activity type with **TADA_RunKeyFlagFunctions**. By
setting clean = TRUE, the function will remove all results flagged as
suspect.

For more detail on which flagging functions are included as key flag
functions enter the following in the console:

-   ?TADA_RunKeyFlagFunctions

For details on flagging functions run as part of
TADA_RunKeyFlagFunctions enter each of the following in the console:

-   ?TADA_FlagResultUnit

-   ?TADA_FlagFraction

-   ?TADA_FlagSpeciation

-   ?TADA_FindQCActivities

-   ?TADA_FlagMeasureQualifierCode

How many results were removed by **TADA_RunKeyFlagFunctions**? How many
results were flagged in each category (note: a result may be flagged
more in more than one category).

```{r}
MS_WQP_PrioritySites7<- TADA_RunKeyFlagFunctions(MS_WQP_PrioritySites6, clean = TRUE)
```

```{r flag.answers}
MS_WQP_PrioritySites_keyflags <- TADA_RunKeyFlagFunctions(MS_WQP_PrioritySites6, clean = FALSE)

sort(unique(MS_WQP_PrioritySites_keyflags$TADA.ResultDepthHeightMeasure.MeasureValueDataTypes.Flag)) 


# notes for building code chunk to assess results of key flag functions
# "TADA.ActivityType.Flag" != "Non_QC"
# "TADA.SuspectCoordinates.Flag" != "Pass"
# "TADA.SampleFraction.Flag" != "Not Reviewed", "Pass"
# "TADA.MethodSpeciation.Flag" != "Not Reviewed", "Pass"
# "TADA.ResultUnit.Flag" != "Not Reviewed", "Pass"
# "TADA.ResultMeasureValueDataTypes.Flag"  != "Numeric"
# "TADA.DetectionQuantitationLimitMeasure.MeasureValueDataTypes.Flag" 
# "TADA.ResultDepthHeightMeasure.MeasureValueDataTypes.Flag"
# "TADA.ActivityDepthHeightMeasure.MeasureValueDataTypes.Flag"
# "TADA.ActivityTopDepthHeightMeasure.MeasureValueDataTypes.Flag"
# "TADA.ActivityBottomDepthHeightMeasure.MeasureValueDataTypes.Flag" 
```

identify potential dup sites

```{r}
MS_WQP_PrioritySites8 <- MS_WQP_PrioritySites7 |>
  TADA_FindPotentialDuplicatesSingleOrg() |>
  dplyr::filter(TADA.SingleOrgDup.Flag == "Unique") |>
  TADA_FindPotentialDuplicatesMultipleOrgs(
    dist_buffer = 100,
    org_hierarchy = "none"
  ) |>
  dplyr::filter(TADA.ResultSelectedMultipleOrgs == "Y")
```

uses echos logic to append rows for TN & TP summation

```{r}
TotalNP <- TADA_CalculateTotalNP(MS_WQP_PrioritySites8)
```

view final cleaned table

```{r}
TADA_FieldValuesTable(TotalNP, field = "TADA.ComparableDataIdentifier")
```

### Case Study 1: Hypoxia Task Force

For this case study the HTF monitoring group is interested in
determining where to prioritize funds for a list of priority monitoring
sites using NPS grant funds.

Looking at historical 20 years of nutrient data (from 2000 to 2020) for
a certain region of Mississippi, let's determine if there are any
existing overlapping WQP sites with these priority sites of interest. We
are also interested in determining if there are prior water bodies that
have been assessed in prior ATTAINS assessment cycles near these
priority sites.

First, let's look at our cleaned nutrient data on the map. We will run
TADA_CreateAUMLCrosswalk to determine what sites can be matched to
ATTAINS Assessment Units through a geospatial join or if any exists
directly from ATTAINS in this region.

Crosswalk ATTAINS AU and ML sites

```{r}
AUML <- TADA_CreateAUMLCrosswalk(TotalNP, org_id = "21MSWQ")
# View ATTAINS AU and ML sites map
TADA_ViewATTAINS(AUML)
```

Now let's add our priority sites to the map. These lat/longs are
provided from the HTF list of priority sites appendix.

Add dummy variables with priority sites in the WQP data frame to map it
in TADA (edit in excel)

```{r}
extra_sites <- tibble::tibble(
  OrganizationIdentifier = "USGS",
  TADA.MonitoringLocationIdentifier = c("USGS-07288955", "USGS-322023090544500", "USGS-07290000", "21MSWQ_WQX-07290650"),
  TADA.MonitoringLocationName = c("Priority Site 1", "Priority Site 2", "Priority Site 3", "Priority Site 4"),
  TADA.MonitoringLocationTypeName = "River/Stream",
  LatitudeMeasure = c(32.444167, 32.339722, 32.347778, 32.018),
  TADA.LatitudeMeasure = c(32.444167, 32.339722, 32.347778, 32.018),
  LongitudeMeasure = c(-90.914167, -90.9125, -90.696944, -90.877194),
  TADA.LongitudeMeasure = c(-90.914167, -90.9125, -90.696944, -90.877194),
  .source = "overlay"  # optional flag to tell them apart later
) |>
  TADA_CorrectColType()
```

Make sure types match and bind to data

```{r}
MS_WQP_modified <- dplyr::bind_rows(
  TotalNP |>
    dplyr::mutate(.source = ".wqp"),
  extra_sites
) |> TADA_CorrectColType()
```

Identify these priority sites on the map (include one existing AU row as
TADA_CreateAUMLCrosswalk() currently crashes if no AUML from a user
supplied ref is found in the data frame.

```{r}
flag_priority_AURef <- tibble::tibble(
  OrganizationIdentifier = "USGS",
  TADA.MonitoringLocationIdentifier = c("USGS-07290000", "USGS-07288955", "USGS-322023090544500", "USGS-07290000", "21MSWQ_WQX-07290650"),
  TADA.MonitoringLocationName = c("NA", "Priority Site 1", "Priority Site 2", "Priority Site 3", "Priority Site 4"),
  ATTAINS.OrganizationIdentifier = c("21MSWQ", "21MSWQ", "21MSWQ", "21MSWQ", "21MSWQ"),
  ATTAINS.AssessmentUnitIdentifier = c("MS107811", "Flag1", "Flag2", "Flag3", "Flag4"),
  ATTAINS.WaterType = c("STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER")
) |>
  TADA_CorrectColType()
```

recreate the AUML crosswalk with the dummy priority sites

```{r}
AUML2 <- TADA_CreateAUMLCrosswalk(MS_WQP_modified, org_id = "21MSWQ", au_ref = flag_priority_AURef)
# view the map, examine where sites overlap (exact), where there exists nearby sites, and if ATTAINS assessments are found prior.
TADA_ViewATTAINS(AUML2)
```
