---
title: "Participatory (Volunteer) Scientists are Monitoring Waters and Sharing Data via EPA's Water Quality eXchange (WQX)"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Participatory (Volunteer) Scientists are Monitoring Waters and Sharing Data via EPA's Water Quality eXchange (WQX)}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE
)
```

#### Load the EPATADA R Package

```{r libraries, results=F}
# Load necessary libraries
library(EPATADA)
```

#### Find volunteer data in WQX

Find volunteer monitoring organizations who have submitted data to EPA's Water Quality eXchange (WQX) by reviewing the `organization` domain table available [here](https://www.epa.gov/waterdata/storage-and-retrieval-and-water-quality-exchange-domain-services-and-downloads).

```{r get org domain}
# Get the WQX organizations domain
organizations <- read.csv(url("https://cdx.epa.gov/wqx/download/DomainValues/Organization.CSV"))

# Subset to include only "Volunteer" organizations and exclude WQX test/training orgs
volunteer_orgs <- subset(organizations, 
                         Type == "Volunteer" &
                           !grepl("training", 
                                  Name, 
                                  ignore.case = TRUE) &
                           !grepl("test", 
                                  Name, 
                                  ignore.case = TRUE) &
                           !grepl("\\*", 
                                  Name, 
                                  ignore.case = TRUE))

unique(volunteer_orgs$Name)
```

Generate a list of 5 random organization IDs:

```{r, eval = F}
random_volunteer_orgIDs <- sample(volunteer_orgs$ID, size = 5)
```

Prepare list for use in `TADA_DataRetrieval`:

```{r, eval = F}
unlist(random_volunteer_orgIDs)
```

Query the Water Quality Portal (WQP) using `TADA_DataRetrieval` and the 5 random volunteer organizations IDs. We will look for any data available from 2015 to present.

```{r, eval = F}
volunteer_data <- TADA_DataRetrieval(
  startDate = "2015-01-01",
  organization = random_volunteer_orgIDs,
  ask = FALSE,
  applyautoclean = TRUE
)
```

#### Retrieve data from WQP

Alternatively, choose 5 volunteer monitoring organizations and query WQP. We will move forward with this example of volunteer organizations in CT:

```{r}
selected_orgs <- 
  c("CONNRIVERCONSERVANCY",
    "CT_NERR",
    "BANTAMLAKE_WQX",
    "CTVOLMON",
    "CT_NERR")

volunteer_data <- TADA_DataRetrieval(
  # startDate = "2022-01-01",
  organization = selected_orgs,
  ask = FALSE,
  applyautoclean = TRUE
)
```

#### Explore and refine results

Review volunteer monitoring projects in WQX:

```{r projects}
unique(volunteer_data$ProjectName)
```

Generate pie chart:

```{r}
TADA_FieldValuesPie(volunteer_data, field = "MonitoringLocationTypeName")
```

Review and remove sites if coordinates are imprecise or outside US:

```{r, results=F}
volunteer_data <- TADA_FlagCoordinates(volunteer_data, 
                                       clean_outsideUSA = "remove", 
                                       clean_imprecise = TRUE, 
                                       flaggedonly = FALSE)
```

Use `TADA_OverviewMap` to generate a map:

```{r}
TADA_OverviewMap(volunteer_data)
```

Review and remove duplicate results if present:

```{r}
volunteer_data <- TADA_FindPotentialDuplicatesSingleOrg(volunteer_data)
volunteer_data <- dplyr::filter(volunteer_data, TADA.SingleOrgDup.Flag == "Unique")
```

Prepare censored (nondetects and overdetects) results for analysis:

```{r}
volunteer_data <- TADA_SimpleCensoredMethods(
  volunteer_data,
  nd_method = "multiplier",
  nd_multiplier = 0.5,
  od_method = "as-is",
  od_multiplier = "null"
  )
```

Run key TADA quality control flagging functions and remove suspect results:

```{r}
volunteer_data <- TADA_RunKeyFlagFunctions(
  volunteer_data,
  clean = TRUE
  )
```

Flag results above and below thresholds. Review carefully and consider removing.

```{r}
volunteer_data <- TADA_FlagAboveThreshold(volunteer_data, 
                                   clean = FALSE, 
                                   flaggedonly = FALSE)

volunteer_data <- TADA_FlagBelowThreshold(volunteer_data, 
                                   clean = FALSE, 
                                   flaggedonly = FALSE)
```

Harmonize synonyms if found:

```{r}
volunteer_data <- TADA_HarmonizeSynonyms(volunteer_data)
```

Generate table:

```{r}
TADA_FieldValuesTable(volunteer_data, field = "ActivityTypeCode")
```

Generate pie chart:

```{r}
TADA_FieldValuesPie(volunteer_data, field = "OrganizationFormalName")
```

Generate pie chart:

```{r}
TADA_FieldValuesPie(volunteer_data, field = "MonitoringLocationName")
```

Remove non-numeric results:

```{r}
volunteer_data <- TADA_ConvertSpecialChars(
  volunteer_data, 
  col = "TADA.ResultMeasureValue",
  clean = TRUE)
```

Review the number of sites and records for each characteristic:

```{r}
TADA_SummarizeColumn(volunteer_data)
```

Filter data to review a single characteristic:

```{r}
ecoli <- dplyr::filter(
  volunteer_data,
  TADA.ComparableDataIdentifier %in% c(
    "ESCHERICHIA COLI_NA_NA_CFU/100ML"
  )
)
```

Generate scatter plot for E. coli:

```{r}
TADA_GroupedScatterplot(ecoli)
```
