---
title: "TADA Module 3A: Associating WQP Characteristic Names with Applicable ATTAINS State, Tribal or EPA 304(a) Parameter and Use Names"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA Module 3A: Associating WQP Characteristic Names with Applicable ATTAINS State, Tribal or EPA 304(a) Parameter and Use Names}
  %\VignetteEngine{knitr::rmarkdown}
description: The first of multiple Module 3 vignettes (under development) showcasing water quality data analysis functions and example workflows.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = FALSE
)
```

# Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language. The
EPATADA R Package is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to user
feedback. We appreciate your feedback, patience, and interest in these
helpful tools. If you are interested in contributing to EPATADA
development, more information is available
[here](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html). We
welcome collaboration with external partners!

# Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub.

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

```{css, include = F}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

Next, install and load the EPATADA R Package using the remotes package.
Dependency packages will also be downloaded automatically from CRAN. You
may be prompted in the console to update dependencies that have more
recent versions available. If you see this prompt, it is recommended to
update all of them (enter 1 into the console).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

```{r dev_install, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the EPATADA R Package
into your R session.

```{r, results = 'hide'}
library(EPATADA)
```

All TADA R package functions have their own individual help pages,
listed on the [Function
reference](https://usepa.github.io/EPATADA/reference/index.html) page on
the GitHub site. Users can also access the help page for a given
function in R or RStudio using the following format (example below):
`?[name of TADA function]`

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

# Introduction to TADA Module 3

This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to create WQP and ATTAINS data crosswalks that are
needed prior to defining and capturing organization specific criteria
and methodologies. It is the first of several vignettes that are being
developed as part of TADA Module 3.

Specifically, this vignette provides an overview of two functions that
can assist users with:

-   Creating a crosswalk (reference table) between ATTAINS parameter
    names and WQP/TADA characteristic names.

-   Creating a crosswalk (reference table) of all unique combinations of
    ATTAINS uses and and parameters applicable to ATTAINS Assessment
    Units of interest.

These Module 3 functions are being designed with the flexibility for
users from states, tribes, or territories to input organization-specific
information. While TADA functions help generate these crosswalks, users
must review and modify the tables generated in each step of the process
to ensure their accuracy. Additionally, the TADA team has also
incorporated national recommended Clean Water Act (CWA) 304(a) numeric
criteria for optional use in Module 3 functions by leveraging data from
EPA's [Criteria Search Tool
(CST)](https://www.epa.gov/wqs-tech/state-specific-water-quality-standards-effective-under-clean-water-act-cwa).
In the next vignette, TADAModule3_PartB , which is coming soon, this
will allow users to analyze their data against the 304(a) criteria; and
to easily compare results when when using 304(a) criteria vs. their own
organization's criteria.

# Example data from WQP

Let's start with an example data frame that has gone through data
cleaning, wrangling, harmonization, handling of censored data, removal
of suspect results, and other important TADA Module 1 functions. This
process should be completed by users before utilizing Module 3
functions. The example TADA data frame used in this vignette
(Data_HUC8_02070004_Mod1Output) has already been through this process.

```{r Load Example Data Set}
# import example data set (output of Module 1 Workflow)
Data_NCTC <- Data_HUC8_02070004_Mod1Output
```

This example includes results from from multiple states. In the
following sections, for demonstration purposes, we will focus only on
Maryland. The remainder of this vignette will walk through how to fill
out two crosswalk tables (TADA_CreateParamRef and
TADA_CreateUseParamRef) that are needed before we can start assigning
applicable criteria and methodologies information for this specific
ATTAINS organization (Maryland).

# TADA_CreateParamRef() Basics

A crosswalk between WQX/WQP/TADA characteristics and ATTAINS parameters
is needed before we can integrate information from these two data
systems. Before creating the WQP/TADA and ATTAINS parameter crosswalk
table (TADA_CreateParamRef), let's review the unique characteristic,
fraction and speciation combinations from the example data frame
(TADA.ComparableDataIdentifier) and how many results are available for
each within the example data set (Data_NCTC).

```{r Create Unique Counts}
# create table with counts of TADA.ComparableDataIdentifiers
TADA_FieldValuesTable(Data_NCTC, field = "TADA.ComparableDataIdentifier")
```

We can also review the domain values of parameter names listed as causes
in prior assessment cycles (for all organizations) in ATTAINS by using a
function from rATTAINS, which provides functions for downloading tidy
data from the ATTAINS public web services
(<https://github.com/mps9506/rATTAINS>).

```{r Access ATTAINS Domain Values}
# return ATTAINS domain names
rATTAINS::domain_values()

# return ATTAINS parameter domain values
TADA_TableExport(rATTAINS::domain_values(domain_name = "ParameterName"))

# return ATTAINS organization domain values
TADA_TableExport(rATTAINS::domain_values(domain_name = "OrgName"))
```

As the ATTAINS Parameter Names are more general than the TADA Comparable
Data Identifiers, we recommend users to manually provide the crosswalk
or to provide a pre-filled crosswalk with the correct column names
("TADA.ComparableDataIdentifier" and "ATTAINS.ParameterName"). However,
TADA_CreateParamRef() includes an argument input 'fillBy' to assist
users in doing an exact match between TADA Characteristic Names to
ATTAINS Parameter Name. There are roughly 598/1180 unique ATTAINS
Parameter Names that have an exact match. \*\*It is important to note
that even with these exact matches, there are additional complexities
related to the fraction, speciation and units that may be unique to your
organization. For example, some organizations may consider any of
Nitrate/Nitrite, Ammonia as "Nitrogen" as the ATTAINS.ParameterName and
may not care for the speciation in their assessment needs.

Let's go through a few examples.

TADA_CreateParamRef() creates a template which includes all
TADA.ComparableDataIdentifiers from the TADA data frame and provides a
blank column, "ATTAINS.ParameterName" for users to input the
corresponding ATTAINS parameter name.

Setting "excel = TRUE" in TADA_CreateParamRef causes the function to
automatically create an Excel spreadsheet of the reference table for
easy review and editing. If users do not want to work in Excel, they can
use "excel = FALSE" to return a data frame. The data frame can be edited
directly in R if desired, although there are no TADA-specific functions
designed to facilitate this.

When using TADA_CreateParamRef(), users should specify the
organization(s) of interested in the "org_names" argument. This ensures
that the correct number of rows will be created in the reference table.
When "excel = TRUE", specifying the organization(s) also creates a
separate tab in the Excel file which contains the parameter names used
in prior assessment cycles by the specified ATTAINS organization(s).
This will allow users to decide whether to continue to use the same
ATTAINS parameter names from prior assessments that their
organizations(s) have used in the past, or to use other valid parameter
names from the entire ATTAINS domain list. If there is not a suitable
ATTAINS parameter name to match a TADA.ComparableDataIdentifier, users
should contact the ATTAINS team (attains\@epa.gov) for assistance.

For our first example, let's start of with a single organization. In
this example TADA data frame, this spatial area consists of WQP data
from several different states. We will specify Maryland "MDE_EASP" as
our ATTAINS organization identifier of interest. By default, fillBy =
"None" which will not populate any entries for the ATTAINS.ParameterName
column, and users will need to manually fill this crosswalk out.

```{r NCTC_CreateParamRef}
# create TADA parameter reference table for specified organization
NCTC_ParamRef <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  fillBy = "None"
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

# export TADA parameter reference table as csv
TADA_TableExport(NCTC_ParamRef)
```

Now, let's go through an example in which users auto-populate the
crosswalk with exact matches. We will use fillBy = 'All' to indicate we
will use exact matches regardless of if your organization has used this
"ATTAINS.ParameterName" domain value in the past. i.e. Your organization
is fine with using an ATTAINS Parameter Name that has been submitted to
ATTAINS outside of your organization.

```{r NCTC_ParamRef2}
# create TADA parameter reference table for specified organization
NCTC_ParamRef2 <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  fillBy = "All"
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

# export TADA parameter reference table as csv
TADA_TableExport(NCTC_ParamRef2)
```

We can also use fillBy = 'Org' to indicate we will use exact matches
ONLY if your organization has used this "ATTAINS.ParameterName" domain
value in the past.

```{r NCTC_ParamRef3}
# create TADA parameter reference table for specified organization
NCTC_ParamRef3 <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  fillBy = "Org"
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

# export TADA parameter reference table as csv
TADA_TableExport(NCTC_ParamRef3)
```

The code chunk below demonstrates how the parameter reference table can
be modified in R. Once modifications are complete, we can re-run
TADA_CreateParamRef using the modified parameter reference table as the
input for "paramRef". This will update the ATTAINS.FlagParameterName
column in the parameter reference table (see NCTC_ParamRef2 in example
below). This column provides information about whether or not the
organization specified in the "ATTAINS.OrganizationIdentifier" column
has listed this parameter as a cause in prior assessment cycles. It also
flags rows where the TADA.ComparableDataIdentifier as not been assigned
an ATTAINS parameter name.

```{r NCTC_ParamRef_Manual}
# only run the code chunks below if you make edits to the excel file. Please note you must open and save the excel file at least once to reflect all Excel formula based values.

# downloads_path <- file.path(Sys.getenv("USERPROFILE"), "Downloads", "myfileRef.xlsx")
# ParamRef <- openxlsx::read.xlsx(downloads_path, sheet = "CreateParamRef")

# NCTC_ParamRef2 <- TADA_CreateParamRef(
#   Data_NCTC,
#   org_names = c("MDE_EASP"),
#   paramRef = ParamRef,
#   excel = FALSE
#   # uncomment excel = TRUE, overwrite = TRUE to run the excel file
#   # excel = TRUE, overwrite = TRUE
#   )

ParamRef <- dplyr::mutate(NCTC_ParamRef, ATTAINS.ParameterName = dplyr::case_when(
  grepl("PH", TADA.ComparableDataIdentifier) ~ "PH, HIGH",
  TADA.ComparableDataIdentifier == "ZINC_DISSOLVED_NA_UG/L" ~ "ZINC",
  grepl("NITRATE", TADA.ComparableDataIdentifier) ~ "NITROGEN, TOTAL"
))

NCTC_ParamRef_Manual <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = ParamRef
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(NCTC_ParamRef_Manual)
```

We can also use the output of NCTC_ParamRef2 or NCTC_ParamRef3 as a
paramRef argument input in TADA_CreateParamRef().

```{r NCTC_ParamRef4}
NCTC_ParamRef4 <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = NCTC_ParamRef2,
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(NCTC_ParamRef4)
```

Your 'paramRef' is prioritized over the 'fillBy' argument inputs.
However, fillBy will fill in any remaining blanks that you have not
filled in. For example, this 'paramRef' argument input that was provided
matched 'Nitrate' to 'NITROGEN, TOTAL' rather than 'NITRATE', thus even
if you specified 'fillBy = 'All', this crosswalk will prioritize
'NITROGEN, TOTAL' as the ATTAINS.ParameterName. However, by specifying
fillBy = 'All' for any values that have not been crosswalk,
TADA_CreateParamRef() will fill those if there is an exact match, in the
example below, no crosswalk of pH in your ParamRef2 was provided, so
NCTC_ParamRef_Manual2 will crosswalk the TADA Characteristic name to
'PH' as the exact match for ATTAINS parameter name.

```{r NCTC_ParamRef_Final}
ParamRef2 <- dplyr::mutate(NCTC_ParamRef, ATTAINS.ParameterName = dplyr::case_when(
  # grepl("PH", TADA.ComparableDataIdentifier) ~ "PH",
  TADA.ComparableDataIdentifier == "ZINC_DISSOLVED_NA_UG/L" ~ "ZINC",
  grepl("NITRATE", TADA.ComparableDataIdentifier) ~ "NITROGEN, TOTAL"
))

NCTC_ParamRef_Final <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = ParamRef2,
  fillBy = "All",
  excel = TRUE, overwrite = TRUE
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(NCTC_ParamRef_Final)
```

Once the parameter reference table has been reviewed and modified, it
can be saved and reused for analysis of future TADA data frames.

# TADA_CreateUseParamRef() Basics

Our first example will pull in all prior ATTAINS uses (ATTAINS.UseNames)
and parameter names (ATTAINS.ParameterName) specific to the specified
ATTAINS organization (defined by the org_names function argument for
TADA_CreateUseParamRef), which in this case is "MDE_EASP" (Maryland).
Users will review the output and choose which ATTAINS.UseName's are
applicable to their analysis. The default functionality is to "Include"
all ATTAINS use_names (see the "IncludeOrExclude" column) that has an
ATTAINS.UseName associated with it. Later in the data analysis process,
users will be asked to define criteria and methodologies that is
applicable to each ATTAINS use_name and parameter name labeled as
"Include". If a ATTAINS.UseName is not applicable, users should choose
"Exclude" for that ATTAINS parameter name and use name.

For any ATTAINS parameter name(s) that have not been used by an
organization in prior assessment cycles, there will be no prior use
names associated with them. In this case, users will be responsible for
manually assigning the appropriate ATTAINS.UseName applicable to their
org under the column 'ATTAINS.UseName', and for inputting additional
rows as needed if the parameter applies to multiples uses. Users can
also choose to 'auto_assign' all unique ATTAINS.UseName by
ATTAINS.OrganizationName to any ATTAINS.ParameterName missing a use name
associated with it.

In the example below, we can see only "ZINC" and "NITROGEN, TOTAL" were
the ATTAINS.ParameterNames done in prior assessment cycles for MDE_EASP
(see ATTAINS.FlagUseName column in NCTC_UseParamRef). If desired, a user
can manually assign this parameter to any applicable uses (disclaimer:
this is for demonstration purposes only and does not reflect MDE_EASP's
process).

```{r NCTC_UseParamRef}
NCTC_UseParamRef <- TADA_CreateUseParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = NCTC_ParamRef_Final,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(NCTC_UseParamRef)
```

We can choose to assign all unique use names found by your organization
to those ATTAINS.ParameterName without any associated ATTAINS.UseName.

```{r NCTC_UseParamRef2}
NCTC_UseParamRef2 <- TADA_CreateUseParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = NCTC_ParamRef_Final,
  auto_assign = TRUE,
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  excel = TRUE, overwrite = TRUE
)
TADA_TableExport(NCTC_UseParamRef2)
```

If a user includes wants to include additional organizations beyond
"MDE_EASP" , they will need to have defined those organizations in
TADA_CreateParamRef with the org_id argument input. We will create
another example paramRef from TADA_CreateParamRef that would allow us to
pull in this information for more than one organization.

```{r NCTC_UseParamRef3}
# create TADA parameter reference table for specified organization
NCTC_ParamRef4 <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP", "21VASWCB", "21PA", "WVDEP"), fillBy = "Org"
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

ParamRef2 <- dplyr::mutate(NCTC_ParamRef4, ATTAINS.ParameterName = dplyr::case_when(
  grepl("PH", TADA.ComparableDataIdentifier) ~ "PH",
  TADA.ComparableDataIdentifier == "ZINC_DISSOLVED_NA_UG/L" ~ "ZINC",
  grepl("NITRATE", TADA.ComparableDataIdentifier) ~ "NITROGEN, TOTAL"
))

NCTC_ParamRef2_Final <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = ParamRef2
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

NCTC_UseParamRef3 <- TADA_CreateUseParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP", "21VASWCB", "21PA", "WVDEP"),
  paramRef = NCTC_ParamRef2_Final, auto_assign = TRUE,
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(NCTC_UseParamRef3)
```

Instead, users should create a parameter and use reference table,
TADA_CreateUseParamRef(), for the same organizations included in the
parameter name crosswalk, TADA_CreateParamRef(). While aspects of this
process are partially automated, both crosswalks must be manually
reviewed, adjusted, and finalized by the user of these functions.

Users may choose to voluntarily share their reference tables for their
organization with the TADA Team so that others can use it in the future.

```{r NCTC_UseParamRef4}
NCTC_UseParamRef4 <- TADA_CreateUseParamRef(
  Data_NCTC,
  org_id = c("EPA304a", "MDE_EASP"),
  paramRef = NCTC_ParamRef_Final,
  excel = FALSE
)
TADA_TableExport(NCTC_UseParamRef4)
```

In this example, we designate that the ATTAINS.ParameterName "PH" is
equivalent to the TADA.ComparableDataIdentifier 'PH_NA_NA_STD UNITS',
which was not used in prior ATTAINS assessment cycles for **Maryland**.
This results in the 'ATTAINS.UseName' column returning a blank value for
ATTAINS.ParameterName "PH". Therefore, we need to specify the
ATTAINS.UseName(s) that we would like to associate with "PH" for this
organization. In excel, there will be a drop down list of organization
("MDE_EASP") specific ATTAINS.UseNames that were used in prior ATTAINS
assessments. If no excel file is generated, this is an example workflow
of how a user can edit the table in the R environment.

```{r NCTC_UseParamRef_Final}
add_data <- data.frame(
  "ATTAINS.OrganizationIdentifier" = "MDE_EASP",
  "ATTAINS.ParameterName" = rep("PH", 3),
  "ATTAINS.UseName" = c(
    "Aquatic Life and Wildlife", "Water Contact Sports",
    "Seasonal Migratory Fish Spawning and Nursery Subcategory"
  )
)

# The output of this will not reflect changes to the ATTAINS.FlagUseName column. To do so, we need to re run TADA_CreateUseParamRef() with UseParamRef = UseParamRef as an argument.
UseParamRef <- NCTC_UseParamRef3 %>%
  dplyr::left_join(add_data, by = c("ATTAINS.OrganizationIdentifier", "ATTAINS.ParameterName"), keep = FALSE) %>%
  dplyr::mutate(ATTAINS.UseName = dplyr::coalesce(ATTAINS.UseName.x, ATTAINS.UseName.y)) %>%
  dplyr::select(-c(ATTAINS.UseName.x, ATTAINS.UseName.y))

# User now has a saved dataframe which can be used to reflect any updates of Parameter-use ref in the future.
TADA_TableExport(UseParamRef)

# PH will now reflect the changes
NCTC_UseParamRef_Final <- TADA_CreateUseParamRef(
  Data_NCTC,
  paramRef = NCTC_ParamRef_Final,
  useParamRef = UseParamRef, # Edits were made to UseParamRef, updates flag column
  org_id = c("EPA304a", "MDE_EASP"),
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)
TADA_TableExport(NCTC_UseParamRef_Final)
```

# Next steps

Stay tuned! Future vignettes aim to assist with next steps in the
analysis process, including:

-   Defining the water quality criteria and methodologies used for each
    unique combination of ATTAINS assessment unit, use, and parameter.

-   Summarizing results by: (1) individual magnitude excursions.

Disclaimer: The EPATADA Module 3 functions are under active development.
EPATADA functions do not constitute current EPA recommendations, policy
or regulatory requirements. Organizations may optionally choose to use
EPATADA as a a tool in their process. Use of EPATADA is not required.
