---
title: "TADA Module 3: Associating WQP Characteristics with Applicable ATTAINS and EPA 304(a) Parameters and Uses"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA Module 3: Associating WQP Characteristics with Applicable ATTAINS and EPA 304(a) Parameters and Uses}
  %\VignetteEngine{knitr::rmarkdown}
description: The first of multiple vignettes (under development) showcasing TADA Module 3 functions and a recommended workflow.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language. The
EPATADA R Package is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to user
feedback. We appreciate your feedback, patience, and interest in these
helpful tools. If you are interested in contributing to EPATADA
development, more information is available
[here](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html). We
welcome collaboration with external partners!

# Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load the EPATADA R Package using the remotes package.
Dependency packages will also be downloaded automatically from CRAN. You
may be prompted in the console to update dependencies that have more
recent versions available. If you see this prompt, it is recommended to
update all of them (enter 1 into the console).

```{r install_TADA_dev, eval = T, include = F}
# helps with development (knit and checks), this chunk should not appear on website
remotes::install_github("USEPA/EPATADA",
  ref = "Create-Module-3-Reference-Tables",
  dependencies = TRUE
)
```

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the EPATADA R Package
into your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

All TADA R package functions have their own individual help pages,
listed on the [Function
reference](https://usepa.github.io/EPATADA/reference/index.html) page on
the GitHub site. Users can also access the help page for a given
function in R or RStudio using the following format (example below):
`?[name of TADA function]`

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

# Introduction to TADA Module 3

This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to create crosswalk reference tables that are needed
prior to defining and capturing their organization's criteria and
methodologies. It is the first of several vignettes that are being
developed as part of TADA Module 3.

Specifically, this vignette provides an overview of two functions that
can assist users with:

-   Creating a crosswalk reference table between ATTAINS parameter names
    and WQP/TADA characteristic names.

-   Creating a crosswalk reference table of all unique combinations of
    ATTAINS uses and and parameters.

These Module 3 functions are being designed with the flexibility for
users from states, tribes, or territories to input organization-specific
information. While TADA functions help generate these crosswalks, users
must review and modify the tables generated in each step of the process
to ensure their accuracy. Additionally, the TADA team has also
incorporated national recommended Clean Water Act (CWA) 304(a) numeric
criteria for optional use in Module 3 functions by leveraging data from
the [Criteria Search Tool
(CST)](https://www.epa.gov/wqs-tech/state-specific-water-quality-standards-effective-under-clean-water-act-cwa).
In the next vignette, TADAModule3_PartB , which is coming soon, this
will allow users to analyze their data against the 304(a) criteria; and
to easily compare results when when using 304(a) criteria vs. their own
organization's criteria.

# Example data from WQP

Let's start with an example data frame that has gone through data
cleaning, wrangling, harmonization, handling of censored data, removal
of suspect results, and other important TADA Module 1 functions. This
process should be completed by users before utilizing Module 3
functions. The example TADA data frame used in this vignette
(Data_WV_Mod1_Output) has already been through this process.

```{r}
# import example data set (output of Module 1 Workflow)
Data_NCTC <- Data_WV_Mod1_Output
```

Using an example data set from Maryland, we will work through an example
of creating two TADA Module 3 reference tables (TADA_CreateParamRef and
TADA_CreateParamUseRef) that are needed before we can start assigning
applicable criteria and methodologies information for a specific ATTAINS
organization.

# TADA_CreateParamRef() Basics

Before creating the WQP/TADA and ATTAINS parameter crosswalk table
(TADA_CreateParamRef), let's review the unique characteristic, fraction
and speciation combinations (TADA.ComparableDataIdentifier) and how many
results are available for each within the example data set (Data_NCTC).

```{r}
# create table with counts of TADA.ComparableDataIdentifiers
TADA_FieldValuesTable(Data_NCTC, field = "TADA.ComparableDataIdentifier")
```

We can also review the domain values of parameter names listed as causes
in prior assessment cycles (for all organizations) in ATTAINS by using a
function from rATTAINS, which provides functions for downloading tidy
data from the ATTAINS public web services
(<https://github.com/mps9506/rATTAINS>).

```{r, eval = FALSE}
# return ATTAINS domain names
rATTAINS::domain_values()

# return ATTAINS parameter domain values
TADA_CSVExport(rATTAINS::domain_values(domain_name = "ParameterName"))
```

As the ATTAINS parameters are more general than the
TADA.ComparableDataIdenitfiers, it is not possible at this time to
completely automate the assignment of TADA.ComparableIentifiers to
ATTAINS parameters. TADA_CreateParamRef() creates a template which
includes all TADA.ComparableDataIdentifiers from the TADA data frame and
provides a blank column, "ATTAINS.ParameterName" for users to input the
corresponding ATTAINS parameter name. The function automatically pairs
the EPA 304(a) pollutant name with the TADA.ComparableDataIdentifier for
some priority characteristics.

Setting "excel = TRUE" in TADA_CreateParamRef causes the function to
automatically create an Excel spreadsheet of the reference table for
easy review and editing. If users do not want to work in Excel, they can
use "excel = FALSE" to return a data frame. The data frame can be edited
directly in R if desired, although there are no TADA-specific functions
designed to facilitate this.

When using TADA_CreateParamRef(), users should specify the
organization(s) of interested in the "org_names" argument. This ensures
that the correct number of rows will be created in the reference table.
When "excel = TRUE", specifying the organization(s) also creates a
separate tab in the Excel file which contains the parameter names used
in prior assessment cycles by the specified ATTAINS organization(s).
This will allow users to decide whether to continue to use the same
ATTAINS parameter names from prior assessments that their
organizations(s) have used in the past, or to use other valid parameter
names from the entire ATTAINS domain list. If there is not a suitable
ATTAINS parameter name to match a TADA.ComparableDataIdentifier, users
should contact the ATTAINS team (attains\@epa.gov) for assistance.

```{r}
# return ATTAINS organization domain values
TADA_CSVExport(rATTAINS::domain_values(domain_name = "OrgName"))

# create TADA parameter reference table for specified organization
NCTC_ParamRef <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"), # Maryland
  excel = FALSE, # change to TRUE to download excel file
  overwrite = FALSE
)

# export TADA parameter reference table as csv
EPATADA::TADA_CSVExport(ref = NCTC_ParamRef)
```

Once the parameter reference table has been reviewed and modified, it
can be saved and reused for analysis of future TADA data frames.

The code chunk below demonstrates how the parameter reference table can
be modified in R. Once modifications are complete, we can re-run
TADA_CreateParamRef using the modified parameter reference table as the
input for "paramRef". This will update the ATTAINS.FlagParameterName
column in the parameter reference table (see NCTC_ParamRef2 in example
below). This column provides information about whether or not the
organization specified in the "organization_identifier" column has
listed this parameter as a cause in prior assessment cycles. It also
flags rows where the TADA.ComparableDataIdentifier as not been assigned
an ATTAINS parameter name.

```{r}
# only run the code chunks below if you make edits to the excel file.
# downloads_path <- file.path(Sys.getenv("USERPROFILE"), "Downloads", "myfileRef.xlsx")
# ParamRef <- openxlsx::read.xlsx(downloads_path, sheet = "CreateParamRef")

# NCTC_ParamRef3 <- TADA_CreateParamRef(
#   Data_NCTC,
#   org_names = c("MDE_EASP"),
#   paramRef = ParamRef,
#   excel = FALSE # comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
#   #excel = TRUE, overwrite = TRUE
#   )

ParamRef <- dplyr::mutate(NCTC_ParamRef, ATTAINS.ParameterName = dplyr::case_when(
  TADA.CharacteristicName == "PH" ~ "PH",
  TADA.ComparableDataIdentifier == "ZINC_TOTAL_NA_UG/L" ~ "ZINC",
  grepl("NITRATE", TADA.ComparableDataIdentifier) ~ "NITROGEN, TOTAL"
))
TADA_CSVExport(ref = ParamRef)

NCTC_ParamRef2 <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = ParamRef,
  excel = FALSE
) # comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
# excel = TRUE, overwrite = TRUE)
TADA_CSVExport(ref = NCTC_ParamRef2)
```

# TADA_CreateParamUseRef() Basics

Our first example will pull in all prior ATTAINS uses (use_names) and
parameter names (ATTAINS.ParameterName) specific to the specified
ATTAINS organization (defined by the org_names function argument for
TADA_CreateParamUseRef), which in this case is "MDE_EASP" (Maryland).
Users will review the output and choose which use_name's are applicable
to their analysis. The default functionality is to "Include" all ATTAINS
use_names (see the "IncludeOrExclude" column). Later in the data
analysis process, users will be asked to define criteria and
methodologies that is applicable to each ATTAINS use_name and parameter
name labeled as "Include". If a use_name is not applicable, users should
choose "Exclude" for that ATTAINS parameter name and use name.

For any ATTAINS parameter name(s) that have not been used by this
organization in prior assessment cycles, there will be no prior use
names associated with them. In this case, users will be responsible for
manually assigning the appropriate use_name applicable to their org
under the column 'use_name', and for inputting additional rows as needed
if the parameter applies to multiples uses.

In the example below, we can see "PH" was not a prior ATTAINS parameter
name for MDE_EASP (see ATTAINS.FlagUseName column in NCTC_ParamUseRef).
If desired, a user can manually assign this parameter to any applicable
uses (disclaimer: this is for demonstration purposes only and does not
reflect MDE_EASP's process).

```{r}
NCTC_ParamUseRef <- TADA_CreateParamUseRef(
  Data_NCTC,
  org_id = c("MDE_EASP"),
  paramRef = ParamRef,
  excel = FALSE)
TADA_CSVExport(ref = NCTC_ParamUseRef)
```

Users may also choose to include the EPA304a recommended water quality
criteria within the function input "org_names". This option utilizes the
EPA304a.PollutantName column created in TADA_CreateParamRef(). If a user
includes additional organizations beyond "MDE_EASP" and "EPA304a", like
in the example NCTC_ParamUseRef2 below, it will not generate rows for
the parameter and use combinations for those additional orgs ("21VASWCB"
(Virginia), "21PA"(Pennsylvania), WVDEP (West Virginia)) because only
MDE_EASP (Maryland) was defined in TADA_CreateParamRef() for paramRef =
ParamRef. If you would like to include other organizations here, you
must first fill out the TADA_CreateParamRef() for all organizations of
interest.

```{r}
NCTC_ParamUseRef2 <- TADA_CreateParamUseRef(
  Data_NCTC,
  org_id = c("EPA304a", "MDE_EASP", "21VASWCB", "21PA", "WVDEP"),
  paramRef = ParamRef,
  excel = FALSE
)
TADA_CSVExport(ref = NCTC_ParamUseRef2)
```

Instead, users should create a parameter and use reference table,
TADA_CreateParamUseRef(), for the same organizations included in the
parameter name crosswalk, TADA_CreateParamRef(). While aspects of this
process are partially automated, both crosswalks must be manually
reviewed, adjusted, and finalized by the user of these functions.

Users may choose to voluntarily share their reference tables for their
organization with the TADA Team so that others can use it in the future.

```{r}
NCTC_ParamUseRef3 <- TADA_CreateParamUseRef(
  Data_NCTC,
  org_id = c("EPA304a", "MDE_EASP"),
  paramRef = NCTC_ParamRef2,
  excel = FALSE
)
TADA_CSVExport(ref = NCTC_ParamUseRef3)
```

In this example, we designate that the ATTAINS.ParameterName "PH" is
equivalent to the TADA.ComparableDataIdentifier 'PH_NA_NA_STD UNITS',
which was not used in prior ATTAINS assessment cycles for **Maryland**.
This results in the 'use_name' column returning a blank value for
ATTAINS.ParameterName "PH". Therefore, we need to specify the
use_name(s) that we would like to associate with "PH" for this
organization. In excel, there will be a drop down list of organization
("MDE_EASP") specific use_names that were used in prior ATTAINS
assessments. If no excel file is generated, this is an example workflow
of how a user can edit the table in the R environment.

```{r paged.print=TRUE}
add_data <- data.frame("organization_identifier" = "MDE_EASP", "ATTAINS.ParameterName" = rep("PH", 3), "use_name" = c(
  "Aquatic Life and Wildlife", "Water Contact Sports",
  "Seasonal Migratory Fish Spawning and Nursery Subcategory"
))

# The output of this will not reflect changes to the ATTAINS.FlagUseName column. To do so, we need to re run TADA_CreateParamUseRef() with paramUseRef = ParamUseRef as an argument.
ParamUseRef <- NCTC_ParamUseRef3 %>%
  dplyr::left_join(add_data, by = c("organization_identifier", "ATTAINS.ParameterName"), keep = FALSE) %>%
  dplyr::mutate(use_name = dplyr::coalesce(use_name.x, use_name.y)) %>%
  dplyr::select(-c(use_name.x, use_name.y))

# User now has a saved dataframe which can be used to reflect any updates of Parameter-use ref in the future.
TADA_CSVExport(ref = ParamUseRef)

# PH will now reflect the changes
NCTC_ParamUseRef3.1 <- TADA_CreateParamUseRef(
  Data_NCTC,
  paramUseRef = ParamUseRef, # Edits were made to paramUseRef, updates flag column
  org_id = c("EPA304a", "MDE_EASP"),
  paramRef = NCTC_ParamRef2,
  excel = FALSE
)
TADA_CSVExport(ref = NCTC_ParamUseRef3.1)
```

# Next steps

Stay tuned! Future vignettes aim to assist with next steps in the
analysis process, including:

-   Defining the water quality criteria and methodologies used for each
    unique combination of ATTAINS assessment unit, use, and parameter.

-   Summarizing results by: (1) individual magnitude excursions.

Disclaimer: The EPATADA Module 3 functions are under active development.
EPATADA functions do not constitute current EPA recommendations, policy
or regulatory requirements. Organizations may optionally choose to use
EPATADA as a a tool in their process. Use of EPATADA is not required.
