---
title: "TADA Module 3b: Defining Criteria (Magnitude Only) and Summarizing Individual Excursions"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA Module 3: Associating WQP Characteristics with Applicable ATTAINS Parameters and Uses}
  %\VignetteEngine{knitr::rmarkdown}
description: The first of multiple vignettes (under development) showcasing TADA Module 3 functions and a recommended workflow.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to download the EPATADA R package from GitHub, access
and parameterize several important functions, and create basic
visualizations with a sample data set.

**Note: EPATADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.**

**If you are interested in contributing to EPATADA development, more
information is available at:**

[**Contributing**](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)

**We welcome collaboration with external partners.**

## Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load the EPATADA R Package using the remotes package.
Dependency packages will also be downloaded automatically from CRAN. You
may be prompted in the console to update dependencies that have more
recent versions available. If you see this prompt, it is recommended to
update all of them (enter 1 into the console).

```{r install_TADA_dev, eval = T, include = F}
# helps with development (knit and checks), this chunk should not appear on website
remotes::install_github("USEPA/EPATADA",
  ref = "Create-Module-3-AURef",
  dependencies = TRUE
)
```

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the EPATADA R Package
into your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Help pages

All TADA R package functions have their own individual help pages,
listed on the [Function
reference](https://usepa.github.io/EPATADA/reference/index.html) page on
the GitHub site. Users can also access the help page for a given
function in R or RStudio using the following format (example below):
`?[name of TADA function]`

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

## Module 3 Functions in TADA

Disclaimer: The EPATADA Module 3 functions were designed to: (1) assist
users with associating Water Quality Portal monitoring locations with
assessment units and designated uses from ATTAINS and (2) compare Water
Quality Portal results with numeric water quality criteria. EPATADA
functions do not constitute current EPA policy or regulatory
requirements. Organizations may choose to use EPATADA as a a tool in
their decision making processes. Use of EPATADA is not required.

## Introduction

Please see "TADA Module 3a: Associating WQP Characteristics with
Applicable ATTAINS Parameters and Uses" on information on creating the
necessary crosswalk tables prior to this step.

Once a user has completed the necessary crosswalk tables, they can
proceed to defining their organization's criteria and methods for a
parameter and use. The EPA304a standards, if a user has chosen to
include, will be autopopulated. Users are able to modify these methods
as desire.

We will start off by loading in the 3 user created reference files that
were created during the TADA module 3a vignette.

```{r}
Data_NCTC <- Data_HUC8_02070004_Mod1Output

NCTC_ParamRef <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("WVDEP"), # WV
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

ParamRef <- dplyr::mutate(NCTC_ParamRef, ATTAINS.ParameterName = dplyr::case_when(
grepl("PH", TADA.ComparableDataIdentifier) ~ "PH",
grepl("ZINC", TADA.ComparableDataIdentifier) ~ "ZINC",
grepl("NITRATE", TADA.ComparableDataIdentifier) ~ "NITROGEN, TOTAL"
))

NCTC_ParamRef_Final <- TADA_CreateParamRef(
  Data_NCTC,
  org_id = c("WVDEP"),
  paramRef = ParamRef,
  excel = FALSE,
  # excel = TRUE, overwrite = TRUE
)

add_data <- data.frame("organization_identifier" = "WVDEP" ,"ATTAINS.ParameterName" = rep("PH",3), "use_name" = c("Aquatic Life and Wildlife", "Water Contact Sports",
"Seasonal Migratory Fish Spawning and Nursery Subcategory"
))

# The output of this will not reflect changes to the ATTAINS.FlagUseName column. To do so, we need to re run TADA_CreateUseParamRef() with UseParamRef = UseParamRef as an argument.
UseParamRef_Final <- NCTC_UseParamRef3 %>%
  dplyr::left_join(add_data, by = c("organization_identifier", "ATTAINS.ParameterName"), keep = FALSE) %>% 
  dplyr::mutate(use_name = dplyr::coalesce(use_name.x, use_name.y)) %>% 
  dplyr::select(-c(use_name.x, use_name.y))

# PH will now reflect the changes 
NCTC_UseParamRef_Final <- TADA_CreateUseParamRef(
  Data_NCTC,
  #useParamRef = UseParamRef, # Edits were made to UseParamRef, updates flag column
  org_id = c("WVDEP"), 
  paramRef = NCTC_ParamRef_Final, 
  excel = FALSE # comment out 'excel = FALSE' and uncomment 'excel = TRUE, overwrite = TRUE' to run the excel file
  #excel = TRUE, overwrite = TRUE 
  ) 
```


# Create and Define Spatial Reference Tables

If you are interested in only a monitoring location level summary of your WQP data retrieval,
you do not need to perform any monitoring sites to assessment unit crosswalk table. In this scenario,
you will only need to run:

1) TADA_CreateParamRef
2) TADA_CreareUseParamRef
3) TADA_CreateSpatialRef: For this function, you do not need to provide the argument inputs for useAURef or waterUseParamRef

```{r}
NCTC_SpatialRef_ML <- TADA_CreateSpatialRef(
  Data_NCTC,
  org_id = "WVDEP",
  useParamRef = NCTC_UseParamRef_Final,
  excel = FALSE
)
```

Now, if you are interested in summarizing results on an assessment unit level, it is highly recommended to run the two helper function to run through this step by step process on how you may review or specify site-specific criteria prior to creating and defining your Criteria and Methodology table for a parameter, use and water type.

First, we will got through an example of how you can utilize the results directly from TADA_GetATTAINS() as your ML to AU crosswalk. This crosswalk table will be named SiteAURef.

```{r}
GetATTAINSRef <- TADA_GetATTAINS(Data_NCTC, return_sf = FALSE)
SitesAURef <- GetATTAINSRef
```

In this first example, we demonstrate what your output table will look like in TADA_CreateSpatialRef() if you decide to not run the two helper functions: TADA_CreateUseAURef() and TADA_CreateWaterUseParamRef().

These helper functions are not required, but will help to further review your ML to AU crosswalk as well as provide support in filling out site-specific criteria. In this example, our spatial area in our TADA daraframe contains WQP observations from other states (MD, PA, VA). These observations are flagged in the column Flag.AssessmentNote. It is up to the user to decide if these are relevant to their assessments, and if not, they should remove these rows. This column will also help to identify other cases such as no ML to AU matches, no use_names associated with your AUs and more! (Still in development)

```{r}
NCTC_SpatialRef_No_Helper <- TADA_CreateSpatialRef(
  Data_NCTC, 
  org_id = "WVDEP",
  sitesAURef = SitesAURef,
  useParamRef = NCTC_UseParamRef_Final,
  # useAURef = NCTC_UseAURef,
  # waterUseParamRef = NCTC_WaterUseParamRef_Final,
  excel = FALSE
  # excel = TRUE, overwrite = TRUE
  )
```

Now, lets run through an example with the helper functions. We must run these helper functions in order. First, we will run TADA_CreateUseAURef(). This will use the rExpertQuery data query pull of ATTAINS national extracts to pull in prior use_names associated with your assessment units. This helper function is helpful in cases in which a user has new AU or modified AU that are not found in prior ATTAINS cycles. In this case, it is important that users define these use names associated with those AUs as TADA will be unable to tell what use names are appropriate and will require your review.

We note that all AU will contain a match as the AUs came directly from TADA_GetATTAINS(). But lets add on a few rows to demonstrate cases of new AUs. Note: A user may also choose to perform adding new ML to AU as part of the sitesAURef, but they would still need to specify new AU and use names during this step.

```{r}
NCTC_UseAURef <- TADA_CreateUseAURef(
  Data_NCTC, 
  SitesAURef = SitesAURef, 
  org_id = "WVDEP", 
  excel = FALSE
  #TRUE, overwrite = TRUE
  )

add_AU_df <- data.frame(
  organization_identifier = c("WVDEP", "WVDEP"),
  ATTAINS.assessmentunitidentifier = c("NEW_WV_AU1", "NEW_WV_AU1"), 
  ATTAINS.assessmentunitname = c("Example1", "Example1"),
  use_name = c("Public Water Supply", "Warm Water Fishery"),
  waterType = c("STREAM/CREEK/RIVER", "STREAM/CREEK/RIVER"),
  Existing = c("New", "New"),
  geometry = c()
)

NCTC_UseAURef_Final <- dplyr::left_join(NCTC_UseAURef, add_AU_df, by = c(-"geometry"))
```

```{r}
NCTC_SpatialRef2 <- TADA_CreateSpatialRef(
  Data_NCTC, 
  org_id = "WVDEP",
  sitesAURef = SitesAURef,
  #useAURef = NCTC_UseAURef,
  useParamRef = NCTC_UseParamRef_Final,
  #waterUseParamRef = NCTC_WaterUseParamRef_Final,
  excel = TRUE, overwrite = TRUE
  )

NCTC_UseAURef_Final <- TADA_CreateUseAURef(
  Data_NCTC, 
  SitesAURef = SitesAURef, 
  org_id = "WVDEP", 
  excel = FALSE
  ) #TRUE, overwrite = TRUE)

NCTC_WaterUseParamRef_Final <- TADA_CreateWaterUseParamRef(
  Data_NCTC, 
  useAURef = NCTC_UseAURef_Final,
  useParamRef = NCTC_UseParamRef_Final, 
  org_id = "WVDEP", 
  excel = FALSE
  ) #TRUE, overwrite = TRUE)


NCTC_Excursions <- TADA_MagnitudeSummary(
  Data_NCTC,
  #UseAURef = NCTC_UseAURef,
  StandardsRef = NCTC_magnitude_Final,
  overwrite = FALSE
)

NCTC_Excursions2 <- TADA_MagnitudeSummary(
  Data_NCTC,
  UseAURef = NCTC_UseAURef,
  StandardsRef = NCTC_magnitude_Final,
  overwrite = TRUE
)
```

# Defining Criteria - Magnitude Methodology

Next, users will use the reference tables in order to define the
associated methods for each parameter and use combination that they have
defined.

```{r}
NCTC_magnitude_Final <- TADA_DefineMagnitude(
  Data_NCTC, 
  paramRef = NCTC_ParamRef_Final, 
  UseParamRef = NCTC_UseParamRef_Final, 
  AURef = NCTC_SpatialRef, 
  excel=FALSE,
  #excel = TRUE, overwrite = TRUE
  )
# If user makes update in excel file
downloads_path <- file.path(Sys.getenv("USERPROFILE"), "Downloads", "myfileRef.xlsx")
# NCTC_magnitude_Final <- openxlsx::read.xlsx(downloads_path, sheet = "DefineMagnitude")
```
