---
title: "TADA R8 Demo January 2026"
always_all_html: yes
author: "TADA Team"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    self_contained: false
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteIndexEntry{TADA R8 Demo}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{EPATADA}
  %\VignetteEngine{knitr::rmarkdown}
description: An introduction to using the EPATADA R package to retrieve, clean, analyze, and visualize WQP data. This recommended work flow focuses on A) Retrieving, filtering and cleaning data from the WQP. B) Create, review and define up to date assessment unit/WQP monitoring location. C) Assigning uses to assessment units. D) Create a TADA-Compatible criteria table which defines the relevant parameters and uses applicable for analysis of your WQP data.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = T,
  dev.args = list(bg = "transparent")
)
```

# Overview and Setup

**Welcome!**

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to download the TADA R package from GitHub, access and
parameterize several important functions, and create basic
visualizations with a sample data set.

**Note:** TADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.

If you are interested in contributing to TADA development, more
information is available at:

[Contributing](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)

We welcome collaboration with external partners.

Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing TADA because it is only available on GitHub.

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load TADA using the remotes package. TADA R Package
dependencies will also be downloaded automatically from CRAN with the
TADA install. You may be prompted in the console to update dependency
packages that have more recent versions available. If you see this
prompt, it is recommended to update all of them (enter 1 into the
console).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r dev_install, eval = F, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the library() function to load the TADA R Package into your
R session.

```{r library, results = 'hide'}
library(EPATADA)
```

Help pages

All TADA R package functions have their own individual help pages,
listed on the Function reference page on the GitHub site. Users can also
access the help page for a given function in R or RStudio using the
following format (example below): ?[name of TADA function]

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

## A) Retrieving, filtering and cleaning data from the WQP

Query the WQP using TADA_DataRetrieval. TADA_AutoClean is a powerful
function that runs as part of TADA_DataRetrieval when applyautoclean =
TRUE. It performs a variety of tasks, for example:

1.  creating new "TADA" prefixed columns and and capitalizing their
    contents to reduce case sensitivity issues,

2.  converts special characters in value columns,

3.  converts latitude and longitude values to numeric,

4.  replaces "meters" with "m",

5.  replaces deprecated characteristic names with current WQX names,

6.  harmonizes result and detection limit units,

7.  converts depths to meters, and

8.  creates the column TADA.ComparableDataIdentifier by concatenating
    characteristic name, result sample fraction, method speciation, and
    result measure unit.

In this example, we will get bacteria and pH data from Missoula County,
Montana.

```{r load WQP data}
# get MT data
tada.MT <- TADA_DataRetrieval(
  startDate = "2020-01-01",
  endDate = "2022-12-31",
  statecode = "MT",
  characteristicName = c(
    "Dissolved oxygen (DO)",
    "Escherichia",
    "Escherichia coli",
    "pH"
  ),
  countycode = "Missoula County",
  applyautoclean = TRUE,
  ask = FALSE
)
```

### Flag, clean, and visualize

Now, let's use EPATADA functions to review, visualize, and whittle the
returned WQP data down to include only results that are applicable to
our water quality analysis and area of interest.

TADA is primarily designed to accommodate water data from the WQP. Let’s
see what activity media types are represented in the data set. Are there
any media type that are not water in our data frame?

```{r Review and Filter By Media Type}
# Create table with count for each ActivityMediaName
TADA_FieldValuesTable(
  tada.MT,
  field = "ActivityMediaName",
  characteristicName = "null"
)
```

Create an overview map. For each site, we can view the measurement
count, visit count and characteristic counts along with the site ID,
site name, and organization name.

```{r TADA_OverviewMap}
TADA_OverviewMap(tada.MT)
```

Let's take a quick look at all unique values in the
MonitoringLocationIdentifier column and see how how many results are
associated with each.

```{r TADA_FieldValuesTable}
# use TADA_FieldValuesTable to create a table of the number of results per MonitoringLocationIdentifier
sites <- TADA_FieldValuesTable(
  tada.MT,
  field = "MonitoringLocationIdentifier",
  characteristicName = "null"
)

DT::datatable(sites, fillContainer = TRUE)
```

What about OrganizationFormalName?

```{r TADA_FieldValuesTable2 }
# use TADA_FieldValuesTable to create a table of the number of results per MonitoringLocationIdentifier
orgs <- TADA_FieldValuesTable(
  tada.MT,
  field = "OrganizationFormalName",
  characteristicName = "null"
)

DT::datatable(orgs, fillContainer = TRUE)
```

Let's filter our TADA data frame to only include data from the top 4
OrganizationFormalName. The orgs data frame is sorted in descending
order in which 1:4 will select the top 4 OrganizationFormalName values.

```{r}
tada.MT.filter <- tada.MT |>
  dplyr::filter(
    OrganizationFormalName %in% orgs$Value[1:4]
    )
```

Are there sites located within 100 meters of each other?

```{r TADA_FlagCoordinates}
# tada.MT.nearby <- TADA_FindNearbySites(
#   tada.MT.filter,
#   dist_buffer = 100,
#   nhd_res = "Hi",
#   org_hierarchy = "none",
#   meta_select = "random",
#   by_AU = FALSE,
#   catchment = FALSE
# )
# 
# TADA_NearbySitesMap(tada.MT.nearby)
```

Next, let's check if the dataset contains potential duplicate results
from within a single organization or from within multiple organizations
(such as when two or more organizations monitor the same location and
may submit duplicate results).

For the case in which there exists potential duplicates from multiple
organizations, if you would like to prioritize results from one
organization over another, this can be done using the org_hierarchy
argument in `TADA_FindPotentialDuplicatesMultipleOrgs`.

```{r duplicates}
# # find duplicates from single org
# tada.MT.nearby.flag.dups <- TADA_FindPotentialDuplicatesSingleOrg(tada.MT.nearby)
# 
# # Review organizations. You can select one to prioritize in TADA_FindPotentialDuplicatesMultipleOrgs
# unique(tada.MT.nearby.flag.dups$OrganizationIdentifier)
# unique(tada.MT.nearby.flag.dups$OrganizationFormalName)
# 
# # find duplicates across multiple orgs
# tada.MT.nearby.flag.dups <- TADA_FindPotentialDuplicatesMultipleOrgs(
#   tada.MT.nearby.flag.dups,
#   org_hierarchy = c("MDEQ_WQ_WQX", "TSWQC_WQX", "MTWTRSHD_WQX", "USGS-MT")
# )
```

We will select to keep only unique samples from
`TADA_FindPotentialDuplicatesSingleOrg` by filtering for
TADA.SingleOrgDup.Flag equals "Unique".

There are no multiple org duplicates from
`TADA_FindPotentialDuplicatesMultipleOrgs` in this example, but if there
were, duplicates can by removed by filtering for
TADA.ResultSelectedMultipleOrgs equals "Y".

```{r filter}
# tada.MT.clean <- tada.MT.nearby.flag.dups |>
#   dplyr::filter(TADA.SingleOrgDup.Flag == "Unique") |>
#   dplyr::filter(TADA.ResultSelectedMultipleOrgs == "Y")
tada.MT.clean <- tada.MT.filter
```

Remove intermediate variables in R by using 'rm()'. In the remainder of
this workshop, we will work with the clean data set.

```{r}
rm(tada.MT.nearby.flag.dups, tada.MT.nearby, tada.MT.filter)
```

Censored data are measurements for which the true value is not known,
but we can estimate the value based on known lower or upper detection
conditions and limit types. TADA fills missing *TADA.ResultMeasureValue*
and *TADA.ResultMeasure.MeasureUnitCode* values with values and units
from *TADA.DetectionQuantitationLimitMeasure.MeasureValue* and
*TADA.DetectionQuantitationLimitMeasure.MeasureUnitCode*, respectively,
using the `TADA_AutoClean` function.

The TADA package currently has functions that summarize censored data
incidence in the dataset and perform simple substitutions of censored
data values, including x times the detection limit and random selection
of a value between 0 and the detection limit. The user may specify the
methods used for non-detects and over-detects separately in the input to
the `TADA_SimpleCensoredMethods` function. The next step we take in this
example is to perform simple conversions to the censored data in the
dataset: we keep over-detects as is (no conversion made) and convert
non-detect values to 0.5 times the detection limit (half the detection
limit).

In this example, no censored data results were found. The input
tada.MT.clean data is returned.

```{r censored}
tada.MT.clean <- TADA_SimpleCensoredMethods(
  tada.MT.clean,
  nd_method = "multiplier",
  nd_multiplier = 0.5,
  od_method = "as-is",
  od_multiplier = "null"
)
```

`TADA_FindQCActivities` removes results with QA/QC ActivityTypeCode's.

```{r QCdata}
tada.MT.clean <- TADA_FindQCActivities(tada.MT.clean, clean = TRUE)
```

`TADA_ConvertSpecialChars` removes rows where the result value is not
numeric to prepare a dataframe for quantitative analyses. Specifically,
this function removes rows with "Text" and "NA - Not Available" in the
TADA.ResultMeasureValueDataTypes.Flag column, or NA in the
TADA.ResultMeasureValue column. In addition, this function removes
results with QA/QC ActivityTypeCode's. This function also removes any
columns not required for TADA workflow where all values are equal to NA.

```{r SpecialChars}
tada.MT.clean <- TADA_ConvertSpecialChars(tada.MT.clean, col = "TADA.ResultMeasureValue", clean = TRUE)
```

TADA_RunKeyFlagFunctions is a shortcut function to run important TADA
flagging functions. See ?function documentation for TADA_FlagResultUnit,
TADA_FlagFraction, TADA_FlagMeasureQualifierCode, and
TADA_FlagSpeciation for more information.

```{r TADA_RunKeyFlagFunctions}
tada.MT.clean <- TADA_RunKeyFlagFunctions(
  tada.MT.clean
)
```

Let's take another look at all unique values in the
TADA.ComparableDataIdentifier column and see how how many results are
associated with each. TADA.ComparableDataIdentifier concatenates
TADA.CharacteristicName, TADA.ResultSampleFractionText,
TADA.MethodSpeciationName, and TADA.ResultMeasure.MeasureUnitCode.

```{r TADA_FieldValuesTable3}
# use TADA_FieldValuesTable to create a table of the number of results per TADA.ComparableDataIdentifier
chars <- TADA_FieldValuesTable(
  tada.MT.clean,
  field = "TADA.ComparableDataIdentifier",
  characteristicName = "null"
)

chars_before <- unique(tada.MT.clean$TADA.ComparableDataIdentifier)
chars_before

DT::datatable(chars, fillContainer = TRUE)
```

Scroll through the table and check to see if there any synonyms. Are
there any TADA.ResultSampleFractionText, TADA.MethodSpeciationName, and
TADA.ResultMeasure.MeasureUnitCode that may need to be harmonized for
consistency? Let's check using `TADA_HarmonizeSynonyms` so their results
can be directly compared.

Let's give it a try.

```{r}
tada.MT.clean <- TADA_HarmonizeSynonyms(tada.MT.clean)
```

How many unique TADA.ComparableDataIdentifier's do we have now? In this
example, PH has now been harmonized from to "PH_NONE_NONE_NONE". This is
the recommended consistent format for PH harmonization across all states
and tribes. Dissolved Oxygen has also been harmonized.

```{r}
chars_after <- unique(tada.MT.clean$TADA.ComparableDataIdentifier)
chars_after
```

Remove intermediate variables.

```{r}
rm(chars_before, chars_after)
```

```{r thresholds}
tada.MT.clean <- TADA_FlagAboveThreshold(tada.MT.clean, clean = TRUE, flaggedonly = FALSE)

tada.MT.clean <- TADA_FlagBelowThreshold(tada.MT.clean, clean = TRUE, flaggedonly = FALSE)
```

## visualize data

Create a pie chart.

```{r}
TADA_FieldValuesPie(
  tada.MT.clean,
  field = "TADA.CharacteristicName",
  characteristicName = "null"
)
```

## B) Create, review and define up to date assessment unit/WQP monitoring location.

Using `tada.MT.clean` we can pull in the ATTAINS catchment features that
intersect our observations:

```{r, eval = T}
TADA_with_ATTAINS <- TADA_CreateATTAINSAUMLCrosswalk(
  .data = tada.MT.clean,
  return_sf = FALSE,
  return_nearest = TRUE
)
```

For demonstration purposes, let's assume we know for sure this
MonitoringLocationIdentifier, "MTWTRSHD_WQX-COMBITR02" should be
associated with the AssessmentUnitIdentifier "MT76M001_020". We are
however uncertain if there are other sites that we want to associate
with this AssessmentUnitIdentifier. We will use the result from
TADA_with_ATTAINS above and filter our TADA data frame to
AssessmentUnitIdentifier "MT76M001_020" and also keep the\
MonitoringLocationIdentifier, "MTWTRSHD_WQX-COMBITR02".

```{r demo.user, results = 'hide'}
# create an example user supplied crosswalk (user would like to associate this MonitoringLocationIdentifier)
user.supplied.cw <- data.frame(
    AssessmentUnitIdentifier = "MT76M001_020",
    MonitoringLocationIdentifier = "MTWTRSHD_WQX-COMBITR02",
    WaterType = "RIVER"
  )
```

filter our cleaned data frame as specified. We will also filter our data
frame to just DISSOLVED OXYGEN (DO)

```{r filter single AU}
# use TADA_FieldValuesTable to create a table of the number of results per TADA.ComparableDataIdentifier
unique(TADA_with_ATTAINS$assessmentunitidentifier)

TADA_with_ATTAINS_filter <- TADA_with_ATTAINS |>
  dplyr::filter(
    assessmentunitidentifier == "MT76M001_020" | MonitoringLocationIdentifier == "MTWTRSHD_WQX-COMBITR02",
    TADA.CharacteristicName == "DISSOLVED OXYGEN (DO)")

tada.MT.clean <- tada.MT.clean |>
  dplyr::filter(
    MonitoringLocationIdentifier %in% TADA_with_ATTAINS_filter$MonitoringLocationIdentifier,
    TADA.CharacteristicName == "DISSOLVED OXYGEN (DO)")
```

### Run `TADA_CreateAUMLCrosswalk`

Make the AU assignments for the remaining unassigned MLs not defined
from our user.supplied.cw using TADA_CreateATTAINSAUMLCrosswalk.

```{r TADA_CreateAUMLCrosswalk}
# make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLCrosswalk(
  tada.MT.clean,
  au_ref = user.supplied.cw,
  org_id = "MTDEQ",
  fill_ATTAINS_catch = TRUE,
  return_nearest = TRUE,
  batch_upload = TRUE
)
```

Let's view the data associated with MT.AUMLRef.

```{r view.map}
TADA_ViewATTAINS(MT.AUMLRef, ref_icons = TRUE)
```

## C) Assigning uses to assessment units.

We will use the ATTAINS.UseNames that were used in prior
ATTAINS assessment cycles from MTDEQ for this assessment unit.
Users can append additional rows as needed or remove not applicable
use names if a use is no longer applicable for the assessment unit.
See TADA_Module2 vignette for more information.

```{r}
AU_uses <- TADA_AssignUsesToAU(
  tada.MT.clean,
  org_id = "MTDEQ",
  AUMLRef = MT.AUMLRef$ATTAINS_crosswalk
  )
```

## D) Create a TADA-Compatible criteria table which defines the relevant parameters and uses applicable for analysis of your WQP data.

We will load in a pre-filled criteria template completed by MTDEQ.
(R8 partners have also contributed to the TADA_CommunityHub repository
which contains pre-filled templates as well for R8 states and tribes!)

```{r}
# Load the example R8 criteria table
criteria_table <- system.file("extdata", "criteria_table.rda", package = "EPATADA")
load(criteria_table)

MT_criteria <- TADA_DefineCriteriaMethodology(
  tada.MT.clean,
  org_id = "MTDEQ",
  criteriaMethods = criteria_table,
  AUMLRef = MT.AUMLRef$ATTAINS_crosswalk,
  AU_UsesRef = AU_uses
  )
```

View scatterplots for all parameter and use criteria defined in your
table compared to the results from your TADA data frame for DO. 

Note that the scatterpoints are only individual result measurements. Future
TADA development in the package is being made for handling of the full
criteria (magnitude, duration and frequency), seasonal considerations and 
other data sufficiency needs!

```{r}
TADA_Scatter_list <- list()

n <- length(MT_criteria$MagnitudeValueLower)

for(i in 1:n){

fig <- TADA_Scatterplot(
  tada.MT.clean,
  id_cols = "TADA.ComparableDataIdentifier") |>
  plotly::add_lines(
    y = MT_criteria$MagnitudeValueLower[i],
    x = c(min(tada.MT.clean$ActivityStartDate), max(tada.MT.clean$ActivityStartDate)),
    inherit = FALSE,
    showlegend = TRUE,
    line = list(color = "red"),
    #hoverinfo = "none",
    name = paste(MT_criteria$ATTAINS.UseName[i], MT_criteria$MagnitudeValueLower[i],MT_criteria$MagnitudeUnit[i], "as a",  MT_criteria$DurationValue[i], MT_criteria$DurationUnit[i], MT_criteria$DurationMethod[i])
  )

fig <- fig |>
  plotly::layout(
    legend = list(
      orientation = "h", # Horizontal legend
      x = 0,             # X coordinate (0 is left, 1 is right)
      y = -0.7           # Y coordinate (negative value places it outside and below the plot area)
    ),
    margin = list(b = 100) 
  )
TADA_Scatter_list[[i]] <- fig
}
```

View individual scatterplots with defined criteria for different
combination of uses for DO.

```{r}
TADA_Scatter_list[[1]]
TADA_Scatter_list[[2]]
```
