---
title: 'Beginner R user: introduction to using the EPATADA R package to retrieve, clean, analyze, and visualize WQP data'
author: "TADA Team"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = F)
```

## Install and load

First, install and load the remotes package specifying the repo. This is
needed before installing the *EPATADA* R package because it is only
available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes", repos = "http://cran.us.r-project.org")
library(remotes)
```

Next, install (or update) and load the *EPATADA* R package using the
*remotes* R package. Additional dependency R packages that are used
within *EPATADA* will be downloaded automatically. You may be prompted
in the console to update dependency packages that have more recent
versions available. If you see this prompt, it is recommended to update
all of them (enter 1 into the console). Our team is actively developing
*EPATADA*, therefore we highly recommend that you update the package
(and all of its dependencies) each time you use it.

```{r install_EPATADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA", ref = "develop", dependencies = TRUE)
```

```{r install_EPATADA_dev, eval = T, include = F}
remotes::install_github("USEPA/EPATADA", ref = "cybertown2025", dependencies = TRUE
)
```

## Record start time

```{r}
start.time <- Sys.time()
```

## Retrieve and filter

Use HMW to find a watershed of interest:
<https://mywaterway.epa.gov/community/Shepherdstown,%20WV,%20USA/identified-issues>

```{r TADA_DataRetrieval}
bbox <- c(-77.9947, 39.3086, -86.6135, 39.6164)
tada8 <- TADA_DataRetrieval(bBox = bbox)
 
WQP_Shepherdstown <- TADA_DataRetrieval(
  startDate = "2019-01-01",
  endDate = "2024-12-31",
  bBox = bbox,
  # huc = c("0207000411"),
  applyautoclean = TRUE,
  ask = FALSE)
```

```{r TADA_GetATTAINS}
ATTAINS_Shepherdstown <- TADA_GetATTAINS(WQP_Shepherdstown)
```

## Map

```{r TADA_ViewATTAINS}
# View catchments and assessment units on map

ATTAINS_map <- TADA_ViewATTAINS(ATTAINS_Shepherdstown)

ATTAINS_map
```

What characteristics are available?

```{r fieldValues_actmedia}
# Create pie chart for CharacteristicName
CharName_Pie <- TADA_FieldValuesPie(WQP_Shepherdstown, field = "TADA.CharacteristicName")

CharName_Pie
```

How many results are available for each characteristic?

```{r}
# Create table with count for each TADA.CharacteristicName
CharName_FieldValues <- TADA_FieldValuesTable(WQP_Shepherdstown, field = "TADA.CharacteristicName")

CharName_FieldValues
```

Are there any synonyms?

```{r}
WQP_Shepherdstown_harmonized = TADA_HarmonizeSynonyms(WQP_Shepherdstown)
```

Filter: let's focus on a few characteristics of interest.

```{r}
# Select characteristics of interest
WQP_Shepherdstown_subset <- WQP_Shepherdstown %>%
  dplyr::filter(TADA.CharacteristicName == c("PH", # 6.5-8.5 but dependent on application
                                             "SPECIFIC CONDUCTANCE", # 250000 µg/l Human Health (Human Health • Water + Organism)
                                             "IRON", # 1000 µg/l Aquatic Life, Freshwater, Chronic OR 300 µg/l Organoleptic
                                             "CHLORIDE", # 860000 µg/l (Aquatic Life • Freshwater • Acute) OR 230000 µg/l (Aquatic Life • Freshwater • Chronic)
                                             "FECAL COLIFORM",
                                             "ESCHERICHIA COLI"
                                             ))
```

## Clean

## Analyze

## Visualize

## Record end time

```{r analysis time}
end.time <- Sys.time()

end.time - start.time
```
