---
title: "TADA Geospatial Functions"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{TADA Geospatial Functions}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = FALSE,
  fig.width = 8,
  fig.height = 8
)
```

## Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to download the TADA R package from GitHub, access and
parameterize several important functions, and create basic
visualizations with a sample data set. The sample data set contains data
from one week from all EPA Region 5 states.

**Note: TADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.**

**If you are interested in contributing to TADA development, more
information is available at [Contributing](<https://usepa.github.io/TADA/articles/CONTRIBUTING.html>). We welcome
collaboration with external partners.**

## Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing TADA because it is only available on GitHub.

```{r install_remotes, eval = F, results = 'hide', message = FALSE, warning = FALSE}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load TADA using the remotes package. TADA R Package
dependencies will also be downloaded automatically from CRAN with the
TADA install. You may be prompted in the console to update dependency
packages that have more recent versions available. If you see this
prompt, it is recommended to update all of them (enter 1 into the
console).

```{r install_TADA, eval = F, results = 'hide', message = FALSE, warning = FALSE}
remotes::install_github("USEPA/TADA",
  ref = "R5_package_demo",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the TADA R Package into
your R session.

```{r library, results = 'hide', message = FALSE, warning = FALSE}
library(TADA)
```

## Help pages

All TADA R package functions have their own individual help pages,
listed on the [Function
reference](https://usepa.github.io/TADA/reference/index.html) page on
the GitHub site. Users can also access the help page for a given
function in R or RStudio using the following format (example below):
`?TADA::[name of TADA function]`

```{r help_pages, message = FALSE}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

## Geospatial Functions in TADA

This vignette represents functions that provide users the option to
convert TADA Water Quality Portal data into a geospatial sf object as
well as to associate water quality observations with their nearest
state-defined water quality assessment units in ATTAINS.

**A Note About ATTAINS:**

The Assessment, Total Maximum Daily Load (TMDL) Tracking and
Implementation System ([ATTAINS](https://www.epa.gov/waterdata/attains))
is an online platform that organizes and combines each state’s Clean
Water Act reporting data into a single data repository. The geospatial
component of ATTAINS includes spatial representations of each state’s
assessment units as well as their assigned designated uses, their most
recent EPA reporting category (i.e., their impairment status), their
impaired designated uses, and the parameter(s) causing the impairment.

Within an assessment unit, the water quality standards remain the same
and all water features are assessed as one entity. Depending on the
state, these assessment units can be a specific point along a river, a
river reach, an entire river, or even an entire watershed. (In other
words, assessment units can take the form of point, line, and area
features, or some combination of all of them.) Moreover, it is possible
that some assessment units are not geospatially referenced at all,
meaning they are not captured in the ATTAINS geospatial database.

## `TADA_MakeSpatial()`

This function converts any Water Quality Portal (WQP)-style dataframe
with latitude/longitude data into a geospatial sf object. To run the
function, the user supplies a WQP dataframe and the coordinate reference
system that they want the spatial object to be in (the default is WGS
84). For the function to work properly, the input dataframe must have -
at a minimum - WQP observation coordinates in "LongitudeMeasure" and
"LatitudeMeasure" and a "HorizontalCoordinateReferenceSystemDatumName"
column.

#### Using `TADA_MakeSpatial()`

First, we will need to pull in some TADA Water Quality Portal Data:

```{r}
# pH data in Larimer County, Colorado for the year 2020. 
TADA_dataframe <- TADA_DataRetrieval(startDate = "2020-01-01",
                                     endDate = "2020-12-31",
                                     characteristicName = "pH",
                                     countycode = "US:08:069",
                                     applyautoclean = TRUE)
```

Now, we can make the water quality data spatial by running
`TADA_MakeSpatial()`:

```{r}
# default CRS is WGS84 (4326)
TADA_spatial <- TADA_MakeSpatial(data = TADA_dataframe, crs = 4326)
```

This new spatial object is identical to the original TADA dataset, but
now includes a "geometry" column that allows for mapping and additional
geospatial capabilities. Enter `?TADA_MakeSpatial` into the console to
review another example of this function in use and additional
information.

```{r}
leaflet::leaflet() %>% 
  leaflet::addProviderTiles("Esri.WorldTopoMap", 
                            group = "World topo", 
                            options = leaflet::providerTileOptions(updateWhenZooming = FALSE,
                                                                   updateWhenIdle = TRUE)) %>% 
  leaflet::clearShapes() %>% 
  leaflet.extras::addResetMapButton()  %>%
  leaflet::addLegend(position = "bottomright",
                     colors = "black",
                     labels = "Water Quality Observation(s)",
                     opacity = 1) %>%
  leaflet::addCircleMarkers(data = TADA_spatial,
                            color = "grey", fillColor = "black", 
                            fillOpacity = 0.8, stroke = TRUE, weight = 1.5, radius = 6, 
                            popup = paste0("Site ID: ", 
                                           TADA_spatial$MonitoringLocationIdentifier, 
                                           "<br> Site Name: ",
                                           TADA_spatial$MonitoringLocationName))
```

## `TADA_GetATTAINS()`

This function pulls in ATTAINS data from the EPA's ATTAINS Assessment
Geospatial Service and links it to TADA-pulled Water Quality Portal
observations. For the function to work properly, the input dataframe
must have - at a minimum - WQP observation coordinates in
"LongitudeMeasure" and "LatitudeMeasure" columns and a
"HorizontalCoordinateReferenceSystemDatumName" column.

Users also have the option of returning the ATTAINS geospatial sf
objects with their ATTAINS-linked Water Quality Portal dataframe. If
`return_sf = TRUE`, the function returns a list containing the dataframe
plus shapefile objects named `ATTAINS_catchments`, `ATTAINS_lines`,
`ATTAINS_points`, and `ATTAINS_polygons`. Note, if any of these
shapefile objects are empty, this indicates that there are no ATTAINS
objects of that type intersecting any WQP catchment.

Regardless of the user's decision on returning the ATTAINS sf objects,
`TADA_GetATTAINS()` always returns a dataframe containing the original
TADA WQP dataset, plus new columns representing the ATTAINS assessment
unit(s) that fall within the same NHDPlus HR catchment as them. This
means that it is possible for a single TADA WQP observation to have
multiple ATTAINS assessment units linked to it and subsequently more
than one row of data. Such WQP observations can be identified using the
new `index` column (i.e., multiple rows with the same index value are
the same observation).

#### Using `TADA_GetATTAINS()`

Using either our original `TADA_dataframe` *or* the geospatial version
`TADA_spatial`, we can pull in the ATTAINS features that are within the
same NHD HR catchment as our observations:

```{r}
TADA_with_ATTAINS <- TADA_GetATTAINS(data = TADA_dataframe, return_sf = FALSE)

TADA_with_ATTAINS <- TADA_GetATTAINS(data = TADA_spatial, return_sf = FALSE)
```

This new `TADA_with_ATTAINS` object is a modification of the original
TADA Water Quality Portal dataframe that now has additional columns
associated with the ATTAINS assessment unit(s) that lie in the same NHD
HR catchment as them (these columns are prefixed with "ATTAINS").
Moreover, because our `TADA_with_ATTAINS` object contains more rows than
the original TADA dataframe, we can deduce that some Water Quality
Portal observations fall within an NHD HR catchment that contains more
than one ATTAINS assessment unit.

```{r}
TADA_with_ATTAINS_list <- TADA_GetATTAINS(data = TADA_dataframe, return_sf = TRUE)

TADA_with_ATTAINS_list <- TADA_GetATTAINS(data = TADA_spatial, return_sf = TRUE)
```

If we set `return_sf = TRUE` as done to create the
`TADA_with_ATTAINS_list` object above, we also now have all the raw
ATTAINS features associated with these TADA Water Quality Portal
observations stored in a list with the TADA dataframe.

Enter `?TADA_GetATTAINS` into the console to review another example of
this function in use and additional information.

## `TADA_ViewATTAINS()`

This function visualizes the raw ATTAINS features that are linked to the
TADA Water Quality Portal observations that are generated in
`TADA_GetATTAINS()` when `return_sf = TRUE`. For the function to work
properly, the input dataframe must be the list produced from
`TADA_GetATTAINS()` with `return_sf = TRUE`. The map also displays the
Water Quality Portal monitoring locations used to find the ATTAINS
features.

#### Using `TADA_ViewATTAINS()`

Let's view the data associated with our `TADA_with_ATTAINS_list` object!
Enter `?TADA_ViewATTAINS` into the console to review another example
query and additional information.

```{r}
TADA_ViewATTAINS(ATTAINS_list = TADA_with_ATTAINS_list)
```
