---
title: "TADA: WaterSciCon Workshop Demo"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA: WaterSciCon Workshop Demo}
  %\VignetteEngine{knitr::rmarkdown}
description: A demonstration of data discovery for the confluence of the San Juan and Animas Rivers near Farmington, New Mexico. This example focuses on surface water quality with a focus on total dissolved solids and conduvitivty.

editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview

We will walk through the specific use case of query, downloading,
preparing, and QA/QCing data from the San Juan and Animas Rivers near
Farmington, New Mexico with a focus on total dissolved solids and
conductivity.

This vignette is designed to be concise and provide an example of how
you might use TADA functions for basic data discovery and preperation.
It leverages functions from TADA Modules 1 and 2. For more details on
each function, see the [TADA Function
Reference](https://usepa.github.io/EPATADA/reference/index.html) or the
more comprehensive module vignettes:

-   [TADA Module 1 Beginner
    Training](https://usepa.github.io/EPATADA/articles/TADAModule1_BeginnerTraining.html)

-   [TADA Module 1 Advanced
    Training](https://usepa.github.io/EPATADA/articles/TADAModule1_AdvancedTraining.html)

-   [TADA Module 1 Water Quality Portal Data Discovery and
    Cleaning](https://usepa.github.io/EPATADA/articles/TADAModule1.html)

-   [TADA Module
    2](https://usepa.github.io/EPATADA/articles/TADAModule2.html)

## Load Required Packages

The fist step is to load the TADA package and other packages used in our
workflow. (If you are a first-time TADA user, see [TADA Module 1
Beginner
Training](https://usepa.github.io/EPATADA/articles/TADAModule1_BeginnerTraining.html)
for TADA package installation instructions.)

We'll also record the start time for our analysis, beginning after
loading packages.

```{r library, results = 'hide'}
# Load required packages
library(EPATADA)
library(dplyr)
library(yaml)
library(lwgeom)
library(sf)
library(lubridate)
library(knitr)
library(DT)


# Record start time
start.time <- Sys.time()
```

## How's My Waterway

To help define the area of interest, we can start by taking a look at
How's My Waterway by searching for Farmington, New Mexico:
[HMW_Farmington,
NM](https://mywaterway.epa.gov/community/farmington,%20new%20mexico/overview).
We can identify the Assessment Unit Identifiers for the segments of the
San Juan and Animas Rivers closest to their confluence. These are:

**1) San Juan River (Navajo bnd at Hogback to Animas River) -
NM-2401_10**

**2) Animas River (San Juan River to Estes Arroyo) - NM-2403.A_00**

We can also identify the HUC-12s for each segment. This information will
help us build our Water Quality Portal query to find results for the
segements of interest plus other nearby waterbodies. The HUC-12s are
140801050505 (San Juan River) and 140801041006 (Animas River).

## Data Retrieval - Water Chemistry

The water chemistry data we will use are downloaded from the Water
Quality Portal (<https://www.waterqualitydata.us/>). The WQP integrates
publicly available water quality data from the USGS National Water
Information System (**NWIS**) and the EPA Water Quality Exchange
(**WQX**) Data Warehouse. The EPA water quality data originate from the
[Water Quality
Exchange](https://www.epa.gov/waterdata/water-quality-data), the EPA's
repository of water quality monitoring data collected by water resource
management groups across the country. Organizations, including states,
tribes, watershed groups, other federal agencies, volunteer groups, and
universities, submit data to the WQX.

One of the fields that can be used to query the WQP is HUC. As we know
the HUC-12s of interest, we will use them to retrieve WQP data using
`TADA_DataRetrieval`. As we are most interested in total dissovled
solids and conductivity and these are in the "Physical" characteristic
type group in the WQP, we will also include characteristic group type in
our query. We will not set start and end dates as we are interested in
all available data for these locations and characteristics. .

```{r TADA_DataRetrieval}
# Import data from WQP

data <- TADA_DataRetrieval(
  statecode = "NM",
  huc = c("140801050505", "140801041006"),
  characteristicType = "Physical",
  applyautoclean = TRUE
)
```

## AutoClean

As part of `TADA_DataRetrieval`, `TADA_AutoClean` is performed
automatically unless specified otherwise by the user (by setting
applyautoclean = FALSE). When `TADA_AutoClean` is run, the following
functions are performed on the data retrieved from the WQP:

-   `TADA_ConvertSpecialChars` - converts result value columns to
    numeric and flags non-numeric values that could not be converted.

-   `TADA_ConvertResultUnits` - unifies result units for easier quality
    control and review

-   `TADA_ConvertDepthUnits` - converts depth units to a consistent unit
    (meters).

-   `TADA_IDCensoredData` - categorizes detection limit data and
    identifies mismatches in result detection condition and result
    detection limit type.

-   Other helpful actions - converts important text columns to all
    upper-case letters, and uses WQX format rules to harmonize specific
    NWIS metadata conventions (e.g. move characteristic speciation from
    the TADA.ResultMeasure.MeasureUnitCode column to the
    TADA.MethodSpeciationName column)

As a general rule, TADA functions do not change any contents in the
WQP-served columns. Instead, they add new columns with the prefix
"TADA." This allows users to easily review any changes TADA functions
have made.

## Review and Filter Results By Comparable Data Identifier

Now we can review the results and decide which ones we would like to
retain for further QC and analysis. First let's take a look at all
unique values in the TADA.ComparableDataIdentifier column and see how
how many results are associated with each. TADA.ComparableDataIdentifier
concatenates TADA.CharacteristicName, TADA.ResultSampleFractionText,
TADA.MethodSpeciationName, and TADA.ResultMeasure.MeasureUnitCode.

```{r TADA_FieldValuesTable}

# use TADA_FieldValuesTable to create a table of the number of results per TADA.ComparableDataIdentifier
compid_table <- TADA_FieldValuesTable(data, field = "TADA.ComparableDataIdentifier")

DT::datatable(compid_table, fillContainer = TRUE)
```

As we scroll through the table, we can see that there are some repeated
characteristic names. It may be possible that some of these are synonyms
which can be harmonized using `TADA_HarmonizeSynonyms` so their results
can be directly compared. However, for purposes of this demo we will not
explore harmonization and will instead focus on the two
TADA.ComparableDataIdentifiers with the greatest number of results,
"TOTAL DISSOLVED SOLIDS_DISSOLVED_NA_UG/L" and "SPECIFIC
CONDUCTANCE_TOTAL_NA_US/CM @25C". For information on harmonization, see
the [EPATADA Function
Reference](https://usepa.github.io/EPATADA/reference/index.html).

Now, we can filter the data set to retain only the
TADA.ComparableDataIdentifiers of interest.

```{r}
data <- data %>%
  dplyr::filter(TADA.ComparableDataIdentifier %in% c("TOTAL DISSOLVED SOLIDS_DISSOLVED_NA_UG/L", "SPECIFIC CONDUCTANCE_TOTAL_NA_US/CM @25C"))

```

## Water Chemistry Data Preparation and QC

Before performing any analysis, there are some additional data
preparation and QC steps that can be performed using TADA functions.

First, we can check the data for additional conditions which might cause
us to exclude certain results from further analysis. For example, we can
use `TADA_FlagMethod` to check for invalid analytical
method-characteristic combinations; `TADA_FlagSpeciation` to check for
invalid characteristics-method speciation combinations,
`TADA_FlagResultUnit` to check the the validity of each
characteristic-media-result unit combination, and `TADA_FlagFraction` to
check the validity of characteristic-fraction combinations. In this
simple example, we will remove all invalid combinations identified by
each of these functions. When setting up your own workflow, you will
want to review each function and its output more carefully before moving
on.

```{r remove invalid combinations} # Flag and remove results}

data <- data %>%   TADA_FlagMethod(clean = TRUE) %>%   
  TADA_FlagSpeciation(clean = "both") %>%
  TADA_FlagResultUnit(clean = "both") %>% 
  TADA_FlagFraction(clean = TRUE)
```

Another set of TADA flagging functions, `TADA_FlagAboveThreshold` and
`TADA_FlagBelowThreshold`, can be used to check results against national
lower and upper thresholds. As in the previous examples, we will set
these functions to remove any results that fall outside the national
thresholds.

```{r remove outside thresholds} # Flag and remove results}

data <- data %>%   
  TADA_FlagAboveThreshold(clean = TRUE) %>%
  TADA_FlagBelowThreshold(clean = TRUE)
```

We can also review the data set to see if it contains potential
duplicate results from within a single organization or potential
duplicates from within multiple organizations (such as when two or more
organizations monitor the same location and may submit duplicate
results).

We will select to keep only unique samples from
`TADA_FindPotentialDuplicatesSingleOrg` by filtering for
TADA.SingleOrgDup.Flag equals "Unique". Then we will search for
potential duplicates from multiple orgs with
`TADA_FindPotentialDuplicatesMultipleOrgs` and filter only to retain
only one value per potential duplicate by filtering for
TADA.ResultSelectedMultipleOrgs equals "Y".

If you would like to prioritize results from one organization over
another, this can be done using the org_hierarchy argument in
`TADA_FindPotentialDuplicatesMultipleOrgs`. Information on how to this
can be found in the function reference for
[TADA_FindPotentialDuplicatesMultipleOrgs](https://usepa.github.io/EPATADA/reference/TADA_FindPotentialDuplicatesMultipleOrgs.html).
In this example, we will use the default setting "none" where a single
representative is selected randomly from each duplicate group.

```{r remove potential dups} # Flag and remove results}

data <- data %>%   
  TADA_FindPotentialDuplicatesSingleOrg() %>%
  dplyr::filter(TADA.SingleOrgDup.Flag == "Unique") %>% 
  TADA_FindPotentialDuplicatesMultipleOrgs(dist_buffer = 100,   
                                           org_hierarchy = "none") %>% dplyr::filter(TADA.ResultSelectedMultipleOrgs == "Y")
```

Next we can remove QC and qualified samples (with qualifiers identified
as suspect) using `TADA_FindQCActivities` and
`TADA_FlagMeasureQualifierCode`, respectively. See the TADA Module 1
vignettes for more information on how these functions identify and flag
results for removal as you may wish to make different decisions than the
TADA defaults regarding which results to retain.

```{r remove QC and qualified} # Flag and remove results}
data <- data %>% 
  TADA_FindQCActivities(clean = TRUE) %>%   
  TADA_FlagMeasureQualifierCode(clean = TRUE)
```

## Censored Data

We can use a TADA function to identify any censored data in the data
set.

```{r TADA_IDCensoredData}

data <- TADA_IDCensoredData(data)
```

Because there are no censored data in the data set, we do not need to
take any further action. If censored data do exist,
`TADA_SimpleCensoredMethods` could be used. See the TADA Module 1
vignettes or for more information.

## Data Retrieval - Geospatial

All geospatial data used in this example are downloaded from the
Assessment Total Maximum Daily Load (TMDL) Tracking and Implementation
System (**ATTAINS**) (<https://www.epa.gov/waterdata/attains>).
**ATTAINS** is an online system for accessing information about
conditions in the Nation's surface waters.

This information reported to EPA by states is available in **ATTAINS**.
The public information is made available via **ATTAINS** web services.
New TADA functions leverage the **ATTAINS** geospatial services to make
geospatial information including catchment and assessment unit geometry
easily accessible to R users.

We can use the function, `TADA_GetATTAINS` to obtain geospatial data
from **ATTAINS** relevant to the Monitoring Locations included in the
**WQP** data set. See [TADA Module
2](https://usepa.github.io/EPATADA/articles/TADAModule2.html) for a much
more detailed look at the logic behind this and the other TADA
geospatial functions.

```{r TADA_GetAttains}
# Import data from ATTAINS geospatial services

ATTAINS_data <- TADA_GetATTAINS(data)
```

## View Geospatial Features

The `TADA_ViewATTAINS` function allows us to see where the monitoring
locations from the **WQP** data set are relative to **ATTAINS-**indexed
catchments and assessment units. This can be helpful when deciding which
Monitoring Locations should be retained for additional analysis. It can
also allow us to sub

```{r TADA_ViewAttains}
# View catchments and assessment units on map

ATTAINS_map <- TADA_ViewATTAINS(ATTAINS_data)

ATTAINS_map
```

## Monitoring Location Filter and Review

Now that we've associated geospatial data from **ATTAINS** with the
**WQP** data, we
can generate pie charts for each TADA.ComparableDataIdentifier to give
us some information about how many results are available per each
monitoring location group.

```{r TADA_FieldValuesPie}
# Select TADA data with ATTAINS data
data <- ATTAINS_data$TADA_with_ATTAINS %>%
  # Remove geometry to reduce size of data set
  sf::st_drop_geometry() %>%
  # Assing "Other" as name for unnamed assessment units
  dplyr::mutate(ATTAINS.assessmentunitname = 
                  ifelse(is.na(ATTAINS.assessmentunitname),
                         "Other", ATTAINS.assessmentunitname))

# Create pie chart of results for each AUID group for total dissolved solids

TADA_FieldValuesPie(data, field = "ATTAINS.assessmentunitname",
                    characteristicName = "TOTAL DISSOLVED SOLIDS")

# Create pie chart of results for each AUID group for specific conductance

TADA_FieldValuesPie(data, field = "ATTAINS.assessmentunitname",
                    characteristicName = "SPECIFIC CONDUCTANCE")
```

We can also create a table to give us a little more information about
assessment unit each monitoring location belongs to, the number of
results per each parameter and the oldest and most recent sampling date.

```{r}

# Create table of monitoring location identifiers
MonitoringLocations <- data %>%
  dplyr::select(MonitoringLocationName, MonitoringLocationIdentifier, OrganizationFormalName,
                ATTAINS.assessmentunitname, TADA.CharacteristicName, ActivityStartDate) %>%
  dplyr::group_by(MonitoringLocationIdentifier) %>%
  dplyr::mutate(n_TDS = length(TADA.CharacteristicName[TADA.CharacteristicName == "TOTAL DISSOLVED SOLIDS"]), 
                n_SpecificConductance = length(TADA.CharacteristicName[TADA.CharacteristicName == "SPECIFIC CONDUCTANCE"]),
                StartDate = min(ActivityStartDate),
                EndDate = max(ActivityStartDate)) %>%
  dplyr::select(MonitoringLocationIdentifier, 
                ATTAINS.assessmentunitname, 
                n_TDS,
                n_SpecificConductance,
                StartDate,
                EndDate) %>%
  dplyr::distinct()

DT::datatable(MonitoringLocations, fillContainer = TRUE)

```

Now, let's create subsets of the data by assessment unit name. These
will be useful to create some basic data visualizations.

```{r}
data_animas <- data %>%
  dplyr::filter(ATTAINS.assessmentunitname == "Animas River (San Juan River to Estes Arroyo)")

data_sanjuan <- data %>%
  dplyr::filter(ATTAINS.assessmentunitname == "San Juan River (Navajo bnd at Hogback to Animas River)")

data_other <- data %>%
  dplyr::filter(ATTAINS.assessmentunitname == "Other")
```

## Data Visualization

First, let's create a two characteristic scatterplot of total dissolved
solids and specific conductance from the entire data set to begin
visualizing the data, using `TADA_TwoCharacteristicScatterplot`.

```{r two param scatterplot}
# choose two and generate scatterplot
TADA_TwoCharacteristicScatterplot(data, id_cols = "TADA.ComparableDataIdentifier", groups = c("SPECIFIC CONDUCTANCE_TOTAL_NA_US/CM @25C", "TOTAL DISSOLVED SOLIDS_DISSOLVED_NA_UG/L"))
```

While there are some older records (including a specific conductance
result from 1900), we may decide we are interested only in more recent
data. We could filter the data set to exclude any results collected
prior to 2015 and recreate the scatterplot.

```{r two param scatterplot filtered date}
# choose two and generate scatterplot
TADA_TwoCharacteristicScatterplot(data %>% 
                                    dplyr::filter
                                  (ActivityStartDate > "2014-12-31"), 
                                  id_cols = "TADA.ComparableDataIdentifier", 
                                  groups = c("SPECIFIC CONDUCTANCE_TOTAL_NA_US/CM @25C", 
                                             "TOTAL DISSOLVED SOLIDS_DISSOLVED_NA_UG/L"))
```

We could also choose to explore one characteristic from one of the
assessment units more closely. For example, we could create a histograms
for both characteristics for the Animas River only, excluding samples
prior to 2014.

```{r TADA_Histogram}

TADA_Histogram(data %>% 
                 dplyr::filter(ActivityStartDate > "2014-12-31",
                               TADA.CharacteristicName == 
                                 "SPECIFIC CONDUCTANCE",
                               ATTAINS.assessmentunitname == "Animas River (San Juan River to Estes Arroyo)"),
               id_cols = "TADA.ComparableDataIdentifier")

```

Or we could create a boxplot for a particular characteristic and
assessment unit, for example, total dissolved solids for the San Juan
River.

```{r TADA_Boxplot}

TADA_Boxplot(data %>% 
                 dplyr::filter(ActivityStartDate > "2014-12-31",
                               TADA.CharacteristicName == 
                                 "SPECIFIC CONDUCTANCE",
                               ATTAINS.assessmentunitname == "San Juan River (Navajo bnd at Hogback to Animas River)"),
               id_cols = "TADA.ComparableDataIdentifier")
```

We might also want to directly compare the same characteristic overtime
between the two rivers. We can do this by revisiting
`TADA.TwoCharacteristicScatterplot.`

```{r compare locations}

# create two characteristic scatterplot using TADA_TWoCharacteristicScatterplot
twochar_scatter <- TADA_TwoCharacteristicScatterplot(data %>% 
                 dplyr::filter(ActivityStartDate > "2014-12-31",
                               TADA.ComparableDataIdentifier == 
                                "SPECIFIC CONDUCTANCE_TOTAL_NA_US/CM @25C"),
               id_cols = "ATTAINS.assessmentunitname",
               groups = c("San Juan River (Navajo bnd at Hogback to Animas River)", "Animas River (San Juan River to Estes Arroyo)")) %>%
  # remove default plot features that are not applicable for a location comparison
  plotly::layout(yaxis2 = list(overlaying = "y", side = "right", title = "", visible = FALSE),
                 title = TADA_InsertBreaks("SPECIFIC CONDUCTANCE for the San Juan and Animas Rivers Over Time"))

# modify plot titles to reflect group/location names
twochar_scatter <- plotly::plotly_build(twochar_scatter)
twochar_scatter$x$data[[2]]$name <- "San Juan River"
twochar_scatter$x$data[[3]]$name <- "Animas River"

# rebuild graph with corrected legend labels
plotly::plotly_build(twochar_scatter)
```

We can also use `TADA_Stats` to generate a summary table for the entire
data set or subsets of it for each river.

```{r stats data}
# create table with all data
data_stats <- TADA_Stats(data)

DT::datatable(data_stats, fillContainer = TRUE)
```

```{r stats animas}
# create table with animas data
animas_stats <- TADA_Stats(data %>% dplyr::filter(ATTAINS.assessmentunitname == "Animas River (San Juan River to Estes Arroyo)"))

DT::datatable(animas_stats, fillContainer = TRUE)
```

```{r stats sanjuan}
# create table with san juan data
sanjuan_stats <- TADA_Stats(data %>% dplyr::filter(ATTAINS.assessmentunitname == "San Juan River (Navajo bnd at Hogback to Animas River)"))

DT::datatable(sanjuan_stats, fillContainer = TRUE)
```

## Reproducible and Documented

This workflow is reproducible and the decisions at each step are well
documented. This means that it is easy to go back and review every step,
understand the decisions that were made, make changes as necessary, and
run it again.

For example, we could modify the code to repeat the same analysis at a
different location. Or we could run `TADA_ConvertResultUnits` an
additional time outside to convert units for one or more
characteristics.

We could change or add additional characteristics to our analysis. We
could even write additional custom functions or even use functions from
other packages as needed for more complex standards, evaluate trends, or
answer other questions using a TADA data frame.

The code chunk below displays the elapsed time it took to run the code
chunks and create this document, providing an example of how
incorporating TADA in your workflow can increase efficiency.

```{r analysis time}
end.time <- Sys.time()

end.time - start.time
```
