---
title: "TADA Module 3: Integration of Common Water Quality Standards (WQS) Criteria"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA Module 3: Integration of Common Water Quality Standards (WQS) Criteria}
  %\VignetteEngine{knitr::rmarkdown}
description: An overview of TADA Module 3 functions and a recommended workflow.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to download the TADA R package from GitHub, access and
parameterize several important functions, and create basic
visualizations with a sample data set.

**Note: TADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.**

**If you are interested in contributing to TADA development, more
information is available at:**

[**Contributing**](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)

**We welcome collaboration with external partners.**

## Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing TADA because it is only available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load TADA using the remotes package. TADA R Package
dependencies will also be downloaded automatically from CRAN with the
TADA install. You may be prompted in the console to update dependency
packages that have more recent versions available. If you see this
prompt, it is recommended to update all of them (enter 1 into the
console).

\#`{r install_TADA, eval = F, results = 'hide'} # remotes::install_github("USEPA/EPATADA", #   ref = "develop", #   dependencies = TRUE # ) #`

```{r install_TADA, eval = T, results = 'hide'}
# help with knit and pass the checks, but we want to make sure this chunk does not show up in the final PR, just for development.
remotes::install_github("USEPA/EPATADA",
  ref = "Create-Module-3-Reference-Tables",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the TADA R Package into
your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Help pages

All TADA R package functions have their own individual help pages,
listed on the [Function
reference](https://usepa.github.io/EPATADA/reference/index.html) page on
the GitHub site. Users can also access the help page for a given
function in R or RStudio using the following format (example below):
`?[name of TADA function]`

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

## WQS Criteria Functions in TADA

Disclaimer:

Use of Module 3 development and tools in the EPATADA package is to
assist different organization (state, territories and tribal nations) on
their assessment methods under the CWA. It does not constitute current
Agency policy or regulation requirements. This tool is not to be solely
used for decision making process. Module 3 integration of WQS for
assessments is not a regulation and does not impose legally binding
requirements on EPA or the States in reporting water conditions under
biannual reports of the CWA.

Criteria is defined as the Magnitude, duration and frequency associated
with a parameter and designated use.

-   Magnitude is defined as the limiting concentration not to be
    exceeded.

-   Duration is the period/length of time in which observations are
    aggregated.

-   Frequency is defined as the percent or number of times in which a
    magnitude can be exceeded during a time frame.

This vignette represents functions that require users to:

-   Perform a parameter crosswalk between ATTAINS parameter names and
    WQP/TADA parameter names, define the parameter and designated use of
    a water body associated with each state and participating tribe's
    WQS criteria, and define the geospatial components to match WQP
    monitoring sites to an Assessment Unit.

-   Define the Criteria and Methodologies used for assessments and
    determine recommended impairment decisions for each parameter and
    designated use of a water body based on these defined WQS criteria
    conditions for each AU. Users will have two options of either
    summarizing individual magnitude excursions or for the full Criteria
    (Magnitude, Duration and Frequency) associated with each parameter
    and use.

The EPA has derived its "national recommended Clean Water Act (CWA)
304(a) criteria". Using data from the Criteria Search Tool (CST), the
TADA team has developed initial efforts to incorporate these metrics as
the standards that can be used to determine magnitude excursions and
full criteria exceedance for a parameter and use.

However, the EPA304a standards do not account for more each
state-specific or site-specific conditions in which each state may
derive its own criteria. The functions in Module 3 TADA aims to be
flexible to capture these criteria conditions specific to each state or
tribal nation's definition. Users will be required to modify and review
the appropriate crosswalk tables for each step of the process. While
efforts have been made to reduce manual inputs, initial workflow may
require much additions and removals during each step.

There will be

**Additional Description:**

**...**

## Example with Data_NCTCShepherdstown_HUC12

**Introduction:**

The goal is to assist states, territory and tribal nations with
assessment and monitoring of waters against their WQS. The plan is to
focus on a high level summary of assessments. We will start with
capturing the magnitude component of criteria to determine impairment
decisions for a parameter and use. We will also incorporate the ability
to compare an org's standards with priority characteristics using EPA
304a standards.

Future functionalities will work to incorporate the Duration and
Frequency portion of criteria along with monitoring components
(comparing prior years assessments when possible) of WQS and other
functionalities such as with handling of equation based standards,
additional components such as low flow waters, comparing against other
state's standards when possible, capturing additional details of EPA
304a standards, and more.

**Creating the 3 Reference Tables**

We will work through an example of creating the three user reference
tables that will be needed in order to help users with defining the
criteria magnitude assessments for a parameter and use using the
Shepherdstown example.

Clean, harmonize and perform QAQC on the data frame for assessment.

```{r}
#### Note: Additional Mod 1 and Mod 2 items to consider prior to running Mod 3 function: 
# Consider handling non - detects
# Run other flagging functions during this step.
# Can we create a mod 1 function that fully automates these common cleaning during this step of the process for assessments? Think through a more streamlined process.
# Checking if media type is water is also important.
# TADA_AnalysisFilter() - just keep surface water info.
# Finding Nearby sites, flag coordinates above or below thresholds, flag methods...
# Split into 2 vignettes or Consider using the Mod 1 and Mod 2 vignette examples, and use that example with this mod 3 vignette. Mod 1 and Mod 2 can be ran in any order, but they might have slightly different steps. Try to consider examples that will work both ways - Shepherdstown example could be a difficult one to do so.
# what example should we use? A state, or a tribe? Need to consider that it has enough geospatial data to retrieve. Maybe we can do one example for each state and a tribe.
# ask for an org in our working group to be open to use as an example. Need to get their approval.

# Mod 2 is where the final output of ML and AU should be done.

Data_NCTC <- TADA_RunKeyFlagFunctions(Data_NCTCShepherdstown_HUC12)
Data_NCTC2 <- TADA_HarmonizeSynonyms(Data_NCTC)
```

Filter data frame by TADA.Characteristics for a few priority
characteristics. Do some exploratory data analysis to view the number of
counts of TADA.ComparableDataIdentifier. Users can reference the domain
values of parameter names listed as a cause in prior assessment cycles
with the rATTAINS library function: domain_values.

```{r}
Data_NCTC2 <- dplyr::filter(Data_NCTC2, TADA.CharacteristicName %in% c("ZINC", "PH","NITRATE"))

TADA_FieldValuesTable(Data_NCTC2, field = "TADA.ComparableDataIdentifier")

rATTAINS::domain_values(domain_name = "ParameterName")
```

To begin the ATTAINS assessment process, you will create a blank
template of ATTAINS.ParameterName which needs crosswalk to a
TADA.ComparableDataIdentifier. Users should specify the name(s) of their
orgs that will be used for assessment.

By specifying the name of the ATTAINS organization(s) used, if creating an excel
spreadsheet (using excel = TRUE), this will create a separate tab in the
excel file which contains the parameter names used in prior assessment
cycles filtered by those ATTAINS organization(s). This will allow users to
decide whether to continue to use the same ATTAINS parameter names from
prior assessments that their org(s) have used in the past, or to use
other valid parameter names found in the ATTAINS domain list. If there
are no ATTAINS parameter name that properly captures the
TADA.ComparableDataIdentifier, users should consider contacting the
ATTAINS team first before proceeding.

```{r}
NCTC_ParamRef <- TADA_CreateParamRef(
  Data_NCTC2, 
  org_names = c("Maryland"),
  excel = FALSE # comment out this line 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
  #excel = TRUE, overwrite = TRUE 
  )
```

Users who have already completed this crosswalk can re-use this
reference file in the future. In this example, we will fill out the
excel blank template, read it into the R environment, and reuse this
parameter reference data frame that we have named "ParamRef".

If a user modifies this reference table directly in the R environment
only and not in the excel file, the user will need to re-run the
function using the "ParamRef" file to reflect the flagging of parameter
names.

```{r}
# only run the code chunks below if you make edits to the excel file.
# downloads_path <- file.path(Sys.getenv("USERPROFILE"), "Downloads", "myfileRef.xlsx")
# ParamRef <- openxlsx::read.xlsx(downloads_path, sheet = "CreateParamRef")

# NCTC_ParamRef3 <- TADA_CreateParamRef(
#   Data_NCTC2, 
#   org_names = c("Maryland"),
#   paramRef = ParamRef,
#   excel = FALSE # comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
#   #excel = TRUE, overwrite = TRUE 
#   )

ParamRef <- dplyr::mutate(NCTC_ParamRef, ATTAINS.ParameterName = dplyr::case_when(
TADA.CharacteristicName == "PH" ~ "PH",
TADA.CharacteristicName == "ZINC" ~ "ZINC",
TADA.ComparableDataIdentifier == "NITRATE_FILTERED_AS N_MG/L" ~ "NITROGEN, TOTAL"
))

NCTC_ParamRef2 <- TADA_CreateParamRef(
  Data_NCTC2, 
  org_names = c("Maryland"),
  paramRef = ParamRef,
  excel = FALSE)# comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE) 
```

Users may also include states outside of their spatial region if
desired, depending on a user's interest. Users will be responsible for
filling out this crosswalk for org(s)/state(s) that are outside their own org(s).
Although, future efforts will focus on creating a repository of criteria
and methods along with crosswalks done from other TADA users.

```{r}
NCTC_ParamRef3 <- TADA_CreateParamRef(
  Data_NCTC2, 
  org_names = c("Maryland", "Florida"),
  paramRef = ParamRef, 
  excel = FALSE
  )
```

# Three cases of TADA_CreateParamUseRef().

First example will pull in all prior use_names for each ATTAINS
parameter name specific to the defined org_names argument input
"Maryland". Users will review this list and choose which use_name are
appropriate to define for assessment. The default is "Include" under the
"IncludeOrExclude" column. All use_names that will need a definition for
magnitude criteria and methodology should be labeled as "Include". If a
use_name is not appropriate, users should choose "Exclude" for that
ATTAINS parameter name and use name.

For any ATTAINS parameter name(s) that have not been used by this org in
prior assessment cycles, there will be no prior use names associated
with them. In this case, users will be responsible for
inputting additional rows as needed and choosing the appropriate
use_name applicable to their org under the column 'use_name'. In this
example, we can see "PH" was not a prior ATTAINS parameter name for
Maryland, thus additional rows should be inputted to reflect what
parameter and use combinations are needed to be defined for assessments
(disclaimer: this is for demonstration purposes only and does not
reflect Maryland's Criteria and Methodologies for PH).

```{r}
NCTC_ParamUseRef <- TADA_CreateParamUseRef(
  Data_NCTC2, 
  org_names = c("Maryland"),
  paramRef = ParamRef, 
  excel = FALSE # comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
  #excel = TRUE, overwrite = TRUE 
  )
```

User can include the EPA304a standards with the input "org_names", which
utilizes the CST.PollutantName column created in TADA_CreateParamRef().
If a user includes additional orgs besides "Maryland" and "EPA304a", it
will not generate rows for those parameter and use combinations for
those additional orgs as only Maryland was defined in
TADA_CreateParamRef().

```{r}
NCTC_ParamUseRef2 <- TADA_CreateParamUseRef(
  Data_NCTC2, 
  org_names = c("EPA304a", "Maryland", "Virginia", "Pennsylvania", "West Virginia"), 
  paramRef = ParamRef, 
  excel = FALSE # comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
  #excel = TRUE, overwrite = TRUE 
  )

# User should only run TADA_CreateParamUseRef() which contains the orgs of interest. If there are other orgs that are interested, this should have been defined in TADA_CreateParamRef() in which users will need to be aware of the appropriate parameter name crosswalk (if users share their criteria and methods in the future, this could be more feasible)
NCTC_ParamUseRef3 <- TADA_CreateParamUseRef(
  Data_NCTC2, 
  org_names = c("EPA304a", "Maryland"), 
  paramRef = NCTC_ParamRef2, 
  excel = FALSE # comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
  #excel = TRUE, overwrite = TRUE 
  )
```

```{r}
NCTC_ParamUseRef4 <- TADA_CreateParamUseRef(
  Data_NCTC2, 
  org_names = c("EPA304a", "Maryland","Florida"), 
  paramRef = NCTC_ParamRef3, 
  excel = FALSE # comment out 'excel = FALSE' and uncomment excel = TRUE, overwrite = TRUE to run the excel file
  #excel = TRUE, overwrite = TRUE 
  )
```

Users must review the matching done with TADA_GetATTAINS of monitoring
sites to AU. Users may also provide their own AU monitoring sites match
as a reference table. For any new WQP monitoring sites found in the TADA
dataframe, users will need to determine the appropriate match.

If there are site-specific criteria that only applies to some monitoring
locations within an AU, whether it's an ecoregion, for a specific
parameter and use, or if there are separate criteria for "warm-waters"
versus "cold-waters" or "trout-based" versus other species based, this
should be defined here. Users can further customize specific exclusions
that pertains to their water quality assessments. Users will be
responsible for

```{r}
#NCTC_AUIDRef <- TADA_CreateAUIDRef(Data_NCTC2, overwrite = TRUE)
```
