---
title: "TADA: Assessment Unit Use Case Demo"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA: Assessment Unit Use Case Demo}
  %\VignetteEngine{knitr::rmarkdown}
description: An overview of a TADA use case for a single Assessment Unit.
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview

We will walk through the specific use case of query, downloading,
preparing, QA/QCing data from a single Assessment Unit with multiple
Monitoring Locations sampled by several organizations. Three
characteristics, pH, water temperature, and total nitrogen are used in
this example. Two of the characteristics, pH and water temperature, will
be compared to a state water quality standard from Illinois. The third
characteristic, total nitrogen, does not have an Illinois water quality
standard, but is used to demonstrate some helpful functions for data
exploration and visualization.

This vignette is designed to be concise and provide an example of how
you might use TADA functions for basic data discovery and analysis with
regards to a single Assessment Unit and a small subset of
characteristics. It leverages functions from TADA Modules 1 and 2. For
more details on each function, see the [TADA Function
Reference](https://usepa.github.io/EPATADA/reference/index.html) or the
more comprehensive module vignettes:

-   [TADA Module 1 Beginner
    Training](https://usepa.github.io/EPATADA/articles/TADAModule1_BeginnerTraining.html)

-   [TADA Module 1 Advanced
    Training](https://usepa.github.io/EPATADA/articles/TADAModule1_AdvancedTraining.html)

-   [TADA Module 1 Water Quality Portal Data Discovery and
    Cleaning](https://usepa.github.io/EPATADA/articles/TADAModule1.html)

-   [TADA Module
    2](https://usepa.github.io/EPATADA/articles/TADAModule2.html)

## Load Required Packages

The fist step is to load the TADA package and other packages used in our
workflow. (If you are a first-time TADA user, see [TADA Module 1
Beginner
Training](https://usepa.github.io/EPATADA/articles/TADAModule1_BeginnerTraining.html)
for TADA package installation instructions.)

We'll also record the start time for our analysis, beginning after
loading packages.

```{r library, results = 'hide', message = FALSE, warning = FALSE}
# Load required packages
library(EPATADA)
library(dplyr)
library(yaml)
library(lwgeom)
library(sf)
library(lubridate)
library(knitr)
library(DT)


# Record start time
start.time <- Sys.time()
```

## Data Retrieval - Water Chemistry

The water chemistry data we will use are downloaded from the Water
Quality Portal (<https://www.waterqualitydata.us/>). The WQP integrates
publicly available water quality data from the USGS National Water
Information System (**NWIS**) and the EPA Water Quality Exchange
(**WQX**) Data Warehouse. The EPA water quality data originate from the
[Water Quality
Exchange](https://www.epa.gov/waterdata/water-quality-data), the EPA's
repository of water quality monitoring data collected by water resource
management groups across the country. Organizations, including states,
tribes, watershed groups, other federal agencies, volunteer groups, and
universities, submit data to the WQX.

We are interested in Assessment Unit "IL_I-84". This is a Mississippi
River Assessment Unit. Because we don't have a comprehensive list of the
Monitoring Locations in this Assessment Unit, we will need to create a
broad query to retrieve all results that might be associated the
IL_I-84. One way to do this is to include all HUC10s that are relevant
to the Assessment Unit.

We can use `TADA_DataRetrieval` to retrieve WQP data. In this example,
we have specified lists of HUCs and Characteristic Types to reduce the
size of the query since we have a general idea of the location of our
assessment unit of interest and know which characteristics we'd like to
explore. One quick way to find this information is to view IL_I-84's
Waterbody Report on How's My Waterway:
[WaterbodyReport/IL_EPA/IL_I-84/2022](https://mywaterway.epa.gov/waterbody-report/IL_EPA/IL_I-84/2022).

We'll also need to set start and end dates. For this example we'll use a
ten year range between 2010 and 2020. We will also set the
Characteristic Type to "Nutrient" and "Physical", because these types
contain nitrogen, pH and water temperature Characteristics. It is
possible to query by a specific Characteristic Name, but different
organizations may input the same name in a variety of ways. Therefore,
we recommend keeping your initial queries as broad as possible so all
relevant results are downloaded and useful data from synonymous
Characteristics are not missed. There are TADA functions that can be
used to harmonize these synonyms.

```{r TADA_DataRetrieval}
# Import data from WQP

data <- TADA_DataRetrieval(
  statecode = "IL",
  startDate = "2010-01-01",
  endDate = "2020-12-31",
  huc = c("0714010505", "0714010504", "0714010508", "0714010501", "0714010503"),
  applyautoclean = TRUE
)
```

## AutoClean

As part of `TADA_DataRetrieval`, `TADA_AutoClean` is performed
automatically unless specified otherwise by the user (by setting
applyautoclean = FALSE). When `TADA_AutoClean` is run, the following
functions are performed on the data retrieved from the WQP:

-   `TADA_ConvertSpecialChars` - converts result value columns to
    numeric and flags non-numeric values that could not be converted.

-   `TADA_ConvertResultUnits` - unifies result units for easier quality
    control and review

-   `TADA_ConvertDepthUnits` - converts depth units to a consistent unit
    (meters).

-   `TADA_IDCensoredData` - categorizes detection limit data and
    identifies mismatches in result detection condition and result
    detection limit type.

-   Other helpful actions - converts important text columns to all
    upper-case letters, and uses WQX format rules to harmonize specific
    NWIS metadata conventions (e.g. move characteristic speciation from
    the TADA.ResultMeasure.MeasureUnitCode column to the
    TADA.MethodSpeciationName column)

As a general rule, TADA functions do not change any contents in the
WQP-served columns. Instead, they add new columns with the prefix
"TADA." This allows users to easily review any changes TADA functions
have made.

## Data Retrieval - Geospatial

All geospatial data used in this example are downloaded from the
Assessment Total Maximum Daily Load (TMDL) Tracking and Implementation
System (**ATTAINS**) (<https://www.epa.gov/waterdata/attains>).
**ATTAINS** is an online system for accessing information about
conditions in the Nation's surface waters.

This information reported to EPA by states is available in **ATTAINS**.
The public information is made available via **ATTAINS** web services.
New TADA functions leverage the **ATTAINS** geospatial services to make
geospatial information including catchment and assessment unit geometry
easily accessible to R users.

We can use the function, `TADA_GetATTAINS` to obtain geospatial data
from **ATTAINS** relevant to the Monitoring Locations included in the
**WQP** data set. See [TADA Module
2](https://usepa.github.io/EPATADA/articles/TADAModule2.html) for a much
more detailed look at the logic behind this and the other TADA
geospatial functions.

```{r TADA_GetAttains}
# Import data from ATTAINS geospatial services

ATTAINS_data <- TADA_GetATTAINS(data)
```

## View Geospatial Features

The `TADA_ViewATTAINS` function allows us to see where the monitoring
locations from the **WQP** data set are relative to **ATTAINS-**indexed
catchments and assessment units. This can be helpful when deciding which
Monitoring Locations should be retained for additional analysis. For
this demo, we are focusing on a single Assessment Unit, IL-84, with
multiple monitoring locations (*MonitoringLocationIdentifier*).

```{r TADA_ViewAttains}
# View catchments and assessment units on map

ATTAINS_map <- TADA_ViewATTAINS(ATTAINS_data)

ATTAINS_map
```

## Monitoring Location Filter and Review

Now that we've associated geospatial data from **ATTAINS** with the
**WQP** data, we can filter the data set to retain only results that
were collected in the specified Assessment Unit. We can also generate a
new table to give us some information about the individual monitoring
locations within the assessment unit. After this filter step is
complete, we can remove the "ATTAINS." prefixed columns to reduce the
size of the data set, as we won't need them for the remaining steps in
this example.

We can also create a table with some basic information about the
Monitoring Locations in Assessment Unit IL-84 and a pie chart to display
the relative number of results contributed by each organization.

```{r Data Retrieval - Geospatial}
# Filter data for specified assessment unit

AUID_data <- ATTAINS_data$TADA_with_ATTAINS %>%
  dplyr::filter(ATTAINS.assessmentunitidentifier == "IL_I-84")

Analysis_data <- ATTAINS_data$TADA_with_ATTAINS %>%
  dplyr::filter(ATTAINS.assessmentunitidentifier == "IL_I-84") %>%
  dplyr::select(-contains("ATTAINS.")) %>%
  sf::st_drop_geometry() %>%
  TADA_RetainRequired()


# Create table of monitoring location identifiers

MonitoringLocations <- Analysis_data %>%
  dplyr::select(MonitoringLocationName, MonitoringLocationIdentifier, OrganizationFormalName) %>%
  dplyr::distinct()

DT::datatable(MonitoringLocations, fillContainer = TRUE)


# Create pie of results by organization
Orgs_pie <- TADA_FieldValuesPie(Analysis_data, field = "OrganizationFormalName")

Orgs_pie
```

```{r Hidden Calcuations, echo = FALSE}
# Determine number of organizations

norgs <- length(unique(Analysis_data$OrganizationFormalName))

# Determine number of monitoring locations

nmls <- length(unique(Analysis_data$MonitoringLocationIdentifier))

# Determine N results
nres <- length(Analysis_data$TADA.ResultMeasureValue)

# Determine earliest sample date

earlydate <- min(Analysis_data$ActivityStartDate)

# Determine latest sample date

latedate <- max(Analysis_data$ActivityStartDate)
```

We can see that there are `r nmls` monitoring locations from `r norgs`
organizations with results from `r earlydate` to `r latedate`. There are
`r nres` results in the data frame.

## Water Chemistry Data Preparation and QC

Before we can begin comparisons between the water chemistry results and
a water quality standard, there are some additional data preparation and
QC steps that can be performed using TADA functions.

First, we can check the data for additional conditions which might cause
us to exclude certain results from further analysis. For example, we can
use `TADA_FlagMethod` to check for invalid analytical
method-characteristic combinations; `TADA_FlagSpeciation` to check for
invalid characteristics-method speciation combinations,
`TADA_FlagResultUnit` to check the the validity of each
characteristic-media-result unit combination, and `TADA_FlagFraction` to
check the validity of characteristic-fraction combinations. In this
simple example, we will remove all invalid combinations identified by
each of these functions. When setting up your own workflow, you will
want to review each function and its output more carefully before moving
on.

```{r remove invalid combinations}
# Flag and remove results

Analysis_data <- Analysis_data %>%
  TADA_FlagMethod(clean = TRUE) %>%
  TADA_FlagSpeciation(clean = "both") %>%
  TADA_FlagResultUnit(clean = "both") %>%
  TADA_FlagFraction(clean = TRUE)
```

Another set of TADA flagging functions, `TADA_FlagAboveThreshold` and
`TADA_FlagBelowThreshold`, can be used to check results against national
lower and upper thresholds. As in the previous examples, we will set
these functions to remove any results that fall outside the national
thresholds.

```{r remove outside thresholds}
# Flag and remove results

Analysis_data <- Analysis_data %>%
  TADA_FlagAboveThreshold(clean = TRUE) %>%
  TADA_FlagBelowThreshold(clean = TRUE)
```

We can also review the data set to see if it contains potential
duplicate results from within a single organization or potential
duplicates from within multiple organizations (such as when two or more
organizations monitor the same location and may submit duplicate
results).

We will select to keep only unique samples from
`TADA_FindPotentialDuplicatesSingleOrg` by filtering for
TADA.SingleOrgDup.Flag equals "Unique". Then we will search for
potential duplicates from multiple orgs with
`TADA_FindPotentialDuplicatesMultipleOrgs` and filter only to retain
only one value per potential duplicate by filtering for
TADA.ResultSelectedMultipleOrgs equals "Y".

If you would like to prioritize results from one organization over
another, this can be done using the org_hierarchy argument in
`TADA_FindPotentialDuplicatesMultipleOrgs`. Information on how to this
can be found in the function reference for
[TADA_FindPotentialDuplicatesMultipleOrgs](https://usepa.github.io/EPATADA/reference/TADA_FindPotentialDuplicatesMultipleOrgs.html).
In this example, we will use the default setting "none" where a single
representative is selected randomly from each duplicate group.

```{r remove potential dups}
# Flag and remove results

Analysis_data <- Analysis_data %>%
  TADA_FindPotentialDuplicatesSingleOrg() %>%
  dplyr::filter(TADA.SingleOrgDup.Flag == "Unique") %>%
  TADA_FindPotentialDuplicatesMultipleOrgs(
    dist_buffer = 100,
    org_hierarchy = "none"
  ) %>%
  dplyr::filter(TADA.ResultSelectedMultipleOrgs == "Y")
```

Next we can remove QC and qualified samples (with qualifiers identified
as suspect) using `TADA_FindQCActivities` and
`TADA_FlagMeasureQualifierCode`, respectively. See the TADA Module 1
vignettes for more information on how these functions identify and flag
results for removal as you may wish to make different decisions than the
TADA defaults regarding which results to retain.

```{r remove QC and qualified}
# Flag and remove results

Analysis_data <- Analysis_data %>%
  TADA_FindQCActivities(clean = TRUE) %>%
  TADA_FlagMeasureQualifierCode(clean = TRUE)
```

## Censored Data

Censored data are measurements for which the true value is not known,
but we can estimate the value based on lower or upper detection
conditions and limit types. TADA fills missing *TADA.ResultMeasureValue*
and *TADA.ResultMeasure.MeasureUnitCode* values with values and units
from *TADA.DetectionQuantitationLimitMeasure.MeasureValue* and
*TADA.DetectionQuantitationLimitMeasure.MeasureUnitCode*, respectively,
using the `TADA_AutoClean` function.

The TADA package currently has functions that summarize censored data
incidence in the dataset and perform simple substitutions of censored
data values, including x times the detection limit and random selection
of a value between 0 and the detection limit. The user may specify the
methods used for non-detects and over-detects separately in the input to
the `TADA_SimpleCensoredMethods` function. The next step we take in this
example is to perform simple conversions to the censored data in the
dataset: we keep over-detects as is (no conversion made) and convert
non-detect values to 0.5 times the detection limit (half the detection
limit).

See the TADA Module 1 vignettes for more detailed information on the
logic behind `TADA_SimpleCensoredMethods`**.**

After running `TADA_SimpleCensoredMethods`, we will also filter the data
to remove any remaining NAs.

```{r SummarizeCensoredData}
Analysis_data <- TADA_SimpleCensoredMethods(Analysis_data,
  nd_method = "multiplier",
  nd_multiplier = 0.5,
  od_method = "as-is",
  od_multiplier = "null"
) %>%
  dplyr::filter(!is.na(TADA.ResultMeasureValue))
```

## Harmonize Results

The `TADA_GetSynonymRef` function generates a synonym reference table
that is specific to the input dataframe. Users can review how their
input data relates to standard TADA values for the following elements:
TADA.CharacteristicName, TADA.ResultSampleFractionText,
TADA.MethodSpeciationName, and TADA.ResultMeasure.MeasureUnitCode and
edit if desired.

Users can also edit the reference file to meet their needs if desired.
The download argument can be used to save the harmonization file to your
current working directory when `download = TRUE`, the default is
`download = FALSE`. In this example, we will modify the harmonization
file to group results with the fraction "NA" and results with the
fraction "TOTAL" for pH and water temperature.

The `TADA_HarmonizeSynonyms` function then compares the input dataframe
to the TADA Synonym Reference Table and makes conversions where target
characteristics/fractions/speciations/units are provided. This function
also appends a column called TADA.Harmonized.Flag, indicating which
results had metadata changed/converted in this function. The purpose of
this function is to make similar data consistent and therefore easier to
compare and analyze.

```{r TADA_HarmonizeSynonyms}
UniqueHarmonizationRef <- TADA_GetSynonymRef(Analysis_data)

UniqueHarmonizationRef_edit <- UniqueHarmonizationRef %>%
  dplyr::mutate(
    Target.TADA.CharacteristicName = ifelse(TADA.CharacteristicName == "TEMPERATURE, WATER", "TEMPERATURE", TADA.CharacteristicName),
    Target.TADA.ResultSampleFractionText = ifelse(TADA.CharacteristicName == "PH", NA, Target.TADA.ResultSampleFractionText),
    HarmonizationGroup = ifelse(TADA.CharacteristicName == "PH", "PH", HarmonizationGroup),
    HarmonizationGroup = ifelse(Target.TADA.CharacteristicName == "TEMPERATURE", "TEMPERATURE",
      HarmonizationGroup
    )
  )

Harmonized_data <- TADA_HarmonizeSynonyms(Analysis_data,
  ref = UniqueHarmonizationRef_edit
)
```

## Total Nitrogen and Total Phosphorus Calculations

`TADA_CalculateTotalNP` uses the [Nutrient Aggregation
logic](https://echo.epa.gov/trends/loading-tool/resources/nutrient-aggregation)
to add together specific subspecies to obtain a total. TADA adds one
more equation to the mix: total particulate nitrogen + total dissolved
nitrogen. The function uses as many subspecies as possible to calculate
a total for each given site, date, and depth group, but it will estimate
total nitrogen with whatever subspecies are present. This function
creates NEW total nutrient measurements (total nitrogen unfiltered as N
and total phosphorus unfiltered as P) and adds them to the dataframe.

Users can use the default summation worksheet (see
`TADA_GetNutrientSummationRef`) or customize it to suit their needs. The
function also requires a daily aggregation value, either minimum,
maximum, or mean. The default is 'max', which means that if multiple
measurements of the same subspecies-fraction-speciation-unit occur on
the same day at the same site and depth, the function will pick the
maximum value to use in summation calculations. In this example, we will
use the default summation worksheet to obtain total nitrogen.

```{r, TADA_CalculateTotalNP}
Harmonized_data <- TADA_CalculateTotalNP(Harmonized_data, daily_agg = "max")
```

## Parameter Level Filtering

Now we can filter to our three parameters of interest and use
`TADA_Stats` take a look at some basic statistics for each, including
location count, measurement count, min, max, and more.

```{r parameter level filtering}
# review unique identifiers
unique(Harmonized_data$TADA.ComparableDataIdentifier)

# filter for three comparable data identifiers of interest
Filtered_data <- Harmonized_data %>%
  dplyr::filter(TADA.ComparableDataIdentifier %in% c("TEMPERATURE_NA_NA_DEG C", "PH_NA_NA_NA", "TOTAL NITROGEN, MIXED FORMS_FILTERED, LAB_NA_MG/L"))

# generate stats table
Filtered_data_stats <- TADA_Stats(Filtered_data)

DT::datatable(Filtered_data_stats, fillContainer = TRUE)
```

## Standard Comparisons With Scatterplot and Pie Chart

Next, we will create a two characteristic scatterplot of pH and
temperature to begin visualizing the data, using
`TADA_TwoCharacteristicScatterplot`.

```{r two param scatterplot}
# choose two and generate scatterplot
TADA_TwoCharacteristicScatterplot(Harmonized_data, id_cols = "TADA.ComparableDataIdentifier", groups = c("TEMPERATURE_NA_NA_DEG C", "PH_NA_NA_NA"))
```

Let's focus on just one characteristic, pH. We can filter the data frame
to retain only pH samples. We can also add a column to indicate whether
each result falls within the water quality standard range for pH (6.5 -
9). From this pH-only frame we can create a table with just a few
columns for easier review, and a single characteristic scatterplot using
`TADA_Scatterplot`. In this example, horizontal lines have been added to
indicate the upper and lower ends of the water quality standard. These
are not a default option in `TADA_Scatterplot` but can be added via the
`plotly` package.

```{r ph analysis and visualization}
# comparison to standard for ph
pH_Standard <- Filtered_data %>%
  dplyr::filter(TADA.ComparableDataIdentifier == "PH_NA_NA_NA") %>%
  dplyr::mutate(MeetsStandard = ifelse(TADA.ResultMeasureValue >= 6.5 & TADA.ResultMeasureValue <= 9, "Yes", "No"))


# subset of pH data with fewer rows
pH_Table <- pH_Standard %>%
  dplyr::select(
    MonitoringLocationIdentifier, OrganizationFormalName, ActivityStartDate, TADA.ResultMeasureValue,
    MeetsStandard
  )

DT::datatable(pH_Table, fillContainer = TRUE)

# pH scatterplot
pH_Scatter <- TADA_Scatterplot(pH_Standard, id_cols = "TADA.ComparableDataIdentifier") %>%
  plotly::add_lines(
    y = 6.5,
    x = c(min(pH_Standard$ActivityStartDate), max(pH_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  ) %>%
  plotly::add_lines(
    y = 9,
    x = c(min(pH_Standard$ActivityStartDate), max(pH_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  )

pH_Scatter
```

Comparison of temperature to water quality standards is a little more
complicated as the maximum temperature varies by season and there are
multiple components of the temperature standard. For this example, we
will focus on the "shall never exceed" seasonal temperature standards.

These are:

17.7 deg C for January, February, March, and December

33.7 deg C for April, May, June, July, August, September, October, and
November.

We can use `lubridate` and `dplyr` functions to assign the appropriate
standard to each result by using `lubridate::month` to identify the
month in which each sample was collected and `dplyr::mutate` to create a
new column for the temperature standard. Then the results can be
compared to the standard.

We can use `TADA_Scatterplot` again to visualize the data. Due to the
seasonal variation in the standard and the ten-year date range in our
data set, we may want to consider other visualizations to visually
review the number of results not meeting the standard. We can use
`TADA_FieldValuesPie` on the MeetsStandard field we created to see how
many samples met or did not meet the standard.

Creating a table also allows for easy filtering of results to identify
the results that did not meet the standard and review during which years
and seasons they occurred.

```{r temp analysis and visualization}
# comparison to standard for temperature
Temp_Standard <- Filtered_data %>%
  dplyr::filter(TADA.ComparableDataIdentifier == "TEMPERATURE_NA_NA_DEG C") %>%
  dplyr::mutate(
    MonthForAnalysis = lubridate::month(ActivityStartDate),
    TempStandard = ifelse(MonthForAnalysis %in% c(1, 2, 3, 12), 17.7, 33.7),
    MeetsStandard = ifelse(TADA.ResultMeasureValue < TempStandard,
      "Yes", "No"
    )
  )

# create scatterplot for temperature
Temp_Scatter <- TADA_Scatterplot(Temp_Standard, id_cols = "TADA.ComparableDataIdentifier")

Temp_Scatter

# create pie chart for temperature

Temp_Pie <- TADA_FieldValuesPie(Temp_Standard, field = "MeetsStandard")

Temp_Pie

# create subset table for temperature
Temp_Table <- Temp_Standard %>%
  dplyr::select(
    MonitoringLocationIdentifier, OrganizationFormalName, ActivityStartDate, TADA.ResultMeasureValue, TempStandard,
    MeetsStandard
  )

DT::datatable(Temp_Table, fillContainer = TRUE)
```

## Additional Data Exploration

We may also want to use TADA functions for exploration and visualization
of characteristics without water quality standards. For this example, we
will use "TOTAL NITROGEN, MIXED FORMS_FILTERED, LAB_NA_MG/L".

To better understand the distribution of results, we can use the TADA
functions `TADA_Histogram` and `TADA_Boxplot`.

`TADA_Histogram` can be useful for identifying the overall shape of the
data.

```{r boxplot and histogram}
# filter dataframe to comparable data identifier of interest

Nitrogen_data <- dplyr::filter(Filtered_data, TADA.ComparableDataIdentifier == "TOTAL NITROGEN, MIXED FORMS_FILTERED, LAB_NA_MG/L")

# generate a histogram
Nitrogen_Histogram <- TADA_Histogram(Nitrogen_data, id_cols = "TADA.ComparableDataIdentifier")

# view histogram
Nitrogen_Histogram
```

`TADA_Boxplot` can be useful for identifying skewness and percentiles.

```{r boxplot, fig.width=8, fig.height=6, fig.fullwidth=TRUE}
Nitrogen_Boxplot <- TADA_Boxplot(Nitrogen_data, id_cols = "TADA.ComparableDataIdentifier")


Nitrogen_Boxplot
```

Filtering the results of `TADA_Stats` for only "TOTAL NITROGEN, MIXED
FORMS_TOTAL RECOVERABLE_NA_MG/L" and selecting a smaller subset of
columns or creating a single characteristic scatterplot may also provide
useful information.

```{r stats and scatterplot}
# create table with nitrogen stats
Nitrogen_stats <- Filtered_data_stats %>%
  dplyr::filter(TADA.ComparableDataIdentifier == "TOTAL NITROGEN, MIXED FORMS_FILTERED, LAB_NA_MG/L") %>%
  dplyr::select(Location_Count, Measurement_Count, Min, Max, Mean)

DT::datatable(Nitrogen_stats, fillContainer = TRUE)

# create nitrogen scatterplot
Nitrogen_Scatterplot <- TADA_Scatterplot(Nitrogen_data, id_cols = "TADA.ComparableDataIdentifier")

Nitrogen_Scatterplot
```

## Reproducible and Documented

Major benefits to this type are workflow are that it is reproducible and
the decisions at each step are well documented. This means that it is
easy to go back and review every step, understand the decisions that
were made, make changes as necessary, and run it again. If someone asks
"How were data filtered prior to analysis?" or "How did you identify
additional Monitoring Locations within the Assessment Unit?", you can
refer to the code and function documentation to provide answers.

For example, we could change the Assessment Unit of interest, modify the
relevant code chunks (potentially including the WQP query if the new
Assessment Unit is in a different HUC or state) and repeat the same
analysis for the same characteristics at a different location. We could
also modify the harmonization reference to group additional comparable
data (if appropriate) or add a step to run `TADA_ConvertResultUnits` an
additional time outside of `TADA_AutoClean` if we wanted to convert
temperature results to deg F for analysis.

Or we could change or add additional characteristics to our analysis. We
could even write additional custom functions or even use functions from
other packages as needed for more complex standards, evaluate trends, or
answer other questions using a TADA data frame.

The code chunk below displays the elapsed time it took to run the code
chunks and create this document, providing an example of how
incorporating TADA in your workflow can increase efficiency.

```{r analysis time}
end.time <- Sys.time()

end.time - start.time
```
