---
title: "TADA Cybertown Workshop June 2025"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA Cybertown Workshop June 2025}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
description: An introduction to using the EPATADA R package to retrieve, clean, analyze, and visualize WQP data.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = F
)
```

```{css, include = F}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

## Accessing vignette

A [vignette](https://r-pkgs.org/vignettes.html) is a long-form guide to
a package often written as a R Markdown document, such as this one. It
provides detailed explanations of functions and showcases an example
workflow. This vignette can be created as an html document or other
format using the knit option on the top of the RStudio toolbar. Users
can also access this vignette on the EPATADA GitHub page found
([here](https://usepa.github.io/EPATADA/articles/TADACybertown2025.html)).

## Load libraries

First, please install the EPATADA R package by following the
instructions [here](https://github.com/USEPA/EPATADA). Next, load the
EPATADA R Package.

```{r load EPATADA}
library(EPATADA)
```

## Record start time

```{r}
start.time <- Sys.time()
```

## Retrieve

Query the WQP using TADA_DataRetrieval. TADA_AutoClean is a powerful
function that runs as part of TADA_DataRetrieval when applyautoclean =
TRUE. It performs a variety of tasks, for example:

1.  creating new "TADA" prefixed columns and and capitalizing their
    contents to reduce case sensitivity issues,

2.  converts special characters in value columns,

3.  converts latitude and longitude values to numeric,

4.  replaces "meters" with "m",

5.  replaces deprecated characteristic names with current WQX names,

6.  harmonizes result and detection limit units,

7.  converts depths to meters, and

8.  creates the column TADA.ComparableDataIdentifier by concatenating
    characteristic name, result sample fraction, method speciation, and
    result measure unit.

In this example, we will first leverage EPA's How's My Waterway (HMW)
application to discover an ATTAINS Assessment Unit of interest ([example
waterbody
report](https://mywaterway.epa.gov/waterbody-report/CT_DEP01/CT6400-00-1-L5_01/2022)).
Then, we will use the Assessment Unit ID to query ATTAINS geospatial
services for the associated shapefile (polygon area of the Assessment
Unit). Now we can use this shapefile (only works for polygons for now)
as our input for the new aoi_sf query option included in
TADA_DataRetrieval. This allows us to download WQP data within the
Assessment Unit (our area of interest/AOI).

```{r TADA_DataRetrieval}
query.params <- list(
  where = "assessmentunitidentifier IN ('CT6400-00-1-L5_01')",
  outFields = "*",
  f = "geojson"
)

url <- "https://gispub.epa.gov/arcgis/rest/services/OW/ATTAINS_Assessment/MapServer/2/query?"

poly.response <- httr::GET(url, query = query.params)

poly.geojson <- httr::content(poly.response, as = "text", encoding = "UTF-8")

poly.sf <- sf::st_read(poly.geojson, quiet = TRUE)

WQP_raw <- TADA_DataRetrieval(
  startDate = "null",
  endDate = "null",
  aoi_sf = poly.sf,
  countrycode = "null",
  countycode = "null",
  huc = "null",
  siteid = "null",
  siteType = "null",
  tribal_area_type = "null",
  tribe_name_parcel = "null",
  characteristicName = "null",
  characteristicType = "null",
  sampleMedia = "null",
  statecode = "null",
  organization = "null",
  project = "null",
  providers = "null",
  bBox = "null",
  maxrecs = 350000,
  ask = FALSE,
  applyautoclean = TRUE
)
```

Remove intermediate variables in R by using 'rm()'.

```{r}
rm(poly.response, poly.sf, query.params, poly.geojson, url)
```

## Flag, clean, and visualize

Now, let's use EPATADA functions to review, visualize, and whittle the
returned WQP data down to include only results that are applicable to
our water quality analysis and area of interest.

TADA is primarily designed to accommodate water data from the WQP. Let’s
see what activity media types are represented in the data set. Are there
any media type that are not water in our data frame?

```{r Review and Filter By Media Type}
# Create table with count for each ActivityMediaName
TADA_FieldValuesTable(
  WQP_raw,
  field = "ActivityMediaName",
  characteristicName = "null"
)
```

Create an overview map.

```{r TADA_OverviewMap}
TADA_OverviewMap(WQP_raw)
```

Let's take a quick look at all unique values in the
MonitoringLocationIdentifier column and see how how many results are
associated with each.

```{r TADA_FieldValuesTable}
# use TADA_FieldValuesTable to create a table of the number of results per MonitoringLocationIdentifier
sites <- TADA_FieldValuesTable(
  WQP_raw,
  field = "MonitoringLocationIdentifier",
  characteristicName = "null"
)

DT::datatable(sites, fillContainer = TRUE)
```

Are there sites located within 100 meters of each other?

```{r TADA_FlagCoordinates}
WQP_clean <- TADA_FindNearbySites(
  WQP_raw,
  dist_buffer = 100,
  nhd_res = "Hi",
  org_hierarchy = "none",
  meta_select = "random",
  by_AU = FALSE,
  catchment = FALSE
)

TADA_NearbySitesMap(WQP_clean)
```

Now let's review all unique values in the TADA.ComparableDataIdentifier
column and see how how many results are associated with each.
TADA.ComparableDataIdentifier concatenates TADA.CharacteristicName,
TADA.ResultSampleFractionText, TADA.MethodSpeciationName, and
TADA.ResultMeasure.MeasureUnitCode.

```{r TADA_FieldValuesTable2}
# use TADA_FieldValuesTable to create a table of the number of results per TADA.ComparableDataIdentifier
chars <- TADA_FieldValuesTable(
  WQP_clean,
  field = "TADA.ComparableDataIdentifier",
  characteristicName = "null"
)

DT::datatable(chars, fillContainer = TRUE)
```

Remove intermediate variables in R by using 'rm()'.

```{r}
rm(chars, sites, WQP_flag_review, WQP_flag)
```

Next, let's check if the dataset contains potential duplicate results
from within a single organization or from within multiple organizations
(such as when two or more organizations monitor the same location and
may submit duplicate results).

If you would like to prioritize results from one organization over
another, this can be done using the org_hierarchy argument in
`TADA_FindPotentialDuplicatesMultipleOrgs`.

```{r duplicates}
# find duplicates from single org
WQP_flag <- TADA_FindPotentialDuplicatesSingleOrg(WQP_clean)

# Review organizations. You can select one to prioritize in TADA_FindPotentialDuplicatesMultipleOrgs
unique(WQP_flag$OrganizationIdentifier)
unique(WQP_flag$OrganizationFormalName)

# find duplicates across multiple orgs
WQP_flag <- TADA_FindPotentialDuplicatesMultipleOrgs(
  WQP_flag,
  org_hierarchy = c("CT_DEP01", "USGS-CT", "CTVOLMON", "NALMS")
)
```

Let's review the duplicates:

```{r review}
WQP_flag_review <- WQP_flag |>
  dplyr::select(
    MonitoringLocationName,
    TADA.MonitoringLocationIdentifier,
    # TADA.MultipleOrgDuplicate,
    # TADA.MultipleOrgDupGroupID,
    # TADA.ResultSelectedMultipleOrgs,
    TADA.SingleOrgDupGroupID,
    TADA.SingleOrgDup.Flag,
    TADA.ComparableDataIdentifier,
    ResultMeasureValue,
    TADA.ResultMeasure.MeasureUnitCode,
    TADA.MonitoringLocationName,
    TADA.NearbySites.Flag,
    TADA.NearbySiteGroup,
    OrganizationIdentifier
  ) |>
  dplyr::filter(TADA.SingleOrgDupGroupID != "Not a duplicate") |>
  dplyr::distinct()
```

We will select to keep only unique samples from
`TADA_FindPotentialDuplicatesSingleOrg` by filtering for
TADA.SingleOrgDup.Flag equals "Unique".

There are no multiple org duplicates from
`TADA_FindPotentialDuplicatesMultipleOrgs` in this example, but if there
were, duplicates can by removed by filtering for
TADA.ResultSelectedMultipleOrgs equals "Y".

```{r filter}
WQP_clean <- WQP_flag |>
  dplyr::filter(TADA.SingleOrgDup.Flag == "Unique") |>
  dplyr::filter(TADA.ResultSelectedMultipleOrgs == "Y")
```

Remove intermediate variables in R by using 'rm()'. In the remainder of
this workshop, we will work with the clean data set.

```{r}
rm(WQP_flag, WQP_flag_review)
```

Censored data are measurements for which the true value is not known,
but we can estimate the value based on known lower or upper detection
conditions and limit types. TADA fills missing *TADA.ResultMeasureValue*
and *TADA.ResultMeasure.MeasureUnitCode* values with values and units
from *TADA.DetectionQuantitationLimitMeasure.MeasureValue* and
*TADA.DetectionQuantitationLimitMeasure.MeasureUnitCode*, respectively,
using the `TADA_AutoClean` function.

The TADA package currently has functions that summarize censored data
incidence in the dataset and perform simple substitutions of censored
data values, including x times the detection limit and random selection
of a value between 0 and the detection limit. The user may specify the
methods used for non-detects and over-detects separately in the input to
the `TADA_SimpleCensoredMethods` function. The next step we take in this
example is to perform simple conversions to the censored data in the
dataset: we keep over-detects as is (no conversion made) and convert
non-detect values to 0.5 times the detection limit (half the detection
limit).

```{r censored}
WQP_clean <- TADA_SimpleCensoredMethods(
  WQP_clean,
  nd_method = "multiplier",
  nd_multiplier = 0.5,
  od_method = "as-is",
  od_multiplier = "null"
)
```

`TADA_FindQCActivities` removes results with QA/QC ActivityTypeCode's.

```{r QCdata}
WQP_clean <- TADA_FindQCActivities(WQP_clean, clean = TRUE)
```

`TADA_ConvertSpecialChars` removes rows where the result value is not
numeric to prepare a dataframe for quantitative analyses. Specifically,
this function removes rows with "Text" and "NA - Not Available" in the
TADA.ResultMeasureValueDataTypes.Flag column, or NA in the
TADA.ResultMeasureValue column. In addition, this function removes
results with QA/QC ActivityTypeCode's. This function also removes any
columns not required for TADA workflow where all values are equal to NA.

```{r SpecialChars}
WQP_clean <- TADA_ConvertSpecialChars(WQP_clean, col = "TADA.ResultMeasureValue", clean = TRUE)
```

TADA_RunKeyFlagFunctions is a shortcut function to run important TADA
flagging functions. See ?function documentation for TADA_FlagResultUnit,
TADA_FlagFraction, TADA_FlagMeasureQualifierCode, and
TADA_FlagSpeciation for more information.

```{r TADA_RunKeyFlagFunctions}
WQP_clean <- TADA_RunKeyFlagFunctions(
  WQP_clean,
  clean = TRUE
)
```

Another set of TADA flagging functions, `TADA_FlagAboveThreshold` and
`TADA_FlagBelowThreshold`, can be used to check results against national
lower and upper thresholds. For these, we will set clean = FALSE and
flaggedonly = TRUE so that it returns only flagged results in the review
dataframe returned. We will keep these in our "clean" dataframe for now.

```{r thresholds}
WQP_flag_reviewabove <- TADA_FlagAboveThreshold(WQP_clean, clean = FALSE, flaggedonly = TRUE)

WQP_flag_reviewbelow <- TADA_FlagBelowThreshold(WQP_clean, clean = FALSE, flaggedonly = TRUE)
```

Remove intermediate variables.

```{r}
rm(WQP_flag_reviewabove, WQP_flag_reviewbelow)
```

Let's take another look at all unique values in the
TADA.ComparableDataIdentifier column and see how how many results are
associated with each. TADA.ComparableDataIdentifier concatenates
TADA.CharacteristicName, TADA.ResultSampleFractionText,
TADA.MethodSpeciationName, and TADA.ResultMeasure.MeasureUnitCode.

```{r TADA_FieldValuesTable3}
# use TADA_FieldValuesTable to create a table of the number of results per TADA.ComparableDataIdentifier
chars <- TADA_FieldValuesTable(
  WQP_clean,
  field = "TADA.ComparableDataIdentifier",
  characteristicName = "null"
)

chars_before <- unique(WQP_clean$TADA.ComparableDataIdentifier)

DT::datatable(chars, fillContainer = TRUE)
```

Scroll through the table and check to see if there any synonyms. It may
be possible that some of these can be automatically harmonized using
`TADA_HarmonizeSynonyms` so their results can be directly compared.

Let's give it a try.

```{r}
WQP_clean <- TADA_HarmonizeSynonyms(WQP_clean)
```

How many unique TADA.ComparableDataIdentifier's do we have now? In this
example, there were no synonyms.

```{r}
chars_after <- unique(WQP_clean$TADA.ComparableDataIdentifier)
```

Remove intermediate variables.

```{r}
rm(chars_before, chars_after)
```

Create a pie chart.

```{r}
TADA_FieldValuesPie(
  WQP_clean,
  field = "TADA.CharacteristicName",
  characteristicName = "null"
)
```

## Select characteristic

Let's filter the data and focus on a one characteristic of interest.

```{r}
# Select characteristics of interest
WQP_clean_subset <- WQP_clean |>
  dplyr::filter(TADA.CharacteristicName %in% "ESCHERICHIA COLI")
```

Remove intermediate variables. We will focus on the subset from now on.

```{r}
rm(WQP_clean, chars)
```

## Integrate ATTAINS and map

In this section, we will associate geospatial data from **ATTAINS** with
the **WQP** data. Our initial WQP data pull was done using a shapefile
for Assessment Unit CT6400-00-1-L5_01. TADA functions can pull in
additional ATTAINS meta data for this assessment unit. We can also
generate a new table to give us some information about the individual
monitoring locations within this assessment unit.

-   TADA_CreateATTAINSAUMLCrosswalk() automates matching of WQP
    monitoring locations with ATTAINS assessment units that fall within
    (intersect) the same NHDPlus catchment
    ([details](https://usepa.github.io/EPATADA/articles/TADAModule2.html))
-   The function uses high resolution NHDPlus catchments by default
    because 80% of state submitted assessment units in ATTAINS were
    developed based on high res NHD; users can select med-res if
    applicable to their use case

```{r Data Retrieval - Geospatial}
WQP_clean_subset_spatial <- TADA_CreateATTAINSAUMLCrosswalk(
  WQP_clean_subset,
  return_nearest = TRUE,
  return_sf = TRUE
)

# Adds ATTAINS info to df
WQP_clean_subset <- WQP_clean_subset_spatial$TADA_with_ATTAINS
```

View catchments and assessment units on map

```{r TADA_ViewATTAINS}
ATTAINS_map <- TADA_ViewATTAINS(WQP_clean_subset_spatial,
  ref_icons = FALSE
)

ATTAINS_map
```

Remove intermediate variables:

```{r}
rm(ATTAINS_map)
```

Create table of monitoring location identifiers and AUs.

```{r}
ML_AU_crosswalk <- WQP_clean_subset |>
  dplyr::select(TADA.MonitoringLocationIdentifier, ATTAINS.AssessmentUnitIdentifier, ATTAINS.AssessmentUnitName, TADA.CharacteristicName) |>
  dplyr::distinct()
```

Remove intermediate variables. Let's keep going with WQP_clean_subset.

```{r}
rm(ML_AU_crosswalk, WQP_clean_subset_spatial)
```

`TADA_RetainRequired` removes all duplicate columns where TADA has
created a new column with a TADA prefix. It retains all TADA prefixed
columns as well as other original fields that are either required by
other TADA functions or are commonly used filters.

```{r}
WQP_clean_subset <- TADA_RetainRequired(WQP_clean_subset)
```

## Exploratory analysis

Review unique TADA.ComparableDataIdentifier's

```{r}
unique(WQP_clean_subset$TADA.ComparableDataIdentifier)
```

Let's check if any results are above the EPA 304A recommended maximum
criteria magnitude (see: [2012 Recreational Water Quality Criteria Fact
Sheet](https://www.epa.gov/sites/default/files/2015-10/documents/rec-factsheet-2012.pdf)).

[![EPA 2012 recreational water quality criteria (RWQC) recommendations
for protecting human health in all coastal and non-coastal waters
designated for primary contact recreation use. EPA provides two sets of
recommended criteria. The RWQC consist of three components: magnitude,
duration and frequency. The magnitude of the bacterial indicators are
described by both a geometric mean (GM) and a statistical threshold
value (STV) for the bacteria samples. The waterbody GM should not be
greater than the selected GM magnitude in any 30-day interval. The STV
approximates the 90th percentile of the water quality distribution and
is intended to be a value that should not be exceeded by more than 10
percent of the samples in the same 30-day interval. The table summarizes
the magnitude component of the
recommendations.](images/bacteria.png)](chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://www.epa.gov/sites/default/files/2015-10/documents/rec-factsheet-2012.pdf)

If interested, you can find other state, tribal, and EPA 304A criteria
in [EPA's Criteria Search
Tool](https://www.epa.gov/wqs-tech/state-specific-water-quality-standards-effective-under-clean-water-act-cwa).

Let's check if any individual results exceed 320 CFU/100mL (the
magnitude component of the EPA recommendation 2 criteria for ESCHERICHIA
COLI).

```{r}
# add column with comparison to criteria mag (excursions)
WQP_clean_subset <- WQP_clean_subset |>
  sf::st_drop_geometry() |>
  dplyr::mutate(meets_criteria_mag = ifelse(TADA.ResultMeasureValue <= 320, "Yes", "No"))

# review
WQP_clean_subset_review <- WQP_clean_subset |>
  dplyr::select(
    MonitoringLocationIdentifier, OrganizationFormalName, ActivityStartDate, TADA.ResultMeasureValue,
    meets_criteria_mag
  )

DT::datatable(WQP_clean_subset_review, fillContainer = TRUE)
```

Generate stats table. Review percentiles. Less than 5% of results fall
above 10 CFU/100mL, and over 98% of results fall below 265.2 CFU/100m.

```{r stats}
WQP_clean_subset_stats <- WQP_clean_subset |>
  sf::st_drop_geometry() |>
  TADA_Stats()
```

Generate a scatterplot. One result value is above the threshold.

```{r}
TADA_Scatterplot(WQP_clean_subset, id_cols = "TADA.ComparableDataIdentifier") |>
  plotly::add_lines(
    y = 320,
    x = c(min(WQP_clean_subset$ActivityStartDate), max(WQP_clean_subset$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  )
```

Generate a histogram.

```{r boxplot and histogram}
TADA_Histogram(WQP_clean_subset, id_cols = "TADA.ComparableDataIdentifier")
```

`TADA_Boxplot` can be useful for identifying skewness and percentiles.

```{r boxplot, fig.width=8, fig.height=6, fig.fullwidth=TRUE}
TADA_Boxplot(WQP_clean_subset, id_cols = "TADA.ComparableDataIdentifier")
```

`TADA_RetainRequired` removes extra columns. Enter ?TADA_RetainRequired
into the console for more details.

```{r}
WQP_clean_subset_final <- TADA_RetainRequired(WQP_clean_subset)
```

## Record end time

```{r analysis time}
end.time <- Sys.time()

end.time - start.time
```

Reproducible and Documented

This workflow is reproducible and the decisions at each step are well
documented. This means that it is easy to go back and review every step,
understand the decisions that were made, make changes as necessary, and
run it again.
