---
title: "TADA R8 Demo January 2026"
always_all_html: yes
author: "TADA Team"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    self_contained: false
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteIndexEntry{TADA R8 Demo January 2026}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{EPATADA}
  %\VignetteEngine{knitr::rmarkdown}
description: An introduction to using the EPATADA R package to retrieve, clean, analyze, and visualize WQP data. This recommended work flow focuses on A) Retrieving, filtering and cleaning data from the WQP. B) Create, review and define up to date assessment unit/WQP monitoring location. C) Assigning uses to assessment units. D) Create a TADA-Compatible criteria table which defines the relevant parameters and uses applicable for analysis of your WQP data.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = T,
  dev.args = list(bg = "transparent")
)
```

# Overview and Setup

**Welcome!**

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to download the TADA R package from GitHub, access and
parameterize several important functions, and create basic
visualizations with a sample data set.

**Note:** TADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.

If you are interested in contributing to TADA development, more
information is available at:

[Contributing](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)

We welcome collaboration with external partners.

Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing TADA because it is only available on GitHub.

```{r install_remotes, results = 'hide'}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load TADA using the remotes package. TADA R Package
dependencies will also be downloaded automatically from CRAN with the
TADA install. You may be prompted in the console to update dependency
packages that have more recent versions available. If you see this
prompt, it is recommended to update all of them (enter 1 into the
console).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

```{r dev_install, eval = F, results = 'hide', include = F}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the library() function to load the TADA R Package into your
R session.

```{r library, results = 'hide'}
library(EPATADA)
```

Help pages

All TADA R package functions have their own individual help pages,
listed on the Function reference page on the GitHub site. Users can also
access the help page for a given function in R or RStudio using the
following format (example below): ?[name of TADA function]

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

## A) Retrieving, filtering and cleaning data from the WQP

Query the WQP using TADA_DataRetrieval. TADA_AutoClean is a powerful
function that runs as part of TADA_DataRetrieval when applyautoclean =
TRUE. It performs a variety of tasks, for example:

1.  creating new "TADA" prefixed columns and and capitalizing their
    contents to reduce case sensitivity issues,

2.  converts special characters in value columns,

3.  converts latitude and longitude values to numeric,

4.  replaces "meters" with "m",

5.  replaces deprecated characteristic names with current WQX names,

6.  harmonizes result and detection limit units,

7.  converts depths to meters, and

8.  creates the column TADA.ComparableDataIdentifier by concatenating
    characteristic name, result sample fraction, method speciation, and
    result measure unit.

In this example, we will get dissolved oxygen, bacteria and pH data from 
Missoula County, Montana.

```{r load WQP data}
# get MT data
tada.MT <- TADA_DataRetrieval(
  startDate = "2020-01-01",
  endDate = "2022-12-31",
  statecode = "MT",
  characteristicName = c(
    "Dissolved oxygen (DO)",
    "Escherichia",
    "Escherichia coli",
    "pH"
  ),
  countycode = "Missoula County",
  applyautoclean = TRUE,
  ask = FALSE
)
```

### Flag, clean, and visualize

Now, let's use EPATADA functions to review, visualize, and whittle the
returned WQP data down to include only results that are applicable to
our water quality analysis and area of interest.

TADA is primarily designed to accommodate water data from the WQP. Let’s
see what activity media types are represented in the data set. Are there
any media type that are not water in our data frame?

```{r Review and Filter By Media Type}
# Create table with count for each ActivityMediaName
TADA_FieldValuesTable(
  tada.MT,
  field = "ActivityMediaName",
  characteristicName = "null"
)
```

Create an overview map. For each site, we can view the measurement
count, visit count and characteristic counts along with the site ID,
site name, and organization name.

```{r TADA_OverviewMap}
TADA_OverviewMap(tada.MT)
```

Let's take a quick look at all unique values in the
MonitoringLocationIdentifier column and see how how many results are
associated with each.

```{r TADA_FieldValuesTable}
# use TADA_FieldValuesTable to create a table of the number of results per MonitoringLocationIdentifier
sites <- TADA_FieldValuesTable(
  tada.MT,
  field = "MonitoringLocationIdentifier",
  characteristicName = "null"
)

DT::datatable(sites, fillContainer = TRUE)
```

What about OrganizationFormalName?

```{r TADA_FieldValuesTable2 }
# use TADA_FieldValuesTable to create a table of the number of results per MonitoringLocationIdentifier
orgs <- TADA_FieldValuesTable(
  tada.MT,
  field = "OrganizationFormalName",
  characteristicName = "null"
)

DT::datatable(orgs, fillContainer = TRUE)
```

Or TADA.CharacteristicName?

```{r TADA_FieldValuesTable3 }
# use TADA_FieldValuesTable to create a table of the number of results per TADA.CharacteristicName
chars <- TADA_FieldValuesTable(
  tada.MT,
  field = "TADA.CharacteristicName"
)

DT::datatable(chars, fillContainer = TRUE)
```

Let's filter our TADA data frame to only include data from the top 4
OrganizationFormalName. The orgs data frame is sorted in descending
order in which [1:4] will select the top 4 OrganizationFormalName values.

```{r}
tada.MT.filter <- tada.MT |>
  dplyr::filter(
    OrganizationFormalName %in% orgs$Value[1:4]
    )
```

Next, let's check if the dataset contains potential duplicate results
from within a single organization or from within multiple organizations
(such as when two or more organizations monitor the same location and
may submit duplicate results).

For the case in which there exists potential duplicates from multiple
organizations, if you would like to prioritize results from one
organization over another, this can be done using the org_hierarchy
argument in `TADA_FindPotentialDuplicatesMultipleOrgs`.

```{r duplicates}
# find duplicates from single org
tada.MT.singleorg.dups <- TADA_FindPotentialDuplicatesSingleOrg(tada.MT.filter)

# Review organizations. You can select one to prioritize in TADA_FindPotentialDuplicatesMultipleOrgs
unique(tada.MT.singleorg.dups$OrganizationIdentifier)
unique(tada.MT.singleorg.dups$OrganizationFormalName)

# find duplicates across multiple orgs
tada.MT.multipleorgs.dups <- TADA_FindPotentialDuplicatesMultipleOrgs(
  tada.MT.singleorg.dups,
  org_hierarchy = c("MDEQ_WQ_WQX", "TSWQC_WQX", "MTWTRSHD_WQX", "USGS-MT")
)
```

We will select to keep only unique samples from
`TADA_FindPotentialDuplicatesSingleOrg` where TADA.SingleOrgDup.Flag equals "Unique".

```{r filter}
tada.MT.clean <- tada.MT.multipleorgs.dups |>
   dplyr::filter(TADA.SingleOrgDup.Flag == "Unique") |>
   dplyr::filter(TADA.ResultSelectedMultipleOrgs == "Y")
```

Remove intermediate variables in R by using 'rm()'. In the remainder of
this workshop, we will work with the clean data set.

```{r}
rm(tada.MT.nearby.flag.dups, tada.MT.nearby, tada.MT.filter)
```

Censored data are measurements for which the true value is not known,
but we can estimate the value based on known lower or upper detection
conditions and limit types. TADA fills missing *TADA.ResultMeasureValue*
and *TADA.ResultMeasure.MeasureUnitCode* values with values and units
from *TADA.DetectionQuantitationLimitMeasure.MeasureValue* and
*TADA.DetectionQuantitationLimitMeasure.MeasureUnitCode*, respectively,
using the `TADA_AutoClean` function.

The TADA package currently has functions that summarize censored data
incidence in the dataset and perform simple substitutions of censored
data values, including x times the detection limit and random selection
of a value between 0 and the detection limit. The user may specify the
methods used for non-detects and over-detects separately in the input to
the `TADA_SimpleCensoredMethods` function. The next step we take in this
example is to perform simple conversions to the censored data in the
dataset: we keep over-detects as is (no conversion made) and convert
non-detect values to 0.5 times the detection limit (half the detection
limit).

In this example, no censored data results were found. The input
tada.MT.clean data is returned.

```{r censored}
tada.MT.clean <- TADA_SimpleCensoredMethods(
  tada.MT.clean,
  nd_method = "multiplier",
  nd_multiplier = 0.5,
  od_method = "as-is",
  od_multiplier = "null"
)
```

`TADA_FindQCActivities` identifies results with QA/QC identifies results with QA/QC ActivityTypeCodes. When clean = TRUE, it removes QA/QC results.

```{r QCdata}
tada.MT.clean <- TADA_FindQCActivities(tada.MT.clean, clean = TRUE)
```

TADA_RunKeyFlagFunctions is a shortcut function to run important TADA
flagging functions. See ?function documentation for TADA_FlagResultUnit,
TADA_FlagFraction, TADA_FlagMeasureQualifierCode, and
TADA_FlagSpeciation for more information.

```{r TADA_RunKeyFlagFunctions}
tada.MT.clean <- TADA_RunKeyFlagFunctions(
  tada.MT.clean,
  clean = TRUE
)
```

Let's take another look at all unique values in the
TADA.ComparableDataIdentifier column and see how how many results are
associated with each. TADA.ComparableDataIdentifier concatenates
TADA.CharacteristicName, TADA.ResultSampleFractionText,
TADA.MethodSpeciationName, and TADA.ResultMeasure.MeasureUnitCode.

```{r TADA_FieldValuesTable4}
# use TADA_FieldValuesTable to create a table of the number of results per TADA.ComparableDataIdentifier
chars <- TADA_FieldValuesTable(
  tada.MT.clean,
  field = "TADA.ComparableDataIdentifier",
  characteristicName = "null"
)

chars_before <- unique(tada.MT.clean$TADA.ComparableDataIdentifier)
chars_before

DT::datatable(chars, fillContainer = TRUE)
```

Scroll through the table and check to see if there any synonyms. Are
there any TADA.ResultSampleFractionText, TADA.MethodSpeciationName, and
TADA.ResultMeasure.MeasureUnitCode that may need to be harmonized for
consistency? Let's check using `TADA_HarmonizeSynonyms` so their results
can be directly compared.

Let's give it a try.

```{r}
tada.MT.clean <- TADA_HarmonizeSynonyms(tada.MT.clean)
```

How many unique TADA.ComparableDataIdentifiers do we have now? In this
example, PH has now been harmonized to "PH_NONE_NONE_NONE". Dissolved Oxygen has also been harmonized.

```{r}
chars_after <- unique(tada.MT.clean$TADA.ComparableDataIdentifier)
```

Remove intermediate variables.

```{r}
rm(chars_before, chars_after)
```

The functions TADA_FlagAboveThreshold and TADA_FlagBelowThreshold are used to flag results falling above and below the WQX national thresholds, respectively. When clean = TRUE, the flagged results are removed from the TADA data frame.

```{r thresholds}
tada.MT.clean <- TADA_FlagAboveThreshold(tada.MT.clean, clean = TRUE, flaggedonly = FALSE)

tada.MT.clean <- TADA_FlagBelowThreshold(tada.MT.clean, clean = TRUE, flaggedonly = FALSE)
```

### visualize data

Create a pie chart to display the count of results for each TADA.CharacteristicName.

```{r}
TADA_FieldValuesPie(
  tada.MT.clean,
  field = "TADA.CharacteristicName",
  characteristicName = "null"
)
```

## B) Create, review and define up to date assessment unit/WQP monitoring location.

Module 2: Geospatial Functions in TADA and assignment of uses to Assessment Units.

The **`TADA_CreateAUMLCrosswalk`** function efficiently creates a
crosswalk between ATTAINS assessment units and WQP monitoring locations.
It uses three prioritized data sources:

1.  **User-Supplied Crosswalk:** An optional crosswalk provided by the
    user (e.g., see example **`user.supplied.cw`** in next code chunk).

2.  **ATTAINS Crosswalk:** Utilizes **`TADA_GetATTAINSAUMLCrosswalk`**
    to incorporate crosswalk information stored by participating
    organizations in ATTAINS. There may not be an ATTAINS crosswalk
    available for all organizations as storing this information in
    ATTAINS is optional for states and some tribes are still in the
    process of developing their crosswalks.

3.  **ATTAINS Catchments/Geospatial Join:** Employs
    **`TADA_CreateATTAINSAUMLCrosswalk`** to connect monitoring
    locations to assessment units using ATTAINS catchments through a
    geospatial join. This function converts WQP monitoring locations
    into a geospatial sf object and associates them with their
    intersecting NHDPlus high resolution catchments containing
    [entity-defined assessment units in
    ATTAINS](https://www.epa.gov/waterdata/integrated-reporting-georeferencing-pilot-report).

**Process**

-   The function prioritizes these sources in the order listed. It
    automatically attempts to assign unassigned monitoring locations
    using the next available data source. For example, if any monitoring
    locations remain unassigned after checking the user-supplied
    crosswalk, the function will check the ATTAINS crosswalk, then the
    geospatial join, as needed.

For demonstration purposes, let's assume we know for sure this
MonitoringLocationIdentifier, "MTWTRSHD_WQX-COMBITR02" should be
associated with the AssessmentUnitIdentifier "MT76M001_020". 
Let's create a crosswalk to assign "MTWTRSHD_WQX-COMBITR02" to the
assessment unit "MT76M001_020" and give it the WaterType "RIVER".

### Define a user supplied crosswalk of Assessment Units and Monitoring Location(s)

```{r demo.user, results = 'hide'}
# user would like to associate this MonitoringLocationIdentifier to this AU.
user.supplied.cw <- data.frame(
    AssessmentUnitIdentifier = "MT76M001_020",
    MonitoringLocationIdentifier = "MTWTRSHD_WQX-COMBITR02",
    WaterType = "RIVER"
  )

DT::datatable(user.supplied.cw, fillContainer = TRUE)
```

### Run `TADA_GetATTAINSCrosswalk`

Now, let's check to see if MTDEQ has submitted a prior crosswalk to ATTAINS
for the assessment unit identifier "MT76M001_020". 

Does our TADA data frame contain any of these monitoring locations?

```{r TADA_GetATTAINSCrosswalk}
ATTAINS.cw <- TADA_GetATTAINSAUMLCrosswalk(
  org_id = "MTDEQ"
) |>
  dplyr::filter(ATTAINS.AssessmentUnitIdentifier == "MT76M001_020")

any(unique(tada.MT.clean$MonitoringLocationIdentifier) %in% ATTAINS.cw)
```

### Run `TADA_CreateAUMLCrosswalk`

We are uncertain about any other assignments of assessment units
to monitoring locations. Let's match the monitoring locations catchments overlay
from tada.MT.clean with assessment units catchments with a geospatial join 
from ATTAINS by using TADA_CreateATTAINSAUMLCrosswalk to identify which assessment units
intersect with each monitoring location. 

Recall, TADA_CreateATTAINSAUMLCrosswalk prioritizes the user supplied crosswalk,
user.supplied.cw. If any monitoring locations remain unassigned after checking 
the user-supplied crosswalk, the function will check the ATTAINS crosswalk 
(see ATTAINS.cw above), then lastly, the geospatial join, as needed.

```{r TADA_CreateAUMLCrosswalk}
# make AU assignments for unassigned MLs
MT.AUMLRef <- TADA_CreateAUMLCrosswalk(
  tada.MT.clean,
  au_ref = user.supplied.cw,
  org_id = "MTDEQ",
  fill_ATTAINS_catch = TRUE,
  return_nearest = TRUE,
  batch_upload = FALSE,
)
```

The output of TADA_CreateAUMLCrosswalk is a list of data frames:
1) "TADA_with_ATTAINS"   
2) "ATTAINS_catchments"  
3) "ATTAINS_points"     
4) "ATTAINS_lines"
5) "ATTAINS_polygons"    
6) "ATTAINS_crosswalk"  

The data frame, TADA_with_ATTAINS contains all of the clean TADA data as well as the added ATTAINS columns. The other ATTAINS_catchments, ATTAINS_points, ATTAINS_lines, and ATTAINS_polygons data frames contain geometry for mapping. The ATTAINS_crosswalk data frame contains a simple crosswalk of monitoring locations and assessment units.

Now, let's view the assessment units and monitoring locations on a map to review the assessment unit/monitoring location assignments.

```{r view.map}
TADA_ViewATTAINS(MT.AUMLRef, ref_icons = TRUE)
```

## C) Assigning uses to assessment units.

Let's filter our TADA data frame to a single assessment unit "MT76M001_020",
and for a single TADA.CharacteristicName, "DISSOLVED OXYGEN (DO)" for
demonstration purposes for the remainder of this example.

```{r}
# get subset DO data for one AU
tada.MT.clean.DO <- MT.AUMLRef$TADA_with_ATTAINS |>
  sf::st_drop_geometry() |>
  dplyr::filter(TADA.CharacteristicName == "DISSOLVED OXYGEN (DO)",
                ATTAINS.AssessmentUnitIdentifier == "MT76M001_020")

tada.MT.AUMLRef <- MT.AUMLRef$ATTAINS_crosswalk |>
  dplyr::filter(ATTAINS.AssessmentUnitIdentifier == "MT76M001_020")
```

### From Prior ATTAINS Assessment Cycles

We will use the ATTAINS.UseNames that were used in prior
ATTAINS assessment cycles from MTDEQ for this assessment unit.
Users can append additional rows as needed or remove not applicable
use names if a use is no longer applicable for the assessment unit.
See TADA_Module2 vignette for more information.

```{r}
AU.uses <- TADA_AssignUsesToAU(
  tada.MT.clean.DO,
  org_id = "MTDEQ",
  AUMLRef = tada.MT.AUMLRef
  )

DT::datatable(AU.uses, fillContainer = TRUE)
```

### From a User Supplied Crosswalk (Uses to AU)

Users may also choose to supply their own list of use_names
if desired rather than sourcing it from the prior ATTAINS assessment cycle.
To ensure we use the correct column names, we can run TADA_AssignUsesToAU()
with no argument function inputs. Note: The TADA.AssessmentUnitStatus column is auto populated
and does not need to be supplied from a user supplied crosswalk. For 
ATTAINS.WaterType, please ensure this matches the ATTAINS.WaterType from 
the output of MT.AUMLRef$ATTAINS_crosswalk for your assessment unit.

```{r}
names(TADA_AssignUsesToAU())

AU.uses.user.supplied <- data.frame(
  ATTAINS.OrganizationIdentifier = c("MTDEQ", "MTDEQ"),
  ATTAINS.AssessmentUnitIdentifier = c("MT76M001_020", "MT76M001_020"),
  ATTAINS.UseName = c("Example: Aquatic Life, cold waters", "Example: Human Health"),
  ATTAINS.WaterType = c("RIVER", "RIVER"),
  IncludeOrExclude = c("Include", "Include")
)

AU.uses2 <- TADA_AssignUsesToAU(
  tada.MT.clean,
  org_id = "MTDEQ",
  AUMLRef = tada.MT.AUMLRef,
  AU_UsesRef = AU.uses.user.supplied
)

DT::datatable(AU.uses2, fillContainer = TRUE)
```


## D) Create a TADA-Compatible criteria table which defines the relevant parameters and uses applicable for analysis of your TADA data frame.

The EPATADA Module 3 functions were designed to: 

(1)   Assist users with associating Water Quality Portal monitoring locations with assessment units and designated uses       from ATTAINS and 
(2)   Compare Water Quality Portal results with numeric water quality criteria from a TADA-Compatible criteria table.

Users can choose to 

1)    Fill in and populate their own criteria table **manually**, 
2)    use an **auto_assign method** of creating a crosswalk of ATTAINS parameter names
      to WQX Characteristic and to CST pollutant names along with a crosswalk between
      ATTAINS use names and CST uses. This allows for populating the magnitude 
      components from the Criteria Search Tool for each ATTAINS parameter and use,
      if one is found for your organization (duration, frequency and other 
      methodology components will still need user inputs and a thorough review 
      of the crosswalk between the CST and ATTAINS parameter and uses will be required),
3)    or pull in their organization's criteria table from a list of pre-filled 
      template from the **TADACommunityHub*** if one exists (currently all R8 states                                          and tribes have participated in submitting their organization's criteria 
      table to this repository!).

\* The TADACommunityHub is a collaborative hub where TADA users share and maintain information for custom workflows (e.g., state, tribal, territory, EPA, or other entities). For example, it includes user-contributed criteria and methodologies templates to support reproducible and efficient analyses.

### TADACommunityHub criteria table

For this demo, We will load in a pre-filled criteria template completed by MTDEQ.

```{r}
# Load the example R8 criteria table
criteria_table <- system.file("extdata", "criteria_table.rda", package = "EPATADA")
load(criteria_table)

MT_criteria <- TADA_DefineCriteriaMethodology(
  tada.MT.clean.DO,
  org_id = "MTDEQ",
  criteriaMethods = criteria_table,
  AUMLRef = tada.MT.AUMLRef,
  AU_UsesRef = AU.uses,
  displayUniqueId = TRUE
  )

DT::datatable(MT_criteria, fillContainer = TRUE)
```

### Autoassign criteria table

Alternatively, if your organization has not submitted a criteria table
to the TADACommunityHub repository, TADA_DefineCriteriaMethodology has
an auto_assign function input which allows populating the magnitude values
from the Criteria Search Tool (CST) to past ATTAINS.ParameterName and
ATTAINS.UseName for your organization from prior submission to ATTAINS
assessment cycles. Filter it to just "DISSOLVED OXYGEN (DO)".

Note that duration, frequency and other criteria components will still need
to be filled out with auto_assign. However, if users are only interested in
individual result measurement excursions, they can proceed to use this table as is.

```{r}
criteria_table_auto <- TADA_DefineCriteriaMethodology(
  tada.MT.clean,
  org_id = "MTDEQ",
  auto_assign = T,
  criteriaMethods = NULL,
  AUMLRef = tada.MT.AUMLRef,
  AU_UsesRef = AU.uses,
  displayUniqueId = TRUE
  ) |>
  dplyr::filter(TADA.CharacteristicName == "DISSOLVED OXYGEN (DO)")

DT::datatable(criteria_table_auto, fillContainer = TRUE)
```

We will continue using the criteria table for MTDEQ that was submitted
to the TADACommunityHub.

### Editing the criteria table in excel

If further edits are desired to the criteria table, users can export
the criteria table into an excel spreadsheet and read it back into
the R environment once any edits are made.

```{r, eval = FALSE}
MT_criteria_excel <- TADA_DefineCriteriaMethodology(
  tada.MT.clean.DO,
  org_id = "MTDEQ",
  criteriaMethods = criteria_table,
  AUMLRef = tada.MT.AUMLRef,
  AU_UsesRef = AU.uses,
  displayUniqueId = TRUE,
  excel = T,
  overwrite = T
  )
```

### View and analyze scatter plot excursions for DO for your assessment unit

Now, lets view some scatter plots for the criteria defined in your criteria table
which will be compared to the results from your TADA data frame for DO for a 
single assessment unit, "MT76M001_020". We will join your TADA data frame to the
criteria table using the MTDEQ table sourced from the TADACommunityHub.

```{r}
# join use and criteria to tada df (many to many relationships may occur)
tada.MT.clean.DO2 <- tada.MT.clean.DO |>
   dplyr::left_join(
     MT_criteria,
     relationship = 'many-to-many',
     dplyr::join_by(TADA.ComparableDataIdentifier)
     )

# break into separate data frames for each unique combination of DurationValue, DurationMethod, DurationUnit, and UniqueSpatialCriteria for analysis and figures
tada.MT.do.subsets <- tada.MT.clean.DO2 |>
  dplyr::group_by(DurationValue, DurationMethod, DurationUnit,
                  UniqueSpatialCriteria) |>
  dplyr::group_split()
```

Note that the scatter points are only individual result measurements. 
Future TADA development in the package is being made for handling of the 
full criteria (magnitude, duration and frequency), seasonal considerations
and other data sufficiency needs!

```{r}
TADA_scatters <- list()

n <- length(tada.MT.do.subsets)

for (i in 1:n){
  # need to set this up as function to plot multiple
  desc <- paste0(
    "Lower bound for ",
    tada.MT.do.subsets[[i]]$ATTAINS.UseName[1],
    " (",
    tada.MT.do.subsets[[i]]$DurationValue[1],
    " ",
    tada.MT.do.subsets[[i]]$DurationUnit[1],
    " ",
    tada.MT.do.subsets[[i]]$DurationMethod[1],
    " for ",
    tada.MT.do.subsets[[i]]$UniqueSpatialCriteria[1],
    ")"
  )
  
  TADA_scatters[[i]] <- TADA_Scatterplot(tada.MT.do.subsets[[i]]) |>
     plotly::add_lines(
        x = c(
      min(tada.MT.do.subsets[[i]]$ActivityStartDate, na.rm = TRUE),
      max(tada.MT.do.subsets[[i]]$ActivityStartDate, na.rm = TRUE)
    ),
     y = (tada.MT.do.subsets[[i]]$MagnitudeValueLower[1]),
    inherit = FALSE,
    name = paste(strwrap(desc, width = 20), collapse = "<br>"),
    line = list(color = "red", dash = "dash")
  ) 
  
  names(TADA_scatters)[i] <- desc
}
```

Our criteria table for DO contained multiple rows. Let's view the different 
criteria and determine which are appropriate for our assessment unit.
Let's assume the scatter plots for 
"Lower bound for Aquatic Life (1 n-day arithmetic min for Cold water, other life stages present)" and
"Lower bound for Aquatic Life (7 n-day arithmetic mean for Cold water, other life stages present)" 
are appropriate criteria to use for our assessment unit.

```{r}
names(TADA_scatters)
# TADA_scatters
TADA_scatters[[1]]
TADA_scatters[[5]]
```
