---
title: "TADA Module 2: Geospatial Functions"
format: html
editor: visual
author: "TADA Team"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{TADA Module 2: Geospatial Functions}
  %\VignetteEngine{knitr::rmarkdown}
description: An overview of TADA Module 2 functions and a recommended workflow.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Welcome!

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://bookdown.org/yihui/rmarkdown/) document walks
users through how to download the TADA R package from GitHub, access and
parameterize several important functions, and create basic
visualizations with a sample data set.

**Note: TADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.**

**If you are interested in contributing to TADA development, more
information is available at:**

[**Contributing**](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)

**We welcome collaboration with external partners.**

## Install and load packages

First, install and load the remotes package specifying the repo. This is
needed before installing TADA because it is only available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)
```

Next, install and load TADA using the remotes package. TADA R Package
dependencies will also be downloaded automatically from CRAN with the
TADA install. You may be prompted in the console to update dependency
packages that have more recent versions available. If you see this
prompt, it is recommended to update all of them (enter 1 into the
console).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
```

Finally, use the **library()** function to load the TADA R Package into
your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Help pages

All TADA R package functions have their own individual help pages,
listed on the [Function
reference](https://usepa.github.io/EPATADA/reference/index.html) page on
the GitHub site. Users can also access the help page for a given
function in R or RStudio using the following format (example below):
`?[name of TADA function]`

```{r help_pages, eval = F}
# Access help page for TADA_DataRetrieval
?TADA_DataRetrieval
```

## Geospatial Functions in TADA

This vignette represents functions that provide users the option to
convert TADA Water Quality Portal data into a geospatial sf object as
well as to associate water quality observations with their [intersecting
NHD catchments containing entity-defined water quality assessment units
in
ATTAINS](https://www.epa.gov/waterdata/integrated-reporting-georeferencing-pilot-report).

**A Note About ATTAINS:**

The Assessment, Total Maximum Daily Load (TMDL) Tracking and
Implementation System ([ATTAINS](https://www.epa.gov/waterdata/attains))
is an online platform that organizes and combines each state and
participating tribe's Clean Water Act reporting data into a single data
repository. The geospatial component of ATTAINS includes spatial
representations of each entity's surface water assessment units as well
as their assigned designated uses, their most recent EPA reporting
category (i.e., their impairment status), their impaired designated
uses, and the parameter(s) causing the impairment.

Within an assessment unit, the criteria or thresholds used to assess
water quality typically remain the same and all water features are
assessed as one entity (although there are some exceptions, for example
if a single assessment unit crosses multiple ecoregions). Depending on
the state or tribe, these assessment units can be a specific point or
series of points along a waterbody such as a river or lake, a river
reach (line), an entire waterbody such as a river or lake (polygon), or
even an entire watershed. In other words, assessment units can take the
form of point, line, and area (polygon) features, or some combination of
all of them. Moreover, it is possible that some assessment units are not
geospatially referenced at all, meaning they are not captured in the
ATTAINS geospatial database.

## `TADA_MakeSpatial()`

This function converts any Water Quality Portal (WQP)-style dataframe
with latitude/longitude data into a geospatial shapefile object. To run
the function, the user supplies a WQP dataframe and the coordinate
reference system that they want the spatial object to be in (the default
is WGS 84). For the function to work properly, the input dataframe must
have - at a minimum - WQP observation coordinates in "LongitudeMeasure"
and "LatitudeMeasure" and a
"HorizontalCoordinateReferenceSystemDatumName" column.

#### Using `TADA_MakeSpatial()`

First, we will need to pull in some TADA Water Quality Portal Data:

```{r}
# pH data in Larimer County, Colorado for the year 2020.
TADA_dataframe <- TADA_DataRetrieval(
  startDate = "2020-01-01",
  endDate = "2020-12-31",
  characteristicName = "pH",
  countycode = "US:08:069",
  applyautoclean = TRUE
)
```

Now, we can make the water quality data spatial by running
`TADA_MakeSpatial()`:

```{r}
# default CRS is WGS84 (4326)
TADA_spatial <- TADA_MakeSpatial(.data = TADA_dataframe, crs = 4326)
```

This new spatial object is identical to the original TADA dataset, but
now includes a "geometry" column that allows for mapping and additional
geospatial capabilities. Enter `?TADA_MakeSpatial` into the console to
review another example of this function in use and additional
information.

```{r}
leaflet::leaflet() %>%
  leaflet::addProviderTiles("Esri.WorldTopoMap",
    group = "World topo",
    options = leaflet::providerTileOptions(
      updateWhenZooming = FALSE,
      updateWhenIdle = TRUE
    )
  ) %>%
  leaflet::clearShapes() %>%
  leaflet.extras::addResetMapButton() %>%
  leaflet::addLegend(
    position = "bottomright",
    colors = "black",
    labels = "Water Quality Observation(s)",
    opacity = 1
  ) %>%
  leaflet::addCircleMarkers(
    data = TADA_spatial,
    color = "grey", fillColor = "black",
    fillOpacity = 0.8, stroke = TRUE, weight = 1.5, radius = 6,
    popup = paste0(
      "Site ID: ",
      TADA_spatial$MonitoringLocationIdentifier,
      "<br> Site Name: ",
      TADA_spatial$MonitoringLocationName
    )
  )
```

## `TADA_GetATTAINS()`

This function pulls in ATTAINS data from the EPA's ATTAINS Assessment
Geospatial Service and links it to TADA-pulled Water Quality Portal
observations. For the function to work properly, the input dataframe
must have - at a minimum - WQP observation coordinates in
"LongitudeMeasure" and "LatitudeMeasure" columns and a
"HorizontalCoordinateReferenceSystemDatumName" column.

By default, `TADA_GetATTAINS()` returns a dataframe with ATTAINS-linked
Water Quality Portal entries. Users have the added option of returning
the intersecting ATTAINS geospatial shapefile objects with their
ATTAINS-linked Water Quality Portal dataframe. If `return_sf = TRUE`,
the function returns a list containing the dataframe and shapefile
objects named `ATTAINS_catchments`, `ATTAINS_lines`, `ATTAINS_points`,
and `ATTAINS_polygons`. Note, if any of these shapefile objects are
empty, this indicates that there are no ATTAINS objects of that type
intersecting any WQP-linked ATTAINS catchment.

Regardless of the user's decision on returning the ATTAINS shapefile
objects, `TADA_GetATTAINS()` always returns a dataframe (or dataframes
if `fill_catchments = TRUE`, see section ***Filling in missing ATTAINS
assessment units**)* containing the original TADA WQP dataset, plus new
columns representing the ATTAINS assessment unit(s) that fall within the
same NHDPlus HiRes catchment as them. This means that it is possible for
a single TADA WQP observation to have multiple ATTAINS assessment units
linked to it and subsequently more than one row of data. Such WQP
observations can be identified using the new index column (i.e.,
multiple rows with the same index value are the same observation).

#### Using `TADA_GetATTAINS()`

Using either our original `TADA_dataframe` *or* the geospatial version
`TADA_spatial`, we can pull in the ATTAINS catchment features that
intersect our observations:

```{r}
TADA_with_ATTAINS <- TADA_GetATTAINS(.data = TADA_dataframe, return_sf = FALSE)

# Can also be performed on the spatial data:
# TADA_with_ATTAINS <- TADA_GetATTAINS(.data = TADA_spatial, return_sf = FALSE)
```

This new `TADA_with_ATTAINS` object is a modification of the original
TADA Water Quality Portal dataframe that now has additional columns
associated with the ATTAINS assessment unit(s) that lie in the same NHD
HiRes catchment as them (these columns are prefixed with "ATTAINS").
Moreover, because our `TADA_with_ATTAINS` object contains more rows than
the original TADA dataframe, we can deduce that some Water Quality
Portal observations fall within an NHD catchment that contains more than
one ATTAINS assessment unit.

```{r}
TADA_with_ATTAINS_list <- TADA_GetATTAINS(.data = TADA_dataframe, return_sf = TRUE)

# Can also be performed on the spatial data:
# TADA_with_ATTAINS_list <- TADA_GetATTAINS(.data = TADA_spatial, return_sf = TRUE)
```

If we set `return_sf = TRUE` as done to create the
`TADA_with_ATTAINS_list` object above, we also now have all the raw
intersecting ATTAINS features associated with these ATTAINS catchment
observations stored in a list along with the TADA dataframe.

Now, let's select specific columns from the `TADA_with_ATTAINS`
dataframe, and create a new dataframe with ONLY the unique combinations
of WQP MonitoringLocationIdentifier's and ATTAINS Assessment Unit
Identifiers.

```{r}
TADA_with_ATTAINS_subset <- TADA_with_ATTAINS %>%
  dplyr::select(
      "LongitudeMeasure", "LatitudeMeasure", "MonitoringLocationTypeName",
      "MonitoringLocationIdentifier", "ATTAINS.assessmentunitidentifier",
      "ATTAINS.overallstatus", "ATTAINS.isassessed", "ATTAINS.isimpaired",
      "ATTAINS.organizationid", "ATTAINS.assessmentunitname",
      "ATTAINS.reportingcycle", "ATTAINS.waterbodyreportlink"
  ) %>%
  dplyr::distinct(.keep_all = FALSE)
```

#### Filling in missing ATTAINS assessment units

As you can see in the above examples, not all WQP observations have an
intersecting ATTAINS catchment: see that in the returned dataframe, some
WQP observations have NAs where there should be ATTAINS information. In
these instances, the user can optionally fill in catchment information
from the NHD by entering `fill_catchments = TRUE`:

```{r}
TADA_with_ATTAINS_filled <- TADA_GetATTAINS(TADA_dataframe, fill_catchments = TRUE, return_sf = TRUE)
```

When `fill_catchments = TRUE`, the returned object splits observations
into two dataframes: WQP observations with ATTAINS catchment data, and
WQP observations without ATTAINS catchment data. Instead of listing
ATTAINS information in `TADA_without_ATTAINS`, it links basic
information about the catchment including its unique identifier,
catchment area, and the resolution of the NHD used. As a default,
`TADA_GetATTAINS()` will use the NHD HiRes (`resolution = "Hi"`) for
filling in missing ATTAINS catchments. However, the user can choose to
change the resolution to the NHDPlus V2 by setting `resolution = "Med".`

Moreover, when `return_sf = TRUE` as above, the function will
additionally return the raw catchment features associated with the
observations in `TADA_without_ATTAINS` in a new shapefile called
`without_ATTAINS_catchments`.

#### Arguments for `TADA_GetATTAINS()`

-   `.data`: Your input TADA-style Water Quality Portal data.

-   `fill_catchments`: If TRUE, it will find intersecting NHD catchments
    to fill in information for samples not covered by ATTAINS.

-   `resolution`: Specifies which version of the NHD to use if filling
    catchments: the "Med", or "Hi". The default option is "Hi".

-   `return_sf`: If TRUE, returns spatial data in addition to tabular
    data.

Enter `?TADA_GetATTAINS` into the console to review another example of
this function in use and additional information.

## `TADA_ViewATTAINS()`

This function visualizes the raw ATTAINS features that are linked to the
TADA Water Quality Portal observations that are generated in
`TADA_GetATTAINS()` when `return_sf = TRUE`. For the function to work
properly, the input dataframe must be the list produced from
`TADA_GetATTAINS()` with `return_sf = TRUE`. The map also displays the
Water Quality Portal monitoring locations used in `TADA_GetATTAINS()`.

#### Using `TADA_ViewATTAINS()`

Let's view the data associated with our `TADA_with_ATTAINS_list` object!
Enter `?TADA_ViewATTAINS` into the console to review another example
query and additional information.

```{r, fig.height=7}
TADA_ViewATTAINS(.data = TADA_with_ATTAINS_list)
```

When `fill_catchments = TRUE`, `TADA_ViewATTAINS()` will also map the
`without_ATTAINS_catchments`:

```{r, fig.height=7}
TADA_ViewATTAINS(.data = TADA_with_ATTAINS_filled)
```

Enter `?TADA_ViewATTAINS` into the console to review another example of
this function in use and additional information.
