---
title: 'Beginner R user: introduction to using the EPATADA R package to retrieve, clean, analyze, and visualize WQP data'
author: "TADA Team"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = F)
```

## Install and load

First, install and load the remotes package specifying the repo. This is
needed before installing the *EPATADA* R package because it is only
available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes", repos = "http://cran.us.r-project.org")
library(remotes)
```

Next, install (or update) and load the *EPATADA* R package using the
*remotes* R package. Additional dependency R packages that are used
within *EPATADA* will be downloaded automatically. You may be prompted
in the console to update dependency packages that have more recent
versions available. If you see this prompt, it is recommended to update
all of them (enter 1 into the console). Our team is actively developing
*EPATADA*, therefore we highly recommend that you update the package
(and all of its dependencies) each time you use it.

```{r install_EPATADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA", ref = "develop", dependencies = TRUE)
```

```{r install_EPATADA_dev, eval = T, include = F}
remotes::install_github("USEPA/EPATADA", ref = "cybertown2025", dependencies = TRUE
)
```

## Record start time

```{r}
start.time <- Sys.time()
```

## Retrieve

Query the WQP using a bounding box (lat/long extent). 

```{r TADA_DataRetrieval}
WQP_Shepherdstown_raw <- TADA_DataRetrieval(
  startDate = "2019-01-01",
  endDate = "2024-12-31",
  bBox = c(-78.2241,38.9185,-77.1041,39.7487),
  applyautoclean = TRUE,
  ask = FALSE)
```

## Filter and clean

```{r}
WQP_Shepherdstown_clean <- TADA_AnalysisDataFilter(
  WQP_Shepherdstown_raw,
  clean = TRUE,
  surface_water = TRUE,
  ground_water = FALSE,
  sediment = FALSE
)
```

Find and remove duplicates

```{r}
WQP_Shepherdstown_clean <- TADA_FindPotentialDuplicatesSingleOrg(
  WQP_Shepherdstown_clean)
WQP_Shepherdstown_clean <- dplyr::filter(WQP_Shepherdstown_clean, TADA.SingleOrgDup.Flag == "Unique")
```

Censored data substitutions 
```{r}
WQP_Shepherdstown_clean <- TADA_SimpleCensoredMethods(
  WQP_Shepherdstown_clean,
  nd_method = "multiplier",
  nd_multiplier = 0.5,
  od_method = "as-is",
  od_multiplier = "null"
)
```
addressing conflicting detection condition and detection limit type information

```{r}
# WQP_Shepherdstown_clean <- TADA_FindPotentialDuplicatesMultipleOrgs(
#   WQP_Shepherdstown_clean
# )
# WQP_Shepherdstown_clean <- dplyr::filter(
#   WQP_Shepherdstown_clean,
#   TADA.ResultSelectedMultipleOrgs == "Y"
# )
```

```{r}
WQP_Shepherdstown_clean <- TADA_AutoFilter(WQP_Shepherdstown_clean)
```

```{r}
WQP_Shepherdstown_clean <- TADA_RunKeyFlagFunctions(
  WQP_Shepherdstown_clean,
  clean = TRUE
)
```

address na results
```{r}
WQP_Shepherdstown_clean <- WQP_Shepherdstown_clean[!is.na(WQP_Shepherdstown_clean$TADA.ResultMeasure.MeasureUnitCode),]
```

flag above and below thresholds
```{r}
WQP_Shepherdstown_clean <- TADA_FlagAboveThreshold(WQP_Shepherdstown_clean, clean = TRUE, flaggedonly = FALSE)
```

```{r}
WQP_Shepherdstown_clean <- TADA_FlagBelowThreshold(WQP_Shepherdstown_clean, clean = TRUE, flaggedonly = FALSE)
```

Harmonize synonyms
```{r}
WQP_Shepherdstown_clean <- TADA_HarmonizeSynonyms(WQP_Shepherdstown_clean)
```

calculate TN and TP
```{r}
WQP_Shepherdstown_clean <- TADA_CalculateTotalNP(WQP_Shepherdstown_clean, daily_agg = "max")

```

What characteristics are available?

```{r fieldValues_actmedia}
# Create pie chart for CharacteristicName
TADA_FieldValuesPie(WQP_Shepherdstown_clean, field = "TADA.CharacteristicName")
```

Filter: let's focus on a few characteristics of interest.

```{r}
# Select characteristics of interest
WQP_Shepherdstown_subset <- WQP_Shepherdstown_clean %>%
  dplyr::filter(TADA.CharacteristicName %in% c(
    "PH", "SPECIFIC CONDUCTANCE", "IRON", "CHLORIDE", "FECAL COLIFORM", "ESCHERICHIA COLI", "TOTAL NITROGEN"))
```

## Monitoring Location Filter and Review

Now that we've associated geospatial data from **ATTAINS** with the
**WQP** data, we can filter the data set to retain only results that
were collected in specified Assessment Units. We can also generate a
new table to give us some information about the individual monitoring
locations within the assessment units. After this filter step is
complete, we can remove the "ATTAINS." prefixed columns to reduce the
size of the data set, as we won't need them for the remaining steps in
this example.

We can also create a table with some basic information about the
Monitoring Locations in Assessment Units of interest and a pie chart to display the relative number of results contributed by each organization.

```{r Data Retrieval - Geospatial}
WQP_Shepherdstown_subset <- TADA_GetATTAINS(
  WQP_Shepherdstown_subset, 
  fill_catchments = FALSE,
  return_sf = FALSE, 
  return_nearest = TRUE)
```

Create table of monitoring location identifiers and AUs
```{r}
ML_AU_crosswalk <- WQP_Shepherdstown_subset %>%
  dplyr::select(MonitoringLocationName, TADA.MonitoringLocationIdentifier, ATTAINS.assessmentunitidentifier, ATTAINS.assessmentunitname, TADA.CharacteristicName) %>%
  dplyr::distinct()
```


Filter data for specified assessment unit
```{r}
unique(WQP_Shepherdstown_subset$ATTAINS.assessmentunitname)

WQP_Shepherdstown_subset_AUs <- WQP_Shepherdstown_subset %>%
  dplyr::filter(ATTAINS.assessmentunitname == "Shenandoah River"| ATTAINS.assessmentunitname == "Rattlesnake Run")

unique(WQP_Shepherdstown_subset_AUs$ATTAINS.assessmentunitidentifier)
```

## Generate maps

```{r TADA Overview Map}
TADA_OverviewMap(WQP_Shepherdstown_subset_AUs)
```

Check coordinates
```{r}
WQP_Shepherdstown_subset_AUs = TADA_FlagCoordinates(WQP_Shepherdstown_subset_AUs, clean_outsideUSA = "change sign", clean_imprecise = FALSE)

WQP_Shepherdstown_subset_AUs <- WQP_Shepherdstown_subset_AUs[!is.na(WQP_Shepherdstown_subset_AUs$LongitudeMeasure),]
WQP_Shepherdstown_subset_AUs <- WQP_Shepherdstown_subset_AUs[!is.na(WQP_Shepherdstown_subset_AUs$LatitudeMeasure),]
```

Remove extra cols
```{r}
WQP_Shepherdstown_subset_AUs = TADA_RetainRequired(WQP_Shepherdstown_subset_AUs)
```

## Exploratory visualizations

```{r TADA_GetATTAINS}
# ATTAINS_Shepherdstown <- TADA_GetATTAINS(WQP_Shepherdstown_subset_AUs, return_sf = TRUE, return_nearest = TRUE)

```

```{r TADA_ViewATTAINS}
# # View catchments and assessment units on map
# 
# ATTAINS_map <- TADA_ViewATTAINS(WQP_Shepherdstown_subset_AUs)
# 
# ATTAINS_map
```

review unique identifiers
```{r}
unique(WQP_Shepherdstown_subset_AUs$TADA.ComparableDataIdentifier)

```
Generate stats table

```{r stats}
WQP_Shepherdstown_stats <- TADA_Stats(WQP_Shepherdstown_subset_AUs)
```

Generate scatterplot

```{r scatterplot}
TADA_TwoCharacteristicScatterplot(WQP_Shepherdstown_subset_AUs, id_cols = "TADA.ComparableDataIdentifier",  groups = c("SPECIFIC CONDUCTANCE_TOTAL_NA_US/CM @25C", "CHLORIDE_TOTAL_NA_UG/L"))
```

```{r ph analysis and visualization}
# comparison to standard for ph
pH_Standard <- WQP_Shepherdstown_subset_AUs %>%
  dplyr::filter(TADA.ComparableDataIdentifier == "PH_NA_NA_STD UNITS") %>%
  dplyr::mutate(MeetsStandard = ifelse(TADA.ResultMeasureValue >= 6.5 & TADA.ResultMeasureValue <= 9, "Yes", "No"))


# subset of pH data with fewer rows
pH_Table <- pH_Standard %>%
  dplyr::select(
    MonitoringLocationIdentifier, OrganizationFormalName, ActivityStartDate, TADA.ResultMeasureValue,
    MeetsStandard
  )

DT::datatable(pH_Table, fillContainer = TRUE)

# pH scatterplot
pH_Scatter <- TADA_Scatterplot(pH_Standard, id_cols = "TADA.ComparableDataIdentifier") %>%
  plotly::add_lines(
    y = 6.5,
    x = c(min(pH_Standard$ActivityStartDate), max(pH_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  ) %>%
  plotly::add_lines(
    y = 8.5,
    x = c(min(pH_Standard$ActivityStartDate), max(pH_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  )

pH_Scatter
```

SPECIFIC CONDUCTANCE_TOTAL_NA_US/CM @25C


"PH", \# 6.5-8.5 but dependent on application

"SPECIFIC CONDUCTANCE", \# 250000 µg/l Human Health (Human Health •
Water + Organism)

"IRON", \# 1000 µg/l Aquatic Life, Freshwater, Chronic OR 300 µg/l
Organoleptic

"CHLORIDE", \# 860000 µg/l (Aquatic Life • Freshwater • Acute) OR 230000
µg/l (Aquatic Life • Freshwater • Chronic)

"FECAL COLIFORM",

"ESCHERICHIA COLI" ))

```{r Iron analysis and visualization}
# comparison to standard for iron
Iron_Standard <- WQP_Shepherdstown_subset_AUs %>%
  dplyr::filter(TADA.ComparableDataIdentifier == "IRON_TOTAL_NA_UG/L") %>%
  dplyr::mutate(MeetsStandard = ifelse(TADA.ResultMeasureValue <= 300, "Yes", "No"))


# subset of iron data with fewer rows
Iron_Table <- Iron_Standard %>%
  dplyr::select(
    MonitoringLocationIdentifier, OrganizationFormalName, ActivityStartDate, TADA.ResultMeasureValue,
    MeetsStandard
  )

DT::datatable(Iron_Table, fillContainer = TRUE)

# iron scatterplot
Iron_Scatter <- TADA_Scatterplot(Iron_Standard, id_cols = "TADA.ComparableDataIdentifier") %>%
  plotly::add_lines(
    y = 600,
    x = c(min(Iron_Standard$ActivityStartDate), max(Iron_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  ) %>%
  plotly::add_lines(
    y = 8.5,
    x = c(min(Iron_Standard$ActivityStartDate), max(Iron_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  )

Iron_Scatter
```

```{r boxplot and histogram}
# filter dataframe to comparable data identifier of interest

Nitrogen_data <- dplyr::filter(Filtered_data, TADA.ComparableDataIdentifier == "TOTAL NITROGEN, MIXED FORMS_UNFILTERED_AS N_MG/L")

# generate a histogram
Nitrogen_Histogram <- TADA_Histogram(Nitrogen_data, id_cols = "TADA.ComparableDataIdentifier")

# view histogram
Nitrogen_Histogram
```

`TADA_Boxplot` can be useful for identifying skewness and percentiles.

```{r boxplot, fig.width=8, fig.height=6, fig.fullwidth=TRUE}
Nitrogen_Boxplot <- TADA_Boxplot(Nitrogen_data, id_cols = "TADA.ComparableDataIdentifier")


Nitrogen_Boxplot
```

Filtering the results of `TADA_Stats` for only "TOTAL NITROGEN, MIXED
FORMS_TOTAL RECOVERABLE_NA_MG/L" and selecting a smaller subset of
columns or creating a single characteristic scatterplot may also provide
useful information.

```{r stats and scatterplot}
# create table with nitrogen stats
Nitrogen_stats <- Filtered_data_stats %>%
  dplyr::filter(TADA.ComparableDataIdentifier == "TOTAL NITROGEN, MIXED FORMS_UNFILTERED_AS N_MG/L") %>%
  dplyr::select(Location_Count, Measurement_Count, Min, Max, Mean)

DT::datatable(Nitrogen_stats, fillContainer = TRUE)

# create nitrogen scatterplot
Nitrogen_Scatterplot <- TADA_Scatterplot(Nitrogen_data, id_cols = "TADA.ComparableDataIdentifier")

Nitrogen_Scatterplot
```

Generate a new dataframe with only the daily max dissolved oxygen
values.

```{r DO_filtered}
R5ProfileCleanDO_dailymax <- TADA_AggregateMeasurements(R5ProfileCleanDO,
  agg_fun = c("max"),
  clean = TRUE
)
```

We can also explore depth profiles for selected characteristics at
specific site on a single date. There are a few functions that can help
with this. First we can use `TADA_FlagDepthCategory` to place results
into various depth categories (surface, middle, and bottom).

```{r depth category}
dataset_depth <- TADA_FlagDepthCategory(dataset_cens)
```

We can also use another function, `TADA_IDDepthProfiles` to identify
location/date/characteristic combinations in the data set that can be
used for depth profile plots or analysis. The default number of values
required to identify a location/date/characteristic as a depth profile
is 2, but this can be changed by the user. We will specify a larger
value, 5, so that any depth profiles identified will have results from
at least 5 different depths.

```{r id depth profile}
depth_profile_id <- TADA_IDDepthProfiles(dataset_depth, nvalue = 5)
```

**Question 11: How can TADA_IDDepthProfiles() help users use
TADA_DepthProfilePlot most efficiently?**

Now, we can use `TADA_DepthProfilePlot` to plot up to three
characteristics against depth. In this example, we will look at pH,
secchi depth, and pH.

```{r depth plot}
TADA_DepthProfilePlot(dataset_cens,
  groups = c(
    "TEMPERATURE, WATER_NA_NA_DEG C",
    "DEPTH, SECCHI DISK DEPTH_NA_NA_M",
    "PH_NA_NA_NONE"
  ),
  location = "REDLAKE_WQX-ANKE",
  activity_date = "2018-10-04",
  depthcat = TRUE,
  surfacevalue = 2,
  bottomvalue = 2,
  unit = "m"
)
```

## Record end time

```{r analysis time}
end.time <- Sys.time()

end.time - start.time
```
