---
title: 'Beginner R user: introduction to using the EPATADA R package to retrieve, clean, analyze, and visualize WQP data'
author: "TADA Team"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = F)
```

## Install and load

First, install and load the remotes package specifying the repo. This is
needed before installing the *EPATADA* R package because it is only
available on GitHub.

```{r install_remotes, results = 'hide', eval = F}
install.packages("remotes", repos = "http://cran.us.r-project.org")
library(remotes)
```

Next, install (or update) and load the *EPATADA* R package using the
*remotes* R package. Additional dependency R packages that are used
within *EPATADA* will be downloaded automatically. You may be prompted
in the console to update dependency packages that have more recent
versions available. If you see this prompt, it is recommended to update
all of them (enter 1 into the console). Our team is actively developing
*EPATADA*, therefore we highly recommend that you update the package
(and all of its dependencies) each time you use it.

```{r install_EPATADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA", ref = "develop", dependencies = TRUE)
```

```{r install_EPATADA_dev, eval = T, include = F}
remotes::install_github("USEPA/EPATADA", ref = "cybertown2025", dependencies = TRUE
)
```

## Record start time

```{r}
start.time <- Sys.time()
```

## Retrieve and autoclean

Query the WQP using a bounding box (lat/long extent), start date, and
end date. TADA_AutoClean is a powerful function that runs as part of
TADA_DataRetrieval when applyautoclean = TRUE. It performs a variety of
tasks, for example:

1.  creating new "TADA" prefixed columns and and capitalizing their
    contents to reduce case sensitivity issues,

2.  converts special characters in value columns,

3.  converts latitude and longitude values to numeric,

4.  replaces "meters" with "m",

5.  replaces deprecated characteristic names with current WQX names,

6.  harmonizes result and detection limit units,

7.  converts depths to meters, and

8.  creates the column TADA.ComparableDataIdentifier by concatenating
    characteristic name, result sample fraction, method speciation, and
    result measure unit.

```{r TADA_DataRetrieval}
WQP_Shepherdstown_raw <- TADA_DataRetrieval(
  startDate = "2019-01-01",
  endDate = "2024-12-31",
  bBox = c(-78.2241,38.9185,-77.1041,39.7487),
  applyautoclean = TRUE,
  ask = FALSE)
```

## Filter and clean

Now, let's use EPATADA functions to review, visualize, and whittle the
returned WQP data down to include only results that are applicable to
our water quality analysis and area of interest.

The **TADA_AnalysisDataFilter** function can assist in identifying and
filtering surface water, groundwater, and sediment results. If you set
clean = FALSE, this function will categorize and flag (but not remove)
rows in a new *TADA.UseForAnalysis.Flag* column for review. However, the
default functionality (clean = TRUE) is to include surface water and
exclude groundwater and sediment results. This is shown in the example
below.

```{r}
WQP_Shepherdstown_clean <- TADA_AnalysisDataFilter(
  WQP_Shepherdstown_raw,
  clean = TRUE,
  surface_water = TRUE,
  ground_water = FALSE,
  sediment = FALSE
)
```

Next, let's check if the dataset contains potential duplicate results
from within a single organization or from within multiple organizations
(such as when two or more organizations monitor the same location and
may submit duplicate results).

If you would like to prioritize results from one organization over
another, this can be done using the org_hierarchy argument in
`TADA_FindPotentialDuplicatesMultipleOrgs`.

```{r}
# find duplicates from single org
WQP_Shepherdstown_clean <- TADA_FindPotentialDuplicatesSingleOrg(
  WQP_Shepherdstown_clean)

# find duplicates across multiple orgs
WQP_Shepherdstown_clean <- TADA_FindPotentialDuplicatesMultipleOrgs(
  WQP_Shepherdstown_clean, 
  org_hierarchy = "21VASWCB"
)
```

Let's review the duplicates:

```{r}
review_dups <- WQP_Shepherdstown_clean %>%
  dplyr::select(MonitoringLocationName, 
                TADA.MonitoringLocationIdentifier, 
                TADA.MultipleOrgDuplicate, 
                TADA.MultipleOrgDupGroupID, 
                TADA.ResultSelectedMultipleOrgs, 
                TADA.SingleOrgDupGroupID, 
                TADA.SingleOrgDup.Flag, 
                TADA.ComparableDataIdentifier, 
                ResultMeasureValue, 
                TADA.ResultMeasure.MeasureUnitCode, 
                TADA.MonitoringLocationName, 
                TADA.NearbySites.Flag, 
                TADA.NearbySiteGroup, 
                OrganizationIdentifier) %>%
  dplyr::distinct()
```

Remove intermediate variable in R by using 'rm()'

```{r}
rm(review_dups)
```

We will select to keep only unique samples from
`TADA_FindPotentialDuplicatesSingleOrg` by filtering for
TADA.SingleOrgDup.Flag equals "Unique". We can now review multiple
organization duplicates from `TADA_FindPotentialDuplicatesMultipleOrgs`
and filter only to retain only one value by filtering for
TADA.ResultSelectedMultipleOrgs equals "Y".

```{r}
WQP_Shepherdstown_clean <- dplyr::filter(WQP_Shepherdstown_clean, TADA.SingleOrgDup.Flag == "Unique")

WQP_Shepherdstown_clean <- dplyr::filter(
  WQP_Shepherdstown_clean,
  TADA.ResultSelectedMultipleOrgs == "Y"
)
```

Censored data are measurements for which the true value is not known,
but we can estimate the value based on known lower or upper detection
conditions and limit types. TADA fills missing *TADA.ResultMeasureValue*
and *TADA.ResultMeasure.MeasureUnitCode* values with values and units
from *TADA.DetectionQuantitationLimitMeasure.MeasureValue* and
*TADA.DetectionQuantitationLimitMeasure.MeasureUnitCode*, respectively,
using the `TADA_AutoClean` function.

The TADA package currently has functions that summarize censored data
incidence in the dataset and perform simple substitutions of censored
data values, including x times the detection limit and random selection
of a value between 0 and the detection limit. The user may specify the
methods used for non-detects and over-detects separately in the input to
the `TADA_SimpleCensoredMethods` function. The next step we take in this
example is to perform simple conversions to the censored data in the
dataset: we keep over-detects as is (no conversion made) and convert
non-detect values to 0.5 times the detection limit (half the detection
limit).

```{r}
WQP_Shepherdstown_clean <- TADA_SimpleCensoredMethods(
  WQP_Shepherdstown_clean,
  nd_method = "multiplier",
  nd_multiplier = 0.5,
  od_method = "as-is",
  od_multiplier = "null"
)
```

In this example, the metadata is conflicting (detection condition and
detection limit type do not match) for one result. Let's review this.

```{r}
# review unique flags
unique(WQP_Shepherdstown_clean$TADA.CensoredData.Flag)

# review row with conflicting information
conflict <- WQP_Shepherdstown_clean %>%
  dplyr::filter(TADA.CensoredData.Flag == "Conflict between Condition and Limit") %>%
  dplyr::select(c("DetectionQuantitationLimitTypeName", "ResultDetectionConditionText", "TADA.CensoredData.Flag"))

```

Now, since the information is conflicting, we will remove row.

```{r}
WQP_Shepherdstown_clean <- WQP_Shepherdstown_clean %>%
  dplyr::filter(TADA.CensoredData.Flag != "Conflict between Condition and Limit")
```

Remove intermediate variable in R by using 'rm()'.

```{r}
rm(conflict)
```

`TADA_AutoFilter` removes rows where the result value is not numeric to
prepare a dataframe for quantitative analyses. Specifically, this
function removes rows with "Text" and "NA - Not Available" in the
TADA.ResultMeasureValueDataTypes.Flag column, or NA in the
TADA.ResultMeasureValue column. In addition, this function removes
results with QA/QC ActivityTypeCode's. This function also removes any
columns not required for TADA workflow where all values are equal to NA.

```{r}
WQP_Shepherdstown_clean <- TADA_AutoFilter(WQP_Shepherdstown_clean)
```

TADA_RunKeyFlagFunctions is a shortcut function to run important TADA
flagging functions. See ?function documentation for TADA_FlagResultUnit,
TADA_FlagFraction, TADA_FindQCActivities, TADA_FlagMeasureQualifierCode,
and TADA_FlagSpeciation for more information.

```{r}
WQP_Shepherdstown_clean <- TADA_RunKeyFlagFunctions(
  WQP_Shepherdstown_clean,
  clean = TRUE
)
```

Let's filter the data to remove any remaining NAs.

```{r}
WQP_Shepherdstown_clean <- WQP_Shepherdstown_clean[!is.na(WQP_Shepherdstown_clean$TADA.ResultMeasure.MeasureUnitCode),]
```

Another set of TADA flagging functions, `TADA_FlagAboveThreshold` and
`TADA_FlagBelowThreshold`, can be used to check results against national
lower and upper thresholds. As in the previous examples, we will set
clean = TRUE for these functions to remove results that fall outside the
national thresholds.

```{r}
WQP_Shepherdstown_clean <- TADA_FlagAboveThreshold(WQP_Shepherdstown_clean, clean = TRUE, flaggedonly = FALSE)

WQP_Shepherdstown_clean <- TADA_FlagBelowThreshold(WQP_Shepherdstown_clean, clean = TRUE, flaggedonly = FALSE)
```

Let's take a look at all unique values in the
TADA.ComparableDataIdentifier column and see how how many results are
associated with each. TADA.ComparableDataIdentifier concatenates
TADA.CharacteristicName, TADA.ResultSampleFractionText,
TADA.MethodSpeciationName, and TADA.ResultMeasure.MeasureUnitCode.

```{r TADA_FieldValuesTable}
# use TADA_FieldValuesTable to create a table of the number of results per TADA.ComparableDataIdentifier
compid_table <- TADA_FieldValuesTable(WQP_Shepherdstown_clean, field = "TADA.ComparableDataIdentifier")

compid_before = unique(WQP_Shepherdstown_clean$TADA.ComparableDataIdentifier)

DT::datatable(compid_table, fillContainer = TRUE)
```

Remove intermediate variable in R by using 'rm()'.

```{r}
rm(compid_table)
```

As we scroll through the table, we can see that there are some repeated
characteristic names. It may be possible that some of these are synonyms
which can be automatically harmonized using `TADA_HarmonizeSynonyms` so
their results can be directly compared.

Let's give it a try.

```{r}
WQP_Shepherdstown_clean <- TADA_HarmonizeSynonyms(WQP_Shepherdstown_clean)
```

How many unique TADA.ComparableDataIdentifier's do we have now? We
reduced it by over 25!

```{r}
compid_after = unique(WQP_Shepherdstown_clean$TADA.ComparableDataIdentifier)
```

```{r}
rm(compid_before)
rm(compid_after)
```

`TADA_CalculateTotalNP` uses the [Nutrient Aggregation
logic](https://echo.epa.gov/trends/loading-tool/resources/nutrient-aggregation)
to add together specific subspecies to obtain a total. TADA adds one
more equation to the mix: total particulate nitrogen + total dissolved
nitrogen. The function uses as many subspecies as possible to calculate
a total for each given site, date, and depth group, but it will estimate
total nitrogen with whatever subspecies are present. This function
creates NEW total nutrient measurements (total nitrogen unfiltered as N
and total phosphorus unfiltered as P) and adds them to the dataframe.

Users can use the default summation worksheet (see
`TADA_GetNutrientSummationRef`) or customize it to suit their needs. The
function also requires a daily aggregation value, either minimum,
maximum, or mean. The default is 'max', which means that if multiple
measurements of the same subspecies-fraction-speciation-unit occur on
the same day at the same site and depth, the function will pick the
maximum value to use in summation calculations. In this example, we will
use the default summation worksheet to obtain total nitrogen.

```{r}
WQP_Shepherdstown_clean <- TADA_CalculateTotalNP(WQP_Shepherdstown_clean, daily_agg = "max")

```

What characteristics are available now?

```{r fieldValues_actmedia}
# Create pie chart for CharacteristicName
TADA_FieldValuesPie(WQP_Shepherdstown_clean, field = "TADA.CharacteristicName")
```

Let's filter the data and focus on a few characteristics of interest.

```{r}
# Select characteristics of interest
WQP_Shepherdstown_subset <- WQP_Shepherdstown_clean %>%
  dplyr::filter(TADA.CharacteristicName %in% c(
    "PH", "SPECIFIC CONDUCTANCE", "IRON", "CHLORIDE", "FECAL COLIFORM", "ESCHERICHIA COLI", "TOTAL NITROGEN"))
```

## Monitoring Location Filter and Review

In this section, we will associate geospatial data from **ATTAINS** with
the **WQP** data, and filter the dataset to retain only results that
were collected in specified Assessment Units. We can also generate a new
table to give us some information about the individual monitoring
locations within the assessment units.

-   TADA_GetATTAINS() automates matching of WQP monitoring locations
    with ATTAINS assessment units that fall within (intersect) the same
    NHDPlus catchment
    ([details](https://usepa.github.io/EPATADA/articles/TADAModule2.html))
-   The function uses high resolution NHDPlus catchments by default
    because 80% of state submitted assessment units in ATTAINS were
    developed based on high res NHD; users can select med-res if
    applicable to their use case

```{r Data Retrieval - Geospatial}
WQP_Shepherdstown_subset_spatial <- TADA_GetATTAINS(
  WQP_Shepherdstown_subset, 
  fill_catchments = FALSE,
  return_sf = TRUE, 
  return_nearest = TRUE)

# pull TADA df with ATTAINS info added out of the list
Subset_TADAwithATTAINS <- WQP_Shepherdstown_subset_spatial$TADA_with_ATTAINS
```

Create table of monitoring location identifiers and AUs.

```{r}
ML_AU_crosswalk <- Subset_TADAwithATTAINS %>%
  dplyr::select(MonitoringLocationName, TADA.MonitoringLocationIdentifier, ATTAINS.assessmentunitidentifier, ATTAINS.assessmentunitname, TADA.CharacteristicName) %>%
  dplyr::distinct()
```

Review unique assessment unit names.

```{r}
unique(Subset_TADAwithATTAINS$ATTAINS.assessmentunitname)
```

Filter data for specified assessment unit.

```{r}
Subset_TADAwithATTAINS_selectAUs <- Subset_TADAwithATTAINS %>%
  dplyr::filter(
    ATTAINS.assessmentunitname == "Shenandoah River"|
      ATTAINS.assessmentunitname == "Rattlesnake Run")
```

Review unique Assessments Units.

```{r}
unique(Subset_TADAwithATTAINS_selectAUs$ATTAINS.assessmentunitidentifier)
```

## Generate maps

```{r TADA Overview Map}
TADA_OverviewMap(Subset_TADAwithATTAINS_selectAUs)
```

Check coordinates

```{r}
Subset_TADAwithATTAINS_selectAUs = TADA_FlagCoordinates(Subset_TADAwithATTAINS_selectAUs, clean_outsideUSA = "change sign", clean_imprecise = FALSE)

Subset_TADAwithATTAINS_selectAUs <- Subset_TADAwithATTAINS_selectAUs[!is.na(Subset_TADAwithATTAINS_selectAUs$LongitudeMeasure),]
Subset_TADAwithATTAINS_selectAUs <- Subset_TADAwithATTAINS_selectAUs[!is.na(Subset_TADAwithATTAINS_selectAUs$LatitudeMeasure),]
```

`TADA_RetainRequired` removes all duplicate columns where TADA has
created a new column with a TADA prefix. It retains all TADA prefixed
columns as well as other original fields that are either required by
other TADA functions or are commonly used filters.

```{r}
Subset_TADAwithATTAINS_selectAUs = TADA_RetainRequired(Subset_TADAwithATTAINS_selectAUs)
```

## Exploratory visualizations

```{r TADA_ViewATTAINS}
# not working

# # View catchments and assessment units on map
# 
# ATTAINS_map <- TADA_ViewATTAINS(Subset_TADAwithATTAINS_selectAUs)
# 
# ATTAINS_map
```

Review unique TADA.ComparableDataIdentifier's

```{r}
# throws error, why?
# compid_table_2 <- TADA_FieldValuesTable(Subset_TADAwithATTAINS_selectAUs, field = "TADA.ComparableDataIdentifier")
# 
# DT::datatable(compid_table_2, fillContainer = TRUE)

unique(Subset_TADAwithATTAINS_selectAUs$TADA.ComparableDataIdentifier)

#these need more harmonization...
```

Generate stats table

```{r stats}
WQP_Shepherdstown_stats <- TADA_Stats(Subset_TADAwithATTAINS_selectAUs)

# thought we already removed remaining NAs, why are there more now?
```

Generate scatterplot

```{r scatterplot}
TADA_TwoCharacteristicScatterplot(Subset_TADAwithATTAINS_selectAUs, id_cols = "TADA.ComparableDataIdentifier",  groups = c("IRON_TOTAL_NA_UG/L", "IRON_DISSOLVED_NA_UG/L"))
```

SPECIFIC CONDUCTANCE_TOTAL_NA_US/CM @25C

"PH", \# 6.5-8.5 but dependent on application

"SPECIFIC CONDUCTANCE", \# 250000 µg/l Human Health (Human Health •
Water + Organism)

"IRON", \# 1000 µg/l Aquatic Life, Freshwater, Chronic OR 300 µg/l
Organoleptic

"CHLORIDE", \# 860000 µg/l (Aquatic Life • Freshwater • Acute) OR 230000
µg/l (Aquatic Life • Freshwater • Chronic)

"FECAL COLIFORM",

"ESCHERICHIA COLI" ))

```{r Iron analysis and visualization}
# comparison to standard for iron
Iron_Standard <- Subset_TADAwithATTAINS_selectAUs %>%
  dplyr::filter(TADA.ComparableDataIdentifier == "IRON_TOTAL_NA_UG/L") %>%
  dplyr::mutate(MeetsStandard = ifelse(TADA.ResultMeasureValue <= 300, "Yes", "No"))


# subset of iron data with fewer rows
Iron_Table <- Iron_Standard %>%
  dplyr::select(
    MonitoringLocationIdentifier, OrganizationFormalName, ActivityStartDate, TADA.ResultMeasureValue,
    MeetsStandard
  )

DT::datatable(Iron_Table, fillContainer = TRUE)

# iron scatterplot
Iron_Scatter <- TADA_Scatterplot(Iron_Standard, id_cols = "TADA.ComparableDataIdentifier") %>%
  plotly::add_lines(
    y = 600,
    x = c(min(Iron_Standard$ActivityStartDate), max(Iron_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  ) %>%
  plotly::add_lines(
    y = 8.5,
    x = c(min(Iron_Standard$ActivityStartDate), max(Iron_Standard$ActivityStartDate)),
    inherit = FALSE,
    showlegend = FALSE,
    line = list(color = "red"),
    hoverinfo = "none"
  )

Iron_Scatter
```

```{r boxplot and histogram}
# filter dataframe to comparable data identifier of interest

Iron_data <- dplyr::filter(Subset_TADAwithATTAINS_selectAUs, TADA.ComparableDataIdentifier == c("IRON_TOTAL_NA_UG/L", "IRON_DISSOLVED_NA_UG/L"))

# generate a histogram
Iron_Histogram <- TADA_Histogram(Iron_data, id_cols = "TADA.CharacteristicName")

# view histogram
Iron_Histogram
```

`TADA_Boxplot` can be useful for identifying skewness and percentiles.

```{r boxplot, fig.width=8, fig.height=6, fig.fullwidth=TRUE}
Iron_Boxplot <- TADA_Boxplot(Iron_data, id_cols = "TADA.CharacteristicName")


Iron_Boxplot
```

```{r TADA_AggregateMeasurements}
TADA_AggregateMeasurements(x,
  agg_fun = c("max"),
  clean = TRUE
)
```

We can also explore depth profiles for selected characteristics at
specific site on a single date. There are a few functions that can help
with this. First we can use `TADA_FlagDepthCategory` to place results
into various depth categories (surface, middle, and bottom).

```{r TADA_FlagDepthCategory}
x <- TADA_FlagDepthCategory(x)
```

We can also use another function, `TADA_IDDepthProfiles` to identify
location/date/characteristic combinations in the data set that can be
used for depth profile plots or analysis. The default number of values
required to identify a location/date/characteristic as a depth profile
is 2, but this can be changed by the user. We will specify a larger
value, 5, so that any depth profiles identified will have results from
at least 5 different depths.

```{r TADA_IDDepthProfiles}
x <- TADA_IDDepthProfiles(x, nvalue = 5)
```

**Question 11: How can TADA_IDDepthProfiles() help users use
TADA_DepthProfilePlot most efficiently?**

Now, we can use `TADA_DepthProfilePlot` to plot up to three
characteristics against depth. In this example, we will look at pH,
secchi depth, and pH.

```{r depth plot}
TADA_DepthProfilePlot(x,
  groups = c(
    "x"
  ),
  location = "x",
  activity_date = "x",
  depthcat = TRUE,
  surfacevalue = 2,
  bottomvalue = 2,
  unit = "m"
)
```

Show next step of crosswalking results to prepare for ATTAINS submission

```{r}
# create TADA parameter reference table for specified organization
x <- TADA_CreateParamRef(
  x,
  org_id = c("x"), 
  excel = FALSE
  # uncomment excel = TRUE, overwrite = TRUE to run the excel file
  # excel = TRUE, overwrite = TRUE
)

TADA_CreateUseParamRef()
```

## Record end time

```{r analysis time}
end.time <- Sys.time()

end.time - start.time
```

Reproducible and Documented

This workflow is reproducible and the decisions at each step are well
documented. This means that it is easy to go back and review every step,
understand the decisions that were made, make changes as necessary, and
run it again.
