---
title: "WQP Data Harmonization"
author: "Cristina A. Mullin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WQP Data Harmonization}
  %\usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette will walk through how to discover, wrangle, and harmonize
[Water Quality Portal (WQP)](https://www.waterqualitydata.us/) data from
multiple organizations.

## Install and load packages

To install TADA, currently you need to install from GitHub using remotes
(shown) or devtools. dataRetrieval will be downloaded from CRAN, but the
development version can be downloaded directly from GitHub (un-comment).
The following code will also install any packages you do not have, and
load all packages required to run this vignette into your R session.

```{r, results = 'hide', message = FALSE, warning = FALSE}
list.of.packages <- c("plyr", "data.table", "dataRetrieval", "dplyr", "ggplot2", "grDevices", "magrittr", "stringr", "utils", "RColorBrewer", "stats", "lubridate", "remotes", "rlang")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

Load the remotes library before installing TADA or dataRetrieval from
GitHub

```{r, results = 'hide', message = FALSE, warning = FALSE}
# If you have any issues loading the remotes library, uncomment the line below to install the "remotes" package specifying the repo
# install.packages("remotes", repos = "http://cran.us.r-project.org")
library(remotes)
```

Uncomment the lines below to install latest version of TADA and
dataRetrieval from GitHub.

```{r, results = 'hide', message = FALSE, warning = FALSE}
# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
remotes::install_github("USEPA/TADA", dependencies=TRUE)
```

Load the required libraries to run this vignette into your R session

```{r, results = 'hide', message = FALSE, warning = FALSE}
library(plyr)
library(data.table)
library(dplyr)
library(ggplot2)
library(grDevices)
library(magrittr)
library(stringr)
library(utils)
library(RColorBrewer)
library(stats)
library(lubridate)
library(rlang)
library(dataRetrieval)
library(TADA)
```

## Retrieve WQP data

WQP data is retrieved and processed for compatibility with TADA. This
function, TADAdataRetrieval builds on the USGS dataRetrieval package
functions. It joins three WQP profiles (i.e., the station, narrow, and
phys/chem), changes all data in the Characteristic, Speciation,
Fraction, and Unit fields to uppercase, removes true duplicates, removes
data for all non-water media types, and cleans up results with special
characters.

This function uses the same inputs as the dataRetrieval `readWQPdata`
function. `readWQPdata` does not restrict the characteristics pulled
from [Water Quality Portal (WQP)](https://www.waterqualitydata.us/). You
may specify the desired characteristics by using, for instance:
characteristicName = "pH".

Data retrieval filters include:

-   statecode

-   endDate

-   startDate

-   countycode

-   siteid

-   siteType

-   characteristicName

-   ActivityMediaName

Please be aware that this TADAdataRetrieval function automatically runs
the TADA autoclean and MeasureValueSpecialCharacters functions as well,
which are both required for subsequent functions within the TADA R
package to run. These functions alter and/or add the following WQP
columns (enter *?MeasureValueSpecialCharacters* and *?autoclean* into
the console for more details):

-   Alters (e.g., ResultMeasureValue and
    DetectionQuantitationLimitMeasure.MeasureValue fields are converted
    to class numeric)

    -   ResultMeasureValue

    -   DetectionQuantitationLimitMeasure.MeasureValue

-   Adds (All the data cleaning and transformations are done directly to
    the "ResultMeasureValue" and "DetectionLimitMeasureValue" columns,
    however the original "ResultMeasureValue" and
    "DetectionLimitMeasureValue" columns and values from the WQP are
    preserved in these new fields, "ResultMeasureValue.Original" and
    "DetectionLimitMeasureValue.Original". Additionally,
    "TADA.ResultMeasureValue.Flag" and
    "TADA.DetectionLimitMeasureValue.Flag" are created to track and
    changes made to the "ResultMeasureValue" and
    "DetectionLimitMeasureValue" columns; and to provide information
    about the result values that is needed to address censored data
    later on (i.e., nondetections). Specifically, these new columns flag
    if special characters that are included in result values, and
    specifies what the special characters are.

    -   ResultMeasureValue.Original

    -   TADA.ResultMeasureValue.Flag

    -   DetectionLimitMeasureValue.Original

    -   TADA.DetectionLimitMeasureValue.Flag

Downloads using TADAdataRetrieval will have the same columns each time,
but be aware that data are uploaded to the Water Quality Portal by
individual organizations, which may or may not follow the same
conventions. Data and metadata quality are not guaranteed! Make sure to
carefully explore any data and make conservative quality assurance
decisions where information is limited.

Tips:

1.  All the query filters for the WQP work as an AND but within the
    fields there are ORs. For example:

    -   Characteristics: If you choose pH & DO - it's an OR. This means
        you will retrieve both pH OR DO data if available.

    -   States: Similarly, if you choose VA and IL, it's an OR. This
        means you will retrieve both VA OR IL data if available.

    -   Combinations of fields are ANDs, such as State/VA AND
        Characteristic/DO". This means you will receive all DO data
        available in VA.

    -   "Characteristic" and "Characteristic Type" also work as an AND.
        This means that the Characteristic must fall within the
        CharacteristicGroup if both filters are being used, if not you
        will get an error.

2.  The "siteid" is a general term WQP uses to describe both Site IDs
    from USGS databases and Monitoring Location Identifiers (from the
    Water Quality Portal). Each monitoring location in the Water Quality
    Portal (WQP) has a unique Monitoring Location Identifier, regardless
    of the database from which it derives. The Monitoring Location
    Identifier from the WQP is the concatenated Organization Identifier
    plus the Site ID number. Site IDs that only include a number are
    only unique identifiers for monitoring locations within USGS NWIS or
    EPA's WQX databases separately.

Additional resources:

-   Review function documentation by entering the following code into
    the console: ?TADAdataRetrieval

-   [Introduction to the dataRetrieval
    package](https://CRAN.R-project.org/package=dataRetrieval)

-   [General Data Import from Water Quality
    Portal](https://rdrr.io/cran/dataRetrieval/man/readWQPdata.html)

-   [Water Quality Portal Web Services
    Guide](https://www.waterqualitydata.us/webservices_documentation/)

-   [dataRetrieval Tutorial](https://owi.usgs.gov/R/dataRetrieval.html)

Option 1: Use the TADAdataRetrieval function.

```{r}
#You can edit this to define your own WQP query inputs below
TADAProfile <- TADAdataRetrieval(statecode = "UT", characteristicName = c("Ammonia", "Nitrate", "Nitrogen"), startDate = "01-01-2021")

```

Option 2: Alternatively, you can use the data.table::fread function to
read in a web service call for any WQP profile (un-comment).

```{r}
# New_Draft_fullphyschem <- data.table::fread("https://www.waterqualitydata.us/data/Result/search?countrycode=US&statecode=US%3A49&siteid=UTAHDWQ_WQX-4925610&startDateLo=01-01-2015&startDateHi=12-31-2016&mimeType=csv&zip=no&sorted=yes&dataProfile=fullPhysChem&providers=NWIS&providers=STEWARDS&providers=STORET")
```

Option 3: If you need to download a large amount of data from across a
large area, and the TADAdataRetrieval function is not working due to WQP
timeout issues, then the TADABigdataRetrieval function may work better.

This function does multiple synchronous data calls to the WQP
(waterqualitydata.us). It uses the WQP summary service to limit the
amount downloaded to only relevant data, and pulls back data from 100
stations at a time and then joins the data back together and produces a
single TADA compatible dataframe as the output. For large data sets,
that can save a lot of time and ultimately reduce the complexity of
subsequent data processing. Using this function, you will be able to
download all data available from all sites in the contiguous United
States that is available for the time period, characteristicName, and
siteType requested.

See ?TADABigdataRetrieval for more details. WARNING, this can take
multiple hours to run. The total run time depends on your query inputs.

```{r}
#AllWaterTempData <- TADABigdataRetrieval(startDate = "2019-01-01", endDate = "2021-12-31", characteristicName = "Temperature, water", siteType = "Stream")
```

Review all column names in the TADA Profile

```{r}
colnames(TADAProfile)
```

## Depth unit conversions

Converts depth units to a consistent unit.

**ActivityDepthHeightMeasure.MeasureValue** provides depth information.
This is a crucial column for lake data but less often for river data.

Function checks dataset for depth profile data. Where depth profile
columns are populated, the function appends 'Conversion Factor' columns
and populates those columns based on the original unit (MeasureUnitCode
columns) and the target unit, which is defined in the 'unit' argument. A
'Depth Target Unit' column is also appended, indicating the unit all
selected depth data is converted to. When transform = FALSE, the output
includes all 'Conversion Factor' columns and the 'Depth Target Unit'
column. When transform = TRUE, the output includes converted depth data
and the 'Depth Target Unit' column, which acts as a flag indicating
which rows have been converted. Default is transform = TRUE.

The depth profile function can harmonize the depth units across all the
following fields (or only a specific one): "ActivityDepthHeightMeasure",
"ActivityTopDepthHeightMeasure", "ActivityBottomDepthHeightMeasure",
"ResultDepthHeightMeasure"). The default is all.

Allowable values for 'unit' are either 'm' (meter), 'ft' (feet), or 'in'
(inch). 'unit' accepts only one allowable value as an input. Default is
unit = "m".

See additional function documentation for additional function options by
entering the following code in the console: ?DepthProfileData

```{r}
#converts all depth profile data to meters
TADAProfileClean1 <- DepthProfileData(TADAProfile, unit = "m", transform = TRUE)
```

## Result unit conversions

Converts all results to WQX target units. WQX target units are pulled
from the MeasureUnit domain table:

-   <https://cdx.epa.gov/wqx/download/DomainValues/MeasureUnit.CSV>

See additional function documentation for additional function options by
entering the following code in the console: ?WQXTargetUnits

```{r}
#Converts all results to WQX target units
TADAProfileClean2 <- WQXTargetUnits(TADAProfileClean1, transform = TRUE)
```

## Statistically aggregated data

Checks for and removes statistically aggregated high frequency (i.e.,
continuous) data, if present.

The Water Quality Portal (WQP) is not designed to store high-frequency
sensor data. However, sometimes data providers choose to aggregate their
continuous data and submit it to WQP as one value. This type of data may
not be suitable for integration with discrete water quality data for
assessments. Therefore, this function uses metadata submitted by data
providers to flags rows with aggregated continuous data.

-   This is done by flagging results where the
    ResultDetectionConditionText = "Reported in Raw Data (attached)"

    -   When clean = TRUE, rows with aggregated continuous data are
        removed from the dataset and no column will be appended

    -   Default is clean = TRUE

-   See function documentation for additional function options by
    entering the following code in the console: ?DepthProfileData

```{r}
TADAProfileClean3 <- AggregatedContinuousData(TADAProfileClean2, clean = TRUE)
```

## WQX QAQC Service Result Flags

Run the following result functions to address invalid method, fraction,
speciation, and unit metadata by characteristic. The default is clean =
TRUE, which will remove invalid results. You can change this to clean =
FALSE to flag results, but not remove them.

See documentation for more details:

-   ?InvalidMethod

    -   When Clean = FALSE, this function adds the following column to
        your dataframe: WQX.AnalyticalMethodValidity. This column flags
        invalid CharacteristicName,
        ResultAnalyticalMethod/MethodIdentifier, and
        ResultAnalyticalMethod/MethodIdentifierContext combinations in
        your dataset either "Nonstandardized", "Invalid", or "Valid".

    -   When clean = TRUE, "Invalid" rows are removed from the dataset
        and no column will be appended.

-   ?InvalidSpeciation

    -   When clean = FALSE, this function adds the following column to
        your dataframe: WQX.MethodSpeciationValidity. This column flags
        each CharacteristicName and MethodSpecificationName combination
        in your dataset as either "Nonstandardized", "Invalid", or
        "Valid".

    -   When clean = TRUE, "Invalid" rows are removed from the dataset
        and no column will be appended.

-   ?InvalidResultUnit

    -   When clean = FALSE, the following column will be added to your
        dataset: WQX.ResultUnitValidity. This column flags each
        CharacteristicName, ActivityMediaName, and
        ResultMeasure/MeasureUnitCode combination in your dataset as
        either "Nonstandardized", "Invalid", or "Valid".

    -   When clean = TRUE, "Invalid" rows are removed from the dataset
        and no column will be appended.

-   ?InvalidFraction

    -   When clean = FALSE, this function adds the following column to
        your dataframe: WQX.SampleFractionValidity. This column flags
        each CharacteristicName and ResultSampleFractionText combination
        in your dataset as either "Nonstandardized", "Invalid", or
        "Valid".
    -   When clean = TRUE, "Invalid" rows are removed from the dataset
        and no column will be appended.

```{r}
TADAProfileClean4 <- InvalidMethod(TADAProfileClean3, clean = TRUE)
TADAProfileClean5 <- InvalidFraction(TADAProfileClean4, clean = TRUE)
TADAProfileClean6 <- InvalidSpeciation(TADAProfileClean5, clean = FALSE)
TADAProfileClean7 <- InvalidResultUnit(TADAProfileClean6, clean = FALSE)
```

## WQX national upper and lower thresholds

Run the following code to flag or remove results that are above or below
the national upper and lower bound for each characteristic and unit
combination. See documentation for more details:

-   ?AboveNationalWQXUpperThreshold

    -   When clean = FALSE, the following column is added to your
        dataset: AboveWQXUpperThreshold. This column flags rows with
        data that are above the upper WQX threshold.

    -   When clean = TRUE, data that is above the upper WQX threshold is
        removed from the dataset.

-   ?BelowNationalWQXUpperThreshold

    -   When clean = FALSE, the following column is added to your
        dataset: BelowWQXUpperThreshold. This column flags rows with
        data that are below the lower WQX threshold.

    -   When clean = TRUE, data that is below the lower WQX threshold is
        removed from the dataset.

The default is clean=TRUE, but you can change this to only flag results
if desired. Results will be flagged, but not removed, when clean=FALSE.

```{r}
TADAProfileClean8 <- AboveNationalWQXUpperThreshold(TADAProfileClean7, clean = TRUE)
TADAProfileClean9 <- BelowNationalWQXUpperThreshold(TADAProfileClean8, clean = TRUE)
```

## Potential duplicates

Sometimes multiple organizations submit the exact same data to Water
Quality Portal (WQP), which can affect water quality analyses and
assessments. This function checks for and identifies data that is
identical in all fields excluding organization-specific and comment text
fields. Each pair or group of potential duplicate rows is flagged with a
unique ID. For more information, review the documentation by entering
the following into the console:

-   ?PotentialDuplicateRowID

    -   When clean = FALSE, the following column will be added to you
        dataframe: TADA.PotentialDupRowID. This column flags potential
        duplicate rows of data in your dataset, and assigns each
        potential duplicate combination a unique number linking the two
        potential duplication rows.

    -   When clean = FALSE the first of each group of potential
        duplicate rows will be removed from the dataset and no column is
        appended.

When clean = TRUE, the function retains the first occurrence of each
potential duplicate in the dataset. Default is clean = TRUE.

```{r}
TADAProfileClean10 <- PotentialDuplicateRowID(TADAProfileClean9)
```

## Invalid coordinates

Function identifies and flags invalid coordinate data.

When clean_outsideUSA = FALSE and clean_imprecise = FALSE, a column will
be appended titled "TADA.InvalidCoordinates" with the following flags
(if relevant to dataset).

-   If the latitude is less than zero, the row will be flagged with
    "LAT_OutsideUSA".

-   If the longitude is greater than zero AND less than 145, the row
    will be flagged as "LONG_OutsideUSA".

-   If the latitude or longitude contains the string, "999", the row
    will be flagged as invalid.

-   Finally, precision can be measured by the number of decimal places
    in the latitude and longitude provided. If either does not have any
    numbers to the right of the decimal point, the row will be flagged
    as "Imprecise".

```{r}
TADAProfileClean11 <- InvalidCoordinates(TADAProfileClean10, clean_outsideUSA = FALSE, clean_imprecise = FALSE)
```

## Review QAPP information

Check data for an approved QAPP

This function checks to see if there is any information in the column
"QAPPApprovedIndicator". Some organizations submit data for this field
to indicate if the data produced has an approved Quality Assurance
Project Plan (QAPP) or not. In this field, Y indicates yes, N indicates
no.

This function has two default inputs: clean = TRUE and cleanNA = FALSE.
These defaults remove rows of data where the QAPPApprovedIndicator
equals "N".

Users could alternatively remove both N's and NA's using the inputs
clean = TRUE and cleanNA = TRUE.

If both clean = FALSE and cleanNA = FALSE, the function will not do
anything.

```{r}
TADAProfileClean12 <- QAPPapproved(TADAProfileClean11, clean = TRUE, cleanNA = FALSE)
```

Check to see if a QAPP Doc is Available

This function checks data submitted under the "ProjectFileUrl" column to
determine if a QAPP document is available to review. When clean = FALSE,
a column will be appended to flag results that do have an associated
QAPP document URL provided. When clean = TRUE, rows that do not have an
associated QAPP document are removed from the dataset and no column will
be appended. This function should only be used to remove data if an
accompanying QAPP document is required to use data in assessments.

```{r}
TADAProfileClean13 <- QAPPDocAvailable(TADAProfileClean12, clean = FALSE)
```

## Filter data by field

In this section a TADA user will want to review the unique values in
specific fields and may choose to remove data with particular values.

To start, review the list of fields and the number of unique values in
each field.

```{r}
FilterFields(TADAProfileClean13)
```

Next, choose a field from the list to see the unique values in that
field, as well as the number of times each value appears in the dataset.
We'll start with ActivityTypeCode.

Here is a list of other fields to review:

1.  **ResultCommentText** often has details relating to additional QA.
2.  **MeasureQualifierCode** Contains information about data flags 3.
    Other codes may designate suspect data or other flags which may be
    described in detail in **ResultLaboratoryCommentText** or another
    column

```{r, fig.width=6, fig.height=2, fig.fullwidth=TRUE}
FilterFieldReview("ActivityTypeCode", TADAProfileClean13)
```

The ActivityTypeCode field has four unique values -- "Sample-Routine",
"Quality Control Sample-Field Replicate", "Field Msr/Obs", and "Quality
Control Sample-Field Blank." In this example we want to remove quality
control values in the ActivityTypeCode field, therefore, we'll specify
that we want to remove the "Quality Control Sample-Field Replicate" and
"Quality Control Sample-Field Blank" values in the ActivityTypeCode
field.

```{r}
TADAProfileClean14 <- dplyr::filter(TADAProfileClean13, !(ActivityTypeCode %in% c("Quality Control Sample-Field Replicate", "Quality Control Sample-Field Blank", "Quality Control Sample-Lab Duplicate", "Quality Control Sample-Equipment Blank")))
```

We've completed our review of the ActivityTypeCode field.

Let's move on to a different field and see if there are any values that
we want to remove -- we'll look at the values in the
ResultStatusIdentifier field.

```{r, fig.width=6, fig.height=2, fig.fullwidth=TRUE}
FilterFieldReview("ActivityMediaSubdivisionName", TADAProfileClean14)
```

The ActivityMediaSubdivisionName field has two unique values, "Surface
Water" and "Groundwater." In this example we want to remove the
"Groundwater" values.

```{r}
TADAProfileClean15 <- dplyr::filter(TADAProfileClean14, !(ActivityMediaSubdivisionName %in% c("Groundwater", "Bulk deposition")))
```

## Filter data by field, subset by parameter

In this section a TADA user will want to select a parameter, review the
unique values associated with that parameter in specific fields, and
choose to remove particular values.

To start, review the list of parameters in the dataset. (The list is
sorted from highest to lowest counts. Only the first few rows are
displayed to save space on the page)

```{r}
FilterParList(TADAProfileClean15)
```

Next, select a parameter. Let's explore the fields associated with
Nitrogen:

```{r}
FilterParFields(TADAProfileClean15, "NITROGEN")
```

Selecting a parameter generates the list below, which is subset by the
selected parameter, of fields and the number of unique values in each
field.

Then choose a field from the list. In this example we'll remove certain
values from the HydrologicEvent field.

```{r, fig.width=6, fig.height=2, fig.fullwidth=TRUE}
FilterParFieldReview("HydrologicEvent", TADAProfileClean15, "NITROGEN")
```

The HydrologicEvent field has three unique values. In this example we
want to remove samples collected during "Storm" events. Therefore, we'll
specify that we want to remove rows where the CharacteristicName is
"NITROGEN" and the HydrologicEvent field is "Storm."

```{r}
TADAProfileClean16 <- dplyr::filter(TADAProfileClean15, !(CharacteristicName %in% "NITROGEN" & HydrologicEvent %in% "Storm"))
```

## Transform Characteristic, Speciation, and Unit values to TADA Standards

The HarmonizeRefTable function generates a harmonization reference table
that is specific to the input dataset. Users can review how their input
data relates to standard TADA values for the following elements:

-   CharacteristicName

-   ResultSampleFractionText

-   MethodSpecicationName

-   ResultMeasure.MeasureUnitCode

The HarmonizeData function then compares the input dataset to the TADA
Harmonization Reference Table. The purpose of the function is to make
similar data consistent and therefore easier to compare and analyze.

Users can also edit the reference file to meet their needs if desired.
The download argument can be used to save the harmonization file to your
current working directory when download = TRUE, the default is download
= FALSE.

Optional outputs include:

1.  the dataset with Harmonization columns appended,

2.  the dataset with CharacteristicName, ResultSampleFractionText,
    MethodSpecificationName, and ResultMeasure.MeasureUnitCode converted
    to TADA standards or

3.  the four fields converted with most Harmonization Reference Table
    columns appended. Default is transform = TRUE and flag = TRUE.

Here are some examples of how the HarmonizeData function can be used:

1.  **ResultSampleFractionText** specifies forms of constituents. In
    some cases, a single **CharacteristicName** will have both "Total"
    and "Dissolved" forms specified, which should not be combined. In
    these cases, each CharacteristicName and ResultSampleFractionText
    combination is given a different identifier. This identifier can be
    used later on to identify comparable data groups for calculating
    statistics and creating figures for each combination.

2.  Some variables have different names but represent the same
    constituent (e.g., "Total Kjeldahl nitrogen (Organic N & NH3)" and
    "Kjeldahl nitrogen"). The HarmonizeData function gives a consistent
    name (and identifier) to synonyms.

```{r}
UniqueHarmonizationRef <- HarmonizationRefTable(TADAProfileClean16, download = FALSE)
TADAProfileClean17 <- HarmonizeData(TADAProfileClean16, ref = UniqueHarmonizationRef, transform = TRUE, flag = TRUE)
```
