---
title: "WQX 3.0 Migration"
always_allow_html: yes
author: "TADA Team"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    self_contained: false
    toc: true
    fig_caption: yes
    fig_height: 8
    fig_width: 8
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{EPATADA}
  %\VignetteIndexEntry{WQX 3.0 Migration}
  %\VignetteEngine{knitr::rmarkdown}
description: An overview of TADA Module 2 functions and a recommended workflow focused on A) creating an ATTAINS assessment unit/WQP monitoring location crosswalk, B) reviewing assessment units and monitoring locations on a map, C) submitting an up to date assessment unit/WQP monitoring location crosswalk to ATTAINS, and D) assigning uses to assessment units.
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include = F}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  eval = T,
  dev.args = list(bg = "transparent")
)
```

## WQX 3.0 data is now live on the Water Quality Portal Beta

This vignette walks through how to download WQX 3.0 data from the WQP
and rename the columns back to their WQX 2.0 equivalents. Check out the
[WQX3.0 Data Now Available on the Water Quality
Portal](https://waterdata.usgs.gov/blog/wqx3/) blog for more information
about the new WQX 3.0 profiles and WQP beta web services! In addition,
USGS provides a summary of the current status of WQP functions in
dataRetrieval
[here](https://doi-usgs.github.io/dataRetrieval/articles/Status.html).

# Overview and Setup

**Welcome!**

Thank you for your interest in Tools for Automated Data Analysis (TADA).
TADA is an open-source tool set built in the R programming language.
This [RMarkdown](https://yihui.org/rmarkdown/) document walks users
through how to download the TADA R package from GitHub, access and
parameterize several important functions, and create basic
visualizations with a sample data set.

**Note:** TADA is still under development. New functionality is added
weekly, and sometimes we need to make bug fixes in response to tester
and user feedback. We appreciate your feedback, patience, and interest
in these helpful tools.

If you are interested in contributing to TADA development, more
information is available at:

[Contributing](https://usepa.github.io/EPATADA/articles/CONTRIBUTING.html)

We welcome collaboration with external partners.

## Install and Load the EPATADA R Package

First, install and load the remotes package specifying the repo. This is
needed before installing EPATADA because it is only available on GitHub
(not CRAN).

```{r install_remotes, results = 'hide'}
install.packages("remotes", repos = "http://cran.us.r-project.org")
library(remotes)
```

Next, install and load EPATADA using the remotes package. USGS's
dataRetrieval and other TADA R Package dependencies will also be
downloaded automatically from CRAN with the TADA install. If desired,
the development version of dataRetrieval can be downloaded directly from
GitHub (un-comment).

```{r install_TADA, eval = F, results = 'hide'}
remotes::install_github("USEPA/EPATADA",
  ref = "develop",
  dependencies = TRUE
)
# remotes::install_github("USGS-R/dataRetrieval", dependencies=TRUE)
```

Finally, use the **library()** function to load the TADA R Package into
your R session.

```{r library, results = 'hide'}
library(EPATADA)
```

## Retrieve WQX 3.0 Data from WQP Beta

Let's use USGS's dataRetrieval R package to query the [WQP
beta](https://www.waterqualitydata.us/beta/) web services and retrieve
the new WQX 3.0 full physical chemical profile.

```{r WQPquery, eval = F}
wqx3_fullPhysChem <- dataRetrieval::readWQPdata(
  statecode = "Illinois",
  countycode = "DeWitt",
  characteristicName = "Nitrogen",
  service = "ResultWQX3",
  dataProfile = "fullPhysChem",
  ignore_attributes = TRUE
)
```

```{r use-example-data, include = F}
utils::data("wqx3_fullPhysChem", package = "EPATADA")
```

We use the `readWQPdata` and `whatWQPsites` functions from the USGS
`dataRetrieval` R package within `TADA_DataRetrieval`. Currently,
`TADA_DataRetrieval` supports only WQX 2.2 (most recent legacy
profiles). We are evaluating the new WQX 3.0 beta services and planning
to transition the EPATADA R package to utilize these services soon.

## Review Column Name Changes

To facilitate this transition, we have developed a function that
converts all column headers in a WQX 3.0 profile back to their WQX 2.2
equivalents. This enables us to effectively test for backward
compatibility. This new EPATADA function, `TADA_RenametoLegacy`,
leverages a crosswalk file that is available on the [Water Quality
Portal Quick Reference
Guide](https://www.epa.gov/waterdata/water-quality-portal-quick-reference-guide).

```{r WQX3schemafile, eval = F}
wqx3schema <- readr::read_csv("https://www.epa.gov/system/files/other-files/2025-07/schema_outbound_wqx3.0.csv", show_col_types = FALSE)
```

After retrieving the WQX 3.0 full physical chemical profile, we can use
`TADA_RenametoLegacy` to change the column names back to WQX 2.2 where
applicable.

```{r rename}
wqx3_legacynames <- EPATADA::TADA_RenametoLegacy(wqx3_fullPhysChem)
```

Now, we can test other EPATADA package functions as usual, for example,
we can run `TADA_AutoClean`:

```{r autoclean}
wqx3_legacynames <- TADA_AutoClean(wqx3_legacynames)
```

The WQP and WQX teams (USGS and EPA) are interested in hearing feedback
from users on the new WQX 3.0 schema (data profiles) and WQP beta web
services (Contact WQX Helpdesk
[WQX\@epa.gov](mailto:WQX@epa.gov){.email}). In addition, please reach
out if you run into any issues or questions related to the WQX 2.2 to
3.0 crosswalk file (available on the [Water Quality Portal Quick
Reference
Guide](https://www.epa.gov/waterdata/water-quality-portal-quick-reference-guide))
we developed to assist you with your transition to the new services or
data profiles. This EPATADA R package function, , `TADA_RenametoLegacy`,
is provided for WQX 3.0 schema and WQP beta service testing purposes
only. We plan to transition the EPATADA R Package to use the WQX 3.0
schema and new WQP services soon. Stay tuned.
